### 2024-04-30 NEURIPS24/appendix.tex
update paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Proofs}
\label{appendix:proofs}
Denote by $\msg(s)\triangleq \{ (t,s(t)): \eqsp t\in \msi \} $ the graph of a time series $s: \msi \to \Rset^d$ and $ \phi.\msg(s)\triangleq\{ \phi(t,s(t)): \eqsp t\in \msi\} $ the action of  $\phi\in \mcd(\Rset^{d+1}) $ on $\msg(s)$.
\begin{theorem}
    \label{theorem:representation_proof}
Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi,\msj$ two intervals of $\Rset$.
 There exist $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\gamma\in  \mcd(\Rset) $ such that $\gamma(\msi)=\msj $ and $\Phi_f\in \mcd(\Rset^{d+1})$,
 \begin{equation}% TO add more detail on functionnal space
    \msg(s)= \Pi_{\gamma,f}.\msg(\mathbf{s}_0),\eqsp \Pi_{\gamma,f}=\Psi_\gamma\circ\Phi_f.
 \end{equation}
 Moreover, for any $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\bar{\gamma}\in  \mcd(\Rset) $, there exists a continously differentiable time series $\bar{s}$ such that 
 $\msg(\bar{s})= \Pi_{\bar{\gamma},\bar{f}}.\msg(\mathbf{s}_0)$
\end{theorem}
%\sam{preuve todo, classement des variétés à bord, donne un homeomorphisme sur la restriction}
\begin{proof}
  Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi=(a,b),\msj=(\alpha,\beta)$ two intervals of $\Rset$.
  By setting $\gamma: t\in \Rset \mapsto (\beta-\alpha)(t-a)/(b-a)+\alpha\in \Rset $, we have $ \gamma(\msi)=\msj$ and $\gamma \in \mcd(\Rset) $.
   By defining $f:(t,x)\in\Rset^{d+1}\mapsto x-\mathbf{s}_0(t)+s\circ \gamma(t) $, the map $\Phi_f\in \mcd(\Rset^{d+1})$,
    indeed, its inverse is $\Phi_f^{-1}:(t,x)\in\Rset^{d+1}\mapsto (t,x+\mathbf{s}_0(t)-s(t)) $ and is continuously differentiable.
     Moreover, we have $\Pi_{\gamma,f}.\msg(\mathbf{s}_0)=\{(\gamma(t),s\circ \gamma(t)):\eqsp t\in\msi \}=\msg(s) $.

    Let $\bar{f}\in \rmC^0(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^0(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
    We have :
    \begin{align}
      \Pi_{\gamma,f}.\msg(\mathbf{s}_0)&=\{(\gamma(t),f(t,\mathbf{s}_0(t))),\eqsp t\in \msi \} \\
      &\label{eq:proof1_last_eq}=\{(t,f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right),\eqsp t\in \gamma(\msi) \} \eqsp .
    \end{align}
    By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^0(\gamma(\msi), \Rset^d) $ by composition of continuous functions
    and $ \msg(\bar{s})= \Pi_{\gamma,f}.\msg(\mathbf{s}_0)$ by \eqref{eq:proof1_last_eq}, which concludes the proof.
\end{proof}
\begin{lemma}
  If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0},
   there exist $\gamma \in \msd(\Rset) $ and $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ such that $\phi^v=\Psi_\gamma\circ\Phi_f $.
\end{lemma}
\begin{proof}
  Let $v$ be a vector field generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0}.
 We remark that the first coordinate of the velocity field $v_\tau$ denoted by $v_\tau^{\text{time}}$ only depends on the time variable $t$ for any $\tau\in[0,1]$.
 Thus, when computing the first coordinate of the deformation $\phi^v$, denoted by $\gamma$, we integrate \eqref{eq:LDDMM_dynamic} with $v_\tau$ replaced by $v_\tau^{\text{time}}$,
  thus $\gamma$ is independant of the variable $x$. Moreover, $\gamma\in \mcd(\Rset)$ since a Gaussian kernel induced an Hilbert space $\msv$ satisfying $|f|_V\leq |f|_\infty+ |\dd f|_\infty  $ for any $f\in \msv$ by \citep[Theorem 9]{glaunes2005transport}.
  For the same reason, we have $\phi^v\in \mcd(\Rset^{d+1})$, and thus its last coordinates denoted by $f$ belongs to $\rmC^1(\Rset^{d+1},\Rset^d)$, and by construction $\phi^v=\Psi_\gamma\circ\Phi_f $.
\end{proof}

\section{Oriented varifold}

\label{appendix:varifold}
In this section, we introduce the \textit{oriented varifold} associated with curves.
For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}. 
We associate to $\gamma\in \rmC^1((a,b),\Rset^{d+1})$ an \textit{oriented varifold} $\mu_\gamma$, i.e. a distribution on the space $\Rset^{d+1}\times \mathbb{S}^{d}$ defined as follows, for any smooth test function $\omega :\Rset^{d+1}\times \mathbb{S}^{d}\to \Rset  $,
\begin{equation}
  \mathbb{E}_{Y\sim \mu_\gamma}\left[\omega(Y)\right]=\mu_\gamma(\omega)=\int_a^b \omega\left(\gamma(t),\frac{\dot{\gamma}(t)}{|\dot{\gamma}(t)|}\right)|\dot{\gamma}(t)|\dd t \eqsp .
\end{equation}
Denoting by $\msw$ the space of smooth test function, we have that $\mu_\gamma $ belongs to its dual $\msw^*$.
 Thus, a distance on $\msw^*$ is sufficient to set a distance on oriented varifolds associated to curve and thus on $\rmC^1((a,b),\Rset^{d+1})$ by the identification $\gamma\to \mu_\gamma $.
Remark that in (TS-LDDMM), $\gamma$ should be the parametrization of a time series' graph $\msg(s)$, i.e. $\gamma: t\in \msi \to (t,s(t))\in\Rset^{d+1} $ denoting by $s:\msi \to \Rset^d$ the time series.
However, in practice, we work with discrete objects.
 That is why, we set $W$ as an RKHS to use its representation theorem.
 More specifically \citep[Proposition 2 \& 4]{kaltenmark2017general} encourages us to consider a kernel $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$ such that there exist two positive and continuously differentiable kernels $k_{\pos}$ and $k_{\dir}$, 
 such that for any $(x,\overrightarrow{u}),(y,\overrightarrow{v}) \in (\Rset^{d+1} \times \mathbb{S}^d)^2$
 \begin{equation}
     k((x,\overrightarrow{u}),(y,\overrightarrow{v})) = k_{\pos}(x,y)k_{\dir}(\overrightarrow{u},\overrightarrow{v}) \eqsp,
   \end{equation}
 with moreover $k_\dir>0$ and $ k_\pos$ which admits an RKHS $\msw_\pos$ dense in the space of continous function on $\Rset^{d+1}$ vanishing at infinite \cite{carmeli2010vector}.

Given such a kernel $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$
verifying \citep[Proposition 2 \& 4]{kaltenmark2017general}, we have that for any $(x,v)\in \Rset^{d+1}\times \mathbb{S}^{d}$, $\delta_{(x,\overrightarrow{v})}$ belongs to $\msw^*$ as a distribution and that the dual metric $\langle\cdot,\cdot \rangle_{\msw^*} $ satisfies for any $(x_1,v_1),(x_2,v_2)\in \left(\Rset^{d+1}\times \mathbb{S}^{d}\right)^2$,
\begin{equation}
  \langle\delta_{(x_1,\overrightarrow{v}_1)},\delta_{(x_2,\overrightarrow{v}_2)} \rangle_{\msw^*} =k((x_1,\overrightarrow{v}_1),(x_2,\overrightarrow{v}_2)) \eqsp .
\end{equation}
Thus, given two sets of triplets $X=(l_i,x_i,\overrightarrow{v}_i)_{i\in[T_0-1]}\in (\Rset\times \Rset^{d+1}\times \mathbb{S}^d)^{T_0-1},Y=(l_i',y_i,\overrightarrow{w}_i)_{i\in[T_1]}\in (\Rset\times \Rset^{d+1}\times \mathbb{S}^d)^{T_1-1}$ and denoting by 
\begin{equation}
  \label{eq:def_mu_X}
  \mu_X=\sum_{i=1}^{T_0} l_i \delta_{(x_i,\overrightarrow{v}_i)},\mu_Y=\sum_{i=1}^{T_1} l_i' \delta_{(y_i,\overrightarrow{w}_i)} \eqsp,
\end{equation}
 we have,
\begin{equation}
   |\mu_X-\mu_Y |_{\msw^*}^2 = \sum_{i,j = 1}^{T_0-1}l_i k((x_i,\overrightarrow{v_i}),(x_i,\overrightarrow{v_i}^0))l_j - 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l_i k((x_i,\overrightarrow{v_i}),(y_i,\overrightarrow{w_i}))l_j' + \sum_{i,j = 1}^{T_1-1}l_i' k((y_i,\overrightarrow{w_i}),(y_i,\overrightarrow{w_i}))l_j' \eqsp.
\end{equation}
Then, using the identification $X\to \mu_X, Y\to \mu_Y$, we can define a distance on sets of triplets as $d_{\msw^*,3}(X,Y)=|\mu_X-\mu_Y|_{\msw^*}^2$.

Now, we aim to discretize the oriented varifold $\mu_\msg $ related to a time series' graph $\msg(s)$ by using a set of triplets.
This is carried out by using a discretized version of $\msg(s) $, i.e. $\tilde{\msg}=(g_i=(t_i,s(t_i)))_{i\in[T]}\in(\Rset^{d+1})^T$, in the following way: 
For any  $i\in[T-1]$, denoting the center and length of the $i^{th}$ segment $[g_i,g_{i+1}]$ by
$c_i = (g_i + g_{i+1})/2$, $l_i = \| g_{i+1}-g_{i}\|$, and the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$ by
 $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$, we define the set of triplets $X(\tilde{\msg})=(l_i,c_i,\overrightarrow{v_i})_{i\in[T-1]}$ and its related oriented varifold $\mu_{X(\tilde{\msg})}= \sum_{i=1}^{T-1}l_i \delta_{c_i,\overrightarrow{v_i}}$ as in \eqref{eq:def_mu_X}.
 This is a valid discretization of the oriented varifold $\mu_\msg$ according to \citep[Proposition 1]{kaltenmark2017general}: $\mu_{X(\tilde{\msg})}$ converges towards $\mu_\msg$ as the size of the descretization mesh $\sup_{i\in[T-1]} |t_{i+1}-t_i| $ converges to 0.

 Finally, we define a distance on discretized time series' graphs $\tilde{\msg}_1,\tilde{\msg}_2$ as $d_{\msw^*}(\tilde{\msg}_1,\tilde{\msg}_2)=d_{\msw^*,3}(X(\tilde{\msg}_1),X(\tilde{\msg}_2)) $.

\subsection{Varifold kernels}
\label{appendix:kernel_implementation}
Denote the one-dimensional Gaussian kernel by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
In the implementation, we use the following kernels, for any $((t_1,x_1),(t_2,x_2))\in (\Rset^{d+1})^2, ((w_1,v_1),(w_2,v_2))\in (\mathbb{S}^{d})^2 $,
\begin{equation}
  k_{\pos}(x,y)=K_{\sigma_{\pos,t}}^{(1)}(t_1,t_2)K_{\sigma_{\pos,x}}^{(d)}(x_1,x_2),\quad k_{\pos}(x,y)=K_{\sigma_{\dir,t}}^{(1)}(w_1,w_2)K_{\sigma_{\dir,x}}^{(d)}(v_1,v_2)\eqsp ,
\end{equation}
where $\sigma_{\pos,t}, \sigma_{\pos,x}, \sigma_{\dir,t}, \sigma_{\dir,x}>0 $ are hyperparameters.
 In practice, we select $\sigma_{\pos,x}\approx \sigma_{\dir,x} \approx 1$ when the times series are centered and normalized. Otherwise we select $\sigma_{\pos,x}\approx \sigma_{\dir,x} \approx \bar{\sigma}_{s}$ with $\bar{\sigma}_{s}$ the average standard deviation of the time series.
  We choose $\sigma_{\pos,t}\approx \sigma_{\dir,t} = m f_e$ with $f_e$ the sampling frequency of the time series and $m\in [5]$ an integer depending on the time change between the starting and the target time series graph.
  The more significant the time change, the higher $m$ should be. The intuition comes from the fact that the width $\sigma_{\pos,t}, \sigma_{\dir,t}$ rules the time windows used to perform the comparison, and $\sigma_{\pos,x}, \sigma_{\dir,x}$ affects the space window.
   The size of the windows should be selected depending on the variations in the data.

\section{Tuning the hyperparameters of the TS-LDDMM kernel given in \eqref{eq:kernel_TAS}}
\label{appendix:kernel_TS_LDDMM}
The parameter $\sigma_{T,0}$ should be chosen \textit{large} compared the sampling frequency $f_e$ and compared to average standard deviation $\bar{\sigma}_s$ of the time series, e.g $\sigma_{T,0}=100$ as $\bar{\sigma}_s\approx f_e\approx  1 $.
It makes the time transformation smoother. If $\sigma_{T,0}$ is too small, for instance, $\sigma_{T,0}=f_e $, the effect of the time deformation is too localized, and there are not enough samples to make it visible.

The parameter $\sigma_{T,1}$ should be of the same order as $f_e$: two different points in time can have various space transformations.
 $\sigma_x$ should be of the same order of $\bar{\sigma}_s$: two points with a big difference regarding space compared to $\bar{\sigma}_s$ can have very different space transformations.

 We take $c_0\approx 10 c_1 $, we want to encourage time transformation before space transformation. We take $(c_0,c_1)=(1,0.1)$ in all experiments.

\section{Numerical details}
A report of all the hyperparameters selected is given in \Cref{appendix:table:set_up_exp_practical}.
\subsection{Optimization details of \eqref{eq:general_optimization_problem}}
\label{appendix:optimizers_details}
\paragraph{Initialization}
At the initialization of \eqref{eq:general_optimization_problem}, all the momentums parameter are set to $0$ and the graph of reference is set to the graph of a time series in the dataset having a mediane samples size.

\paragraph{Gradient descent.}
The chosen gradient descent method is "adabelief" \cite{zhuang2020adabelief} implemented in the library OPTAX\footnote{https://optax.readthedocs.io/en/latest/}.
There are two main parameters in the gradient descent: the number of steps nb\_steps, and the maximum value of step size $\eta_M$.
The stepsize has a particular scheduling: 
\begin{itemize}
  \item Warmup period on $0.1 \times$ nb\_steps steps: the stepsize increases linearly from $0$ to $\eta_M$. The goal is to learn progressively the parameters.
   If the stepsize is too large at the start, smaller steps at the end can't make up for the mistakes made at the beginning. 
  \item Fine tuning periode on $ 0.9  \times$ nb\_steps : the stepsize decreases from $\eta_M$ to $0$ with a cosine decay implemented in the OPTAX scheduler,
   i.e. the decreasing factor as the form $0.5  (1 + \cos(\pi  t/T))$. 
\end{itemize}
The sharper the deformations, the larger the number of steps and the maximum value of step size should be selected.
We suggest nb\_steps=300, $\eta_M=0.1 $ for small deformations and nb\_steps=800, $\eta_M=0.3 $ for big ones (time dilation with a factor $\lambda\geq 2$).

\subsection{Synthetic experiments}
\label{appendix:numerics_synthetic}
For any deformations generation in both experiments (well-specified and misspecified), we take $\sigma_{T,0},\sigma_{T,1},\sigma_x=(100,1,1)$ and $c_0,c_1=(1,0.1)$ for the kernel $K_\msg$
and $\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}=(2,1,2,0.6) $ for the varifold kernels $k_\pos,k_\dir$ related to the loss $\scrl$.

In both experiments, we have nb\_steps=300 abd $\eta_M=0.1$.

\subsection{Mouse experiments}
The number of steps is larger in the second experiment (before/after injection) because the deformations are sharper.

\subsection{Classification experiments}
\label{appendix:classification_implementation}

We defined a default parametrization for all classifiers.

For classifiers: \href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.deep_learning.CNNClassifier.html#aeon.classification.deep_learning.CNNClassifier}{CNN}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.networks.ResNetNetwork.html#aeon.networks.ResNetNetwork}{ResNet}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.feature_based.Catch22Classifier.html#aeon.classification.feature_based.Catch22Classifier}{Catch22}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.distance_based.KNeighborsTimeSeriesClassifier.html#aeon.classification.distance_based.KNeighborsTimeSeriesClassifier}{DTW-KNN}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.convolution_based.RocketClassifier.html#aeon.classification.convolution_based.RocketClassifier}{Rocket}
we used the aeon\footnote{https://www.aeon-toolkit.org/en/stable/index.html} implementations with their default settings.

For Tloss-SVC we used the implementation provided on github\footnote{https://github.com/mqwfrog/ULTS} 
with the following parameters for learning representations: batch\_size: 10, channels: 40, depth: 10, nb\_steps: 200, in\_channels: 1, 
kernel\_size: 3, lr: 0.001, nb\_random\_samples: 10, negative\_penalty: 1, out\_channels: 320, reduced\_size: 160. 
We used the Support Vector Classifier (\href{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html#sklearn.svm.SVC}{SVC})
from scikit-learn with thee regularization term C: 1. Others parameters are set to default.

For TS-LDDMMM-SVC, all kernels' parameters et optimizer parameter are presented in \Cref{appendix:table:set_up_exp_practical}.
As well, we used the Support Vector Classifier from scikit-learn with thee regularization term C: 1. Others parameters are set to default.

 \begin{table}[hbt!]
  \caption{Parameters used in all the experiments.
   For synthetic data, $K_\msg$ refers to the kernel used in the generation, which is the same for the estimation only in the well-specified case. $\bar{l}$ refers to the average time series length and $N_d$ refers to the number of dimensions.}
  \centering
  \begin{tabular}{lllll}

            \toprule 
            objects & Optimizer &$k_\pos,k_\dir$ & $K_\msg $ \\
            \midrule
            Parameter & (nb\_steps,$\eta_M$) &$(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$ & $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)$ \\
            \midrule
            Synthetic data well-specified &(300,0.1) &$(2,1,2,0.6)$& $(1,0.1,100,1,1)$   \\
            Synthetic data misspecified  & (300,0.1)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
            Mouse before injection  & (400,0.3)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
            Mouse before/after injection  & (800,0.3)  & $ (5,1,5,0.6)$ &$(1,0.1,150,1,1)$  \\
            classification  & (400,0.1)  & $(\max(2,0.03\bar{l}),N_d,\max(2,0.03\bar{l}),0.6)$  & $(1,0.1,0.33\bar{l},1,N_d)$  \\

            \bottomrule
          \end{tabular}
        \label{appendix:table:set_up_exp_practical}
    \end{table}
\section{Mouse respiratory dataset}

\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{pictures/mice_exp.png}
  \caption{A: Illustration of a double-chamber plethysmograph. The term \textit{dpt} stands for differential 
  pressure transducer which measures the pressure in each compartment, the pressure then being converted to flow. 
  B: Nasal airflow (top) and lung volume (bottom). During inspiration, airflow is positive (grey) and during
  expiration, airflow is negative (pink).}
  \label{fig:mice_exp}
\end{figure}

\label{appendix:mouse_dataset}
Ventilation is a simple physiological function that ensures a vital supply of oxygen and the elimination of CO2. 
Acetylcholine (Ach) is a neurotransmitter that plays an important role in muscular activity, notably for breathing. 
Indeed, muscle contraction information passes from the brain to the muscle through the nervous system. Achs are located 
in synapses of the nervous system (central and peripheral) and skeletal muscles. They ensure the information transmission 
from nerve to nerve. However, the transmission cannot end without the hydrolysis of Ach by the enzyme Acetylcholinesterase 
(AchE), allowing nerves to return to their resting state. Inhibition of (AchE) with, for instance, nerve gas, pesticide, 
or drug intoxication leads to respiratory arrests.

The dataset comes from the experiment \cite{nervo2019respiratory}, where they studied the consequences of partial 
deficits in AChE and AChE inhibition on mice respiration. AchE inhibition was induced with an 
irritant molecule called physostigmine (an AchE inhibitor). Mice nasal airflows were sampled at 
2000Hz with a Double Chamber plethysmograph \cite{hoymann2012lung}, as depicted in \Cref{fig:mice_exp}-A). The flow is expressed in 
$ml.s^{-1}$; it has a positive value during inspiration and a negative value expiration \Cref{fig:mice_exp}-B). 
Among the mice population, we selected 7 control mice (\textbf{wt}) and 7 ColQ mice (\textbf{colq}), which do not have 
AChE anchoring in muscles and some tissues. 
As described in \cite{nervo2019respiratory}, mice experiments were as follows:
\begin{enumerate}
  \item The mouse is placed in a DCP for 15 or 20 min to serve as an internal control.
  \item The mouse is removed from the DCP and injected with physostigmine.
  \item The mouse is placed back into the DCP, and its nasal flow is recorded for 35 or 40 min.
\end{enumerate}

Respiratory cycles were extracted following procedure \cite{germain2023unsupervised}. We removed 
respiratory cycles whose duration exceeds 1 second; the average respiratory cycle duration is 
300 ms. We randomly sampled 10 respiratory cycles per minute and mouse. It leads to a dataset of 
12,732 (time, genotype)-annotated respiratory cycles.

\section{Classification datasets}
\label{appendix:classification_dataset}

All datasets were taken from UCR/UEA archives \cite{dau2019ucr,bagnall2018uea}. Among all available 
datasets\footnote{https://timeseriesclassification.com}, we selected 4 datasets related to time series 
shape comparison. All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/} 
which already includes the train test split. Essential dataset information is summarized in \Cref{appendix:table:datasets}.

\begin{table}[hbt!]
  \centering
  \caption{Time series datasets summary for shape based classification.}
  \begin{tabular}{lllllll}
    \toprule
    Dataset & Train size & test size & Lenghth & Number of classes & Number of dimensions & Type \\
    \midrule
    \href{https://timeseriesclassification.com/description.php?Dataset=ArrowHead}{ArrowHead} & 36 & 175 & 251 & 3 & 1 & IMAGE \\
    \href{https://timeseriesclassification.com/description.php?Dataset=ECG200}{ECG200} & 100 & 100 & 96 & 2 & 1 & ECG \\
    \href{https://timeseriesclassification.com/description.php?Dataset=GunPoint}{GunPoint} & 50 & 150 & 150 & 2 & 1 & MOTION \\
    \href{https://timeseriesclassification.com/description.php?Dataset=NATOPS}{NATOPS} & 180 & 180 & 51 & 6 & 24 & MOTION \\
    \bottomrule
  \end{tabular}
  \label{appendix:table:datasets}
\end{table}

%  [11:35, 31/01/2024] Thibaut Germain: avant/après :
%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,150,1,1)
%  Kl = TSGaussGaussKernel(5,1,5,0.6)
%  dataloss = VarifoldLoss(Kl)
%  schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
%  [11:35, 31/01/2024] Thibaut Germain: avant :
%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,100,1,1)
%  Kl = TSGaussGaussKernel(2,1,2,0.6)
%  dataloss = VarifoldLoss(Kl)
%  schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
% Let $s$ be a time series and $\msg(\tilde{s}) = (g_i)_{i \in [n]}$ be its discretized graph.
%  The graph $\msg(s)$ can be approximated as 
% the union of piecewise linear segments between time-consecutive samples, 
% and we associate to the approximated graph the oriented varifold:

% \begin{equation}
%   \mu_{\msg(\tilde{s})} : w \in \msw \mapsto \sum_{i=1}^{n-1} l_i\delta_{(c_i,\overrightarrow{v_i})}(w)
% \end{equation}
% where $\msw$ is a set of real valued test functions defined on $\Rset^{d+1} \times \mathbb{S}^d$, 
% $\delta_{(c_i,\overrightarrow{v_i})}(w) = w(c_i,\overrightarrow{v_i})$ is a Dirac delta function, 
% $c_i = (g_i + g_{i+1})/2$ (resp. $v_i = \| g_{i+1}-g_{i}\|$) is the center (resp. length) of the $i^{th}$ segment, 
% and $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$ is the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$.

% Assuming that the test functions' space $\msw$ is the RKHS associated with a $C^1$ positive definite kernel $k$ 
% that verifies \citep[Proposition 2 \& 4]{kaltenmark2017general}, the oriented varifold belongs to the dual space $\msw^*$. Additionally, 
% we can define a distance $d_{\msw^*}$ between signals' graph sample sets, which, thanks to the kernel reproducing property,
% has an explicit formulation:

% \section{Continuous normalizing flows}
% %Modeling Continuous Stochastic Processes with Dynamic Normalizing Flows time series generative
% CONTUNYIYS NORMALIZING FLOW bofbof
% Here, we present the framework of Continuous Normalizing Flows (CNF) and show how LDDMM can be seen as a particular case.
% \cite{salman2018deep} was already inspired from the work of shape analysis to present deep diffeomorphic normalizing flows.
% Discussion with Alain :
% Given a time-dependant vector field $v:[0,1]\times \Rset^{d'} \to \Rset^{d'} $, the integration of this vector field gives a flow $ \phi:[0,1]\times \Rset^{d'} \to \Rset^{d'}$, defined via the ordinary differential equation (ODE),
% \begin{equation}
%   \frac{\dd }{\dd \tau}\phi_\tau(x)=v(\tau,\phi_\tau(x)), \phi_0(x)=x \eqsp.
%  \end{equation}
% In \cite{chen2018neural}, they introduce Continous Normalizing flows for generative modeling.
%  By modeling $v=f_\theta$ is a neural networks and choosing a prior $X_0\sim\rho_0$ on the inital value, the law of $\phi_1(X_0)$ has the pullback density $\phi_1\# \rho_0$ and the parameter $\theta$ is optimized to maximize the log-likelihood of a given dataset $(y_i)_{i\in[N]}\in (\Rset^{d'})^N$ which is assumed to follow the law of $\phi_1(X_0)$.
% Instead of modeling the law of the end point $\phi_1(X_0) $, we can try to model the law of the whole path $(\phi_\tau(X_0))_{\tau\in[0,1]} $ having a density $p:[0,1]\times \Rset^d\to \Rset_{>0}$ such that $p_1=q, p_0=\rho_0$ where $q$ is the empirical distribution in the dataset, as proposed in \cite{lipman2022flow}.
% In so doing, the problem is ill-posed since $p_t$ can be whatever as soon as $0<t<1$, that is why we try to constraint the learned velocity field $v=f_\theta$ such that the path law $p$ is tractable.
% One way of doing it is to look for geodesics \eqref{eq:geodesics_original} as proposed in LDDMM, it corresponds to the case where $q=\sum_{i=1}^N \delta_{y_i}/N$ and $\rho_0=\sum_{i=1}^N \delta_{x_i}/N$ are empirical distribution and $y_i=\phi^v_1(x_i) $ for a given velocity field $v$.

% The goal is to target $\nu\in \mathcal{P}(\rmC^0([0,1]\times \Rset^d, \Rset^d)) $ by generating randomness on a initial value $x \sim \delta_x$
%  and by integrating a velocity field $v_\theta \in \msw \subset \rmC^0([0,1]\times \Rset^d, \Rset^d)$: 
%  \begin{equation}
%   \frac{\dd h}{\dd \tau}=v_\theta(\tau,h(\tau,x)), h(0,x)=x \eqsp,
%  \end{equation}
%  denoting by $\mu$ the law of $(h(\tau,x))_{\tau \in [0,1]}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \subsection{TTS}
% To circumvant this issue, we propose two methods :
%       \begin{enumerate}
%         \item \label{enum:first_method} [Time And Space (TAS)] Using a specific form of the RKHS's kernel in LDDMM, to stick to the classical shape analysis framework.
%         \item \label{enum:second_method} [Time Then Space (TTS)] Tackling the time axis and space axis separetely in the transformation.
%          This distangles time and space variability in the patterns, increasing the interpretability of the representation.
%           Although the related optimization is more complex than method \ref{enum:first_method}, the results is more robust to the choice of hyper-parameter related to the RKHS's kernel.
%       \end{enumerate}
%        In the following, we present both methods to highlights their pro and cons.

% As described in \Cref{section:optimization}, the method (TAS) learns $\Pi_{\gamma,f}$ directly, while (TTS) find the temporal reparametrisation $\Phi_f$ then the space transformation $\Psi_\gamma$.
%     % If we define a diffeomorpshim $\Phi_\gamma$ as $(\gamma,\Id_d) $, we have $\Phi_\gamma.G(s) =G(s).\gamma^{-1}$.
%     %  To increase the variability, we allow some deformations on the graph of the time series after being reparametrized in time, which allow us to recover every possible time seriess.
%     % Remark that, $G(\mathbf{s}_0\circ \psi^{-1})=\{(t,\mathbf{s}_0\circ \psi^{-1}(t)), t\in \msj \}= \{(\psi(t),\mathbf{s}_0(t)) ,\eqsp t\in \msi \}  $.
%     % [The action of the diffeomorpshim group 
%     % \begin{align}
%     %     &\mcd_*=\mcd(\Rset,\Rset)\times \{\phi\in \mcd(\Rset^{d+1},\Rset^{d+1}): \\
%     %     &\phi(t,x)=(t,f(t,x)),
%     %      (t,x)\in \Rset \times \Rset^{d}, f \in C^0(\Rset^{d+1}, \Rset^d) \} 
%     % \end{align}
%     % on the space of time series graph $\{G^{(a,b)}(s), s\in C^0((a,b), \Rset), (a,b)\in \Rset_+^2, a<b  \}$
%     % is homogene.]
%      %In the next section we define a metric and a norm on this space in order to find a representation $(\phi,\psi)$ with a minimal norm, and we show in (...) how to recover a proxy.
%      \paragraph{The representation as the solution of an optimization}
%      By \Cref{theorem:representation}, for any $\mathbf{s}_0\in C^0(\msi,\Rset^d)$, we can find $ F=(\gamma_j,f_j)_{j\in[N]}$ such that $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$ for any $j\in[N]$.
%      However, to get a meaningfull representation and to tackle the unicity problem, we should optimize the choice of $\mathbf{s}_0 $ and $F$, formally speaking, denoting by $R(\cdot) $ a regularization norm on $\mcd(\Rset^{d+1})$ and $L$ a loss on sets which will be specify later, we aim to solve
%      \begin{equation}
%       \label{eq:minimization}
%       \underset{\mathbf{s_0},(\gamma_j,f_j)_{j\in[N]}}{ \argmin} \sum_{j=1}^N \lambda R(\Pi_{\gamma_j,f_j})+ L(\msg(s^j),\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)) \eqsp ,
%     \end{equation} 
%     where $\lambda>0$ is a penalization factor related to $L$ to relax the condition $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$.

%     In practice, as already proposed in shape analysis, we perform an alternative minimization between $\mathbf{s_0}$ and $(\gamma_j,f_j)_j$.
%     The optimization on $(\gamma_j,f_j)_j$ is parrallelisable, for each couple $(\mathbf{s}_0,s^j)$, we should learn diffemorphisms of minimal norm satisfying a condition. 
%     To this end, we adapt the geodesic shooting method using LDDMM presented in \cite{durrleman2013sparse} to our special case.
%       Before to describe how to optimize the whole in \Cref{section:optimization}, we introduce LDDMM in the next section to specify the norm on diffemorphisms.

% \paragraph{Kernel choices}
% As depicted on figure (mettre ref), we can not use any kernel $K$ to apply the previous methodology to learn deformations on time series' graphs.

% In the following, we expose two choice of RKHS's kernel which answers this issue : (TAS) and (TTS).
%  We rely on the Gaussian kernel, denoting by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
% To apply (TAS) in LDDMM with $d'=d+1$, we use the following anisotropic Gaussian kernel  :
% \begin{align}
%   \label{eq:kernel_TAS}
%   &K_{\text{TAS}}((t,x),(t',x'))=\begin{pmatrix}
%     c_0K & 0 \\
%     0 & c_1 D 
%     \end{pmatrix} \eqsp , \\
%     &D=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
% \end{align}
% parametrized by the widths $\sigma_{T,0},\sigma_{T,1},\sigma_x>0$ and the constants $c_0,c_1>0$.
%  If we use this kernel, by considering \eqref{eq:specific_form}, we remark that the first coordinate of the velocity field $v_\tau$ only depends on the time variable $t$.
%   It implies that the final deformation will have the form $\phi^v=(\gamma,f) $ with  $f\in C^1(\Rset^{d+1},\Rset^d) $, and $\gamma\in \mcd(\Rset)$
%   since the first coordinate follows a LDDMM-ODE \eqref{eq:LDDMM_dynamic}.

% To apply (TTS), we use LDDMM on the time axis with $d'=1$ by taking $K_{\text{TTS}}^{\text{time}}=K_{\sigma_{T,0}}^{(1)}$, then on the space axis with $d'=d$ by choosing
% $K_{\text{TTS}}^{\text{space}}=K_{\sigma_{T,1}}^{(1)}\times K_{\sigma_x}^{(d)} $ where the widths are the same than in \eqref{eq:kernel_TAS}. The alternate optimization will be more detailed in \Cref{section:optimization}.

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2024-04-30 NEURIPS24/neurips_2024_to_submit.tex
update paper

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

% ready for submission
\usepackage{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage[capitalize,noabbrev]{cleveref}

\title{Shapes analysis for time series.}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}
\input{header.tex}
\input{def.tex}
\newcommand\sam[1]{\textcolor{blue}{#1}}
\newcommand\thi[1]{\textcolor{orange}{#1}}
\begin{document}

\maketitle

\begin{abstract}
    Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
    Individuals have specificities depending on their genes or disease
     while sharing, on average, a similar pattern in the features of their corresponding time series. For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
    In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
    Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
    Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
     we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
     $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
     Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
     These features can be post-processed with any statistical or machine learning tools. 
     We demonstrate the advantages of our representation compared to existing methods using synthetic data and real-world examples, motivated by applications in medicine.
\end{abstract}

\input{paper.tex}

\bibliographystyle{plain}
\bibliography{ref.bib}

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
\end{ack}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\input{appendix.tex}
\section*{NeurIPS Paper Checklist}

%%% BEGIN INSTRUCTIONS %%%
The checklist is designed to encourage best practices for responsible machine learning research, addressing issues of reproducibility, transparency, research ethics, and societal impact. Do not remove the checklist: {\bf The papers not including the checklist will be desk rejected.} The checklist should follow the references and precede the (optional) supplemental material.  The checklist does NOT count towards the page
limit.

Please read the checklist guidelines carefully for information on how to answer these questions. For each question in the checklist:
\begin{itemize}
    \item You should answer \answerYes{}, \answerNo{}, or \answerNA{}.
    \item \answerNA{} means either that the question is Not Applicable for that particular paper or the relevant information is Not Available.
    \item Please provide a short (1–2 sentence) justification right after your answer (even for NA). 
   % \item {\bf The papers not including the checklist will be desk rejected.}
\end{itemize}

{\bf The checklist answers are an integral part of your paper submission.} They are visible to the reviewers, area chairs, senior area chairs, and ethics reviewers. You will be asked to also include it (after eventual revisions) with the final version of your paper, and its final version will be published with the paper.

The reviewers of your paper will be asked to use the checklist as one of the factors in their evaluation. While "\answerYes{}" is generally preferable to "\answerNo{}", it is perfectly acceptable to answer "\answerNo{}" provided a proper justification is given (e.g., "error bars are not reported because it would be too computationally expensive" or "we were unable to find the license for the dataset we used"). In general, answering "\answerNo{}" or "\answerNA{}" is not grounds for rejection. While the questions are phrased in a binary way, we acknowledge that the true answer is often more nuanced, so please just use your best judgment and write a justification to elaborate. All supporting evidence can appear either in the main paper or the supplemental material, provided in appendix. If you answer \answerYes{} to a question, in the justification please point to the section(s) where related material for the question can be found.

IMPORTANT, please:
\begin{itemize}
    \item {\bf Delete this instruction block, but keep the section heading ``NeurIPS paper checklist"},
    \item  {\bf Keep the checklist subsection headings, questions/answers and guidelines below.}
    \item {\bf Do not modify the questions and only use the provided macros for your answers}.
\end{itemize}

%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item {\bf Claims}
    \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the abstract and introduction do not include the claims made in the paper.
        \item The abstract and/or introduction should clearly state the claims made, including the contributions made in the paper and important assumptions and limitations. A No or NA answer to this question will not be perceived well by the reviewers. 
        \item The claims made should match theoretical and experimental results, and reflect how much the results can be expected to generalize to other settings. 
        \item It is fine to include aspirational goals as motivation as long as it is clear that these goals are not attained by the paper. 
    \end{itemize}

\item {\bf Limitations}
    \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper has no limitation while the answer No means that the paper has limitations, but those are not discussed in the paper. 
        \item The authors are encouraged to create a separate "Limitations" section in their paper.
        \item The paper should point out any strong assumptions and how robust the results are to violations of these assumptions (e.g., independence assumptions, noiseless settings, model well-specification, asymptotic approximations only holding locally). The authors should reflect on how these assumptions might be violated in practice and what the implications would be.
        \item The authors should reflect on the scope of the claims made, e.g., if the approach was only tested on a few datasets or with a few runs. In general, empirical results often depend on implicit assumptions, which should be articulated.
        \item The authors should reflect on the factors that influence the performance of the approach. For example, a facial recognition algorithm may perform poorly when image resolution is low or images are taken in low lighting. Or a speech-to-text system might not be used reliably to provide closed captions for online lectures because it fails to handle technical jargon.
        \item The authors should discuss the computational efficiency of the proposed algorithms and how they scale with dataset size.
        \item If applicable, the authors should discuss possible limitations of their approach to address problems of privacy and fairness.
        \item While the authors might fear that complete honesty about limitations might be used by reviewers as grounds for rejection, a worse outcome might be that reviewers discover limitations that aren't acknowledged in the paper. The authors should use their best judgment and recognize that individual actions in favor of transparency play an important role in developing norms that preserve the integrity of the community. Reviewers will be specifically instructed to not penalize honesty concerning limitations.
    \end{itemize}

\item {\bf Theory Assumptions and Proofs}
    \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (and correct) proof?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include theoretical results. 
        \item All the theorems, formulas, and proofs in the paper should be numbered and cross-referenced.
        \item All assumptions should be clearly stated or referenced in the statement of any theorems.
        \item The proofs can either appear in the main paper or the supplemental material, but if they appear in the supplemental material, the authors are encouraged to provide a short proof sketch to provide intuition. 
        \item Inversely, any informal proof provided in the core of the paper should be complemented by formal proofs provided in appendix or supplemental material.
        \item Theorems and Lemmas that the proof relies upon should be properly referenced. 
    \end{itemize}

    \item {\bf Experimental Result Reproducibility}
    \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper (regardless of whether the code and data are provided or not)?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item If the paper includes experiments, a No answer to this question will not be perceived well by the reviewers: Making the paper reproducible is important, regardless of whether the code and data are provided or not.
        \item If the contribution is a dataset and/or model, the authors should describe the steps taken to make their results reproducible or verifiable. 
        \item Depending on the contribution, reproducibility can be accomplished in various ways. For example, if the contribution is a novel architecture, describing the architecture fully might suffice, or if the contribution is a specific model and empirical evaluation, it may be necessary to either make it possible for others to replicate the model with the same dataset, or provide access to the model. In general. releasing code and data is often one good way to accomplish this, but reproducibility can also be provided via detailed instructions for how to replicate the results, access to a hosted model (e.g., in the case of a large language model), releasing of a model checkpoint, or other means that are appropriate to the research performed.
        \item While NeurIPS does not require releasing code, the conference does require all submissions to provide some reasonable avenue for reproducibility, which may depend on the nature of the contribution. For example
        \begin{enumerate}
            \item If the contribution is primarily a new algorithm, the paper should make it clear how to reproduce that algorithm.
            \item If the contribution is primarily a new model architecture, the paper should describe the architecture clearly and fully.
            \item If the contribution is a new model (e.g., a large language model), then there should either be a way to access this model for reproducing the results or a way to reproduce the model (e.g., with an open-source dataset or instructions for how to construct the dataset).
            \item We recognize that reproducibility may be tricky in some cases, in which case authors are welcome to describe the particular way they provide for reproducibility. In the case of closed-source models, it may be that access to the model is limited in some way (e.g., to registered users), but it should be possible for other researchers to have some path to reproducing or verifying the results.
        \end{enumerate}
    \end{itemize}

\item {\bf Open access to data and code}
    \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results, as described in supplemental material?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that paper does not include experiments requiring code.
        \item Please see the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
        \item While we encourage the release of code and data, we understand that this might not be possible, so “No” is an acceptable answer. Papers cannot be rejected simply for not including code, unless this is central to the contribution (e.g., for a new open-source benchmark).
        \item The instructions should contain the exact command and environment needed to run to reproduce the results. See the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
        \item The authors should provide instructions on data access and preparation, including how to access the raw data, preprocessed data, intermediate data, and generated data, etc.
        \item The authors should provide scripts to reproduce all experimental results for the new proposed method and baselines. If only a subset of experiments are reproducible, they should state which ones are omitted from the script and why.
        \item At submission time, to preserve anonymity, the authors should release anonymized versions (if applicable).
        \item Providing as much information as possible in supplemental material (appended to the paper) is recommended, but including URLs to data and code is permitted.
    \end{itemize}

\item {\bf Experimental Setting/Details}
    \item[] Question: Does the paper specify all the training and test details (e.g., data splits, hyperparameters, how they were chosen, type of optimizer, etc.) necessary to understand the results?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item The experimental setting should be presented in the core of the paper to a level of detail that is necessary to appreciate the results and make sense of them.
        \item The full details can be provided either with the code, in appendix, or as supplemental material.
    \end{itemize}

\item {\bf Experiment Statistical Significance}
    \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item The authors should answer "Yes" if the results are accompanied by error bars, confidence intervals, or statistical significance tests, at least for the experiments that support the main claims of the paper.
        \item The factors of variability that the error bars are capturing should be clearly stated (for example, train/test split, initialization, random drawing of some parameter, or overall run with given experimental conditions).
        \item The method for calculating the error bars should be explained (closed form formula, call to a library function, bootstrap, etc.)
        \item The assumptions made should be given (e.g., Normally distributed errors).
        \item It should be clear whether the error bar is the standard deviation or the standard error of the mean.
        \item It is OK to report 1-sigma error bars, but one should state it. The authors should preferably report a 2-sigma error bar than state that they have a 96\% CI, if the hypothesis of Normality of errors is not verified.
        \item For asymmetric distributions, the authors should be careful not to show in tables or figures symmetric error bars that would yield results that are out of range (e.g. negative error rates).
        \item If error bars are reported in tables or plots, The authors should explain in the text how they were calculated and reference the corresponding figures or tables in the text.
    \end{itemize}

\item {\bf Experiments Compute Resources}
    \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources (type of compute workers, memory, time of execution) needed to reproduce the experiments?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item The paper should indicate the type of compute workers CPU or GPU, internal cluster, or cloud provider, including relevant memory and storage.
        \item The paper should provide the amount of compute required for each of the individual experimental runs as well as estimate the total compute. 
        \item The paper should disclose whether the full research project required more compute than the experiments reported in the paper (e.g., preliminary or failed experiments that didn't make it into the paper). 
    \end{itemize}

\item {\bf Code Of Ethics}
    \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics \url{https://neurips.cc/public/EthicsGuidelines}?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the authors have not reviewed the NeurIPS Code of Ethics.
        \item If the authors answer No, they should explain the special circumstances that require a deviation from the Code of Ethics.
        \item The authors should make sure to preserve anonymity (e.g., if there is a special consideration due to laws or regulations in their jurisdiction).
    \end{itemize}

\item {\bf Broader Impacts}
    \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work performed?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that there is no societal impact of the work performed.
        \item If the authors answer NA or No, they should explain why their work has no societal impact or why the paper does not address societal impact.
        \item Examples of negative societal impacts include potential malicious or unintended uses (e.g., disinformation, generating fake profiles, surveillance), fairness considerations (e.g., deployment of technologies that could make decisions that unfairly impact specific groups), privacy considerations, and security considerations.
        \item The conference expects that many papers will be foundational research and not tied to particular applications, let alone deployments. However, if there is a direct path to any negative applications, the authors should point it out. For example, it is legitimate to point out that an improvement in the quality of generative models could be used to generate deepfakes for disinformation. On the other hand, it is not needed to point out that a generic algorithm for optimizing neural networks could enable people to train models that generate Deepfakes faster.
        \item The authors should consider possible harms that could arise when the technology is being used as intended and functioning correctly, harms that could arise when the technology is being used as intended but gives incorrect results, and harms following from (intentional or unintentional) misuse of the technology.
        \item If there are negative societal impacts, the authors could also discuss possible mitigation strategies (e.g., gated release of models, providing defenses in addition to attacks, mechanisms for monitoring misuse, mechanisms to monitor how a system learns from feedback over time, improving the efficiency and accessibility of ML).
    \end{itemize}

\item {\bf Safeguards}
    \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models that have a high risk for misuse (e.g., pretrained language models, image generators, or scraped datasets)?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper poses no such risks.
        \item Released models that have a high risk for misuse or dual-use should be released with necessary safeguards to allow for controlled use of the model, for example by requiring that users adhere to usage guidelines or restrictions to access the model or implementing safety filters. 
        \item Datasets that have been scraped from the Internet could pose safety risks. The authors should describe how they avoided releasing unsafe images.
        \item We recognize that providing effective safeguards is challenging, and many papers do not require this, but we encourage authors to take this into account and make a best faith effort.
    \end{itemize}

\item {\bf Licenses for existing assets}
    \item[] Question: Are the creators or original owners of assets (e.g., code, data, models), used in the paper, properly credited and are the license and terms of use explicitly mentioned and properly respected?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not use existing assets.
        \item The authors should cite the original paper that produced the code package or dataset.
        \item The authors should state which version of the asset is used and, if possible, include a URL.
        \item The name of the license (e.g., CC-BY 4.0) should be included for each asset.
        \item For scraped data from a particular source (e.g., website), the copyright and terms of service of that source should be provided.
        \item If assets are released, the license, copyright information, and terms of use in the package should be provided. For popular datasets, \url{paperswithcode.com/datasets} has curated licenses for some datasets. Their licensing guide can help determine the license of a dataset.
        \item For existing datasets that are re-packaged, both the original license and the license of the derived asset (if it has changed) should be provided.
        \item If this information is not available online, the authors are encouraged to reach out to the asset's creators.
    \end{itemize}

\item {\bf New Assets}
    \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not release new assets.
        \item Researchers should communicate the details of the dataset/code/model as part of their submissions via structured templates. This includes details about training, license, limitations, etc. 
        \item The paper should discuss whether and how consent was obtained from people whose asset is used.
        \item At submission time, remember to anonymize your assets (if applicable). You can either create an anonymized URL or include an anonymized zip file.
    \end{itemize}

\item {\bf Crowdsourcing and Research with Human Subjects}
    \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants and screenshots, if applicable, as well as details about compensation (if any)? 
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
        \item Including this information in the supplemental material is fine, but if the main contribution of the paper involves human subjects, then as much detail as possible should be included in the main paper. 
        \item According to the NeurIPS Code of Ethics, workers involved in data collection, curation, or other labor should be paid at least the minimum wage in the country of the data collector. 
    \end{itemize}

\item {\bf Institutional Review Board (IRB) Approvals or Equivalent for Research with Human Subjects}
    \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
 .     \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
        \item Depending on the country in which research is conducted, IRB approval (or equivalent) may be required for any human subjects research. If you obtained IRB approval, you should clearly state this in the paper. 
        \item We recognize that the procedures for this may vary significantly between institutions and locations, and we expect authors to adhere to the NeurIPS Code of Ethics and the guidelines for their institution. 
        \item For initial submissions, do not include any information that would break anonymity (if applicable), such as the institution conducting the review.
    \end{itemize}

\end{enumerate}

\end{document}

### 2024-04-30 NEURIPS24/old.tex
update paper

%%%%%%%% ICML 2023 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{tikz}

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2023} with \usepackage[nohyperref]{icml2023} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2024}

% If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{icml2023}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \theoremstyle{plain}
% \newtheorem{theorem}{Theorem}[section]
% \newtheorem{proposition}[theorem]{Proposition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \theoremstyle{definition}
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{assumption}[theorem]{Assumption}
% \theoremstyle{remark}
% \newtheorem{remark}[theorem]{Remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Shapes analysis for time series.}

\input{header}
\input{def}
\newcommand\sam[1]{\textcolor{blue}{#1}}
\newcommand\thi[1]{\textcolor{orange}{#1}}
\begin{document}

\twocolumn[
\icmltitle{Shapes analysis for time series.\\
           International Conference on Machine Learning (ICML 2024)}%a changer

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2023
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Firstname1 Lastname1}{equal,yyy}
\icmlauthor{Firstname2 Lastname2}{equal,yyy,comp}
\icmlauthor{Firstname3 Lastname3}{comp}
\icmlauthor{Firstname4 Lastname4}{sch}
\icmlauthor{Firstname5 Lastname5}{yyy}
\icmlauthor{Firstname6 Lastname6}{sch,yyy,comp}
\icmlauthor{Firstname7 Lastname7}{comp}
%\icmlauthor{}{sch}
\icmlauthor{Firstname8 Lastname8}{sch}
\icmlauthor{Firstname8 Lastname8}{yyy,comp}
%\icmlauthor{}{sch}
%\icmlauthor{}{sch}
\end{icmlauthorlist}

\icmlaffiliation{yyy}{Department of XXX, University of YYY, Location, Country}
\icmlaffiliation{comp}{Company Name, Location, Country}
\icmlaffiliation{sch}{School of ZZZ, Institute of WWW, Location, Country}

\icmlcorrespondingauthor{Firstname1 Lastname1}{first1.last1@xxx.edu}
\icmlcorrespondingauthor{Firstname2 Lastname2}{first2.last2@www.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
%     The most common way to analyze time series is to find a feature representation manually with expert knowledge or automatically with machine learning.
% While unsupervised representation learning on time series proves efficient to reveal cluster structure in a dataset,
%  actual methods lacks precision for analyzing the inter-individual variability whithin a cluster.
Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
 Individuals have specificities depending on their genes or disease
  while sharing, on average, a similar pattern in the features of their corresponding time series. For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
 In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
 Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
 Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
  These features can be post-processed with any statistical or machine learning tools. 
  We demonstrate the advantages of our representation compared to existing methods using synthetic data and real-world examples, motivated by applications in medicine.
  %  After checking the injectivity of the TS-LDDMM representation on synthetic data (potentially to change), we study respiratory shapes of a mouse dataset by applying Principal components analysis to the features $(\alpha_j)_{j\in[N]} $.

  %  Moreover, we show that the method is competitive with supervised methods for classification, as soon as the dataset is homogene enough.
  %   The generalisation of the method to heterogene dataset (several clusters) is relagated for future works.
\end{abstract}

% citep
% \section{TODO}
% %De ce que je comprends , les gens font surtout des variational auto encoder, ou du masking
% %Mentionner que la méthode proposé peut être utilisé avec des réseaux
% %\sam{Approche NTK, fédéré personalisation, projet avec Axel, Lien entre continuous normalizing flow}
% Plan intro: 
% \begin{itemize}
%     \item Representation of patterns in time series has several applications in biology, but halso for classification, deep learning finds features automaticly.
%      Taking or not taking into account time can be done with DTW. However, either the representation is the result of a black box and need post analysis to be understand,
%       which is difficult to apply statiscal tools on it, or the representation is desgned by hand to select relevant features, but the capacity of analysis is limited by the knowledge of the designer.
%      \item In shape analysis, representation of shapes for statistical analysis in a long time issue, which has been solved mainly by the LDDMM framework. Representing shapes as diffeomorpshims of a common shape. 
%      \item However, this tool has not been yet applied on time series, even if there is already some tools developped for oriented curves. 
%      \item In this work, we propose to apply LDDMM to time series and to show its interest to derive an interpretable and unsupervised representation of patterns in time series.
%      We proposed some tools developped in the shape analysis community to the researcher interested in machine learning and time series ones.
%      \item First, we introduce our assumptions regarding the dataset and the main problem of diffeomorpshim learning. Then, we show how to solve it with LDDMM .
%       We expose how to apply it to time series by selecting wisely a RKHS kernel, motivating it with a representation theorem.
%       experiments on synthetic data are presented to show that the parameter of representation are identifiable when the time seriess are not too sharp and the kernels parameter well tuned. 
%       Then, we show how to apply the method on an unsupervised case of moouse respeitaory dataset and a supervised one on ... .

% \end{itemize}

% One option to tackle this issue is to derive ...
% feature representation of time series which depends on the problem at hand.
% which is parti... -> which is of prime interest in ...

\section{Introduction}
%https://arxiv.org/abs/1811.00075
%Appliqué sur les "UWaveGestureLibrary","ArticularyWordRecognition","Cricket" pour l'archive UEA
%"ArrowHead", "BME", "ECG200","FacesUCR","GunPoint" dans le cas univarié

%Idée de parler d'inter individualité, spécialiser davantage
%k shape, shapelets , ye2009time
%\thi{à reformuler pour faire moins stupide}
A time series is a set of values indexed by time \cite{trirat2024universal}.
This type of data is prevalent in biology and medicine for recording physiological time series \cite{wang2016research,bar2012studying,germain2023unsupervised}, as well as in the broader context of the Internet of Things sensors \cite{fathy2018large} and measurements in physical systems \cite{luo2021deep}.
 However, due to the nature of such data, which have different sample sizes and incorporate a temporal evolution, classical statistical analysis, and machine learning methods need refinement \cite{anderson2011statistical}.
 One option to tackle this issue is to derive a feature representation of time series, which depends on the problem at hand \cite{trirat2024universal}.

 When the goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology,
  the features encode the specificity of an individual compared to another. 
  In physiology, studying the different \textit{shapes} in a time series related to biological phenomena and their variations according to individual or pathology is common.
  However, a \textit{shape} has no clear definition; it is more an intuitive way to speak about the silhouette of a pattern in a time series. In this paper, we refer to as the shape of a time series, the graph of this signal.
  %For instance, the shape of a heartbeat in an Electrocardiogram (ECG) -the pattern- has been well studied \cite{madona2021pqrst} with the PQRST decomposition (see figure\sam{fill}).
  %There exist several methods to detect and segment patterns in a time series (\sam{mettre ref thibuat} ).
   %The goal is then to study their shape according to the property of the individual.

   In most clinical studies, depending on expert knowledge, people inspect hand-crafted features such as the mean, wavelets decomposition, or temporal correlation \cite{anderson2011statistical} to keep interpretability.
   However, this approach is time-consuming and potentially biased by priors, especially in exploratory research.
   Given the success of neural networks \cite{bishop1994neural}, features can now be automatically extracted, depending only on the data, by choosing an objective to optimize and a neural architecture tailored to the problem \cite{ismail2019deep,blazquez2021review,torres2021deep}, even in an unsupervised way \cite{meng2023unsupervised}.
   Moreover, depending on the method, it is still possible to have an interpretability of the features.
    For instance, among individuals, a contrastive clustering method can be applied to spot community structures \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
    More generally, interpretability \cite{yuan2019wave2vec,zhang2018visual,ismail2020benchmarking} is possible by visualizing the shapes of representative patterns to which the model is sensitive.
    Other methods, based on similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k}, also propose to perform clustering on shapes in time series or to find some representatives \cite{ye2009time}.
    Although a community structure with representatives can be learned in an unsupervised way,
     studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem.
     In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.% \thi{In an individual cluster ?}

% Alternatively, one can represent time series with an embedding that can be processed using conventional tools. % or we can represent the time series with an embedding which can be treated by classical tools.

% Depending on the tasks, e.g classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep} or clustering , a meaningful vector representation of time series can be derived.
%  Initially, in this research, handcrafted features were computed from time series, such as the mean, wavelets \cite{percival2000wavelet}, or Fourier coefficients \cite{bloomfield2004fourier}, depending on expert knowledge.
%   However, this approach is time-consuming and potentially biased by priors.
%   Observing the success of neural networks, features are automatically extracted by choosing an objective to optimize and a neural architecture tailored to the problem.
%    Taking a set of time series as input, features can be identified by solving specific supervised tasks—classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep}— or in an unsupervised manner by deriving labels from the inputs \cite{meng2023unsupervised}.
%    \thi{Rajouter une image, et une phrase par rapport au temps}
%    Although the deep learning approach is particularly effective for supervised tasks, annotated datasets are rare, making unsupervised methods more practical.
%    These methods have proven effective in identifying different patterns in the data and providing a vector representation that segregates well into clusters \cite{meng2023unsupervised}.
%     However, they may lack precision on small and homogeneous datasets with only one typical shape, often arising from clinical experiments addressing a specific phenomenon (refer to figures).
%      While current unsupervised methods separate representations into clusters, our focus is on the inter-individual variability within one cluster.
%       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.

This type of question has already been addressed in shape analysis \cite{vaillant2004statistics}, a community focusing on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
The main idea is to represent these different objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
 Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
  %Thus, instead of working on the manifold $\mathcal{M}$, we can work in the tangent space, which is an Euclidean space at $\mathbf{m}_0$.

 \citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
 However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
 For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.

 To tackle this issue, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
 Then, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
  The idea is to represent each element $(\msg(s^j))_{j\in[N]}$ of the dataset as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism.
  Then, the diffeomorphism is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
   The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
    Finally, these features encoding the shapes can feed any statistical or machine learning model.
   Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
   However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g. \Cref{fig:diffeo}.
    We solve this issue by specifying the class of diffeomorphisms to consider and showing how to learn them.
     This change is fruitful in representing time transformation as illustrated in \Cref{fig:transport}.

     Our contributions can be summarized as follows:
     \begin{itemize}
       \item We propose an unsupervised method (TS-LDDMM) to analyze inter-individual variability of shapes in a time series dataset.
       We especially motivate our extension of LDDMM to time series by introducing a theoretical framework.
       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method with respect to its hyperparameters, also providing guidelines for tuning.
       We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
        We illustrate the quantitative interest of the representation on classification tasks on real shape-based datasets.

     \end{itemize}

    %To the best of our knowledge, these theoretical and practical works have never been done before.
      % On synthetic data, we show the sensibility of the results according to the kernel parameters and give guidelines to tune it.
      % In an unsupervised framework, we vectorize respitory patterns of a mouse dataset and analyze the variations depending on the drug taken by the mouse.
      % Finally, we compare to the state-of-the-art methods in classification an ... classifier taking as inputs the representation given by LDDMM.

  %    \paragraph*{Related papers}
  %    %https://arxiv.org/pdf/2106.00750.pdf look to the related papers
  %  In the shape analysis community, a famous paper \cite{srivastava2010shape} addresses the problem of representation of curves $C:\msi \to \Rset^d$ by using the Square-Root Velocity (SRV) representation.
  %      This structure is particularly relevant to compute geodesics and distances between curves when the space is quotiented by time reparametrization.
  %       However, in our case the time axis deserves a particular attention, since the curves are derived from time-series. 
  %       %While \cite{srivastava2010shape} tackles the computations of geodesics and distances on quotient curves space, our aim is to get a vector representation of each pattern. 
  %       Very recently, in \cite{heo2024logistic}, this representation was used to compute PCA and to perform classification on quotient curves space, we follow the same ideas, but we focus on time series and we use LDDMM instead of SRV.
 % optimization framework with a well chosen neural architecture in order to learn automatically the features, which has been very 
% To understand these complex data, we are looking to simpler representations toperform classical statiscal analysis or to perform efficient downstream tasks of machine learning, the time series representation res time series to a real valued vector representation. in order to apply classical statiscal or machine learning tools depending on the needs such as Principal Component Analysis (PCA), logistic regression or deep learning methods for classification or anomaly detection.

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{"./pictures/diffeo.jpeg"}
  \vspace{-2.5em}
  \caption{A time series' graph $\msg=\{(t,s(t)): \eqsp t\in\msi\} $ can lose its structure after applying a general diffeomorphism $\phi.\msg$: a time value can be related to two values on the space axis.}
  \label{fig:diffeo}
  \vspace{-1em}
\end{figure}

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{"./pictures/transport.jpeg"}
  \vspace{-2.5em}
  \caption{LDDMM and TS-LDDMM are applied to ECG data.
  We observe that LDDMM, using a general Gaussian kernel, does not learn the time translation of the first spike but changes the space values, i.e., one spike disappears before emerging at a translated position. At the same time, TS-LDDMM handles the time change in the shape.
  This difference of \textit{deformations} implies differences in features \textit{representations}.   }
  \label{fig:transport}
  \vspace{-1em}
\end{figure*}

\section{Notations}
We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$,
by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
 by $||u||_\infty=\inf_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
and by $\Nset_{>0}$ is the set of positive integers. 
%We are in a sub class of curve quotiented by temporal reparametrisation since a time series' graph can be understood as the equivalent class of t ->(t,s(t))

%Reproducing Kernel Hilbert Space (RKHS)
\section{Methodology}
%\paragraph{Assumptions}
%Denoting by $C^0(\msi,\msj) $ the space of continous functions between two sets $\msi,\msj$,

We consider in this paper observations which consist in a population of $N$ multivariate time series, for any $j\in[N]$, $s^j \in \rmC^1(\msi_j,\Rset^{d})$. 
 Note that the time series are not defined on the same time interval.
To ease the presentation of our methodology, we suppose that we have access to the continuous time series $(s^j)_{j\in[N]}$, while in practice, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
 However, this difference does not change the overall rationale of our method and its implementation up to minor changes, which we summarize in \eqref{eq:general_optimization_problem}.

 We assume the time series population is globally homogeneous regarding their "shapes" even if inter-individual variability exists.
 Intuitively speaking, the "shape" of a time series $s:\msi\to \Rset^d$ is encoded in its graphs $\msg(s)$ defined as the set $\{(t,s(t)):\eqsp t\in\msi \} $ and not only in its values $s(\msi)=\{s(t):\eqsp t\in\msi \} $ since the time axis is crucial. % on the shape as illustrated in Figure\alain{ (mettre ref)}.
 As a motivating use-case, $s^j$ can be the time series of a heartbeat extracted from an individual's electrocardiogram (ECG), see \Cref{fig:transport}.
 The homogeneity in a resulting dataset comes from the fact that humans have similar shapes of heartbeat \cite{ye2012heartbeat,madona2021pqrst}.
 % the shape is also referred as morphology of the time series. 
  %  [More specifically, the heterogeneity in such dataset is often due to local or glocal dilatation of the time series regarding time, the heartbeat is quicker or slower depending on the individual.
  %  In maths, denoting by $\mcd(\msi,\msj) $ the set of diffeomorpshim defined on the sets $\msi$ to $\msj$,
  %   there exist a common pattern $\mathbf{s}_0: \msi \to \Rset^d$ and individual time reparametrisations $ \psi_j\in\mcd(\msi,\msi_j) $ such that,
  %   for any $j\in[N]$, $s^j=\mathbf{s}_0\circ \psi_j^{-1} $.]
%  \paragraph{Goals}%\sam{potentiellement changer l'ordre de time and space}

In this paper, we aim to study the inter-individual variability in the dataset by finding a relevant representation of each time series.
Inspired from the framework of shape analysis \cite{vaillant2004statistics}, addressing similar problems in morphology,
 we propose to represent each time series' graph $\msg(s^j)$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$, related to a time series $\mathbf{s}_0:\msi \to\Rset^d$, by a diffeomorphism $\phi_j$ on $\Rset^{d+1}$, for any $j\in[N]$,
\begin{equation}
 \label{eq:transformation}
 \phi_j.\msg(\mathbf{s}_0)=\{\phi_j\left(t,\mathbf{s}_0(t)\right), \eqsp t\in \msi \} \eqsp.
\end{equation}
$\bfs_0$ will be understood as the typical representative shape common to the collection of time series $(s^j)_{j\in[N]}$.
As $\bfs_0$ is supposed to be fixed, then the representation of the time series $(s^j)_{j\in[N]}$ boils down to the one of the transformation $(\phi_j)_{j\in[N]}$.
We aim to learn $\msg(\bfs_0)$ and $(\phi_j)_{j\in[N]} $.

First, we introduce the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework.
 Then, we explain how to learn a discretization of the graph $\msg(\bfs_0)$ and the diffeomorphisms $(\phi_j)_{j\in[N]} $ by using LDDMM and a gradient descent minimization.
  Finally, we tackle the specificity of graph time series by deriving a representation theorem on diffeomorphisms, which enables us to select the kernel needed in LDDMM, and thus, we propose TS-LDDMM. 
%As $\bfs_0$ is supposed to be fixed, then the representation of the time series $(s_j)_{j\in[N]}$ boils down to the one of the transformation $(\phi_j)_{j\in[N]}$.
 %In \Cref{section:optimization} the choice of the common time series $\mathbf{s}_0$ will be optimized to have $\phi_j$ of minimal norm. %T is given in the following part.

%  To learn the diffeomorphisms $(\phi_j)_{j\in[N]}$, we propose to use in this paper Large Deformation Diffeomorphic Metric Mapping (LDDMM).
%   In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
% We now describe more precisely LDDMM and in a next step the procedure to learn $\mathbf{s}_0$.

% \sam{to improve ?}
% Compared to the litterature of Unsupervised Representation Learning (URL) \cite{meng2023unsupervised}, the homogeneity assumption can be quite restrictive, however this is quite common in Shape Analysis \cite{vaillant2004statistics} (SA).
%     While in URL the goal is to find features accounting for the clustering structure in the heterogene data and to transfer this representation for different ML tasks, in SA the goal is to analyze the inter-individual variability with precision to derive clinical conclusions.
%     The generalization of the method presented in this paper to heterogene dataset is possible, but relagated for future works.
  %These velocity fields are then learned by minimizing the norm of this RKHS.

%In all the following, we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
%\sam{Transformer , en : dans les ensembles}
\subsection{Learning diffeomorphism with LDDMM} %Thèse de Nicolas Charon
\label{section:LDDMM}
% \sam{Introduire flow normalisant et continuous normalisant, faire des exps ?}
To learn the diffeomorphisms $(\phi_j)_{j\in[N]}$, we leverage LDDMM, initially introduced in \cite{beg2005computing}.
  In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).

The basic problem that we consider in this section is the following. Given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$\footnote{Note that we denote by $d'\in\nset$ the ambient space}, a set of starting points $\mathbf{x}=(x_{i})_{i\in[T_1]}$ in $\Rset^{d'}$, we aim to find a diffeomorphism $\phi$ such that the finite set of points $\mathbf{y}$ is similar in a certain sense to the set of finite sets of transformed points $\phi \cdot \mathbf{x} =(\phi(x_i))_{i\in[T_1]} $.
 The function $\phi$ is occasionally referred to as a \textit{deformation}.

 In our practical context, $\mathbf{y}$ represents one of the \textit{discretized graphs} of the time series $(s^j)_{j\in[N]}$, defined as, for any $j\in[N]$,
 \begin{equation}
  \label{eq:descretized_graph}
  \msg(\tsig^j)=(t_i^j,\tsig^j_i)_{ i\in[n_j]}\in (\Rset^{d+1})^{n_j}
 \end{equation}
  and $\mathbf{x}$ corresponds to a discretization of the reference graph $\msg(\mathbf{s}_0)$, i.e., 
  \begin{equation}
    \label{eq:descretized_referenced_graph}
    \mathbf{x}=\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} \eqsp . 
  \end{equation} 
% LDDMM is designed to analyzed an existing dataset, while NF and CNF are made to generalize a dataset for data augmentation.
\paragraph{Representing diffeomorpshims as deformations.}
Such \textit{deformations} $\phi$ are constructed via differential flow equations, for any $x_0\in \Rset^{d'} $ and $\tau\in[0,1]$:
\begin{align}
  \label{eq:LDDMM_dynamic}
    &\frac{\dd X(\tau)}{\dd \tau}= v_\tau(X(\tau)), \quad X(0)=x_0\eqsp ,\\
    &\phi^v_\tau(x_0)=X(\tau), \quad \phi^v=\phi^v_1  \eqsp ,
\end{align}
where the velocity field is $v:\tau\in [0,1]\mapsto v_\tau\in \msv $
and $\msv$ is a Hilbert space of continuously differentiable function
on $\Rset^{d'}$.  If
$||\dd u ||_{\infty}+|| u ||_{\infty}\leq ||u ||_\msv $ for any
$u\in \msv$ and
$v\in \rml^2([0,1],\msv)=\{v\in \rmC^0([0,1],\msv): \int_0^1 ||
v_\tau||^2_\msv \dd \tau<\infty \} $, by \citep[Theorem 5]{glaunes2005transport}
$\phi^v$ exists and belongs to $\mcd(\Rset^{d'})$, where we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
 Therefore, for any choice of $v$, $\phi^v$ defines a valid deformation. 
This offers a general recipe to construct diffeomorphism given a functional space $\msv$.

With this in mind, the velocity field $v$ fitting the data can be
estimated by minimizing 
$v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
 However, two computational challenges arise.
  First, this optimization problem is ill-posed, and a penalty term is needed to obtain a unique solution.
   In addition, we have to find a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, which allows us to solve this minimization problem efficiently.

%\paragraph*{Regularizing the problem to derive uniqueness.}
It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the space of deformations.
Following this geometric point of view, geodesics can be constructed on the space of deformations by using the norm of $\msv$. 
More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
 the following squared norm can be defined
 \begin{equation}
  \label{eq:geodesics_original}
    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
 \end{equation}
     as the minimal "energy" needed to perform the deformation $g$.
By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
To circumvent this issue, another characterization of geodesics can be considered.
 As in Riemannian geometry, 
 instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
 In other words, an Exponential map on deformations is wanted:
given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
 \begin{equation}
  \label{eq:geodesics_shooting}
  v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
 \end{equation}
since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
% Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.
%  Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
%  Moreover, we have $\mathscr{R}(\phi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in L^2([0,1],\msv)$.
%  %Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
%  This is presented with more details in the next paragraph.

\paragraph{Discrete parametrization of diffeomorpshim.}
%\alain{adapt this without mentioning applicaiton to time seriess}
%\sam{Est ce qu'on devrait pas introduire la dicretization du graphe ici et dire que l'on peut utiliser le théoreme de représentation}
In this part, $\msv$ is chosen as an RKHS \cite{berlinet2011reproducing} generated by a smooth kernel $K$ (e.g., Gaussian). % $K(x,y)\triangleq \exp(-|x-y|^2/\sigma^2)\operatorname{Id}_{d'}$ for any $(x,y)\in(\Rset^{d'})^2$, where $\sigma^2>0$ is the kernel width and $\operatorname{Id}_{d'}$ is the identity matrix.
We follow \cite{durrleman2013sparse} and define a 
 discrete parameterization of the velocity fields to perform geodesics shooting \eqref{eq:geodesics_shooting}.  
  The initial velocity field $v_0$ is chosen as a finite linear combination of the RKHS basis vector fields, 
$\mathbf{n}_0$ control points $\msx_0=(x_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0}$ and momentum vectors $\alpha_0=(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0} $ are defined such that for any $x\in \Rset^{d'}$, 
  \begin{equation}
    \label{eq:def_v0}
    v_0\left(\alpha_0,\msx_0\right)(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_{k,0})\alpha_{k,0} \eqsp .
  \end{equation}
   In our applications, the control points $(x_{k,0})_{k\in[\mathbf{n}_0]}$ can be understood as the discretized graph $(t_k,\mathbf{s}_0(t_k))_{k\in[\mathbf{n}_0]}$ of a starting time series $\mathbf{s}_0$. 
  With this parametrization of $v_0$, \citet{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same
  structure along time, such that for any $x\in\Rset^{d'}$ and $\tau\in[0,1]$, 
  \begin{equation}
    \label{eq:specific_form}
    v_\tau(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_k(\tau))\alpha_{k}(\tau) \eqsp ,
  \end{equation}
  \begin{equation} 
    \label{eq:integration}
      \left\{
        \begin{aligned}
        & \frac{\dd x_k(\tau)}{\dd \tau}=v_\tau(x_k(\tau)) \eqsp \\
        &\frac{\dd \alpha_k(\tau)}{\dd \tau}=-\sum_{k=1}^{\mathbf{n}_0} \dd_{x_k(\tau)} K(x_k(\tau),x_l(\tau))\alpha_{l}(\tau)^\top \alpha_{k}(\tau) \eqsp  \\
        & \alpha_k(0)=\alpha_{k,0},\quad x_k(0)=x_{k,0} \eqsp , k\in[\mathbf{n}_0] 
        \end{aligned}
        \right .
  \end{equation}
  These equations are derived from the hamiltonian $H:(\alpha_k,x_k)_{k\in [\mathbf{n}_0]}\mapsto \sum_{k,l=1}^{\mathbf{n}_0} \alpha_{k}^\top K(x_k,x_l)\alpha_{l}  $, such that
  the velocity norm is preserved $||v_\tau||_{\msv}=||v_0||_\msv $ for any $\tau\in [0,1]$.
   By \eqref{eq:integration}, the velocity field related to a geodesic $v^*$ is fully parametrized by its initial control points and momentum $(x_{k,0},\alpha_{k,0})_{k\in[\mathbf{n}_0]}$.
   Thus, given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$, a set of starting points $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$ in $\Rset^{d'}$, a RKHS's kernel $K:\Rset^{d'}\times \Rset^{d'}\to \Rset^{d'\times d'}$, a distance on sets $\mathscr{L}$, 
 a numerical integration scheme of ODE and a penalty factor $\lambda>0$, the basic geodesic shooting step minimizes the following function using a gradient descent method:
   \begin{equation}
    \label{eq:relaxation}
    \mathcal{F}_{\mathbf{x},\mathbf{y}}: (\alpha_k)_{k\in [T_1]}\mapsto \mathscr{L}\left(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2 \eqsp,  
   \end{equation}
   where $v_0$ is defined by \eqref{eq:def_v0} and $\varphi^{\{v_0\}}.\mathbf{x}$ is the result of the numerical integration of \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_{k\in[T_1]} $.

   %\sam{Parler de l'unicté?}
   \paragraph{Coming back to the representation problem \eqref{eq:transformation}.}
   Now, we detail how to learn $(\phi^j)_{j\in[N]}$ and a discretization of the reference graph $\msg(\mathbf{s}_0)$.
    Setting $\mathbf{y}_j=\msg(\tsig^j)$ for any $j\in[N]$ and $\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} $ as given in \eqref{eq:descretized_graph}-\eqref{eq:descretized_referenced_graph} with $\mathbf{n}_0=\operatorname{median}((n_j)_{j\in[N]})$, the representation problem given in \eqref{eq:transformation} boils down solving:
    \begin{equation}
      \label{eq:general_optimization_problem}
      \argmin_{\tilde{\msg}_0,(\alpha_k^j)_{k\in [\mathbf{n}_0]}^{j\in[N]}} \sum_{j=1}^N \mathcal{F}_{\tilde{\msg}_0,\mathbf{y}_j}\left((\alpha_k^j)_{k\in [\mathbf{n}_0]}\right)\eqsp ,
    \end{equation}
    which is carried out by a gradient descent on the control points $\tilde{\msg}_0$ and the momentums $\mathbf{\alpha}_j=(\alpha_k^j)_{k\in [\mathbf{n}_0]}$ for any $j\in[N]$, initialized by a dataset's time series graph of size $\mathbf{n}_0$ and by $0_{(d+1)\mathbf{n}_0}$ respectively.
    The optimization hyperparameter details are given in \Cref{appendix:optimizers_details}.
   The result of the minimization $\tilde{\msg}_0$ is then considered as the $\mathbf{n}_0$-samples of a common time series $\mathbf{s}_0$ and the momentums $\mathbf{\alpha}_j$ encoding $\phi_j$ yields a feature vector in $\Rset^{d \mathbf{n}_0} $ of $s^j$ for any $j\in[N]$.
   %, which will be the representation of the time series when choosing the time series' graph sets as inputs, is dependent on the choice of kernel $K$ and the set distance $L$.
   Finally, the vectors $(\mathbf{\alpha}_j)_{j\in[N]}$ can be analyzed with any statistical or machine learning tools such as Principal Components Analysis (PCA), Latent Discriminant Analysis (LDA), longitudinal data analysis and so on.

   %Before to specify the choice of the RKHS kernel $K$ and the loss on sets $\mathscr{L}$ in \Cref{section:time_series_specificity}, we first discuss the connection of LDDMM with normalizing flows. % to highlight the interest of the LDDMM approach.

%Conversely to NF or CNF, by \eqref{eq:integration}, different momentums parameters raise different final deformations $\phi^{\{v_0\}}$, this injectivity properties offers interpretability of the representation and gives sens to PCA or LDA post-processing as exposed in \Cref{section:experiments}.

   %If a diffeomorpshim $\Phi$ is given and we want to represent it as a deformation $\phi^v$ with $v$ of minimal ernergy, then
  %$(\alpha_{k,0})_k$ is the only parameters to learn since $(x_{k,0})_k$ is fixed.

%   \begin{algorithm}[tb]
%     \caption{Geodesic shooting with LDDMM}
%     \label{alg:example}
%  \begin{algorithmic}
%     \STATE {\bfseries Input:} a RKHS's kernel $K$, a loss on sets $L$, $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$, $\mathbf{y}=(y_i)_{i\in[T_2]}$ two sets of $\Rset^{d'}$, initial momentums 
%     $(\alpha_{k,0})_{k\in [T_1]} \in (\Rset^{d'})^{T_1}$ and a number of gradient iterations $n_{\text{iter}}$.

%     \STATE $f: (\alpha_k)_k\mapsto L\left(\phi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2  $ with $v_0$ and $\phi^{\{v_0\}}$
%      respectively defined by \eqref{eq:def_v0} and \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_k $. 
%      \sam{Peut être à définir dans un autre algo ?}
%     \STATE $\alpha_{k,0}^*\gets \alpha_{k,0}, \eqsp k\in[T_1]$
%     \FOR{$i=1$ {\bfseries to} $n_{\text{iter}}$}
%     \STATE $\alpha_{k,0}^*\gets \text{AdamStep}(f,\alpha_{k,0}^*$) , where $\text{AdamStep}$ is given in appendix (mettre ref)
%     \ENDFOR
%     \STATE {\bfseries Return:} $(\alpha_{k,0}^*)_{k\in [T_1]}$

%  \end{algorithmic}
%  \end{algorithm}

   \subsection{Application of LDDMM to time series analysis: TS-LDDMM}
%\tcr{Mettre l'applicaiton clair pour les times series}

        \label{section:time_series_specificity}
        In this section, we tailor the LDDMM framework to handle time series data.
          The reason is that applying a general diffeomorphism $\phi$ from $\Rset^{d+1}$ to a time series' graph $\msg(s)$ can result in a set $\phi.\msg(s)$ that does not correspond to the graph of any time series, as illustrated in the \Cref{fig:diffeo}.

To address this challenge, we need to identify an RKHS kernel $K:\Rset^{d+1}\times \Rset^{d+1}\to \Rset^{(d+1)^2}$ that generates deformations preserving the structure of the time series graph. This goal motivates us to clarify, in \Cref{theorem:representation}, the specific representation of diffeomorphisms we require before presenting a class of kernels that produce deformations with this representation.

Similarly, selecting a loss function on sets $\mathscr{L}$ that considers the temporal evolution in a time series' graph is crucial for meaningful comparisons with time series data. Consequently, we introduce the oriented Varifold distance.

       \paragraph{A representation separating space and time.}
       We prove that two time series graphs can always be linked by a time transformation composed of a space transformation. Moreover, a time series graph transformed by this kind of transformation is always a time series graph.
        We define $\Psi_\gamma\in \mcd(\Rset^{d+1}) : (t,x)\in\Rset^{d+1}\to (\gamma(t),x)$ for any $\gamma\in \mcd(\Rset)$ and $\Phi_f:  (t,x)\in\Rset^{d+1}\to (t,f(t,x)) $ for any $f\in \rmC^1(\Rset^{d+1},\Rset^d)$. 
        We have the following representation theorem.
        All proofs are given in \Cref{appendix:proofs}.

  Denote by $\msg(s)\triangleq \{ (t,s(t)): \eqsp t\in \msi \} $ the graph of a time series $s: \msi \to \Rset^d$ and $ \phi.\msg(s)\triangleq\{ \phi(t,s(t)): \eqsp t\in \msi\} $ the action of  $\phi\in \mcd(\Rset^{d+1}) $ on $\msg(s)$.
    \begin{theorem}
        \label{theorem:representation}
    Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi,\msj$ two intervals of $\Rset$.
     There exist $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\gamma\in  \mcd(\Rset) $ such that $\gamma(\msi)=\msj $ and $\Phi_f\in \mcd(\Rset^{d+1})$,
     \begin{equation}
        \msg(s)= \Pi_{\gamma,f}.\msg(\mathbf{s}_0),\eqsp \Pi_{\gamma,f}=\Psi_\gamma\circ\Phi_f.
     \end{equation}
     Moreover, for any $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\bar{\gamma}\in  \mcd(\Rset) $, there exists a continously differentiable time series $\bar{s}$ such that 
     $\msg(\bar{s})= \Pi_{\bar{\gamma},\bar{f}}.\msg(\mathbf{s}_0)$
    \end{theorem}
    %\sam{preuve todo, classement des variétés à bord, donne un homeomorphisme sur la restriction}
    % \begin{proof}
    %   Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi=(a,b),\msj=(\alpha,\beta)$ two intervals of $\Rset$.
    %   % $\msg(\mathbf{s}_0)$ and $\msg(s)$ are two smooth connected 1-dimensional manifold of $\Rset^{d+1}$, thus, by \cite[Appendix
    %   % Classfying 1-Manifolds]{milnor1965topology}, there exists $\Pi\in \mcd(\Rset^{d+1})$ such that $\msg(s)=\Pi.\msg(\mathbf{s}_0)$.
    %   %  Therefore, denoting by $\Pi_1: \Rset^{d+1}\mapsto \Rset$, we define $\gamma(t)\triangleq \Pi_1(t,\mathbf{s}_0(t))$ for any $t\in \msi$, then for any $t\in [b,+\infty)$,
    %   %   \begin{equation}
    %   %     \gamma(t)=\\Pi_1(b,\mathbf{s}_0(b))+(t-b)\frac{\dd\gamma}{\dd^- t}(b) \eqsp,
    %   %   \end{equation}
    %   %   and for any $t\in (-\infty,a]$,
    %   %   \begin{equation}
    %   %      \gamma(t)=\Pi_1(a,\mathbf{s}_0(a))+(t-a)\frac{\dd\gamma}{\dd^+ t}(a)\eqsp ,
    %   %   \end{equation}
    %   %    where $ \frac{\dd\gamma}{\dd^- t},\frac{\dd\gamma}{\dd^+ t}$ are respectively the left and right derivative.

    %   By setting $\gamma: t\in \Rset \mapsto (\beta-\alpha)(t-a)/(b-a)+\alpha\in \Rset $, we have $ \gamma(\msi)=\msj$ and $\gamma \in \mcd(\Rset) $.
    %    By defining $f:(t,x)\in\Rset^{d+1}\mapsto x-\mathbf{s}_0(t)+s\circ \gamma(t) $, the map $\Phi_f\in \mcd(\Rset^{d+1})$,
    %     indeed, its inverse is $\Phi_f^{-1}:(t,x)\in\Rset^{d+1}\mapsto (t,x+\mathbf{s}_0(t)-s(t)) $ and is continuously differentiable.
    %      Moreover, we have $\Pi_{\gamma,f}.\msg(\mathbf{s}_0)=\{(\gamma(t),s\circ \gamma(t)):\eqsp t\in\msi \}=\msg(s) $.

    %     Let $\bar{f}\in \rmC^0(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^0(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
    %     We have :
    %     \begin{align}
    %       \Pi_{\gamma,f}.\msg(\mathbf{s}_0)&=\{(\gamma(t),f(t,\mathbf{s}_0(t))),\eqsp t\in \msi \} \\
    %       &\label{eq:proof1_last_eq}=\{(t,f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right),\eqsp t\in \gamma(\msi) \} \eqsp .
    %     \end{align}
    %     By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^0(\gamma(\msi), \Rset^d) $ by composition of continuous functions
    %     and $ \msg(\bar{s})= \Pi_{\gamma,f}.\msg(\mathbf{s}_0)$ by \eqref{eq:proof1_last_eq}, which concludes the proof.
    % \end{proof}
    \begin{remark}
      that for any $\gamma \in \mcd(\Rset) $ and $s\in \rmC^0(\msi,\Rset^d)$,
   \begin{equation}
    \{(\gamma(t),s(t)),\eqsp t\in \msi \}=\{(t,s\circ \gamma^{-1}(t)):\eqsp t\in\gamma(\msi) \}\eqsp .
   \end{equation}
   As a result, $\Psi_\gamma $ can be understood as a temporal reparametrization and $\Phi_f$ encodes the transformation about the space.
 \end{remark}

    \paragraph{Choice for the kernel associated with the RKHS $\msv$}
    \label{paragraph:kernel_V}
    As depicted on \Cref{fig:diffeo}-\ref{fig:transport}, we can not use any kernel $K$ to apply the previous methodology to learn deformations on time series' graphs.
    We describe and motivate our choice in this paragraph.
     Denote the one-dimensional Gaussian kernel by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
    To solve the geodesic shooting problem \eqref{eq:relaxation} on $\Rset^{d+1}$, we consider for $\msv$ the RKHS associated with the kernel defined for any $(t,x),(t',x')\in (\Rset^{d+1})^2$:
    \begin{align}
      \label{eq:kernel_TAS}
      &K_{\msg}((t,x),(t',x'))=\begin{pmatrix}
        c_0K_{\text{time}} & 0 \\
        0 & c_1 K_{\text{space}} 
        \end{pmatrix} \eqsp , \\
        &K_{\text{space}}=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K_{\text{time}}=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
    \end{align}
    parametrized by the widths $\sigma_{T,0},\sigma_{T,1},\sigma_x>0$ and the constants $c_0,c_1>0$.
This choice for $K_\msg$ is motivated by the representation \Cref{theorem:representation} and the following result. 
    \begin{lemma}
      If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0},
       there exist $\gamma \in \msd(\Rset) $ and $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ such that $\phi^v=\Psi_\gamma\circ\Phi_f $.
    \end{lemma}
    % \begin{proof}
    %   Let $v$ be a vector field generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0}.
    %  We remark that the first coordinate of the velocity field $v_\tau$ denoted by $v_\tau^{\text{time}}$ only depends on the time variable $t$ for any $\tau\in[0,1]$.
    %  Thus, when computing the first coordinate of the deformation $\phi^v$, denoted by $\gamma$, we integrate \eqref{eq:LDDMM_dynamic} with $v_\tau$ replaced by $v_\tau^{\text{time}}$,
    %   thus $\gamma$ is independant of the variable $x$. Moreover, $\gamma\in \mcd(\Rset)$ since a Gaussian kernel induced an Hilbert space $\msv$ satisfying $|f|_V\leq |f|_\infty+ |\dd f|_\infty  $ for any $f\in \msv$ by \citep[Theorem 9]{glaunes2005transport}.
    %   For the same reason, we have $\phi^v\in \mcd(\Rset^{d+1})$, and thus its last coordinates denoted by $f$ belongs to $\rmC^1(\Rset^{d+1},\Rset^d)$, and by construction $\phi^v=\Psi_\gamma\circ\Phi_f $.
    % \end{proof
    \begin{remark}
      With this choice of kernel, the features associated to the time transformation can be extracted from the momentums $(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}$ in \eqref{eq:def_v0} by taking the coordinates related to time.
      However, the features related to the space transformation are not only in the space coordinates since the related kernel $K_{\text{space}}$ depends on time as well.
     \end{remark}
    In \Cref{appendix:kernel_TS_LDDMM}, we give guidelines for selecting the hyperparameters $(\sigma_{T,0},\sigma_{T,1},\sigma_x,c_0,c_1)$.

    %  and then we can perform a statistical analysis on the representations $(\psi_j,\phi^j)_{j\in[N]}$,
    % such as applying PCA decomposition \cite{vaillant2004statistics}, carring out longitudinal analysis \cite{guigui2022parallel} and so on (\sam{mettre ref}).
    % % Cette hypothèse intéressante car ça permet de gérer des signaux de taille variables, tout les signaux se repréentent comme ça 
    % % Espace homogène, injectivité de t ,s(t), permet de garder la structure temporel en compte
    % \sam{Can we show unicity by assuming that $\phi,\psi$ minimize a norm ?}
 % KEOPS tp pour lddmm

%and their empiric samples version $\tilde{G}_j=(\{(t_i,s^j_i), \eqsp i\in [n_i]\})_j $

%  We translate the notation from the continous to discrete by adding a tilde :
%   $\tilde{G}^j=\triangleq \{ (t_i,\tilde{s}_i^j),\eqsp i\in[n_i] \}$.

% \paragraph{A problem of minimization}
% In the next section, we will present that diffeomorpshim on space an time can be parametrized as $\phi^{v^S},\phi^{v^T}$
%  with parameters $(v^S,v^T) \in \Theta_S\times \Theta_T$. Then, pattern will be represented by the parameters $v^S,v^T$ solutions of the following minimization :
% \begin{align}
%   \label{eq:minimization}
%   &\argmin_{v^S \in \Theta_S,\eqsp v^T\in \Theta_T } \mathbf{D}(v^S,v^T)+\lambda \mathbf{N}(v^S,v^T)\eqsp , \\
%   & \mathbf{D}(v^S,v^T)=L(\phi^{v^S}.G^\msj(\mathbf{s_0}\circ (\psi^{v^T})^{-1}),G^\msj(s)) \eqsp , 
% \end{align} 
% where $L$ is a loss on graphs and $\mathbf{N}$ is a term of regularization.

% $s_i: t\in[0,T_0]\mapsto s(t)\in \Rset^d $ is only known through its samples $(S_i^j(t_i)=s_i)_{i\in[N]} $ at timestamps $(t_i)_{i\in[T]}$.

%\sam{améliorer transition}
\paragraph{Loss}

This section specifies the distance function $\scrl$ introduced in the loss function defined in \eqref{eq:relaxation}.

In practice, we can only access discretized graphs of time series, $(t_i^j,\tsig^j_i)_{i\in[n_j]}$ for any $j\in[N]$, that are potentially of 
different sizes $n_j$ and sampled at different timestamps $(t_i^j)_{i\in[n_j]}$ for any $j\in[N]$.
 Usual metrics, such as the Euclidean distance, are not appealing as they 
make the underlying assumptions of equal size sets and the existence of a pairing between points.
 Distances between measures on 
sets (taking the empirical distribution), such as Maximum Mean Discaprency (MMD) \cite{dziugaite2015training,borgwardt2006integrating}, alleviate those issues; however, MMD only accounts for positional information 
and lacks information about the time evolution between sampled points.
 A classical data fidelity metric from shape analysis 
corresponding to the distance between \textit{oriented varifolds} associated with curves alleviates this last issue \cite{kaltenmark2017general}.  
Intuitively, an oriented varifold is a measure that accounts for positional and tangential information about the underlying 
curves at sample points. More details and information about \textit{oriented varifolds} can be found in \Cref{appendix:varifold}.

More precisely, given two sets $\msg_0=(g_i^0)_{i\in[T_0]},\msg_1=(g_i^1)_{i\in[T_1]}\in (\Rset^{d+1})^{T_1}$ and a kernel\footnote{$\mathbb{S}^d=\{x\in\Rset^{d+1}:\eqsp |x|=1\}$} $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$
verifying \citep[Proposition 2 \& 4]{kaltenmark2017general}, for any $\xi\in\{0,1\}$ and $i\in[T_\xi-1]$, denoting the center and length of the $i^{th}$ segment $[g_i^\xi,g_{i+1}^\xi]$ by
$c_i^\xi = (g_i^\xi + g_{i+1}^\xi)/2$, $l_i^\xi = \| g_{i+1}^\xi-g_{i}^\xi\|$, and 
$\overrightarrow{v_i}^\xi = (g_{i+1}^\xi-g_{i}^\xi)/l_i^\xi$, the varifold distance between $\msg_0$ and $\msg_1$  is defined as,
\begin{align}
  d_{\msw^*}^2(\msg_0,\msg_1) &= \sum_{i,j = 1}^{T_0-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^0_j,\overrightarrow{v_j}^0))l^0_j \\
  &- 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^1_j,\overrightarrow{v_j}^1))l^1_j \\
  &+ \sum_{i,j = 1}^{T_1-1}l^1_i k((c^1_i,\overrightarrow{v_i}^1),(c^1_j,\overrightarrow{v_j}^1))l^1_j 
\end{align}

In practice, we set the kernel $k$ as the product of two anisotropic Gaussian kernels, $k_{\pos}$ and $k_{\dir}$, 
such that for any $(x,\overrightarrow{u}),(y,\overrightarrow{v}) \in (\Rset^{d+1} \times \mathbb{S}^d)^2$
\begin{equation}
    k((x,\overrightarrow{u}),(y,\overrightarrow{v})) = k_{\pos}(x,y)k_{\dir}(\overrightarrow{u},\overrightarrow{v}) \eqsp.
  \end{equation}
  The specific kernels $k_{\pos},k_{\dir}$ that we use in our experiments are given \Cref{appendix:kernel_implementation}.
   Note that the loss kernel $k$ has nothing to do with the velocity field kernel denoted by $K_\msg$ or $K$ specified in \Cref{paragraph:kernel_V}.
Finally, we define the data fidelity loss function, $\scrl$, as $ d_{\msw^*}^2$, which is differentiable with regards to its first variable.
For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}.

%\sam{TO improve by Thibaut
%\thi{parler du fait qu'on a pas pairing, grace au difféo on conserve la structure}
%To relax the geodesic problem \Cref{eq:geodesics_original}, the end point of the geodesics is compared to the target through a loss $\mathscr{L}$ as proposed in \eqref{eq:relaxation}.
%Since we are dealing with discrete time series $\tsig^j$ of different samples size $n_j$, we need a loss on set measures and not only on vectors.
% A natural choice would be the Maximum Mean Discrepency (MMD) \cite{dziugaite2015training,borgwardt2006integrating}, for any $X=(x_i)_{i\in[T_1]},Y=(y_i)_{i\in[T_2]}\subset Rset^{d'}$, given a kernel $k:\Rset^{d'}\times \Rset^{d'} \to \Rset$,
% \begin{align}
%  &\operatorname{MMD}^k(\mu_X,\mu_Y)=\frac{1}{T_1^2}\sum_{i,j=1}^{T_1} k(x_i,x_j)+\frac{1}{T_2^2}\sum_{i,j=1}^{T_2} k(y_i,y_j) \\
%  &-\frac{2}{T_1T_2}\sum_{i=1}^{T_1}\sum_{j=1}^{T_2} k(x_i,y_j) \eqsp , 
% \end{align}
% where $\mu_X=\sum_{i=1}^{T_1} \delta_{x_i}/T_1$, $\mu_Y=\sum_{j=1}^{T_2} \delta_{y_j}/T_2$.
%However, this distance is agnostic of the time evolutions in time series' graph $(t_i^j,\tsig^j_i)_{i\in[n_j]}$, there is no orientation of the set taken into account.
% That is why, we prefer to choose the Varifold distance $d_V(\cdot,\cdot)$ \cite{kaltenmark2017general} which was specifically designed for oriented curves.
%  The Varifold distance extends the MMD by adding tangent vectors to the input:
%   instead of considering a set $X=(x_i)_{i\in[T_1]}$, we extend the input as $\overrightarrow{X}=(x_i,v_i)_{i\in[T_1]}$ where $v_i$ is the tangent vector at $x_i$ such that
%    \begin{equation}
%      d_V(X,Y)=\operatorname{MMD}^{K_p}(\mu_{\overrightarrow{X}},\mu_{\overrightarrow{X}}) \eqsp ,
%    \end{equation}
%    where in practice the tangent vectors are taken as discrete derivative and $K_p$ is a kernel on $\Rset^{2d'}$ such that for any $(x,v),(y,w)\in (\Rset^{2d})^2$, 
%    we have, 
%    \begin{equation}
%      K_p((x,v),(y,w))=k(x,v)k_T(v,w) \eqsp ,
%    \end{equation}
%   where $k_T $ is a kernel on the tangent vector space.

\paragraph{Relation to Continuous Normalizing Flows.}

One particular popular choice to address the problem of learning a diffeomorphism or a velocity field is Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} (NF) or their continuous counterpart \cite{chen2018neural,grathwohl2019scalable,salman2018deep} (CNF).
However, we do not rely on this class of learning algorithms for several reasons. Indeed, existing and simple normalizing flows are not suitable for the type of data that we are interested in this paper \cite{feng2023multi,deng2020modeling}. 
  In addition,  they are primarily designed to have tractable Jacobian functions, while we do not require such property in our applications. 
Finally, the use of a differential flow solution of an ODE
\eqref{eq:LDDMM_dynamic} trick is also at the basis of CNF, which
then consists of learning a velocity field to address in fitting the data through a loss aiming to address the problem
at hand. Nevertheless, the main difference between CNF and LDDMM lies in the
parametrization of the velocity field. LDDMM uses kernels to
derive closed form formula and enhance interpretability while
NF and CNF take advantage of deep neural networks to scale with
large dataset in high dimensions.

%Wavelet-Based Multiscale Flow For Realistic Image Deformation in the Large Diffeomorphic Deformation Model Framework
% 3.1 +3.2 + specialité de la décomposition (t,f(t,x))
% \section{Specificity to time series}
% We are looking to time seriess 
% Each time series $s: t\in[0,T]\mapsto s(t) $ is only known through its samples $(s(t_i)=s_i)_{i\in[N]} $ at timestamps $(t_i)_{i\in[N]}$.
%is described by its graph $G(s)=\{(y,s(t)), \} $

% \section{optimzation
% % We are not doing Pre-ESPAce car l'on s'intéresse à la reparamétrisation en temps, au lieu de calculer la transfo en temps après avoir
% % push en espace, on le fait avant pour que tout soit défini depuis s_0, méthode itérative car ça marche mieux, moins de calcul de gradient
% %Natural smoothing with TTS
% \label{section:optimization}
%     \begin{center}
%         \begin{tikzpicture}
%             \node at (0,0) {$G_f^I$};
%             \draw [-stealth](0.3,0) -- (1.3,0);
%             \node at (1.6,0) {$G_f^J$};
%             \draw [-stealth](1.6,-0.3) -- (1.6,-1.3);
%             \node at (1.6,-1.6) {$G_g^J$};
%             \draw [-stealth](1.3,-1.6) -- (0.3,-1.6);
%             \node at (0,-1.6) {$G_g^I$};
%             \draw [-stealth](0,-1.3) -- (0,-0.3);

%             \node at (0.8,0.3) {$T$};
%             \node at (0.8,-1.9) {$T^{-1}$};
%             \node at (1.9,-0.8) {$S_J$};
%             \node at (-0.4,-0.8) {$S_I^{-1}$};
%         \end{tikzpicture}
%         \end{center}

%         where $S_I^{-1} = T^{-1} \circ S_J^{-1} \circ T$

\section{Experiments}
\label{section:experiments}

First, we show on synthetic data that the proposed representation is identifiable provided that the hyperparameters and the reference graph are wisely selected, i.e.,
 the parameter $v_0^*$ generating a deformation $\varphi^{\{v_0^*\}}$ of a time series graph $\msg$ can be estimated from the data $\msg,\varphi^{\{v_0^*\}}.\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation}.
 Secondly, we illustrate the qualitative interest of TS-LDDMM in studying inter-individual variability on a clinical dataset.
  Thirdly, we demonstrate the quantitative performance of our representation by performing classification on shape-based datasets.
  The method is implemented on Python using the library JAX\footnote{https://github.com/google/jax}. The code was compiled on a server with NVIDIA RTX A2000 12GB GPU, Intel(R) Xeon(R) Gold 5220R CPU @ 2.20GHz, and 250 GB of RAM. The code will be available on Github.
\subsection{Synthetic experiments}
\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{pictures/samples.pdf}
    \vspace{-2.5em}
    \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
    \label{fig:exemple_synthetic}
    \vspace{-1em}
\end{figure}

\begin{table}
    \caption{Values of $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ as $\mathbf{\alpha}^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
      \centering
         \begin{tabular}{lrrrrrrr}
         \toprule
         $\sigma_{T,0} \backslash \sigma_x$  & 1 & 10 & 50 & 100 & 200 & 300 \\
         \midrule
         0.1 & 2e+0 & 3e-4  & 1e-5&4e-6&7e-4&4e-3 \\
        1 & 4e-2 & 1e-4  & 1e-5&4e-6&7e-4 &4e-3  \\
         100 & 4e-2 & 2e-4  & 1e-5&4e-6&7e-4&4e-3  \\
         \bottomrule
         \end{tabular}
      \label{table:synthetic2}
      \vspace{-1em}
  \end{table}

% \begin{subfigure}{0.7\linewidth}
%   \centering
%   \resizebox{\columnwidth}{!}{%
%      \begin{tabular}{lrrrr}
%      \toprule
%      $\sigma_{T,0} \backslash \sigma_x$  & 0.1 & 1 & 100 \\
%      \midrule
%      1 & 2e+0 & 4e-2 & 4e-2 \\
%      10 & 3e-4 & 1e-4  & 2e-4 \\
%      50 & 1e-5 & 1e-5  & 1e-5 \\
%      100 & 4e-6 & 4e-6 & 4e-6 \\
%      200 & 7e-4 & 7e-4 & 7e-4 \\
%      300 & 4e-3 & 4e-3 & 4e-3 \\
%      \bottomrule
%      \end{tabular}
%      %
%      }
%  \caption{Values of $\scrl(y^*,\varphi^{\{\hat{v}_0\}}.\msx)$ where $y^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
% \end{subfigure}
First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
Given $m_{s}\in \Nset_{>0}$ and $t_{a},s_{a}>0$, we randomly generate initial momentums $\mathbf{\alpha}^*=(\alpha_k^*)_{k\in[\mathbf{n}_0]}$ with the following sampling, called Gen($m_s,t_a,s_a$):
 For any $k\in[\mathbf{n}_0]$, $\mathbf{\alpha}_k'$ is sampled according to a Gaussian normal distribution $\mathcal{N}(0_{d+1},I_{d+1})$.
Then, $(\alpha_k')_{k\in[\mathbf{n}_0]}$ is regularized by a rolling average of size $m_{s}$, we get $\bar{\mathbf{\alpha}}'=(\bar{\alpha}_k')_{k\in[\mathbf{n}_0]}$.
Finally, we normalize $\bar{\mathbf{\alpha}}'$ to derive $\mathbf{\alpha}^*$ such that $|([\alpha_k^*]_t)_{k\in[\mathbf{n}_0]}|=t_{\text{amp}}$ and $|([\alpha_k^*]_s)_{k\in[\mathbf{n}_0]}|=s_{\text{amp}}$ for any $k\in[\mathbf{n}_0]$, denoting by $[\alpha_k^*]_t,[\alpha_k^*]_s$ the time and space coordinates of $\alpha_k^*$ respectively.
Note that the regularizing step $(\mathbf{\alpha}_k')_{k\in[\mathbf{n}_0]}\to \bar{\mathbf{\alpha}}' $ is necessary to obtain realistic deformations which take into account the regularity induced by the RKHS $\msv$.
%  uniformly on the set $\mss_M=\{ (\alpha_k)_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}:\eqsp |\alpha_k|=M\}  $ with $M>0,\mathbf{n}_0\in \Nset_{>0}$,
%  then we regularize $\mathbf{\alpha}$ by convolving
%   with ... to get the generation parameter $\mathbf{\alpha}^*$.
Then, using $v_0(\mathbf{\alpha}^*,\msx)$ as defined in \eqref{eq:def_v0} with initial momentums $\mathbf{\alpha}^*$ and control points $\msx$, we apply the induced deformation $\varphi^{\{v_0\}} $ by \eqref{eq:integration} to $\msx$ and obtain $\varphi^{\{v_0\}}.\msx$.
Finally, we solve \eqref{eq:relaxation} to recover an estimation $\hat{\mathbf{\alpha}}$ of $\mathbf{\alpha}^*$ and report the average relative error (ARE) $|v_0(\hat{\mathbf{\alpha}},\msx)-v_0(\mathbf{\alpha}^*,\msx)|_\msv/|v_0(\mathbf{\alpha}^*,\msx)|_\msv$ on 50 repetitions.
This procedure is performed for any $m_{s},t_{a},s_{a}\in \{10,50,100\}\times \{5,10,15,20\}^2 $.
 Mean, standard deviation, and maximum of the ARE on all these hyperparameters choices are respectively $\mathbf{0.10, 0.03, 0.17}$.
 Therefore, the estimation procedure \eqref{eq:relaxation} offers a good approximation of the true parameter when the kernel $K_\msg$ is well specified.
 We observe that the estimation is difficult when $t_a\ll s_a$ because the time series can be very noisy as illustrated in \Cref{fig:exemple_synthetic}: this impacts the Varifold loss which is sensitive to tangents.

 Secondly, we demonstrate a weak identifiability when the kernel $K_\msg$ is misspecified: we can reconstruct the graph time series' after deformations even if the hyperparameters of $K_\msg$
  are different during the generation and the estimation.
   The hyperparameters of $K_\msg$ during generation are $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)=(1,0.1,100,1,1)$ and we fix $\sigma_{T,1},c_0,c_1=(1,1,0.1) $ for $K_\msg$ during estimation.
   We aim to understand the impact of $\sigma_{T,1},\sigma_x$ on the reconstruction since they are encoding the smoothness of the transformation according to time and space.

    For any choice of the hyperparameters $\sigma_{T,1},\sigma_x\in \{1,10,50,100,200,300 \}\times \{0.1,1,100\}$ related to $K_\msg$ in the estimation,
     we average $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ on 50 repetitions when $\mathbf{\alpha}^*$ is sampled according to Gen$(10,10,50)$ and $\hat{v}_0=v_0(\hat{\alpha},\msx)$ denoting by $\hat{\alpha}$ the result of the minimization \eqref{eq:relaxation}.
  We observe in \Cref{table:synthetic2} that the reconstruction is almost perfect except in the case when $\sigma_{t,0}=1$ during estimation, while $ \sigma_{t,0}=100$ during generation.
   Compared to $\sigma_{T,0}$, $\sigma_x$ has nearly no impact on the reconstruction.
   In \Cref{appendix:kernel_implementation}-\ref{appendix:kernel_TS_LDDMM}, we propose guidelines to drive future hyperparameters tuning and further discussions related to $\sigma_{T,1},c_0,c_1$. 
\subsection{Qualitative analysis of respiratory behavior in mice}
%We consider a dataset composed of mouse respiratory time series before and after a drug injection.
% A complete presentation of this dataset is given in \Cref{appendix:mouse_dataset}.
%  The mouse are divided in two groups depending on their genotypes : \textit{colq} and \textit{wt}.
%  We aim to study the difference of respiratory shapes according to the genotype.
%   That is why, TS-LDDMM \eqref{eq:general_optimization_problem} is applied to derive the features representations $(\mathbf{\alpha}_j)_{j\in[N]}\in (\Rset^2)^N$ related to $N$=\sam{fill} respiratory time series coming from 14 mouse (7 \textit{colq} and \textit{wt}) before drug injection.
%   Then, a Principal Components Analysis (PCA) is performed on $(\mathbf{\alpha}_j)_{j\in[N]}$
\begin{figure*}[t]
  \centering
  \includegraphics[width=0.95\linewidth]{"./pictures/exp_1_bis.pdf"}
  \vspace{-1.5em}
  \caption{Analysis of the three principal components (PC) related to the respiratory cycles of the mouse before exposure.
  In Figure A), the densities of each genotype according to each PC are displayed. In Figure B), the deformations of the reference graph $S_0$ along each PC are given. In Figure D), the graph of reference $S^j$, also called barycenter, related to each mouse, is displayed according to their coordinates on PC1 and PC3. In Figure C) et E), illustrations of respiratory cycles related to mice coming from the \textbf{wt} and \textbf{colq} group are displayed.  }
  \label{fig:exp_1_PCA}
  %\vspace{-1em}
\end{figure*}

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.95\linewidth]{"./pictures/exp_2.pdf"}
  \vspace{-1em}
  \caption{Analysis of the first Principal Component (PC1) related to the respiratory cycles of the mouse 
  before and after exposure. In Figure A), the densities of each genotype according are displayed. In Figure B), the deformations of the reference graph $S_0$ PC1 is given. In Figure C), respiratory cycles displayed with respect to time and according to their coordinates on PC1 and PC2}
  \label{fig:exp_2_PCA}
  \vspace{-1.5em}
\end{figure*}

This experiment highlights the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
We consider a time series dataset recording the evolution of the respiratory airflow of mice exposed 
to an irritant molecule altering respiratory functions \cite{nervo2019respiratory}. The dataset is divided into two groups, one 
composed of 7 control mice (\textbf{wt}) and the other of 7 mice (\textbf{colq}) deficient in an enzyme 
involved in the control of respiration. For each mouse, the respiratory airflow was recorded for 
15 to 20 minutes before exposure to the irritant molecule and then for 35 to 40 minutes. A complete 
description of the dataset is given in the \Cref{appendix:mouse_dataset}.
By comparing the shape of individual respiratory cycles (inspiration + expiration, see \Cref{fig:exp_1_PCA}-C)), 
we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution 
after exposure to the irritant molecule.

We first compare breathing behaviors before exposure.
Solving \eqref{eq:general_optimization_problem}, we derive the reference respiratory cycle's graph $S_0$ and the TS-LDDMM features representations 
$(\alpha_j)_{j\in[N_1]}$ related to $N_1=700$ respiratory cycles extracted according 
to the procedure \cite{germain2023unsupervised}.
Then, we perform a kernel PCA on the initial velocity field $\left(v_0(\alpha_j,S_0)\right)_{j\in[N_1]}\in \msv^{N_1}$ defined in \eqref{eq:def_v0}.
In \Cref{fig:exp_1_PCA}, we focus on the analysis of the three Principal Components (PC).

As observable from \Cref{fig:exp_1_PCA}-B), principal components refer to different types of deformations. 
By interpreting \Cref{fig:exp_1_PCA}-B): Only PC1 accounts for time warping, PC2 expresses the trade-off between inspiration and expiration duration, and PC3 corresponds to a change in signal amplitude.
 Compared to \textbf{wt} mice, the distribution of \textbf{colq} mice 
TS-LDMMM feature representation along the PC1 axis has a heavy tail and the associated deformation (+3 $\sigma_{\text{PC}}$) 
shows an inspiration with two peaks. As illustrated in \Cref{fig:exp_1_PCA}-A), such respiratory cycles are preponderant
 with \textbf{colq} mice and may be caused by motor impairment due to their enzyme deficiency, \cite{germain2023unsupervised}.
  In addition, 
 the \textbf{colq} mice were smaller than the \textbf{wt} mice due to a delay in growth caused by their lack of an enzyme. 
 This difference can be seen on PC3 since the volumes of air (area under the curve) inspired and exhaled are
  smaller for the smaller mice. In correlation, the distribution of \textbf{wt} mice TS-LDDMM feature representations 
  along the PC3 axis have a heavy tail corresponding to large air volume as depicted by the deformation 
  (+3 $\sigma_{\text{PC}}$) in \Cref{fig:exp_1_PCA}-B). Finally, \Cref{fig:exp_1_PCA}-D) shows that PC1 and PC3 capture the main differences between
   the two groups as their respective reference graphs $S^j$ are located in different parts of the space. 
%In Figure \Cref{fig:exp_1_PCA}, we display coordinates densities per genotype and PC.

We perform a second experiment to analyze the evolution of breathing behaviors when mice are exposed to the irritant 
molecule. We follow the same procedure as before. However, we take $N_2=1400$ with 25\% (resp. 75\%) 
before (resp. after) exposure. In \Cref{fig:exp_2_PCA}, we focus on the first principal component PC since
it encodes the effect of the irritant molecule as depicted in \Cref{fig:exp_2_PCA}-C) (the exposure occurs at 20 minutes).
\Cref{fig:exp_2_PCA}-B) shows that the deformation (+3 $\sigma_{\text{PC}}$) leads to longer respiratory cycles that include pauses,
as observed in \cite{germain2023unsupervised}. As well, \Cref{fig:exp_2_PCA}-A) shows that TS-LDDMM features distributions are less spread 
out for \textbf{colq} mice compared to \textbf{wt} mice. Indeed, the irritant molecule inhibits the action of the 
deficient enzyme, \textbf{wt} mice strongly react to the irritant molecule, whereas \textbf{colq} mice are better 
adapted due to their deficiency.
%Secondly, we compare breathing behaviors before and after exposure to observe the impact of the irritant molecule.
% We follow the same procedure as for before exposure, but we take $N_2=1400$ respiratory cycles extracted according 
% to the procedure CITE....
% In \Cref{fig:exp_2_PCA}, we focus on the first Principal Component (PC) since it encodes the effect of the irritant molecule as demonstrated on Figure \ref{fig:exp_2_PCA}-C) (exposure at time 20).
% We observe on \Cref{fig:exp_2_PCA}-B) that after exposure the mouse have longuer breath such that the expiration is longuer than inspiration.
%  Moreover, we deduce from Figure \ref{fig:exp_2_PCA}-A) that \textbf{colq} are more constant in their breath compared to \textbf{wt} after exposure.
%
% In Figure REF, we 
% also display the reference respiratory cycle S0 and its deformations in the principal component directions.
%  Additionally, we learn each mouse's reference respiratory cycle and represent them in the first and third PC coordinates system in Figure REF.

\subsection{Quantitative performances of the TS-LDDMM representation in classification}
Combined with a Support Vector Classifier (SVC) \cite{hsu2003practical}, TS-LDDMM representation can be used for 
classification tasks using the kernel associated with the initial velocity space $\msv$.
We compare TS-LDDMM-SVC classification performances with another SVC using representation 
learned with T-loss \cite{franceschi2019unsupervised}, an unsupervised deep learning feature 
representation method for time series. We also include fully supervised methods in deep learning 
-ResNet, CNN \cite{ismail2019deep}- and machine learning: Catch22 \cite{lubba2019catch22}, Rocket
\cite{dempster2020rocket}, Dynamic Time Wrapping k-Nearest Neighbors (DTW-kNN) 
\cite{muller2007dynamic}. Methods are compared using f1-score on several shape-based UCR/UEA datasets 
\cite{dau2019ucr,bagnall2018uea} introduced in \Cref{appendix:classification_dataset}. All implementation details are given 
in \Cref{appendix:classification_implementation}.
\Cref{table:classification} presents the reuslts. TS-LDDMM-SVC consistently outperforms the other unsupervised methods. It is ranked 1,3,4,3 for all methods combined, 
demonstrating its competitiveness as an unsupervised method on time series dataset homogeneous regarding shape.
%TS-LDDMM representation can be used in a Support Vector Classifier (SVC) \cite{hsu2003practical} using the kernel of the space of initial velocity $\msv$.
%We compare its capacity of representation to an SVC using T-Loss representation \cite{franceschi2019unsupervised}, which is an unsupervised representation deep learning method for time series,
%and to others fully supervised methods using deep learning -ResNet, CNN \cite{ismail2019deep}- or not: Catch22 \cite{lubba2019catch22}, Rocket \cite{dempster2020rocket}, Dynamic Time Wrapping \cite{muller2007dynamic} k-Nearest Neighbors (DTW-kNN).
%These methods are compared using f1-score on several datasets introduced in \Cref{appendix:classification_dataset}. All details of implementations are given in \Cref{appendix:classification_implementation}.
%TS-LDDMM-SVC is always ranked 2 or 3 behind a fully supervised method, demonstrating its competitiveness as an unsupervised method.
%The more homogeneous the shapes in the dataset are, the better TS-LDDMM-SVC performs.

\begin{table}[t]
  \vspace{-0.5em}
  \caption{Classification results in f1-score (U: unsupervised, S: supervised, DL: deep learning, ML: machine learning). \textbf{x} best unsupervised method, \underline{x} best supervised method.}
  \resizebox{\linewidth}{!}{%
  \begin{tabular}{lllrrrr}
    \toprule
     &  &  & ArrowHead & ECG200 & GunPoint & NATOPS \\
    \midrule
    \multirow[m]{3}{*}{U} & \multirow[t]{3}{*}{} & TS-LDDMM-SVC & \textbf{0.84} & \textbf{0.82} & \textbf{0.94} & \textbf{0.93} \\
     &  & T-loss-SVC & 0.57 & 0.76 & 0.82 & 0.88 \\
     &  & DTW-kNN & 0.70 & 0.75 & 0.91 & 0.88 \\
    \cline{1-7} 
    \multirow[m]{5}{*}{S} & \multirow[m]{2}{*}{DL} & CNN & 0.70 & 0.79 & 0.85 & \underline{0.96} \\
     &  & ResNet & 0.77 & 0.87 & 0.97 & 0.95 \\
    \cline{2-7}
     & \multirow[m]{2}{*}{ML} & Catch22 & 0.73 & 0.81 & 0.96 & 0.89 \\
     &  & Rocket & \underline{0.81} & \underline{0.91} & \underline{1.00} & 0.88 \\
    \bottomrule
    \end{tabular}
    \label{table:classification}
  %
}
\vspace{-2em}
  \end{table}
\section{Conclusion}
In this paper, we propose a feature representation method, TS-LDDMM, designed for 
shape comparison in homogeneous time series datasets. We show on a real dataset 
its ability to study, with high interpretability, the inter-individual shape 
variability. As an unsupervised approach, it is user-friendly and enables knowledge 
transfer for different supervised tasks such as classification. Although TS-LDDMM 
is already competitive for classification, its performances can be leveraged on 
more heterogeneous datasets using a hierarchical clustering extension, which is relagated for future work.

\section{Societal impact}
This paper presents work whose goal is to advance the field of Machine Learning. There are many potential societal consequences of our work, none which we feel must be specifically highlighted here.
% \section{Acknowledgements}

%The TS-LDDMM feature representation proposed in this paper enables us to study the inter-individual variability of shapes in a time series dataset with high interpretability.
%Its unsupervised aspect is user-friendly and enables knowledge transfer for different supervised tasks such as classification.
%Although the method is already competitive,
% its capacity can be leveraged on more heterogeneous datasets using a hierarchical clustering extension.
% This is relegated to future works.

%Applying NF or CNF to our
 %  framework with high dimensional time series is relagated for future works.

% \begin{figure}[ht]
% \vskip 0.2in
% \begin{center}
% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
% \caption{Historical locations and number of accepted papers for International
% Machine Learning Conferences (ICML 1993 -- ICML 2008) and International
% Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
% produced, the number of accepted papers for ICML 2008 was unknown and instead
% estimated.}
% \label{icml-historical}
% \end{center}
% \vskip -0.2in
% \end{figure}

% \begin{algorithm}[tb]
%    \caption{Bubble Sort}
%    \label{alg:example}
% \begin{algorithmic}
%    \STATE {\bfseries Input:} data $x_i$, size $m$
%    \REPEAT
%    \STATE Initialize $noChange = true$.
%    \FOR{$i=1$ {\bfseries to} $m-1$}
%    \IF{$x_i > x_{i+1}$}
%    \STATE Swap $x_i$ and $x_{i+1}$
%    \STATE $noChange = false$
%    \ENDIF
%    \ENDFOR
%    \UNTIL{$noChange$ is $true$}
% \end{algorithmic}
% \end{algorithm}

\bibliography{ref}
\bibliographystyle{icml2023}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn

\section{Proofs}
\label{appendix:proofs}
Denote by $\msg(s)\triangleq \{ (t,s(t)): \eqsp t\in \msi \} $ the graph of a time series $s: \msi \to \Rset^d$ and $ \phi.\msg(s)\triangleq\{ \phi(t,s(t)): \eqsp t\in \msi\} $ the action of  $\phi\in \mcd(\Rset^{d+1}) $ on $\msg(s)$.
\begin{theorem}
    \label{theorem:representation_proof}
Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi,\msj$ two intervals of $\Rset$.
 There exist $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\gamma\in  \mcd(\Rset) $ such that $\gamma(\msi)=\msj $ and $\Phi_f\in \mcd(\Rset^{d+1})$,
 \begin{equation}% TO add more detail on functionnal space
    \msg(s)= \Pi_{\gamma,f}.\msg(\mathbf{s}_0),\eqsp \Pi_{\gamma,f}=\Psi_\gamma\circ\Phi_f.
 \end{equation}
 Moreover, for any $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\bar{\gamma}\in  \mcd(\Rset) $, there exists a continously differentiable time series $\bar{s}$ such that 
 $\msg(\bar{s})= \Pi_{\bar{\gamma},\bar{f}}.\msg(\mathbf{s}_0)$
\end{theorem}
%\sam{preuve todo, classement des variétés à bord, donne un homeomorphisme sur la restriction}
\begin{proof}
  Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi=(a,b),\msj=(\alpha,\beta)$ two intervals of $\Rset$.
  By setting $\gamma: t\in \Rset \mapsto (\beta-\alpha)(t-a)/(b-a)+\alpha\in \Rset $, we have $ \gamma(\msi)=\msj$ and $\gamma \in \mcd(\Rset) $.
   By defining $f:(t,x)\in\Rset^{d+1}\mapsto x-\mathbf{s}_0(t)+s\circ \gamma(t) $, the map $\Phi_f\in \mcd(\Rset^{d+1})$,
    indeed, its inverse is $\Phi_f^{-1}:(t,x)\in\Rset^{d+1}\mapsto (t,x+\mathbf{s}_0(t)-s(t)) $ and is continuously differentiable.
     Moreover, we have $\Pi_{\gamma,f}.\msg(\mathbf{s}_0)=\{(\gamma(t),s\circ \gamma(t)):\eqsp t\in\msi \}=\msg(s) $.

    Let $\bar{f}\in \rmC^0(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^0(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
    We have :
    \begin{align}
      \Pi_{\gamma,f}.\msg(\mathbf{s}_0)&=\{(\gamma(t),f(t,\mathbf{s}_0(t))),\eqsp t\in \msi \} \\
      &\label{eq:proof1_last_eq}=\{(t,f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right),\eqsp t\in \gamma(\msi) \} \eqsp .
    \end{align}
    By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^0(\gamma(\msi), \Rset^d) $ by composition of continuous functions
    and $ \msg(\bar{s})= \Pi_{\gamma,f}.\msg(\mathbf{s}_0)$ by \eqref{eq:proof1_last_eq}, which concludes the proof.
\end{proof}
\begin{lemma}
  If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0},
   there exist $\gamma \in \msd(\Rset) $ and $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ such that $\phi^v=\Psi_\gamma\circ\Phi_f $.
\end{lemma}
\begin{proof}
  Let $v$ be a vector field generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0}.
 We remark that the first coordinate of the velocity field $v_\tau$ denoted by $v_\tau^{\text{time}}$ only depends on the time variable $t$ for any $\tau\in[0,1]$.
 Thus, when computing the first coordinate of the deformation $\phi^v$, denoted by $\gamma$, we integrate \eqref{eq:LDDMM_dynamic} with $v_\tau$ replaced by $v_\tau^{\text{time}}$,
  thus $\gamma$ is independant of the variable $x$. Moreover, $\gamma\in \mcd(\Rset)$ since a Gaussian kernel induced an Hilbert space $\msv$ satisfying $|f|_V\leq |f|_\infty+ |\dd f|_\infty  $ for any $f\in \msv$ by \citep[Theorem 9]{glaunes2005transport}.
  For the same reason, we have $\phi^v\in \mcd(\Rset^{d+1})$, and thus its last coordinates denoted by $f$ belongs to $\rmC^1(\Rset^{d+1},\Rset^d)$, and by construction $\phi^v=\Psi_\gamma\circ\Phi_f $.
\end{proof}

\section{Oriented varifold}

\label{appendix:varifold}
In this section, we introduce the \textit{oriented varifold} associated with curves.
For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}. 
We associate to $\gamma\in \rmC^1((a,b),\Rset^{d+1})$ an \textit{oriented varifold} $\mu_\gamma$, i.e. a distribution on the space $\Rset^{d+1}\times \mathbb{S}^{d}$ defined as follows, for any smooth test function $\omega :\Rset^{d+1}\times \mathbb{S}^{d}\to \Rset  $,
\begin{equation}
  \mathbb{E}_{Y\sim \mu_\gamma}\left[\omega(Y)\right]=\mu_\gamma(\omega)=\int_a^b \omega\left(\gamma(t),\frac{\dot{\gamma}(t)}{|\dot{\gamma}(t)|}\right)|\dot{\gamma}(t)|\dd t \eqsp .
\end{equation}
Denoting by $\msw$ the space of smooth test function, we have that $\mu_\gamma $ belongs to its dual $\msw^*$.
 Thus, a distance on $\msw^*$ is sufficient to set a distance on oriented varifolds associated to curve and thus on $\rmC^1((a,b),\Rset^{d+1})$ by the identification $\gamma\to \mu_\gamma $.
Remark that in (TS-LDDMM), $\gamma$ should be the parametrization of a time series' graph $\msg(s)$, i.e. $\gamma: t\in \msi \to (t,s(t))\in\Rset^{d+1} $ denoting by $s:\msi \to \Rset^d$ the time series.
However, in practice, we work with discrete objects.
 That is why, we set $W$ as an RKHS to use its representation theorem.
 More specifically \citep[Proposition 2 \& 4]{kaltenmark2017general} encourages us to consider a kernel $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$ such that there exist two positive and continuously differentiable kernels $k_{\pos}$ and $k_{\dir}$, 
 such that for any $(x,\overrightarrow{u}),(y,\overrightarrow{v}) \in (\Rset^{d+1} \times \mathbb{S}^d)^2$
 \begin{equation}
     k((x,\overrightarrow{u}),(y,\overrightarrow{v})) = k_{\pos}(x,y)k_{\dir}(\overrightarrow{u},\overrightarrow{v}) \eqsp,
   \end{equation}
 with moreover $k_\dir>0$ and $ k_\pos$ which admits an RKHS $\msw_\pos$ dense in the space of continous function on $\Rset^{d+1}$ vanishing at infinite \cite{carmeli2010vector}.

Given such a kernel $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$
verifying \citep[Proposition 2 \& 4]{kaltenmark2017general}, we have that for any $(x,v)\in \Rset^{d+1}\times \mathbb{S}^{d}$, $\delta_{(x,\overrightarrow{v})}$ belongs to $\msw^*$ as a distribution and that the dual metric $\langle\cdot,\cdot \rangle_{\msw^*} $ satisfies for any $(x_1,v_1),(x_2,v_2)\in \left(\Rset^{d+1}\times \mathbb{S}^{d}\right)^2$,
\begin{equation}
  \langle\delta_{(x_1,\overrightarrow{v}_1)},\delta_{(x_2,\overrightarrow{v}_2)} \rangle_{\msw^*} =k((x_1,\overrightarrow{v}_1),(x_2,\overrightarrow{v}_2)) \eqsp .
\end{equation}
Thus, given two sets of triplets $X=(l_i,x_i,\overrightarrow{v}_i)_{i\in[T_0-1]}\in (\Rset\times \Rset^{d+1}\times \mathbb{S}^d)^{T_0-1},Y=(l_i',y_i,\overrightarrow{w}_i)_{i\in[T_1]}\in (\Rset\times \Rset^{d+1}\times \mathbb{S}^d)^{T_1-1}$ and denoting by 
\begin{equation}
  \label{eq:def_mu_X}
  \mu_X=\sum_{i=1}^{T_0} l_i \delta_{(x_i,\overrightarrow{v}_i)},\mu_Y=\sum_{i=1}^{T_1} l_i' \delta_{(y_i,\overrightarrow{w}_i)} \eqsp,
\end{equation}
 we have,
\begin{equation}
   |\mu_X-\mu_Y |_{\msw^*}^2 = \sum_{i,j = 1}^{T_0-1}l_i k((x_i,\overrightarrow{v_i}),(x_i,\overrightarrow{v_i}^0))l_j - 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l_i k((x_i,\overrightarrow{v_i}),(y_i,\overrightarrow{w_i}))l_j' + \sum_{i,j = 1}^{T_1-1}l_i' k((y_i,\overrightarrow{w_i}),(y_i,\overrightarrow{w_i}))l_j' \eqsp.
\end{equation}
Then, using the identification $X\to \mu_X, Y\to \mu_Y$, we can define a distance on sets of triplets as $d_{\msw^*,3}(X,Y)=|\mu_X-\mu_Y|_{\msw^*}^2$.

Now, we aim to discretize the oriented varifold $\mu_\msg $ related to a time series' graph $\msg(s)$ by using a set of triplets.
This is carried out by using a discretized version of $\msg(s) $, i.e. $\tilde{\msg}=(g_i=(t_i,s(t_i)))_{i\in[T]}\in(\Rset^{d+1})^T$, in the following way: 
For any  $i\in[T-1]$, denoting the center and length of the $i^{th}$ segment $[g_i,g_{i+1}]$ by
$c_i = (g_i + g_{i+1})/2$, $l_i = \| g_{i+1}-g_{i}\|$, and the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$ by
 $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$, we define the set of triplets $X(\tilde{\msg})=(l_i,c_i,\overrightarrow{v_i})_{i\in[T-1]}$ and its related oriented varifold $\mu_{X(\tilde{\msg})}= \sum_{i=1}^{T-1}l_i \delta_{c_i,\overrightarrow{v_i}}$ as in \eqref{eq:def_mu_X}.
 This is a valid discretization of the oriented varifold $\mu_\msg$ according to \citep[Proposition 1]{kaltenmark2017general}: $\mu_{X(\tilde{\msg})}$ converges towards $\mu_\msg$ as the size of the descretization mesh $\sup_{i\in[T-1]} |t_{i+1}-t_i| $ converges to 0.

 Finally, we define a distance on discretized time series' graphs $\tilde{\msg}_1,\tilde{\msg}_2$ as $d_{\msw^*}(\tilde{\msg}_1,\tilde{\msg}_2)=d_{\msw^*,3}(X(\tilde{\msg}_1),X(\tilde{\msg}_2)) $.

\subsection{Varifold kernels}
\label{appendix:kernel_implementation}
Denote the one-dimensional Gaussian kernel by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
In the implementation, we use the following kernels, for any $((t_1,x_1),(t_2,x_2))\in (\Rset^{d+1})^2, ((w_1,v_1),(w_2,v_2))\in (\mathbb{S}^{d})^2 $,
\begin{equation}
  k_{\pos}(x,y)=K_{\sigma_{\pos,t}}^{(1)}(t_1,t_2)K_{\sigma_{\pos,x}}^{(d)}(x_1,x_2),\quad k_{\pos}(x,y)=K_{\sigma_{\dir,t}}^{(1)}(w_1,w_2)K_{\sigma_{\dir,x}}^{(d)}(v_1,v_2)\eqsp ,
\end{equation}
where $\sigma_{\pos,t}, \sigma_{\pos,x}, \sigma_{\dir,t}, \sigma_{\dir,x}>0 $ are hyperparameters.
 In practice, we select $\sigma_{\pos,x}\approx \sigma_{\dir,x} \approx 1$ when the times series are centered and normalized. Otherwise we select $\sigma_{\pos,x}\approx \sigma_{\dir,x} \approx \bar{\sigma}_{s}$ with $\bar{\sigma}_{s}$ the average standard deviation of the time series.
  We choose $\sigma_{\pos,t}\approx \sigma_{\dir,t} = m f_e$ with $f_e$ the sampling frequency of the time series and $m\in [5]$ an integer depending on the time change between the starting and the target time series graph.
  The more significant the time change, the higher $m$ should be. The intuition comes from the fact that the width $\sigma_{\pos,t}, \sigma_{\dir,t}$ rules the time windows used to perform the comparison, and $\sigma_{\pos,x}, \sigma_{\dir,x}$ affects the space window.
   The size of the windows should be selected depending on the variations in the data.

\section{Tuning the hyperparameters of the TS-LDDMM kernel given in \eqref{eq:kernel_TAS}}
\label{appendix:kernel_TS_LDDMM}
The parameter $\sigma_{T,0}$ should be chosen \textit{large} compared the sampling frequency $f_e$ and compared to average standard deviation $\bar{\sigma}_s$ of the time series, e.g $\sigma_{T,0}=100$ as $\bar{\sigma}_s\approx f_e\approx  1 $.
It makes the time transformation smoother. If $\sigma_{T,0}$ is too small, for instance, $\sigma_{T,0}=f_e $, the effect of the time deformation is too localized, and there are not enough samples to make it visible.

The parameter $\sigma_{T,1}$ should be of the same order as $f_e$: two different points in time can have various space transformations.
 $\sigma_x$ should be of the same order of $\bar{\sigma}_s$: two points with a big difference regarding space compared to $\bar{\sigma}_s$ can have very different space transformations.

 We take $c_0\approx 10 c_1 $, we want to encourage time transformation before space transformation. We take $(c_0,c_1)=(1,0.1)$ in all experiments.

\section{Numerical details}
A report of all the hyperparameters selected is given in \Cref{appendix:table:set_up_exp_practical}.
\subsection{Optimization details of \eqref{eq:general_optimization_problem}}
\label{appendix:optimizers_details}
\paragraph{Initialization}
At the initialization of \eqref{eq:general_optimization_problem}, all the momentums parameter are set to $0$ and the graph of reference is set to the graph of a time series in the dataset having a mediane samples size.

\paragraph{Gradient descent.}
The chosen gradient descent method is "adabelief" \cite{zhuang2020adabelief} implemented in the library OPTAX\footnote{https://optax.readthedocs.io/en/latest/}.
There are two main parameters in the gradient descent: the number of steps nb\_steps, and the maximum value of step size $\eta_M$.
The stepsize has a particular scheduling: 
\begin{itemize}
  \item Warmup period on $0.1 \times$ nb\_steps steps: the stepsize increases linearly from $0$ to $\eta_M$. The goal is to learn progressively the parameters.
   If the stepsize is too large at the start, smaller steps at the end can't make up for the mistakes made at the beginning. 
  \item Fine tuning periode on $ 0.9  \times$ nb\_steps : the stepsize decreases from $\eta_M$ to $0$ with a cosine decay implemented in the OPTAX scheduler,
   i.e. the decreasing factor as the form $0.5  (1 + \cos(\pi  t/T))$. 
\end{itemize}
The sharper the deformations, the larger the number of steps and the maximum value of step size should be selected.
We suggest nb\_steps=300, $\eta_M=0.1 $ for small deformations and nb\_steps=800, $\eta_M=0.3 $ for big ones (time dilation with a factor $\lambda\geq 2$).

\subsection{Synthetic experiments}
\label{appendix:numerics_synthetic}
For any deformations generation in both experiments (well-specified and misspecified), we take $\sigma_{T,0},\sigma_{T,1},\sigma_x=(100,1,1)$ and $c_0,c_1=(1,0.1)$ for the kernel $K_\msg$
and $\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}=(2,1,2,0.6) $ for the varifold kernels $k_\pos,k_\dir$ related to the loss $\scrl$.

In both experiments, we have nb\_steps=300 abd $\eta_M=0.1$.

\subsection{Mouse experiments}
The number of steps is larger in the second experiment (before/after injection) because the deformations are sharper.

\subsection{Classification experiments}
\label{appendix:classification_implementation}

We defined a default parametrization for all classifiers.

For classifiers: \href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.deep_learning.CNNClassifier.html#aeon.classification.deep_learning.CNNClassifier}{CNN}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.networks.ResNetNetwork.html#aeon.networks.ResNetNetwork}{ResNet}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.feature_based.Catch22Classifier.html#aeon.classification.feature_based.Catch22Classifier}{Catch22}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.distance_based.KNeighborsTimeSeriesClassifier.html#aeon.classification.distance_based.KNeighborsTimeSeriesClassifier}{DTW-KNN}, 
\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.convolution_based.RocketClassifier.html#aeon.classification.convolution_based.RocketClassifier}{Rocket}
we used the aeon\footnote{https://www.aeon-toolkit.org/en/stable/index.html} implementations with their default settings.

For Tloss-SVC we used the implementation provided on github\footnote{https://github.com/mqwfrog/ULTS} 
with the following parameters for learning representations: batch\_size: 10, channels: 40, depth: 10, nb\_steps: 200, in\_channels: 1, 
kernel\_size: 3, lr: 0.001, nb\_random\_samples: 10, negative\_penalty: 1, out\_channels: 320, reduced\_size: 160. 
We used the Support Vector Classifier (\href{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html#sklearn.svm.SVC}{SVC})
from scikit-learn with thee regularization term C: 1. Others parameters are set to default.

For TS-LDDMMM-SVC, all kernels' parameters et optimizer parameter are presented in \Cref{appendix:table:set_up_exp_practical}.
As well, we used the Support Vector Classifier from scikit-learn with thee regularization term C: 1. Others parameters are set to default.

 \begin{table}[hbt!]
  \caption{Parameters used in all the experiments.
   For synthetic data, $K_\msg$ refers to the kernel used in the generation, which is the same for the estimation only in the well-specified case. $\bar{l}$ refers to the average time series length and $N_d$ refers to the number of dimensions.}
  \centering
  \begin{tabular}{lllll}

            \toprule 
            objects & Optimizer &$k_\pos,k_\dir$ & $K_\msg $ \\
            \midrule
            Parameter & (nb\_steps,$\eta_M$) &$(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$ & $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)$ \\
            \midrule
            Synthetic data well-specified &(300,0.1) &$(2,1,2,0.6)$& $(1,0.1,100,1,1)$   \\
            Synthetic data misspecified  & (300,0.1)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
            Mouse before injection  & (400,0.3)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
            Mouse before/after injection  & (800,0.3)  & $ (5,1,5,0.6)$ &$(1,0.1,150,1,1)$  \\
            classification  & (400,0.1)  & $(\max(2,0.03\bar{l}),N_d,\max(2,0.03\bar{l}),0.6)$  & $(1,0.1,0.33\bar{l},1,N_d)$  \\

            \bottomrule
          \end{tabular}
        \label{appendix:table:set_up_exp_practical}
    \end{table}
\section{Mouse respiratory dataset}

\begin{figure}
  \centering
  \includegraphics[width = \linewidth]{pictures/mice_exp.png}
  \caption{A: Illustration of a double-chamber plethysmograph. The term \textit{dpt} stands for differential 
  pressure transducer which measures the pressure in each compartment, the pressure then being converted to flow. 
  B: Nasal airflow (top) and lung volume (bottom). During inspiration, airflow is positive (grey) and during
  expiration, airflow is negative (pink).}
  \label{fig:mice_exp}
\end{figure}

\label{appendix:mouse_dataset}
Ventilation is a simple physiological function that ensures a vital supply of oxygen and the elimination of CO2. 
Acetylcholine (Ach) is a neurotransmitter that plays an important role in muscular activity, notably for breathing. 
Indeed, muscle contraction information passes from the brain to the muscle through the nervous system. Achs are located 
in synapses of the nervous system (central and peripheral) and skeletal muscles. They ensure the information transmission 
from nerve to nerve. However, the transmission cannot end without the hydrolysis of Ach by the enzyme Acetylcholinesterase 
(AchE), allowing nerves to return to their resting state. Inhibition of (AchE) with, for instance, nerve gas, pesticide, 
or drug intoxication leads to respiratory arrests.

The dataset comes from the experiment \cite{nervo2019respiratory}, where they studied the consequences of partial 
deficits in AChE and AChE inhibition on mice respiration. AchE inhibition was induced with an 
irritant molecule called physostigmine (an AchE inhibitor). Mice nasal airflows were sampled at 
2000Hz with a Double Chamber plethysmograph \cite{hoymann2012lung}, as depicted in \Cref{fig:mice_exp}-A). The flow is expressed in 
$ml.s^{-1}$; it has a positive value during inspiration and a negative value expiration \Cref{fig:mice_exp}-B). 
Among the mice population, we selected 7 control mice (\textbf{wt}) and 7 ColQ mice (\textbf{colq}), which do not have 
AChE anchoring in muscles and some tissues. 
As described in \cite{nervo2019respiratory}, mice experiments were as follows:
\begin{enumerate}
  \item The mouse is placed in a DCP for 15 or 20 min to serve as an internal control.
  \item The mouse is removed from the DCP and injected with physostigmine.
  \item The mouse is placed back into the DCP, and its nasal flow is recorded for 35 or 40 min.
\end{enumerate}

Respiratory cycles were extracted following procedure \cite{germain2023unsupervised}. We removed 
respiratory cycles whose duration exceeds 1 second; the average respiratory cycle duration is 
300 ms. We randomly sampled 10 respiratory cycles per minute and mouse. It leads to a dataset of 
12,732 (time, genotype)-annotated respiratory cycles.

\section{Classification datasets}
\label{appendix:classification_dataset}

All datasets were taken from UCR/UEA archives \cite{dau2019ucr,bagnall2018uea}. Among all available 
datasets\footnote{https://timeseriesclassification.com}, we selected 4 datasets related to time series 
shape comparison. All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/} 
which already includes the train test split. Essential dataset information is summarized in \Cref{appendix:table:datasets}.

\begin{table}[hbt!]
  \centering
  \caption{Time series datasets summary for shape based classification.}
  \begin{tabular}{lllllll}
    \toprule
    Dataset & Train size & test size & Lenghth & Number of classes & Number of dimensions & Type \\
    \midrule
    \href{https://timeseriesclassification.com/description.php?Dataset=ArrowHead}{ArrowHead} & 36 & 175 & 251 & 3 & 1 & IMAGE \\
    \href{https://timeseriesclassification.com/description.php?Dataset=ECG200}{ECG200} & 100 & 100 & 96 & 2 & 1 & ECG \\
    \href{https://timeseriesclassification.com/description.php?Dataset=GunPoint}{GunPoint} & 50 & 150 & 150 & 2 & 1 & MOTION \\
    \href{https://timeseriesclassification.com/description.php?Dataset=NATOPS}{NATOPS} & 180 & 180 & 51 & 6 & 24 & MOTION \\
    \bottomrule
  \end{tabular}
  \label{appendix:table:datasets}
\end{table}

%  [11:35, 31/01/2024] Thibaut Germain: avant/après :
%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,150,1,1)
%  Kl = TSGaussGaussKernel(5,1,5,0.6)
%  dataloss = VarifoldLoss(Kl)
%  schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
%  [11:35, 31/01/2024] Thibaut Germain: avant :
%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,100,1,1)
%  Kl = TSGaussGaussKernel(2,1,2,0.6)
%  dataloss = VarifoldLoss(Kl)
%  schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
% Let $s$ be a time series and $\msg(\tilde{s}) = (g_i)_{i \in [n]}$ be its discretized graph.
%  The graph $\msg(s)$ can be approximated as 
% the union of piecewise linear segments between time-consecutive samples, 
% and we associate to the approximated graph the oriented varifold:

% \begin{equation}
%   \mu_{\msg(\tilde{s})} : w \in \msw \mapsto \sum_{i=1}^{n-1} l_i\delta_{(c_i,\overrightarrow{v_i})}(w)
% \end{equation}
% where $\msw$ is a set of real valued test functions defined on $\Rset^{d+1} \times \mathbb{S}^d$, 
% $\delta_{(c_i,\overrightarrow{v_i})}(w) = w(c_i,\overrightarrow{v_i})$ is a Dirac delta function, 
% $c_i = (g_i + g_{i+1})/2$ (resp. $v_i = \| g_{i+1}-g_{i}\|$) is the center (resp. length) of the $i^{th}$ segment, 
% and $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$ is the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$.

% Assuming that the test functions' space $\msw$ is the RKHS associated with a $C^1$ positive definite kernel $k$ 
% that verifies \citep[Proposition 2 \& 4]{kaltenmark2017general}, the oriented varifold belongs to the dual space $\msw^*$. Additionally, 
% we can define a distance $d_{\msw^*}$ between signals' graph sample sets, which, thanks to the kernel reproducing property,
% has an explicit formulation:

% \section{Continuous normalizing flows}
% %Modeling Continuous Stochastic Processes with Dynamic Normalizing Flows time series generative
% CONTUNYIYS NORMALIZING FLOW bofbof
% Here, we present the framework of Continuous Normalizing Flows (CNF) and show how LDDMM can be seen as a particular case.
% \cite{salman2018deep} was already inspired from the work of shape analysis to present deep diffeomorphic normalizing flows.
% Discussion with Alain :
% Given a time-dependant vector field $v:[0,1]\times \Rset^{d'} \to \Rset^{d'} $, the integration of this vector field gives a flow $ \phi:[0,1]\times \Rset^{d'} \to \Rset^{d'}$, defined via the ordinary differential equation (ODE),
% \begin{equation}
%   \frac{\dd }{\dd \tau}\phi_\tau(x)=v(\tau,\phi_\tau(x)), \phi_0(x)=x \eqsp.
%  \end{equation}
% In \cite{chen2018neural}, they introduce Continous Normalizing flows for generative modeling.
%  By modeling $v=f_\theta$ is a neural networks and choosing a prior $X_0\sim\rho_0$ on the inital value, the law of $\phi_1(X_0)$ has the pullback density $\phi_1\# \rho_0$ and the parameter $\theta$ is optimized to maximize the log-likelihood of a given dataset $(y_i)_{i\in[N]}\in (\Rset^{d'})^N$ which is assumed to follow the law of $\phi_1(X_0)$.
% Instead of modeling the law of the end point $\phi_1(X_0) $, we can try to model the law of the whole path $(\phi_\tau(X_0))_{\tau\in[0,1]} $ having a density $p:[0,1]\times \Rset^d\to \Rset_{>0}$ such that $p_1=q, p_0=\rho_0$ where $q$ is the empirical distribution in the dataset, as proposed in \cite{lipman2022flow}.
% In so doing, the problem is ill-posed since $p_t$ can be whatever as soon as $0<t<1$, that is why we try to constraint the learned velocity field $v=f_\theta$ such that the path law $p$ is tractable.
% One way of doing it is to look for geodesics \eqref{eq:geodesics_original} as proposed in LDDMM, it corresponds to the case where $q=\sum_{i=1}^N \delta_{y_i}/N$ and $\rho_0=\sum_{i=1}^N \delta_{x_i}/N$ are empirical distribution and $y_i=\phi^v_1(x_i) $ for a given velocity field $v$.

% The goal is to target $\nu\in \mathcal{P}(\rmC^0([0,1]\times \Rset^d, \Rset^d)) $ by generating randomness on a initial value $x \sim \delta_x$
%  and by integrating a velocity field $v_\theta \in \msw \subset \rmC^0([0,1]\times \Rset^d, \Rset^d)$: 
%  \begin{equation}
%   \frac{\dd h}{\dd \tau}=v_\theta(\tau,h(\tau,x)), h(0,x)=x \eqsp,
%  \end{equation}
%  denoting by $\mu$ the law of $(h(\tau,x))_{\tau \in [0,1]}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \subsection{TTS}
% To circumvant this issue, we propose two methods :
%       \begin{enumerate}
%         \item \label{enum:first_method} [Time And Space (TAS)] Using a specific form of the RKHS's kernel in LDDMM, to stick to the classical shape analysis framework.
%         \item \label{enum:second_method} [Time Then Space (TTS)] Tackling the time axis and space axis separetely in the transformation.
%          This distangles time and space variability in the patterns, increasing the interpretability of the representation.
%           Although the related optimization is more complex than method \ref{enum:first_method}, the results is more robust to the choice of hyper-parameter related to the RKHS's kernel.
%       \end{enumerate}
%        In the following, we present both methods to highlights their pro and cons.

% As described in \Cref{section:optimization}, the method (TAS) learns $\Pi_{\gamma,f}$ directly, while (TTS) find the temporal reparametrisation $\Phi_f$ then the space transformation $\Psi_\gamma$.
%     % If we define a diffeomorpshim $\Phi_\gamma$ as $(\gamma,\Id_d) $, we have $\Phi_\gamma.G(s) =G(s).\gamma^{-1}$.
%     %  To increase the variability, we allow some deformations on the graph of the time series after being reparametrized in time, which allow us to recover every possible time seriess.
%     % Remark that, $G(\mathbf{s}_0\circ \psi^{-1})=\{(t,\mathbf{s}_0\circ \psi^{-1}(t)), t\in \msj \}= \{(\psi(t),\mathbf{s}_0(t)) ,\eqsp t\in \msi \}  $.
%     % [The action of the diffeomorpshim group 
%     % \begin{align}
%     %     &\mcd_*=\mcd(\Rset,\Rset)\times \{\phi\in \mcd(\Rset^{d+1},\Rset^{d+1}): \\
%     %     &\phi(t,x)=(t,f(t,x)),
%     %      (t,x)\in \Rset \times \Rset^{d}, f \in C^0(\Rset^{d+1}, \Rset^d) \} 
%     % \end{align}
%     % on the space of time series graph $\{G^{(a,b)}(s), s\in C^0((a,b), \Rset), (a,b)\in \Rset_+^2, a<b  \}$
%     % is homogene.]
%      %In the next section we define a metric and a norm on this space in order to find a representation $(\phi,\psi)$ with a minimal norm, and we show in (...) how to recover a proxy.
%      \paragraph{The representation as the solution of an optimization}
%      By \Cref{theorem:representation}, for any $\mathbf{s}_0\in C^0(\msi,\Rset^d)$, we can find $ F=(\gamma_j,f_j)_{j\in[N]}$ such that $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$ for any $j\in[N]$.
%      However, to get a meaningfull representation and to tackle the unicity problem, we should optimize the choice of $\mathbf{s}_0 $ and $F$, formally speaking, denoting by $R(\cdot) $ a regularization norm on $\mcd(\Rset^{d+1})$ and $L$ a loss on sets which will be specify later, we aim to solve
%      \begin{equation}
%       \label{eq:minimization}
%       \underset{\mathbf{s_0},(\gamma_j,f_j)_{j\in[N]}}{ \argmin} \sum_{j=1}^N \lambda R(\Pi_{\gamma_j,f_j})+ L(\msg(s^j),\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)) \eqsp ,
%     \end{equation} 
%     where $\lambda>0$ is a penalization factor related to $L$ to relax the condition $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$.

%     In practice, as already proposed in shape analysis, we perform an alternative minimization between $\mathbf{s_0}$ and $(\gamma_j,f_j)_j$.
%     The optimization on $(\gamma_j,f_j)_j$ is parrallelisable, for each couple $(\mathbf{s}_0,s^j)$, we should learn diffemorphisms of minimal norm satisfying a condition. 
%     To this end, we adapt the geodesic shooting method using LDDMM presented in \cite{durrleman2013sparse} to our special case.
%       Before to describe how to optimize the whole in \Cref{section:optimization}, we introduce LDDMM in the next section to specify the norm on diffemorphisms.

% \paragraph{Kernel choices}
% As depicted on figure (mettre ref), we can not use any kernel $K$ to apply the previous methodology to learn deformations on time series' graphs.

% In the following, we expose two choice of RKHS's kernel which answers this issue : (TAS) and (TTS).
%  We rely on the Gaussian kernel, denoting by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
% To apply (TAS) in LDDMM with $d'=d+1$, we use the following anisotropic Gaussian kernel  :
% \begin{align}
%   \label{eq:kernel_TAS}
%   &K_{\text{TAS}}((t,x),(t',x'))=\begin{pmatrix}
%     c_0K & 0 \\
%     0 & c_1 D 
%     \end{pmatrix} \eqsp , \\
%     &D=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
% \end{align}
% parametrized by the widths $\sigma_{T,0},\sigma_{T,1},\sigma_x>0$ and the constants $c_0,c_1>0$.
%  If we use this kernel, by considering \eqref{eq:specific_form}, we remark that the first coordinate of the velocity field $v_\tau$ only depends on the time variable $t$.
%   It implies that the final deformation will have the form $\phi^v=(\gamma,f) $ with  $f\in C^1(\Rset^{d+1},\Rset^d) $, and $\gamma\in \mcd(\Rset)$
%   since the first coordinate follows a LDDMM-ODE \eqref{eq:LDDMM_dynamic}.

% To apply (TTS), we use LDDMM on the time axis with $d'=1$ by taking $K_{\text{TTS}}^{\text{time}}=K_{\sigma_{T,0}}^{(1)}$, then on the space axis with $d'=d$ by choosing
% $K_{\text{TTS}}^{\text{space}}=K_{\sigma_{T,1}}^{(1)}\times K_{\sigma_x}^{(d)} $ where the widths are the same than in \eqref{eq:kernel_TAS}. The alternate optimization will be more detailed in \Cref{section:optimization}.

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018, and modified by Alexandre Bouchard in
% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
% Modified again in 2023 by Sivan Sabato and Jonathan Scarlett.
% Previous contributors include Dan Roy, Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2024-04-30 NEURIPS24/paper.tex
update paper

--- 2024-04-30 NEURIPS24/old.tex
@@
-%%%%%%%% ICML 2023 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
-
-\documentclass{article}
-
-% Recommended, but optional, packages for figures and better typesetting:
-\usepackage{microtype}
-\usepackage{graphicx}
-%\usepackage{subfigure}
-\usepackage{booktabs} % for professional tables
-\usepackage{multirow}
-\usepackage{caption}
-\usepackage{subcaption}
-%\usepackage{tikz}
-
-% hyperref makes hyperlinks in the resulting PDF.
-% If your build breaks (sometimes temporarily if a hyperlink spans a page)
-% please comment out the following usepackage line and replace
-% \usepackage{icml2023} with \usepackage[nohyperref]{icml2023} above.
-\usepackage{hyperref}
-
-% Attempt to make hyperref and algorithmic work together better:
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-% Use the following line for the initial blind version submitted for review:
-\usepackage{icml2024}
-
-% If accepted, instead use the following line for the camera-ready submission:
-% \usepackage[accepted]{icml2023}
-
-% For theorems and such
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{mathtools}
-\usepackage{amsthm}
-
-% if you use cleveref..
-\usepackage[capitalize,noabbrev]{cleveref}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% THEOREMS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \theoremstyle{plain}
-% \newtheorem{theorem}{Theorem}[section]
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
-% \theoremstyle{definition}
-% \newtheorem{definition}[theorem]{Definition}
-% \newtheorem{assumption}[theorem]{Assumption}
-% \theoremstyle{remark}
-% \newtheorem{remark}[theorem]{Remark}
-
-% Todonotes is useful during development; simply uncomment the next line
-%    and comment out the line below the next line to turn off comments
-%\usepackage[disable,textsize=tiny]{todonotes}
-\usepackage[textsize=tiny]{todonotes}
-
-% The \icmltitle you define below is probably too long as a header.
-% Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Shapes analysis for time series.}
-
-\input{header}
-\input{def}
-\newcommand\sam[1]{\textcolor{blue}{#1}}
-\newcommand\thi[1]{\textcolor{orange}{#1}}
-\begin{document}
-
-\twocolumn[
-\icmltitle{Shapes analysis for time series.\\
-           International Conference on Machine Learning (ICML 2024)}%a changer
-
-% It is OKAY to include author information, even for blind
-% submissions: the style file will automatically remove it for you
-% unless you've provided the [accepted] option to the icml2023
-% package.
-
-% List of affiliations: The first argument should be a (short)
-% identifier you will use later to specify author affiliations
-% Academic affiliations should list Department, University, City, Region, Country
-% Industry affiliations should list Company, City, Region, Country
-
-% You can specify symbols, otherwise they are numbered in order.
-% Ideally, you should not use this facility. Affiliations will be numbered
-% in order of appearance and this is the preferred way.
-\icmlsetsymbol{equal}{*}
-
-\begin{icmlauthorlist}
-\icmlauthor{Firstname1 Lastname1}{equal,yyy}
-\icmlauthor{Firstname2 Lastname2}{equal,yyy,comp}
-\icmlauthor{Firstname3 Lastname3}{comp}
-\icmlauthor{Firstname4 Lastname4}{sch}
-\icmlauthor{Firstname5 Lastname5}{yyy}
-\icmlauthor{Firstname6 Lastname6}{sch,yyy,comp}
-\icmlauthor{Firstname7 Lastname7}{comp}
-%\icmlauthor{}{sch}
-\icmlauthor{Firstname8 Lastname8}{sch}
-\icmlauthor{Firstname8 Lastname8}{yyy,comp}
-%\icmlauthor{}{sch}
-%\icmlauthor{}{sch}
-\end{icmlauthorlist}
-
-\icmlaffiliation{yyy}{Department of XXX, University of YYY, Location, Country}
-\icmlaffiliation{comp}{Company Name, Location, Country}
-\icmlaffiliation{sch}{School of ZZZ, Institute of WWW, Location, Country}
-
-\icmlcorrespondingauthor{Firstname1 Lastname1}{first1.last1@xxx.edu}
-\icmlcorrespondingauthor{Firstname2 Lastname2}{first2.last2@www.uk}
-
-% You may provide any keywords that you
-% find helpful for describing your paper; these are used to populate
-% the "keywords" metadata in the PDF but will not be shown in the document
-\icmlkeywords{Machine Learning, ICML}
-
-\vskip 0.3in
-]
-
-% this must go after the closing bracket ] following \twocolumn[ ...
-
-% This command actually creates the footnote in the first column
-% listing the affiliations and the copyright notice.
-% The command takes one argument, which is text to display at the start of the footnote.
-% The \icmlEqualContribution command is standard text for equal contribution.
-% Remove it (just {}) if you do not need this facility.
-
-%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
-\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.
-
-\begin{abstract}
-%     The most common way to analyze time series is to find a feature representation manually with expert knowledge or automatically with machine learning.
-% While unsupervised representation learning on time series proves efficient to reveal cluster structure in a dataset,
-%  actual methods lacks precision for analyzing the inter-individual variability whithin a cluster.
-Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
- Individuals have specificities depending on their genes or disease
-  while sharing, on average, a similar pattern in the features of their corresponding time series.
-For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
- In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
- Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
- Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
-  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
-  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
-  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
-  These features can be post-processed with any statistical or machine learning tools. 
-  We demonstrate the advantages of our representation compared to existing methods using synthetic data and real-world examples, motivated by applications in medicine.
-  %  After checking the injectivity of the TS-LDDMM representation on synthetic data (potentially to change), we study respiratory shapes of a mouse dataset by applying Principal components analysis to the features $(\alpha_j)_{j\in[N]} $.
-
-  %  Moreover, we show that the method is competitive with supervised methods for classification, as soon as the dataset is homogene enough.
-  %   The generalisation of the method to heterogene dataset (several clusters) is relagated for future works.
-\end{abstract}
-
 % citep
@@
 %\thi{à reformuler pour faire moins stupide}
+
+\sam{Aller droit au but, se servir de la section related works pour détailler, notamment enlever le premier paragraphe dire directement que l'on veut faire du unsupervised represnetation learning for time series mais en s'intéressant par ticulièrement à la notion de shape.
+Parler du contexte biomédical, définir la shape. Parler de shape analysis, introduire notre problème de représentation, dire que nous on propose une structure adapté aux time series.}
 A time series is a set of values indexed by time \cite{trirat2024universal}.
@@
   \centering
-  \includegraphics[width=\linewidth]{"./pictures/diffeo.jpeg"}
-  \vspace{-2.5em}
+  \includegraphics[width=0.7\linewidth]{"./pictures/diffeo.jpeg"}
+
   \caption{A time series' graph $\msg=\{(t,s(t)): \eqsp t\in\msi\} $ can lose its structure after applying a general diffeomorphism $\phi.\msg$: a time value can be related to two values on the space axis.}
   \label{fig:diffeo}
-  \vspace{-1em}
+
 \end{figure}
@@
   \includegraphics[width=\linewidth]{"./pictures/transport.jpeg"}
-  \vspace{-2.5em}
+
   \caption{LDDMM and TS-LDDMM are applied to ECG data.
@@
   \label{fig:transport}
-  \vspace{-1em}
+
 \end{figure*}
 
+\section{Related Works}
 \section{Notations}
@@
 %Reproducing Kernel Hilbert Space (RKHS)
+\section{Background on LDDMM}
+\label{section:LDDMM}
+In this part, there is no novelty, we simply expose how to
+ learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
+  In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
+
+The basic problem that we consider in this section is the following.
+Given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$\footnote{Note that we denote by $d'\in\nset$ the ambient space}, a set of starting points $\mathbf{x}=(x_{i})_{i\in[T_1]}$ in $\Rset^{d'}$, we aim to find a diffeomorphism $\phi$ such that the finite set of points $\mathbf{y}$ is similar in a certain sense to the set of finite sets of transformed points $\phi \cdot \mathbf{x} =(\phi(x_i))_{i\in[T_1]} $.
+ The function $\phi$ is occasionally referred to as a \textit{deformation}. In general, these sets $\mathbf{x},\mathbf{y}$ are meshes of continous objects, e.g surfaces, curves, images and so on.
+
+% LDDMM is designed to analyzed an existing dataset, while NF and CNF are made to generalize a dataset for data augmentation.
+\paragraph{Representing diffeomorpshims as deformations.}
+Such \textit{deformations} $\phi$ are constructed via differential flow equations, for any $x_0\in \Rset^{d'} $ and $\tau\in[0,1]$:
+\begin{equation}
+  \label{eq:LDDMM_dynamic}
+    \frac{\dd X(\tau)}{\dd \tau}= v_\tau(X(\tau)), \quad X(0)=x_0\eqsp ,
+    \phi^v_\tau(x_0)=X(\tau), \quad \phi^v=\phi^v_1  \eqsp ,
+\end{equation}
+where the velocity field is $v:\tau\in [0,1]\mapsto v_\tau\in \msv $
+and $\msv$ is a Hilbert space of continuously differentiable function
+on $\Rset^{d'}$.  If
+$||\dd u ||_{\infty}+|| u ||_{\infty}\leq ||u ||_\msv $ for any
+$u\in \msv$ and
+$v\in \rml^2([0,1],\msv)=\{v\in \rmC^0([0,1],\msv): \int_0^1 ||
+v_\tau||^2_\msv \dd \tau<\infty \} $, by \citep[Theorem 5]{glaunes2005transport}
+$\phi^v$ exists and belongs to $\mcd(\Rset^{d'})$, where we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
+ Therefore, for any choice of $v$, $\phi^v$ defines a valid deformation. 
+This offers a general recipe to construct diffeomorphism given a functional space $\msv$.
+
+With this in mind, the velocity field $v$ fitting the data can be
+estimated by minimizing 
+$v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
+ However, two computational challenges arise.
+  First, this optimization problem is ill-posed, and a penalty term is needed to obtain a unique solution.
+   In addition, we have to find a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, which allows us to solve this minimization problem efficiently.
+
+%\paragraph*{Regularizing the problem to derive uniqueness.}
+It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the space of deformations.
+Following this geometric point of view, geodesics can be constructed on the space of deformations by using the norm of $\msv$. 
+More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
+ the following squared norm can be defined
+ \begin{equation}
+  \label{eq:geodesics_original}
+    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
+ \end{equation}
+     as the minimal "energy" needed to perform the deformation $g$.
+By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
+However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
+To circumvent this issue, another characterization of geodesics can be considered.
+ As in Riemannian geometry, 
+ instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
+ In other words, an Exponential map on deformations is wanted:
+given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
+ \begin{equation}
+  \label{eq:geodesics_shooting}
+  v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
+ \end{equation}
+since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
+By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
+% Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.
+%  Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
+%  Moreover, we have $\mathscr{R}(\phi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in L^2([0,1],\msv)$.
+%  %Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
+%  This is presented with more details in the next paragraph.
+
+\paragraph{Discrete parametrization of diffeomorpshim.}
+%\alain{adapt this without mentioning applicaiton to time seriess}
+%\sam{Est ce qu'on devrait pas introduire la dicretization du graphe ici et dire que l'on peut utiliser le théoreme de représentation}
+In this part, $\msv$ is chosen as an RKHS \cite{berlinet2011reproducing} generated by a smooth kernel $K$ (e.g., Gaussian).
+% $K(x,y)\triangleq \exp(-|x-y|^2/\sigma^2)\operatorname{Id}_{d'}$ for any $(x,y)\in(\Rset^{d'})^2$, where $\sigma^2>0$ is the kernel width and $\operatorname{Id}_{d'}$ is the identity matrix.
+We follow \cite{durrleman2013sparse} and define a 
+ discrete parameterization of the velocity fields to perform geodesics shooting \eqref{eq:geodesics_shooting}.  
+  The initial velocity field $v_0$ is chosen as a finite linear combination of the RKHS basis vector fields, 
+$\mathbf{n}_0$ control points $\msx_0=(x_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0}$ and momentum vectors $\alpha_0=(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0} $ are defined such that for any $x\in \Rset^{d'}$,
+  \begin{equation}
+    \label{eq:def_v0}
+    v_0\left(\alpha_0,\msx_0\right)(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_{k,0})\alpha_{k,0} \eqsp .
+  \end{equation}
+   In our applications, the control points $(x_{k,0})_{k\in[\mathbf{n}_0]}$ can be understood as the discretized graph $(t_k,\mathbf{s}_0(t_k))_{k\in[\mathbf{n}_0]}$ of a starting time series $\mathbf{s}_0$.
+  With this parametrization of $v_0$, \citet{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same
+  structure along time, such that for any $x\in\Rset^{d'}$ and $\tau\in[0,1]$, 
+  \begin{equation}
+    \label{eq:specific_form}
+    v_\tau(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_k(\tau))\alpha_{k}(\tau) \eqsp ,
+  \end{equation}
+  \begin{equation} 
+    \label{eq:integration}
+      \left\{
+        \begin{aligned}
+        & \frac{\dd x_k(\tau)}{\dd \tau}=v_\tau(x_k(\tau)) \eqsp, \quad
+        \frac{\dd \alpha_k(\tau)}{\dd \tau}=-\sum_{k=1}^{\mathbf{n}_0} \dd_{x_k(\tau)} K(x_k(\tau),x_l(\tau))\alpha_{l}(\tau)^\top \alpha_{k}(\tau) \eqsp  \\
+        & \alpha_k(0)=\alpha_{k,0},\quad x_k(0)=x_{k,0} \eqsp , k\in[\mathbf{n}_0] 
+        \end{aligned}
+        \right .
+  \end{equation}
+  These equations are derived from the hamiltonian $H:(\alpha_k,x_k)_{k\in [\mathbf{n}_0]}\mapsto \sum_{k,l=1}^{\mathbf{n}_0} \alpha_{k}^\top K(x_k,x_l)\alpha_{l}  $, such that
+  the velocity norm is preserved $||v_\tau||_{\msv}=||v_0||_\msv $ for any $\tau\in [0,1]$.
+   By \eqref{eq:integration}, the velocity field related to a geodesic $v^*$ is fully parametrized by its initial control points and momentum $(x_{k,0},\alpha_{k,0})_{k\in[\mathbf{n}_0]}$.
+   Thus, given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$, a set of starting points $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$ in $\Rset^{d'}$, a RKHS's kernel $K:\Rset^{d'}\times \Rset^{d'}\to \Rset^{d'\times d'}$, a distance on sets $\mathscr{L}$,
+ a numerical integration scheme of ODE and a penalty factor $\lambda>0$, the basic geodesic shooting step minimizes the following function using a gradient descent method:
+   \begin{equation}
+    \label{eq:relaxation}
+    \mathcal{F}_{\mathbf{x},\mathbf{y}}: (\alpha_k)_{k\in [T_1]}\mapsto \mathscr{L}\left(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2 \eqsp,  
+   \end{equation}
+   where $v_0$ is defined by \eqref{eq:def_v0} and $\varphi^{\{v_0\}}.\mathbf{x}$ is the result of the numerical integration of \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_{k\in[T_1]} $.
+
+   %Before to specify the choice of the RKHS kernel $K$ and the loss on sets $\mathscr{L}$ in \Cref{section:time_series_specificity}, we first discuss the connection of LDDMM with normalizing flows.
+% to highlight the interest of the LDDMM approach.
+
+%Conversely to NF or CNF, by \eqref{eq:integration}, different momentums parameters raise different final deformations $\phi^{\{v_0\}}$, this injectivity properties offers interpretability of the representation and gives sens to PCA or LDA post-processing as exposed in \Cref{section:experiments}.
+
+   %If a diffeomorpshim $\Phi$ is given and we want to represent it as a deformation $\phi^v$ with $v$ of minimal ernergy, then
+  %$(\alpha_{k,0})_k$ is the only parameters to learn since $(x_{k,0})_k$ is fixed.
+
+%   \begin{algorithm}[tb]
+%     \caption{Geodesic shooting with LDDMM}
+%     \label{alg:example}
+%  \begin{algorithmic}
+%     \STATE {\bfseries Input:} a RKHS's kernel $K$, a loss on sets $L$, $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$, $\mathbf{y}=(y_i)_{i\in[T_2]}$ two sets of $\Rset^{d'}$, initial momentums 
+%     $(\alpha_{k,0})_{k\in [T_1]} \in (\Rset^{d'})^{T_1}$ and a number of gradient iterations $n_{\text{iter}}$.
+
+%     \STATE $f: (\alpha_k)_k\mapsto L\left(\phi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2  $ with $v_0$ and $\phi^{\{v_0\}}$
+%      respectively defined by \eqref{eq:def_v0} and \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_k $. 
+%      \sam{Peut être à définir dans un autre algo ?}
+%     \STATE $\alpha_{k,0}^*\gets \alpha_{k,0}, \eqsp k\in[T_1]$
+%     \FOR{$i=1$ {\bfseries to} $n_{\text{iter}}$}
+%     \STATE $\alpha_{k,0}^*\gets \text{AdamStep}(f,\alpha_{k,0}^*$) , where $\text{AdamStep}$ is given in appendix (mettre ref)
+%     \ENDFOR
+%     \STATE {\bfseries Return:} $(\alpha_{k,0}^*)_{k\in [T_1]}$
+
+%  \end{algorithmic}
+%  \end{algorithm}
+
 \section{Methodology}
@@
   Finally, we tackle the specificity of graph time series by deriving a representation theorem on diffeomorphisms, which enables us to select the kernel needed in LDDMM, and thus, we propose TS-LDDMM.
+  \sam{smooth this part}
+  In our practical context, $\mathbf{y}$ represents one of the \textit{discretized graphs} of the time series $(s^j)_{j\in[N]}$, defined as, for any $j\in[N]$,
+  \begin{equation}
+   \label{eq:descretized_graph}
+   \msg(\tsig^j)=(t_i^j,\tsig^j_i)_{ i\in[n_j]}\in (\Rset^{d+1})^{n_j}
+  \end{equation}
+   and $\mathbf{x}$ corresponds to a discretization of the reference graph $\msg(\mathbf{s}_0)$, i.e., 
+   \begin{equation}
+     \label{eq:descretized_referenced_graph}
+     \mathbf{x}=\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} \eqsp . 
+   \end{equation}
+     %\sam{Parler de l'unicté?}
+     \paragraph{Coming back to the representation problem \eqref{eq:transformation}.}
+     Now, we detail how to learn $(\phi^j)_{j\in[N]}$ and a discretization of the reference graph $\msg(\mathbf{s}_0)$.
+      Setting $\mathbf{y}_j=\msg(\tsig^j)$ for any $j\in[N]$ and $\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} $ as given in \eqref{eq:descretized_graph}-\eqref{eq:descretized_referenced_graph} with $\mathbf{n}_0=\operatorname{median}((n_j)_{j\in[N]})$, the representation problem given in \eqref{eq:transformation} boils down solving:
+      \begin{equation}
+        \label{eq:general_optimization_problem}
+        \argmin_{\tilde{\msg}_0,(\alpha_k^j)_{k\in [\mathbf{n}_0]}^{j\in[N]}} \sum_{j=1}^N \mathcal{F}_{\tilde{\msg}_0,\mathbf{y}_j}\left((\alpha_k^j)_{k\in [\mathbf{n}_0]}\right)\eqsp ,
+      \end{equation}
+      which is carried out by a gradient descent on the control points $\tilde{\msg}_0$ and the momentums $\mathbf{\alpha}_j=(\alpha_k^j)_{k\in [\mathbf{n}_0]}$ for any $j\in[N]$, initialized by a dataset's time series graph of size $\mathbf{n}_0$ and by $0_{(d+1)\mathbf{n}_0}$ respectively.
+      The optimization hyperparameter details are given in \Cref{appendix:optimizers_details}.
+     The result of the minimization $\tilde{\msg}_0$ is then considered as the $\mathbf{n}_0$-samples of a common time series $\mathbf{s}_0$ and the momentums $\mathbf{\alpha}_j$ encoding $\phi_j$ yields a feature vector in $\Rset^{d \mathbf{n}_0} $ of $s^j$ for any $j\in[N]$.
+     %, which will be the representation of the time series when choosing the time series' graph sets as inputs, is dependent on the choice of kernel $K$ and the set distance $L$.
+     Finally, the vectors $(\mathbf{\alpha}_j)_{j\in[N]}$ can be analyzed with any statistical or machine learning tools such as Principal Components Analysis (PCA), Latent Discriminant Analysis (LDA), longitudinal data analysis and so on.
 %As $\bfs_0$ is supposed to be fixed, then the representation of the time series $(s_j)_{j\in[N]}$ boils down to the one of the transformation $(\phi_j)_{j\in[N]}$.
@@
 \label{section:LDDMM}
-% \sam{Introduire flow normalisant et continuous normalisant, faire des exps ?}
-To learn the diffeomorphisms $(\phi_j)_{j\in[N]}$, we leverage LDDMM, initially introduced in \cite{beg2005computing}.
+\sam{In this part, there is nearly no novelty, we simply expose how} to
+ learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
   In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
@@
 Such \textit{deformations} $\phi$ are constructed via differential flow equations, for any $x_0\in \Rset^{d'} $ and $\tau\in[0,1]$:
-\begin{align}
+\begin{equation}
   \label{eq:LDDMM_dynamic}
-    &\frac{\dd X(\tau)}{\dd \tau}= v_\tau(X(\tau)), \quad X(0)=x_0\eqsp ,\\
-    &\phi^v_\tau(x_0)=X(\tau), \quad \phi^v=\phi^v_1  \eqsp ,
-\end{align}
+    \frac{\dd X(\tau)}{\dd \tau}= v_\tau(X(\tau)), \quad X(0)=x_0\eqsp ,
+    \phi^v_\tau(x_0)=X(\tau), \quad \phi^v=\phi^v_1  \eqsp ,
+\end{equation}
 where the velocity field is $v:\tau\in [0,1]\mapsto v_\tau\in \msv $
@@
         \begin{aligned}
-        & \frac{\dd x_k(\tau)}{\dd \tau}=v_\tau(x_k(\tau)) \eqsp \\
-        &\frac{\dd \alpha_k(\tau)}{\dd \tau}=-\sum_{k=1}^{\mathbf{n}_0} \dd_{x_k(\tau)} K(x_k(\tau),x_l(\tau))\alpha_{l}(\tau)^\top \alpha_{k}(\tau) \eqsp  \\
+        & \frac{\dd x_k(\tau)}{\dd \tau}=v_\tau(x_k(\tau)) \eqsp, \quad
+        \frac{\dd \alpha_k(\tau)}{\dd \tau}=-\sum_{k=1}^{\mathbf{n}_0} \dd_{x_k(\tau)} K(x_k(\tau),x_l(\tau))\alpha_{l}(\tau)^\top \alpha_{k}(\tau) \eqsp  \\
         & \alpha_k(0)=\alpha_{k,0},\quad x_k(0)=x_{k,0} \eqsp , k\in[\mathbf{n}_0] 
@@
         \label{section:time_series_specificity}
-        In this section, we tailor the LDDMM framework to handle time series data.
+        \sam{In this section, we present our theoretical contribution: we tailor the LDDMM framework to handle time series data}.
           The reason is that applying a general diffeomorphism $\phi$ from $\Rset^{d+1}$ to a time series' graph $\msg(s)$ can result in a set $\phi.\msg(s)$ that does not correspond to the graph of any time series, as illustrated in the \Cref{fig:diffeo}.
@@
         0 & c_1 K_{\text{space}} 
-        \end{pmatrix} \eqsp , \\
-        &K_{\text{space}}=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K_{\text{time}}=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
+        \end{pmatrix} \eqsp ,\\
+       & K_{\text{space}}=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K_{\text{time}}=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
     \end{align}
@@
 \begin{align}
-  d_{\msw^*}^2(\msg_0,\msg_1) &= \sum_{i,j = 1}^{T_0-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^0_j,\overrightarrow{v_j}^0))l^0_j \\
-  &- 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^1_j,\overrightarrow{v_j}^1))l^1_j \\
+  &d_{\msw^*}^2(\msg_0,\msg_1) = \sum_{i,j = 1}^{T_0-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^0_j,\overrightarrow{v_j}^0))l^0_j
+  - 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^1_j,\overrightarrow{v_j}^1))l^1_j \\
   &+ \sum_{i,j = 1}^{T_1-1}l^1_i k((c^1_i,\overrightarrow{v_i}^1),(c^1_j,\overrightarrow{v_j}^1))l^1_j 
@@
 large dataset in high dimensions.
+\sam{Parler de méthode adaptatif ici}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{pictures/samples.pdf}
+    \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
     \vspace{-2.5em}
@@
 
-\section{Societal impact}
-This paper presents work whose goal is to advance the field of Machine Learning. There are many potential societal consequences of our work, none which we feel must be specifically highlighted here.
+%\section{Societal impact}
+%This paper presents work whose goal is to advance the field of Machine Learning. There are many potential societal consequences of our work, none which we feel must be specifically highlighted here.
 % \section{Acknowledgements}
@@
 % \end{algorithm}
-
-\bibliography{ref}
-\bibliographystyle{icml2023}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% APPENDIX
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendix
-\onecolumn
-
-\section{Proofs}
-\label{appendix:proofs}
-Denote by $\msg(s)\triangleq \{ (t,s(t)): \eqsp t\in \msi \} $ the graph of a time series $s: \msi \to \Rset^d$ and $ \phi.\msg(s)\triangleq\{ \phi(t,s(t)): \eqsp t\in \msi\} $ the action of  $\phi\in \mcd(\Rset^{d+1}) $ on $\msg(s)$.
-\begin{theorem}
-    \label{theorem:representation_proof}
-Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi,\msj$ two intervals of $\Rset$.
- There exist $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\gamma\in  \mcd(\Rset) $ such that $\gamma(\msi)=\msj $ and $\Phi_f\in \mcd(\Rset^{d+1})$,
- \begin{equation}% TO add more detail on functionnal space
-    \msg(s)= \Pi_{\gamma,f}.\msg(\mathbf{s}_0),\eqsp \Pi_{\gamma,f}=\Psi_\gamma\circ\Phi_f.
- \end{equation}
- Moreover, for any $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\bar{\gamma}\in  \mcd(\Rset) $, there exists a continously differentiable time series $\bar{s}$ such that 
- $\msg(\bar{s})= \Pi_{\bar{\gamma},\bar{f}}.\msg(\mathbf{s}_0)$
-\end{theorem}
-%\sam{preuve todo, classement des variétés à bord, donne un homeomorphisme sur la restriction}
-\begin{proof}
-  Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi=(a,b),\msj=(\alpha,\beta)$ two intervals of $\Rset$.
-  By setting $\gamma: t\in \Rset \mapsto (\beta-\alpha)(t-a)/(b-a)+\alpha\in \Rset $, we have $ \gamma(\msi)=\msj$ and $\gamma \in \mcd(\Rset) $.
-   By defining $f:(t,x)\in\Rset^{d+1}\mapsto x-\mathbf{s}_0(t)+s\circ \gamma(t) $, the map $\Phi_f\in \mcd(\Rset^{d+1})$,
-    indeed, its inverse is $\Phi_f^{-1}:(t,x)\in\Rset^{d+1}\mapsto (t,x+\mathbf{s}_0(t)-s(t)) $ and is continuously differentiable.
-     Moreover, we have $\Pi_{\gamma,f}.\msg(\mathbf{s}_0)=\{(\gamma(t),s\circ \gamma(t)):\eqsp t\in\msi \}=\msg(s) $.
-
-    Let $\bar{f}\in \rmC^0(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^0(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
-    We have :
-    \begin{align}
-      \Pi_{\gamma,f}.\msg(\mathbf{s}_0)&=\{(\gamma(t),f(t,\mathbf{s}_0(t))),\eqsp t\in \msi \} \\
-      &\label{eq:proof1_last_eq}=\{(t,f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right),\eqsp t\in \gamma(\msi) \} \eqsp .
-    \end{align}
-    By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^0(\gamma(\msi), \Rset^d) $ by composition of continuous functions
-    and $ \msg(\bar{s})= \Pi_{\gamma,f}.\msg(\mathbf{s}_0)$ by \eqref{eq:proof1_last_eq}, which concludes the proof.
-\end{proof}
-\begin{lemma}
-  If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0},
-   there exist $\gamma \in \msd(\Rset) $ and $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ such that $\phi^v=\Psi_\gamma\circ\Phi_f $.
-\end{lemma}
-\begin{proof}
-  Let $v$ be a vector field generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0}.
- We remark that the first coordinate of the velocity field $v_\tau$ denoted by $v_\tau^{\text{time}}$ only depends on the time variable $t$ for any $\tau\in[0,1]$.
- Thus, when computing the first coordinate of the deformation $\phi^v$, denoted by $\gamma$, we integrate \eqref{eq:LDDMM_dynamic} with $v_\tau$ replaced by $v_\tau^{\text{time}}$,
-  thus $\gamma$ is independant of the variable $x$.
-Moreover, $\gamma\in \mcd(\Rset)$ since a Gaussian kernel induced an Hilbert space $\msv$ satisfying $|f|_V\leq |f|_\infty+ |\dd f|_\infty  $ for any $f\in \msv$ by \citep[Theorem 9]{glaunes2005transport}.
-  For the same reason, we have $\phi^v\in \mcd(\Rset^{d+1})$, and thus its last coordinates denoted by $f$ belongs to $\rmC^1(\Rset^{d+1},\Rset^d)$, and by construction $\phi^v=\Psi_\gamma\circ\Phi_f $.
-\end{proof}
-
-\section{Oriented varifold}
-
-\label{appendix:varifold}
-In this section, we introduce the \textit{oriented varifold} associated with curves.
-For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}. 
-We associate to $\gamma\in \rmC^1((a,b),\Rset^{d+1})$ an \textit{oriented varifold} $\mu_\gamma$, i.e.
-a distribution on the space $\Rset^{d+1}\times \mathbb{S}^{d}$ defined as follows, for any smooth test function $\omega :\Rset^{d+1}\times \mathbb{S}^{d}\to \Rset  $,
-\begin{equation}
-  \mathbb{E}_{Y\sim \mu_\gamma}\left[\omega(Y)\right]=\mu_\gamma(\omega)=\int_a^b \omega\left(\gamma(t),\frac{\dot{\gamma}(t)}{|\dot{\gamma}(t)|}\right)|\dot{\gamma}(t)|\dd t \eqsp .
-\end{equation}
-Denoting by $\msw$ the space of smooth test function, we have that $\mu_\gamma $ belongs to its dual $\msw^*$.
- Thus, a distance on $\msw^*$ is sufficient to set a distance on oriented varifolds associated to curve and thus on $\rmC^1((a,b),\Rset^{d+1})$ by the identification $\gamma\to \mu_\gamma $.
-Remark that in (TS-LDDMM), $\gamma$ should be the parametrization of a time series' graph $\msg(s)$, i.e.
-$\gamma: t\in \msi \to (t,s(t))\in\Rset^{d+1} $ denoting by $s:\msi \to \Rset^d$ the time series.
-However, in practice, we work with discrete objects.
- That is why, we set $W$ as an RKHS to use its representation theorem.
- More specifically \citep[Proposition 2 \& 4]{kaltenmark2017general} encourages us to consider a kernel $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$ such that there exist two positive and continuously differentiable kernels $k_{\pos}$ and $k_{\dir}$,
- such that for any $(x,\overrightarrow{u}),(y,\overrightarrow{v}) \in (\Rset^{d+1} \times \mathbb{S}^d)^2$
- \begin{equation}
-     k((x,\overrightarrow{u}),(y,\overrightarrow{v})) = k_{\pos}(x,y)k_{\dir}(\overrightarrow{u},\overrightarrow{v}) \eqsp,
-   \end{equation}
- with moreover $k_\dir>0$ and $ k_\pos$ which admits an RKHS $\msw_\pos$ dense in the space of continous function on $\Rset^{d+1}$ vanishing at infinite \cite{carmeli2010vector}.
-
-Given such a kernel $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$
-verifying \citep[Proposition 2 \& 4]{kaltenmark2017general}, we have that for any $(x,v)\in \Rset^{d+1}\times \mathbb{S}^{d}$, $\delta_{(x,\overrightarrow{v})}$ belongs to $\msw^*$ as a distribution and that the dual metric $\langle\cdot,\cdot \rangle_{\msw^*} $ satisfies for any $(x_1,v_1),(x_2,v_2)\in \left(\Rset^{d+1}\times \mathbb{S}^{d}\right)^2$,
-\begin{equation}
-  \langle\delta_{(x_1,\overrightarrow{v}_1)},\delta_{(x_2,\overrightarrow{v}_2)} \rangle_{\msw^*} =k((x_1,\overrightarrow{v}_1),(x_2,\overrightarrow{v}_2)) \eqsp .
-\end{equation}
-Thus, given two sets of triplets $X=(l_i,x_i,\overrightarrow{v}_i)_{i\in[T_0-1]}\in (\Rset\times \Rset^{d+1}\times \mathbb{S}^d)^{T_0-1},Y=(l_i',y_i,\overrightarrow{w}_i)_{i\in[T_1]}\in (\Rset\times \Rset^{d+1}\times \mathbb{S}^d)^{T_1-1}$ and denoting by
-\begin{equation}
-  \label{eq:def_mu_X}
-  \mu_X=\sum_{i=1}^{T_0} l_i \delta_{(x_i,\overrightarrow{v}_i)},\mu_Y=\sum_{i=1}^{T_1} l_i' \delta_{(y_i,\overrightarrow{w}_i)} \eqsp,
-\end{equation}
- we have,
-\begin{equation}
-   |\mu_X-\mu_Y |_{\msw^*}^2 = \sum_{i,j = 1}^{T_0-1}l_i k((x_i,\overrightarrow{v_i}),(x_i,\overrightarrow{v_i}^0))l_j - 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l_i k((x_i,\overrightarrow{v_i}),(y_i,\overrightarrow{w_i}))l_j' + \sum_{i,j = 1}^{T_1-1}l_i' k((y_i,\overrightarrow{w_i}),(y_i,\overrightarrow{w_i}))l_j' \eqsp.
-\end{equation}
-Then, using the identification $X\to \mu_X, Y\to \mu_Y$, we can define a distance on sets of triplets as $d_{\msw^*,3}(X,Y)=|\mu_X-\mu_Y|_{\msw^*}^2$.
-
-Now, we aim to discretize the oriented varifold $\mu_\msg $ related to a time series' graph $\msg(s)$ by using a set of triplets.
-This is carried out by using a discretized version of $\msg(s) $, i.e. $\tilde{\msg}=(g_i=(t_i,s(t_i)))_{i\in[T]}\in(\Rset^{d+1})^T$, in the following way: 
-For any  $i\in[T-1]$, denoting the center and length of the $i^{th}$ segment $[g_i,g_{i+1}]$ by
-$c_i = (g_i + g_{i+1})/2$, $l_i = \| g_{i+1}-g_{i}\|$, and the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$ by
- $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$, we define the set of triplets $X(\tilde{\msg})=(l_i,c_i,\overrightarrow{v_i})_{i\in[T-1]}$ and its related oriented varifold $\mu_{X(\tilde{\msg})}= \sum_{i=1}^{T-1}l_i \delta_{c_i,\overrightarrow{v_i}}$ as in \eqref{eq:def_mu_X}.
- This is a valid discretization of the oriented varifold $\mu_\msg$ according to \citep[Proposition 1]{kaltenmark2017general}: $\mu_{X(\tilde{\msg})}$ converges towards $\mu_\msg$ as the size of the descretization mesh $\sup_{i\in[T-1]} |t_{i+1}-t_i| $ converges to 0.
-
- Finally, we define a distance on discretized time series' graphs $\tilde{\msg}_1,\tilde{\msg}_2$ as $d_{\msw^*}(\tilde{\msg}_1,\tilde{\msg}_2)=d_{\msw^*,3}(X(\tilde{\msg}_1),X(\tilde{\msg}_2)) $.
-
-\subsection{Varifold kernels}
-\label{appendix:kernel_implementation}
-Denote the one-dimensional Gaussian kernel by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
-In the implementation, we use the following kernels, for any $((t_1,x_1),(t_2,x_2))\in (\Rset^{d+1})^2, ((w_1,v_1),(w_2,v_2))\in (\mathbb{S}^{d})^2 $,
-\begin{equation}
-  k_{\pos}(x,y)=K_{\sigma_{\pos,t}}^{(1)}(t_1,t_2)K_{\sigma_{\pos,x}}^{(d)}(x_1,x_2),\quad k_{\pos}(x,y)=K_{\sigma_{\dir,t}}^{(1)}(w_1,w_2)K_{\sigma_{\dir,x}}^{(d)}(v_1,v_2)\eqsp ,
-\end{equation}
-where $\sigma_{\pos,t}, \sigma_{\pos,x}, \sigma_{\dir,t}, \sigma_{\dir,x}>0 $ are hyperparameters.
- In practice, we select $\sigma_{\pos,x}\approx \sigma_{\dir,x} \approx 1$ when the times series are centered and normalized.
-Otherwise we select $\sigma_{\pos,x}\approx \sigma_{\dir,x} \approx \bar{\sigma}_{s}$ with $\bar{\sigma}_{s}$ the average standard deviation of the time series.
-  We choose $\sigma_{\pos,t}\approx \sigma_{\dir,t} = m f_e$ with $f_e$ the sampling frequency of the time series and $m\in [5]$ an integer depending on the time change between the starting and the target time series graph.
-  The more significant the time change, the higher $m$ should be.
-The intuition comes from the fact that the width $\sigma_{\pos,t}, \sigma_{\dir,t}$ rules the time windows used to perform the comparison, and $\sigma_{\pos,x}, \sigma_{\dir,x}$ affects the space window.
-   The size of the windows should be selected depending on the variations in the data.
-
-\section{Tuning the hyperparameters of the TS-LDDMM kernel given in \eqref{eq:kernel_TAS}}
-\label{appendix:kernel_TS_LDDMM}
-The parameter $\sigma_{T,0}$ should be chosen \textit{large} compared the sampling frequency $f_e$ and compared to average standard deviation $\bar{\sigma}_s$ of the time series, e.g $\sigma_{T,0}=100$ as $\bar{\sigma}_s\approx f_e\approx  1 $.
-It makes the time transformation smoother.
-If $\sigma_{T,0}$ is too small, for instance, $\sigma_{T,0}=f_e $, the effect of the time deformation is too localized, and there are not enough samples to make it visible.
-
-The parameter $\sigma_{T,1}$ should be of the same order as $f_e$: two different points in time can have various space transformations.
- $\sigma_x$ should be of the same order of $\bar{\sigma}_s$: two points with a big difference regarding space compared to $\bar{\sigma}_s$ can have very different space transformations.
-
- We take $c_0\approx 10 c_1 $, we want to encourage time transformation before space transformation. We take $(c_0,c_1)=(1,0.1)$ in all experiments.
-
-\section{Numerical details}
-A report of all the hyperparameters selected is given in \Cref{appendix:table:set_up_exp_practical}.
-\subsection{Optimization details of \eqref{eq:general_optimization_problem}}
-\label{appendix:optimizers_details}
-\paragraph{Initialization}
-At the initialization of \eqref{eq:general_optimization_problem}, all the momentums parameter are set to $0$ and the graph of reference is set to the graph of a time series in the dataset having a mediane samples size.
-
-\paragraph{Gradient descent.}
-The chosen gradient descent method is "adabelief" \cite{zhuang2020adabelief} implemented in the library OPTAX\footnote{https://optax.readthedocs.io/en/latest/}.
-There are two main parameters in the gradient descent: the number of steps nb\_steps, and the maximum value of step size $\eta_M$.
-The stepsize has a particular scheduling: 
-\begin{itemize}
-  \item Warmup period on $0.1 \times$ nb\_steps steps: the stepsize increases linearly from $0$ to $\eta_M$. The goal is to learn progressively the parameters.
-   If the stepsize is too large at the start, smaller steps at the end can't make up for the mistakes made at the beginning. 
-  \item Fine tuning periode on $ 0.9  \times$ nb\_steps : the stepsize decreases from $\eta_M$ to $0$ with a cosine decay implemented in the OPTAX scheduler,
-   i.e. the decreasing factor as the form $0.5  (1 + \cos(\pi  t/T))$. 
-\end{itemize}
-The sharper the deformations, the larger the number of steps and the maximum value of step size should be selected.
-We suggest nb\_steps=300, $\eta_M=0.1 $ for small deformations and nb\_steps=800, $\eta_M=0.3 $ for big ones (time dilation with a factor $\lambda\geq 2$).
-
-\subsection{Synthetic experiments}
-\label{appendix:numerics_synthetic}
-For any deformations generation in both experiments (well-specified and misspecified), we take $\sigma_{T,0},\sigma_{T,1},\sigma_x=(100,1,1)$ and $c_0,c_1=(1,0.1)$ for the kernel $K_\msg$
-and $\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}=(2,1,2,0.6) $ for the varifold kernels $k_\pos,k_\dir$ related to the loss $\scrl$.
-
-In both experiments, we have nb\_steps=300 abd $\eta_M=0.1$.
-
-\subsection{Mouse experiments}
-The number of steps is larger in the second experiment (before/after injection) because the deformations are sharper.
-
-\subsection{Classification experiments}
-\label{appendix:classification_implementation}
-
-We defined a default parametrization for all classifiers.
-
-For classifiers: \href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.deep_learning.CNNClassifier.html#aeon.classification.deep_learning.CNNClassifier}{CNN}, 
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.networks.ResNetNetwork.html#aeon.networks.ResNetNetwork}{ResNet}, 
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.feature_based.Catch22Classifier.html#aeon.classification.feature_based.Catch22Classifier}{Catch22}, 
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.distance_based.KNeighborsTimeSeriesClassifier.html#aeon.classification.distance_based.KNeighborsTimeSeriesClassifier}{DTW-KNN},
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.convolution_based.RocketClassifier.html#aeon.classification.convolution_based.RocketClassifier}{Rocket}
-we used the aeon\footnote{https://www.aeon-toolkit.org/en/stable/index.html} implementations with their default settings.
-
-For Tloss-SVC we used the implementation provided on github\footnote{https://github.com/mqwfrog/ULTS} 
-with the following parameters for learning representations: batch\_size: 10, channels: 40, depth: 10, nb\_steps: 200, in\_channels: 1, 
-kernel\_size: 3, lr: 0.001, nb\_random\_samples: 10, negative\_penalty: 1, out\_channels: 320, reduced\_size: 160. 
-We used the Support Vector Classifier (\href{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html#sklearn.svm.SVC}{SVC})
-from scikit-learn with thee regularization term C: 1. Others parameters are set to default.
-
-For TS-LDDMMM-SVC, all kernels' parameters et optimizer parameter are presented in \Cref{appendix:table:set_up_exp_practical}.
-As well, we used the Support Vector Classifier from scikit-learn with thee regularization term C: 1. Others parameters are set to default.
-
- \begin{table}[hbt!]
-  \caption{Parameters used in all the experiments.
-   For synthetic data, $K_\msg$ refers to the kernel used in the generation, which is the same for the estimation only in the well-specified case.
-$\bar{l}$ refers to the average time series length and $N_d$ refers to the number of dimensions.}
-  \centering
-  \begin{tabular}{lllll}
-
-            \toprule 
-            objects & Optimizer &$k_\pos,k_\dir$ & $K_\msg $ \\
-            \midrule
-            Parameter & (nb\_steps,$\eta_M$) &$(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$ & $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)$ \\
-            \midrule
-            Synthetic data well-specified &(300,0.1) &$(2,1,2,0.6)$& $(1,0.1,100,1,1)$   \\
-            Synthetic data misspecified  & (300,0.1)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
-            Mouse before injection  & (400,0.3)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
-            Mouse before/after injection  & (800,0.3)  & $ (5,1,5,0.6)$ &$(1,0.1,150,1,1)$  \\
-            classification  & (400,0.1)  & $(\max(2,0.03\bar{l}),N_d,\max(2,0.03\bar{l}),0.6)$  & $(1,0.1,0.33\bar{l},1,N_d)$  \\
-
-            \bottomrule
-          \end{tabular}
-        \label{appendix:table:set_up_exp_practical}
-    \end{table}
-\section{Mouse respiratory dataset}
-
-\begin{figure}
-  \centering
-  \includegraphics[width = \linewidth]{pictures/mice_exp.png}
-  \caption{A: Illustration of a double-chamber plethysmograph. The term \textit{dpt} stands for differential 
-  pressure transducer which measures the pressure in each compartment, the pressure then being converted to flow. 
-  B: Nasal airflow (top) and lung volume (bottom). During inspiration, airflow is positive (grey) and during
-  expiration, airflow is negative (pink).}
-  \label{fig:mice_exp}
-\end{figure}
-
-\label{appendix:mouse_dataset}
-Ventilation is a simple physiological function that ensures a vital supply of oxygen and the elimination of CO2. 
-Acetylcholine (Ach) is a neurotransmitter that plays an important role in muscular activity, notably for breathing. 
-Indeed, muscle contraction information passes from the brain to the muscle through the nervous system. Achs are located 
-in synapses of the nervous system (central and peripheral) and skeletal muscles. They ensure the information transmission 
-from nerve to nerve. However, the transmission cannot end without the hydrolysis of Ach by the enzyme Acetylcholinesterase 
-(AchE), allowing nerves to return to their resting state. Inhibition of (AchE) with, for instance, nerve gas, pesticide, 
-or drug intoxication leads to respiratory arrests.
-
-The dataset comes from the experiment \cite{nervo2019respiratory}, where they studied the consequences of partial 
-deficits in AChE and AChE inhibition on mice respiration. AchE inhibition was induced with an 
-irritant molecule called physostigmine (an AchE inhibitor). Mice nasal airflows were sampled at 
-2000Hz with a Double Chamber plethysmograph \cite{hoymann2012lung}, as depicted in \Cref{fig:mice_exp}-A). The flow is expressed in 
-$ml.s^{-1}$; it has a positive value during inspiration and a negative value expiration \Cref{fig:mice_exp}-B). 
-Among the mice population, we selected 7 control mice (\textbf{wt}) and 7 ColQ mice (\textbf{colq}), which do not have 
-AChE anchoring in muscles and some tissues. 
-As described in \cite{nervo2019respiratory}, mice experiments were as follows:
-\begin{enumerate}
-  \item The mouse is placed in a DCP for 15 or 20 min to serve as an internal control.
-  \item The mouse is removed from the DCP and injected with physostigmine.
-  \item The mouse is placed back into the DCP, and its nasal flow is recorded for 35 or 40 min.
-\end{enumerate}
-
-Respiratory cycles were extracted following procedure \cite{germain2023unsupervised}. We removed 
-respiratory cycles whose duration exceeds 1 second; the average respiratory cycle duration is 
-300 ms. We randomly sampled 10 respiratory cycles per minute and mouse. It leads to a dataset of 
-12,732 (time, genotype)-annotated respiratory cycles.
-
-\section{Classification datasets}
-\label{appendix:classification_dataset}
-
-All datasets were taken from UCR/UEA archives \cite{dau2019ucr,bagnall2018uea}. Among all available 
-datasets\footnote{https://timeseriesclassification.com}, we selected 4 datasets related to time series 
-shape comparison. All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/} 
-which already includes the train test split. Essential dataset information is summarized in \Cref{appendix:table:datasets}.
-
-\begin{table}[hbt!]
-  \centering
-  \caption{Time series datasets summary for shape based classification.}
-  \begin{tabular}{lllllll}
-    \toprule
-    Dataset & Train size & test size & Lenghth & Number of classes & Number of dimensions & Type \\
-    \midrule
-    \href{https://timeseriesclassification.com/description.php?Dataset=ArrowHead}{ArrowHead} & 36 & 175 & 251 & 3 & 1 & IMAGE \\
-    \href{https://timeseriesclassification.com/description.php?Dataset=ECG200}{ECG200} & 100 & 100 & 96 & 2 & 1 & ECG \\
-    \href{https://timeseriesclassification.com/description.php?Dataset=GunPoint}{GunPoint} & 50 & 150 & 150 & 2 & 1 & MOTION \\
-    \href{https://timeseriesclassification.com/description.php?Dataset=NATOPS}{NATOPS} & 180 & 180 & 51 & 6 & 24 & MOTION \\
-    \bottomrule
-  \end{tabular}
-  \label{appendix:table:datasets}
-\end{table}
-
-%  [11:35, 31/01/2024] Thibaut Germain: avant/après :
-%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,150,1,1)
-%  Kl = TSGaussGaussKernel(5,1,5,0.6)
-%  dataloss = VarifoldLoss(Kl)
-%  schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
-%  [11:35, 31/01/2024] Thibaut Germain: avant :
-%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,100,1,1)
-%  Kl = TSGaussGaussKernel(2,1,2,0.6)
-%  dataloss = VarifoldLoss(Kl)
-%  schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
-% Let $s$ be a time series and $\msg(\tilde{s}) = (g_i)_{i \in [n]}$ be its discretized graph.
-%  The graph $\msg(s)$ can be approximated as 
-% the union of piecewise linear segments between time-consecutive samples, 
-% and we associate to the approximated graph the oriented varifold:
-
-% \begin{equation}
-%   \mu_{\msg(\tilde{s})} : w \in \msw \mapsto \sum_{i=1}^{n-1} l_i\delta_{(c_i,\overrightarrow{v_i})}(w)
-% \end{equation}
-% where $\msw$ is a set of real valued test functions defined on $\Rset^{d+1} \times \mathbb{S}^d$, 
-% $\delta_{(c_i,\overrightarrow{v_i})}(w) = w(c_i,\overrightarrow{v_i})$ is a Dirac delta function, 
-% $c_i = (g_i + g_{i+1})/2$ (resp. $v_i = \| g_{i+1}-g_{i}\|$) is the center (resp. length) of the $i^{th}$ segment, 
-% and $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$ is the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$.
-
-% Assuming that the test functions' space $\msw$ is the RKHS associated with a $C^1$ positive definite kernel $k$ 
-% that verifies \citep[Proposition 2 \& 4]{kaltenmark2017general}, the oriented varifold belongs to the dual space $\msw^*$. Additionally, 
-% we can define a distance $d_{\msw^*}$ between signals' graph sample sets, which, thanks to the kernel reproducing property,
-% has an explicit formulation:
-
-% \section{Continuous normalizing flows}
-% %Modeling Continuous Stochastic Processes with Dynamic Normalizing Flows time series generative
-% CONTUNYIYS NORMALIZING FLOW bofbof
-% Here, we present the framework of Continuous Normalizing Flows (CNF) and show how LDDMM can be seen as a particular case.
-% \cite{salman2018deep} was already inspired from the work of shape analysis to present deep diffeomorphic normalizing flows.
-% Discussion with Alain :
-% Given a time-dependant vector field $v:[0,1]\times \Rset^{d'} \to \Rset^{d'} $, the integration of this vector field gives a flow $ \phi:[0,1]\times \Rset^{d'} \to \Rset^{d'}$, defined via the ordinary differential equation (ODE),
-% \begin{equation}
-%   \frac{\dd }{\dd \tau}\phi_\tau(x)=v(\tau,\phi_\tau(x)), \phi_0(x)=x \eqsp.
-%  \end{equation}
-% In \cite{chen2018neural}, they introduce Continous Normalizing flows for generative modeling.
-%  By modeling $v=f_\theta$ is a neural networks and choosing a prior $X_0\sim\rho_0$ on the inital value, the law of $\phi_1(X_0)$ has the pullback density $\phi_1\# \rho_0$ and the parameter $\theta$ is optimized to maximize the log-likelihood of a given dataset $(y_i)_{i\in[N]}\in (\Rset^{d'})^N$ which is assumed to follow the law of $\phi_1(X_0)$.
-% Instead of modeling the law of the end point $\phi_1(X_0) $, we can try to model the law of the whole path $(\phi_\tau(X_0))_{\tau\in[0,1]} $ having a density $p:[0,1]\times \Rset^d\to \Rset_{>0}$ such that $p_1=q, p_0=\rho_0$ where $q$ is the empirical distribution in the dataset, as proposed in \cite{lipman2022flow}.
-% In so doing, the problem is ill-posed since $p_t$ can be whatever as soon as $0<t<1$, that is why we try to constraint the learned velocity field $v=f_\theta$ such that the path law $p$ is tractable.
-% One way of doing it is to look for geodesics \eqref{eq:geodesics_original} as proposed in LDDMM, it corresponds to the case where $q=\sum_{i=1}^N \delta_{y_i}/N$ and $\rho_0=\sum_{i=1}^N \delta_{x_i}/N$ are empirical distribution and $y_i=\phi^v_1(x_i) $ for a given velocity field $v$.
-
-% The goal is to target $\nu\in \mathcal{P}(\rmC^0([0,1]\times \Rset^d, \Rset^d)) $ by generating randomness on a initial value $x \sim \delta_x$
-%  and by integrating a velocity field $v_\theta \in \msw \subset \rmC^0([0,1]\times \Rset^d, \Rset^d)$: 
-%  \begin{equation}
-%   \frac{\dd h}{\dd \tau}=v_\theta(\tau,h(\tau,x)), h(0,x)=x \eqsp,
-%  \end{equation}
-%  denoting by $\mu$ the law of $(h(\tau,x))_{\tau \in [0,1]}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsection{TTS}
-% To circumvant this issue, we propose two methods :
-%       \begin{enumerate}
-%         \item \label{enum:first_method} [Time And Space (TAS)] Using a specific form of the RKHS's kernel in LDDMM, to stick to the classical shape analysis framework.
-%         \item \label{enum:second_method} [Time Then Space (TTS)] Tackling the time axis and space axis separetely in the transformation.
-%          This distangles time and space variability in the patterns, increasing the interpretability of the representation.
-%           Although the related optimization is more complex than method \ref{enum:first_method}, the results is more robust to the choice of hyper-parameter related to the RKHS's kernel.
-%       \end{enumerate}
-%        In the following, we present both methods to highlights their pro and cons.
-
-% As described in \Cref{section:optimization}, the method (TAS) learns $\Pi_{\gamma,f}$ directly, while (TTS) find the temporal reparametrisation $\Phi_f$ then the space transformation $\Psi_\gamma$.
-%     % If we define a diffeomorpshim $\Phi_\gamma$ as $(\gamma,\Id_d) $, we have $\Phi_\gamma.G(s) =G(s).\gamma^{-1}$.
-%     %  To increase the variability, we allow some deformations on the graph of the time series after being reparametrized in time, which allow us to recover every possible time seriess.
-%     % Remark that, $G(\mathbf{s}_0\circ \psi^{-1})=\{(t,\mathbf{s}_0\circ \psi^{-1}(t)), t\in \msj \}= \{(\psi(t),\mathbf{s}_0(t)) ,\eqsp t\in \msi \}  $.
-%     % [The action of the diffeomorpshim group 
-%     % \begin{align}
-%     %     &\mcd_*=\mcd(\Rset,\Rset)\times \{\phi\in \mcd(\Rset^{d+1},\Rset^{d+1}): \\
-%     %     &\phi(t,x)=(t,f(t,x)),
-%     %      (t,x)\in \Rset \times \Rset^{d}, f \in C^0(\Rset^{d+1}, \Rset^d) \} 
-%     % \end{align}
-%     % on the space of time series graph $\{G^{(a,b)}(s), s\in C^0((a,b), \Rset), (a,b)\in \Rset_+^2, a<b  \}$
-%     % is homogene.]
-%      %In the next section we define a metric and a norm on this space in order to find a representation $(\phi,\psi)$ with a minimal norm, and we show in (...) how to recover a proxy.
-%      \paragraph{The representation as the solution of an optimization}
-%      By \Cref{theorem:representation}, for any $\mathbf{s}_0\in C^0(\msi,\Rset^d)$, we can find $ F=(\gamma_j,f_j)_{j\in[N]}$ such that $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$ for any $j\in[N]$.
-%      However, to get a meaningfull representation and to tackle the unicity problem, we should optimize the choice of $\mathbf{s}_0 $ and $F$, formally speaking, denoting by $R(\cdot) $ a regularization norm on $\mcd(\Rset^{d+1})$ and $L$ a loss on sets which will be specify later, we aim to solve
-%      \begin{equation}
-%       \label{eq:minimization}
-%       \underset{\mathbf{s_0},(\gamma_j,f_j)_{j\in[N]}}{ \argmin} \sum_{j=1}^N \lambda R(\Pi_{\gamma_j,f_j})+ L(\msg(s^j),\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)) \eqsp ,
-%     \end{equation} 
-%     where $\lambda>0$ is a penalization factor related to $L$ to relax the condition $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$.
-
-%     In practice, as already proposed in shape analysis, we perform an alternative minimization between $\mathbf{s_0}$ and $(\gamma_j,f_j)_j$.
-%     The optimization on $(\gamma_j,f_j)_j$ is parrallelisable, for each couple $(\mathbf{s}_0,s^j)$, we should learn diffemorphisms of minimal norm satisfying a condition. 
-%     To this end, we adapt the geodesic shooting method using LDDMM presented in \cite{durrleman2013sparse} to our special case.
-%       Before to describe how to optimize the whole in \Cref{section:optimization}, we introduce LDDMM in the next section to specify the norm on diffemorphisms.
-
-% \paragraph{Kernel choices}
-% As depicted on figure (mettre ref), we can not use any kernel $K$ to apply the previous methodology to learn deformations on time series' graphs.
-
-% In the following, we expose two choice of RKHS's kernel which answers this issue : (TAS) and (TTS).
-%  We rely on the Gaussian kernel, denoting by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
-% To apply (TAS) in LDDMM with $d'=d+1$, we use the following anisotropic Gaussian kernel  :
-% \begin{align}
-%   \label{eq:kernel_TAS}
-%   &K_{\text{TAS}}((t,x),(t',x'))=\begin{pmatrix}
-%     c_0K & 0 \\
-%     0 & c_1 D 
-%     \end{pmatrix} \eqsp , \\
-%     &D=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
-% \end{align}
-% parametrized by the widths $\sigma_{T,0},\sigma_{T,1},\sigma_x>0$ and the constants $c_0,c_1>0$.
-%  If we use this kernel, by considering \eqref{eq:specific_form}, we remark that the first coordinate of the velocity field $v_\tau$ only depends on the time variable $t$.
-%   It implies that the final deformation will have the form $\phi^v=(\gamma,f) $ with  $f\in C^1(\Rset^{d+1},\Rset^d) $, and $\gamma\in \mcd(\Rset)$
-%   since the first coordinate follows a LDDMM-ODE \eqref{eq:LDDMM_dynamic}.
-
-% To apply (TTS), we use LDDMM on the time axis with $d'=1$ by taking $K_{\text{TTS}}^{\text{time}}=K_{\sigma_{T,0}}^{(1)}$, then on the space axis with $d'=d$ by choosing
-% $K_{\text{TTS}}^{\text{space}}=K_{\sigma_{T,1}}^{(1)}\times K_{\sigma_x}^{(d)} $ where the widths are the same than in \eqref{eq:kernel_TAS}.
-The alternate optimization will be more detailed in \Cref{section:optimization}.
-
-\end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
-% by Iain Murray in 2018, and modified by Alexandre Bouchard in
-% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
-% Modified again in 2023 by Sivan Sabato and Jonathan Scarlett.
-% Previous contributors include Dan Roy, Lise Getoor and Tobias
-% Scheffer, which was slightly modified from the 2010 version by
-% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
-% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
-% slightly modified from Prasad Tadepalli's 2007 version which is a
-% lightly changed version of the previous year's version by Andrew
-% Moore, which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2024-04-30 NEURIPS24/paper.tex
update

@@
 \label{section:LDDMM}
+\sam{DIRE l'intérêt de LDDMM}
 In this part, there is no novelty, we simply expose how to
@@
  instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
+ \sam{To CHANGE}
  In other words, an Exponential map on deformations is wanted:
 given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
- \begin{equation}
-  \label{eq:geodesics_shooting}
-  v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
- \end{equation}
-since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
+%  \begin{equation}
+%   \label{eq:geodesics_shooting}
+%   v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
+%  \end{equation}
+% since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
  By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
@@
 
-First, we introduce the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework.
- Then, we explain how to learn a discretization of the graph $\msg(\bfs_0)$ and the diffeomorphisms $(\phi_j)_{j\in[N]} $ by using LDDMM and a gradient descent minimization.
-  Finally, we tackle the specificity of graph time series by deriving a representation theorem on diffeomorphisms, which enables us to select the kernel needed in LDDMM, and thus, we propose TS-LDDMM.
+%First, we introduce the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework.
+% Then, we explain how to learn a discretization of the graph $\msg(\bfs_0)$ and the diffeomorphisms $(\phi_j)_{j\in[N]} $ by using LDDMM and a gradient descent minimization.
+  %Finally, we tackle the specificity of graph time series by deriving a representation theorem on diffeomorphisms, which enables us to select the kernel needed in LDDMM, and thus, we propose TS-LDDMM.
   \sam{smooth this part}
@@
 %\sam{Transformer , en : dans les ensembles}
-\subsection{Learning diffeomorphism with LDDMM} %Thèse de Nicolas Charon
-\label{section:LDDMM}
-\sam{In this part, there is nearly no novelty, we simply expose how} to
- learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
-  In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
-
-The basic problem that we consider in this section is the following.
-Given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$\footnote{Note that we denote by $d'\in\nset$ the ambient space}, a set of starting points $\mathbf{x}=(x_{i})_{i\in[T_1]}$ in $\Rset^{d'}$, we aim to find a diffeomorphism $\phi$ such that the finite set of points $\mathbf{y}$ is similar in a certain sense to the set of finite sets of transformed points $\phi \cdot \mathbf{x} =(\phi(x_i))_{i\in[T_1]} $.
- The function $\phi$ is occasionally referred to as a \textit{deformation}.
-
- In our practical context, $\mathbf{y}$ represents one of the \textit{discretized graphs} of the time series $(s^j)_{j\in[N]}$, defined as, for any $j\in[N]$,
- \begin{equation}
-  \label{eq:descretized_graph}
-  \msg(\tsig^j)=(t_i^j,\tsig^j_i)_{ i\in[n_j]}\in (\Rset^{d+1})^{n_j}
- \end{equation}
-  and $\mathbf{x}$ corresponds to a discretization of the reference graph $\msg(\mathbf{s}_0)$, i.e., 
-  \begin{equation}
-    \label{eq:descretized_referenced_graph}
-    \mathbf{x}=\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} \eqsp . 
-  \end{equation} 
-% LDDMM is designed to analyzed an existing dataset, while NF and CNF are made to generalize a dataset for data augmentation.
-\paragraph{Representing diffeomorpshims as deformations.}
-Such \textit{deformations} $\phi$ are constructed via differential flow equations, for any $x_0\in \Rset^{d'} $ and $\tau\in[0,1]$:
-\begin{equation}
-  \label{eq:LDDMM_dynamic}
-    \frac{\dd X(\tau)}{\dd \tau}= v_\tau(X(\tau)), \quad X(0)=x_0\eqsp ,
-    \phi^v_\tau(x_0)=X(\tau), \quad \phi^v=\phi^v_1  \eqsp ,
-\end{equation}
-where the velocity field is $v:\tau\in [0,1]\mapsto v_\tau\in \msv $
-and $\msv$ is a Hilbert space of continuously differentiable function
-on $\Rset^{d'}$.  If
-$||\dd u ||_{\infty}+|| u ||_{\infty}\leq ||u ||_\msv $ for any
-$u\in \msv$ and
-$v\in \rml^2([0,1],\msv)=\{v\in \rmC^0([0,1],\msv): \int_0^1 ||
-v_\tau||^2_\msv \dd \tau<\infty \} $, by \citep[Theorem 5]{glaunes2005transport}
-$\phi^v$ exists and belongs to $\mcd(\Rset^{d'})$, where we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
- Therefore, for any choice of $v$, $\phi^v$ defines a valid deformation. 
-This offers a general recipe to construct diffeomorphism given a functional space $\msv$.
-
-With this in mind, the velocity field $v$ fitting the data can be
-estimated by minimizing 
-$v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
- However, two computational challenges arise.
-  First, this optimization problem is ill-posed, and a penalty term is needed to obtain a unique solution.
-   In addition, we have to find a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, which allows us to solve this minimization problem efficiently.
-
-%\paragraph*{Regularizing the problem to derive uniqueness.}
-It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the space of deformations.
-Following this geometric point of view, geodesics can be constructed on the space of deformations by using the norm of $\msv$. 
-More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
- the following squared norm can be defined
- \begin{equation}
-  \label{eq:geodesics_original}
-    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
- \end{equation}
-     as the minimal "energy" needed to perform the deformation $g$.
-By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
-However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
-To circumvent this issue, another characterization of geodesics can be considered.
- As in Riemannian geometry, 
- instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
- In other words, an Exponential map on deformations is wanted:
-given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
- \begin{equation}
-  \label{eq:geodesics_shooting}
-  v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
- \end{equation}
-since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
-By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
-% Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.
-%  Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
-%  Moreover, we have $\mathscr{R}(\phi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in L^2([0,1],\msv)$.
-%  %Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
-%  This is presented with more details in the next paragraph.
-
-\paragraph{Discrete parametrization of diffeomorpshim.}
-%\alain{adapt this without mentioning applicaiton to time seriess}
-%\sam{Est ce qu'on devrait pas introduire la dicretization du graphe ici et dire que l'on peut utiliser le théoreme de représentation}
-In this part, $\msv$ is chosen as an RKHS \cite{berlinet2011reproducing} generated by a smooth kernel $K$ (e.g., Gaussian).
-% $K(x,y)\triangleq \exp(-|x-y|^2/\sigma^2)\operatorname{Id}_{d'}$ for any $(x,y)\in(\Rset^{d'})^2$, where $\sigma^2>0$ is the kernel width and $\operatorname{Id}_{d'}$ is the identity matrix.
-We follow \cite{durrleman2013sparse} and define a 
- discrete parameterization of the velocity fields to perform geodesics shooting \eqref{eq:geodesics_shooting}.  
-  The initial velocity field $v_0$ is chosen as a finite linear combination of the RKHS basis vector fields, 
-$\mathbf{n}_0$ control points $\msx_0=(x_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0}$ and momentum vectors $\alpha_0=(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0} $ are defined such that for any $x\in \Rset^{d'}$,
-  \begin{equation}
-    \label{eq:def_v0}
-    v_0\left(\alpha_0,\msx_0\right)(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_{k,0})\alpha_{k,0} \eqsp .
-  \end{equation}
-   In our applications, the control points $(x_{k,0})_{k\in[\mathbf{n}_0]}$ can be understood as the discretized graph $(t_k,\mathbf{s}_0(t_k))_{k\in[\mathbf{n}_0]}$ of a starting time series $\mathbf{s}_0$.
-  With this parametrization of $v_0$, \citet{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same
-  structure along time, such that for any $x\in\Rset^{d'}$ and $\tau\in[0,1]$, 
-  \begin{equation}
-    \label{eq:specific_form}
-    v_\tau(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_k(\tau))\alpha_{k}(\tau) \eqsp ,
-  \end{equation}
-  \begin{equation} 
-    \label{eq:integration}
-      \left\{
-        \begin{aligned}
-        & \frac{\dd x_k(\tau)}{\dd \tau}=v_\tau(x_k(\tau)) \eqsp, \quad
-        \frac{\dd \alpha_k(\tau)}{\dd \tau}=-\sum_{k=1}^{\mathbf{n}_0} \dd_{x_k(\tau)} K(x_k(\tau),x_l(\tau))\alpha_{l}(\tau)^\top \alpha_{k}(\tau) \eqsp  \\
-        & \alpha_k(0)=\alpha_{k,0},\quad x_k(0)=x_{k,0} \eqsp , k\in[\mathbf{n}_0] 
-        \end{aligned}
-        \right .
-  \end{equation}
-  These equations are derived from the hamiltonian $H:(\alpha_k,x_k)_{k\in [\mathbf{n}_0]}\mapsto \sum_{k,l=1}^{\mathbf{n}_0} \alpha_{k}^\top K(x_k,x_l)\alpha_{l}  $, such that
-  the velocity norm is preserved $||v_\tau||_{\msv}=||v_0||_\msv $ for any $\tau\in [0,1]$.
-   By \eqref{eq:integration}, the velocity field related to a geodesic $v^*$ is fully parametrized by its initial control points and momentum $(x_{k,0},\alpha_{k,0})_{k\in[\mathbf{n}_0]}$.
-   Thus, given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$, a set of starting points $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$ in $\Rset^{d'}$, a RKHS's kernel $K:\Rset^{d'}\times \Rset^{d'}\to \Rset^{d'\times d'}$, a distance on sets $\mathscr{L}$,
- a numerical integration scheme of ODE and a penalty factor $\lambda>0$, the basic geodesic shooting step minimizes the following function using a gradient descent method:
-   \begin{equation}
-    \label{eq:relaxation}
-    \mathcal{F}_{\mathbf{x},\mathbf{y}}: (\alpha_k)_{k\in [T_1]}\mapsto \mathscr{L}\left(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2 \eqsp,  
-   \end{equation}
-   where $v_0$ is defined by \eqref{eq:def_v0} and $\varphi^{\{v_0\}}.\mathbf{x}$ is the result of the numerical integration of \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_{k\in[T_1]} $.
 
    %\sam{Parler de l'unicté?}
-   \paragraph{Coming back to the representation problem \eqref{eq:transformation}.}
-   Now, we detail how to learn $(\phi^j)_{j\in[N]}$ and a discretization of the reference graph $\msg(\mathbf{s}_0)$.
-    Setting $\mathbf{y}_j=\msg(\tsig^j)$ for any $j\in[N]$ and $\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} $ as given in \eqref{eq:descretized_graph}-\eqref{eq:descretized_referenced_graph} with $\mathbf{n}_0=\operatorname{median}((n_j)_{j\in[N]})$, the representation problem given in \eqref{eq:transformation} boils down solving:
-    \begin{equation}
-      \label{eq:general_optimization_problem}
-      \argmin_{\tilde{\msg}_0,(\alpha_k^j)_{k\in [\mathbf{n}_0]}^{j\in[N]}} \sum_{j=1}^N \mathcal{F}_{\tilde{\msg}_0,\mathbf{y}_j}\left((\alpha_k^j)_{k\in [\mathbf{n}_0]}\right)\eqsp ,
-    \end{equation}
-    which is carried out by a gradient descent on the control points $\tilde{\msg}_0$ and the momentums $\mathbf{\alpha}_j=(\alpha_k^j)_{k\in [\mathbf{n}_0]}$ for any $j\in[N]$, initialized by a dataset's time series graph of size $\mathbf{n}_0$ and by $0_{(d+1)\mathbf{n}_0}$ respectively.
-    The optimization hyperparameter details are given in \Cref{appendix:optimizers_details}.
-   The result of the minimization $\tilde{\msg}_0$ is then considered as the $\mathbf{n}_0$-samples of a common time series $\mathbf{s}_0$ and the momentums $\mathbf{\alpha}_j$ encoding $\phi_j$ yields a feature vector in $\Rset^{d \mathbf{n}_0} $ of $s^j$ for any $j\in[N]$.
-   %, which will be the representation of the time series when choosing the time series' graph sets as inputs, is dependent on the choice of kernel $K$ and the set distance $L$.
-   Finally, the vectors $(\mathbf{\alpha}_j)_{j\in[N]}$ can be analyzed with any statistical or machine learning tools such as Principal Components Analysis (PCA), Latent Discriminant Analysis (LDA), longitudinal data analysis and so on.
 

### 2024-04-30 NEURIPS24/paper.tex
update

@@
 \label{section:LDDMM}
-\sam{DIRE l'intérêt de LDDMM}
+
 In this part, there is no novelty, we simply expose how to
@@
   In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
+
+  In the next section, the time series are going to be represented by diffeomorphism parameters $(\alpha_j)_{j\in[N]}$.
+  That's why LDDMM is chosen since it offers a parametrization for diffemorphisms which is sparse and interpretable, two features particularly relevant in the biomedical context.
 
@@
 %  \end{algorithm}
+\paragraph{Relation to Continuous Normalizing Flows.}
 
+One particular popular choice to address the problem of learning a diffeomorphism or a velocity field is Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} (NF) or their continuous counterpart \cite{chen2018neural,grathwohl2019scalable,salman2018deep} (CNF).
+However, we do not rely on this class of learning algorithms for several reasons.
+Indeed, existing and simple normalizing flows are not suitable for the type of data that we are interested in this paper \cite{feng2023multi,deng2020modeling}.
+  In addition,  they are primarily designed to have tractable Jacobian functions, while we do not require such property in our applications. 
+Finally, the use of a differential flow solution of an ODE
+\eqref{eq:LDDMM_dynamic} trick is also at the basis of CNF, which
+then consists of learning a velocity field to address in fitting the data through a loss aiming to address the problem
+at hand. Nevertheless, the main difference between CNF and LDDMM lies in the
+parametrization of the velocity field. LDDMM uses kernels to
+derive closed form formula and enhance interpretability while
+NF and CNF take advantage of deep neural networks to scale with
+large dataset in high dimensions.
 \section{Methodology}
@@
 We consider in this paper observations which consist in a population of $N$ multivariate time series, for any $j\in[N]$, $s^j \in \rmC^1(\msi_j,\Rset^{d})$. 
- Note that the time series are not defined on the same time interval.
-To ease the presentation of our methodology, we suppose that we have access to the continuous time series $(s^j)_{j\in[N]}$, while in practice, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
- However, this difference does not change the overall rationale of our method and its implementation up to minor changes, which we summarize in \eqref{eq:general_optimization_problem}.
-
+However, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
+ Note that \textbf{the number of samples $n_j$ is not necessary the same accross individuals} and the timestamps can be \textbf{irregularly sampled}.
+%To ease the presentation of our methodology, we suppose that we have access to the continuous time series $(s^j)_{j\in[N]}$, while in practice, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
+ %However, this difference does not change the overall rationale of our method and its implementation up to minor changes, which we summarize in \eqref{eq:general_optimization_problem}.
  We assume the time series population is globally homogeneous regarding their "shapes" even if inter-individual variability exists.
@@
 %  \paragraph{Goals}%\sam{potentiellement changer l'ordre de time and space}
-
+\paragraph*{The deformation problem.}
 In this paper, we aim to study the inter-individual variability in the dataset by finding a relevant representation of each time series.
 Inspired from the framework of shape analysis \cite{vaillant2004statistics}, addressing similar problems in morphology,
- we propose to represent each time series' graph $\msg(s^j)$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$, related to a time series $\mathbf{s}_0:\msi \to\Rset^d$, by a diffeomorphism $\phi_j$ on $\Rset^{d+1}$, for any $j\in[N]$,
+ we suggest to represent each time series' graph $\msg(s^j)$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$, related to a time series $\mathbf{s}_0:\msi \to\Rset^d$, by a diffeomorphism $\phi_j$ on $\Rset^{d+1}$, for any $j\in[N]$,
 \begin{equation}
@@
   %Finally, we tackle the specificity of graph time series by deriving a representation theorem on diffeomorphisms, which enables us to select the kernel needed in LDDMM, and thus, we propose TS-LDDMM.
-  \sam{smooth this part}
-  In our practical context, $\mathbf{y}$ represents one of the \textit{discretized graphs} of the time series $(s^j)_{j\in[N]}$, defined as, for any $j\in[N]$,
+
+     %\sam{Parler de l'unicté?}
+     \paragraph{Optimization related to \eqref{eq:transformation}.}
+     Defining the \textit{discretized graphs} of the time series $(s^j)_{j\in[N]}$ and a discretization of the reference graph $\msg(\mathbf{s}_0)$  as, for any $j\in[N]$,
      \begin{equation}
       \label{eq:descretized_graph}
-   \msg(\tsig^j)=(t_i^j,\tsig^j_i)_{ i\in[n_j]}\in (\Rset^{d+1})^{n_j}
+      \mathbf{y}_j=\msg(\tsig^j)=(t_i^j,\tsig^j_i)_{ i\in[n_j]}\in (\Rset^{d+1})^{n_j},\quad \tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} \eqsp ,
      \end{equation}
-   and $\mathbf{x}$ corresponds to a discretization of the reference graph $\msg(\mathbf{s}_0)$, i.e., 
-   \begin{equation}
-     \label{eq:descretized_referenced_graph}
-     \mathbf{x}=\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} \eqsp . 
-   \end{equation}
-     %\sam{Parler de l'unicté?}
-     \paragraph{Coming back to the representation problem \eqref{eq:transformation}.}
-     Now, we detail how to learn $(\phi^j)_{j\in[N]}$ and a discretization of the reference graph $\msg(\mathbf{s}_0)$.
-      Setting $\mathbf{y}_j=\msg(\tsig^j)$ for any $j\in[N]$ and $\tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} $ as given in \eqref{eq:descretized_graph}-\eqref{eq:descretized_referenced_graph} with $\mathbf{n}_0=\operatorname{median}((n_j)_{j\in[N]})$, the representation problem given in \eqref{eq:transformation} boils down solving:
+       with $\mathbf{n}_0=\operatorname{median}((n_j)_{j\in[N]})$, the representation problem given in \eqref{eq:transformation} boils down solving:
       \begin{equation}
@@
 
+Nevertheless, \eqref{eq:general_optimization_problem} ask to define a kernel and a loss in order to perform geodesic shooting \ref{eq:relaxation}, which is the purpose of the next subsection.
    \subsection{Application of LDDMM to time series analysis: TS-LDDMM}
@@
         \label{section:time_series_specificity}
-        \sam{In this section, we present our theoretical contribution: we tailor the LDDMM framework to handle time series data}.
+        In this section, we present our theoretical contribution: we tailor the LDDMM framework to handle time series data.
           The reason is that applying a general diffeomorphism $\phi$ from $\Rset^{d+1}$ to a time series' graph $\msg(s)$ can result in a set $\phi.\msg(s)$ that does not correspond to the graph of any time series, as illustrated in the \Cref{fig:diffeo}.
+           Thus, Time series graph have more structure than a simple 1D curve \cite{glaunes2008large} and deserve their special analysis which will prove fruitful as demonstrated in \ref{section:experiments}.% thus this case is not an application of \cite{glaunes2004diffeomorphic}
 
@@
     \end{lemma}
+    \sam{Parler des Cauchy kernel en appndice et du choix de la loss}
     % \begin{proof}
@@
 
-\paragraph{Relation to Continuous Normalizing Flows.}
-
-One particular popular choice to address the problem of learning a diffeomorphism or a velocity field is Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} (NF) or their continuous counterpart \cite{chen2018neural,grathwohl2019scalable,salman2018deep} (CNF).
-However, we do not rely on this class of learning algorithms for several reasons.
-Indeed, existing and simple normalizing flows are not suitable for the type of data that we are interested in this paper \cite{feng2023multi,deng2020modeling}.
-  In addition,  they are primarily designed to have tractable Jacobian functions, while we do not require such property in our applications. 
-Finally, the use of a differential flow solution of an ODE
-\eqref{eq:LDDMM_dynamic} trick is also at the basis of CNF, which
-then consists of learning a velocity field to address in fitting the data through a loss aiming to address the problem
-at hand. Nevertheless, the main difference between CNF and LDDMM lies in the
-parametrization of the velocity field. LDDMM uses kernels to
-derive closed form formula and enhance interpretability while
-NF and CNF take advantage of deep neural networks to scale with
-large dataset in high dimensions.
 \sam{Parler de méthode adaptatif ici}

### 2024-05-07 NEURIPS24/neurips_2024_to_submit.tex
updat

@@
 \begin{abstract}
+    [TO CHANGE]
     Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
@@
     Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
-    Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
-     we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
-     $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
-     Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
-     These features can be post-processed with any statistical or machine learning tools. 
+    % Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
+    %  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
+    %  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
+    %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
+    In this paper, we propose an unsupervised representation learning method (URL) for time series \cite{trirat2024universal,meng2023unsupervised} by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
+    In particular, the method handle time series with \textbf{irregular sampling} and \textbf{different sizes}.
+    This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
+     we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) \cite{beg2005computing,vaillant2004statistics} to the case of time series data by applying difformations to their graph.
+    However, time series graph have more structure than 1D curves:
+     a graph time series transformation by a general diffeomorphism is not always a graph time series. Our contribution arise from this observation.
+    This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.
+    % These features can be post-processed with any statistical or machine learning tools. 
      We demonstrate the advantages of our representation compared to existing methods using synthetic data and real-world examples, motivated by applications in medicine.

### 2024-05-07 NEURIPS24/paper.tex
updat

@@
 
-\sam{Aller droit au but, se servir de la section related works pour détailler, notamment enlever le premier paragraphe dire directement que l'on veut faire du unsupervised represnetation learning for time series mais en s'intéressant par ticulièrement à la notion de shape.
-Parler du contexte biomédical, définir la shape. Parler de shape analysis, introduire notre problème de représentation, dire que nous on propose une structure adapté aux time series.}
-A time series is a set of values indexed by time \cite{trirat2024universal}.
-This type of data is prevalent in biology and medicine for recording physiological time series \cite{wang2016research,bar2012studying,germain2023unsupervised}, as well as in the broader context of the Internet of Things sensors \cite{fathy2018large} and measurements in physical systems \cite{luo2021deep}.
- However, due to the nature of such data, which have different sample sizes and incorporate a temporal evolution, classical statistical analysis, and machine learning methods need refinement \cite{anderson2011statistical}.
- One option to tackle this issue is to derive a feature representation of time series, which depends on the problem at hand \cite{trirat2024universal}.
+%\sam{Aller droit au but, se servir de la section related works pour détailler, notamment enlever le premier paragraphe dire directement que l'on veut faire du unsupervised represnetation learning for time series mais en s'intéressant par ticulièrement à la notion de shape.
+%Parler du contexte biomédical, définir la shape. Parler de shape analysis, introduire notre problème de représentation, dire que nous on propose une structure adapté aux time series.}
 
- When the goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology,
-  the features encode the specificity of an individual compared to another. 
+In this paper, we propose an unsupervised representation learning method (URL) for time series \cite{trirat2024universal,meng2023unsupervised} by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
+In particular, the method handle time series with \textbf{irregular sampling} and \textbf{different sizes}.
+This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
+ we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) \cite{beg2005computing,vaillant2004statistics} to the case of time series data by applying difformations to their graph.
+However, time series graph have more structure than 1D curves:
+ a graph time series transformation by a general diffeomorphism is not always a graph time series.
+This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.
+%A time series is a set of values indexed by time \cite{trirat2024universal}.
+%This type of data is prevalent in biology and medicine for recording physiological time series , as well as in the broader context of the Internet of Things sensors \cite{fathy2018large} and measurements in physical systems \cite{luo2021deep}.
+ %However, due to the nature of such data, which have different sample sizes and incorporate a temporal evolution, classical statistical analysis, and machine learning methods need refinement \cite{anderson2011statistical}.
+ %One option to tackle this issue is to derive a feature representation of time series, which depends on the problem at hand \cite{trirat2024universal}.
+
+ Our goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology \cite{wang2016research,bar2012studying,germain2023unsupervised}.
+  More specifically, we aim to find an unsupervised features representation method which encode the specificity of an individual compared to another. 
   In physiology, studying the different \textit{shapes} in a time series related to biological phenomena and their variations according to individual or pathology is common.
@@
 In this paper, we refer to as the shape of a time series, the graph of this signal.
+
+  Although a community structure with representatives can be learned in an unsupervised way using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}
+  or similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},
+     studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem in URL.
+     %In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.
   %For instance, the shape of a heartbeat in an Electrocardiogram (ECG) -the pattern- has been well studied \cite{madona2021pqrst} with the PQRST decomposition (see figure\sam{fill}).
@@
 
-   In most clinical studies, depending on expert knowledge, people inspect hand-crafted features such as the mean, wavelets decomposition, or temporal correlation \cite{anderson2011statistical} to keep interpretability.
-   However, this approach is time-consuming and potentially biased by priors, especially in exploratory research.
-   Given the success of neural networks \cite{bishop1994neural}, features can now be automatically extracted, depending only on the data, by choosing an objective to optimize and a neural architecture tailored to the problem \cite{ismail2019deep,blazquez2021review,torres2021deep}, even in an unsupervised way \cite{meng2023unsupervised}.
-   Moreover, depending on the method, it is still possible to have an interpretability of the features.
-    For instance, among individuals, a contrastive clustering method can be applied to spot community structures \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
-    More generally, interpretability \cite{yuan2019wave2vec,zhang2018visual,ismail2020benchmarking} is possible by visualizing the shapes of representative patterns to which the model is sensitive.
-    Other methods, based on similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k}, also propose to perform clustering on shapes in time series or to find some representatives \cite{ye2009time}.
-    Although a community structure with representatives can be learned in an unsupervised way,
-     studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem.
-     In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.% \thi{In an individual cluster ?}
+  %  In most clinical studies, depending on expert knowledge, people inspect hand-crafted features such as the mean, wavelets decomposition, or temporal correlation \cite{anderson2011statistical} to keep interpretability.
+  %  However, this approach is time-consuming and potentially biased by priors, especially in exploratory research.
+  %  Given the success of neural networks \cite{bishop1994neural}, features can now be automatically extracted, depending only on the data, by choosing an objective to optimize and a neural architecture tailored to the problem \cite{ismail2019deep,blazquez2021review,torres2021deep}, even in an unsupervised way \cite{meng2023unsupervised}.
+  %  Moreover, depending on the method, it is still possible to have an interpretability of the features.
+  %   For instance, among individuals, a contrastive clustering method can be applied to spot community structures \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
+  %   More generally, interpretability \cite{yuan2019wave2vec,zhang2018visual,ismail2020benchmarking} is possible by visualizing the shapes of representative patterns to which the model is sensitive.
+  %   Other methods, based on similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k}, also propose to perform clustering on shapes in time series or to find some representatives \cite{ye2009time}.
+
+    %Although a community structure with representatives can be learned in an unsupervised way,
+    % studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem.
+     %In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.% \thi{In an individual cluster ?}
 
@@
 
-This type of question has already been addressed in shape analysis \cite{vaillant2004statistics}, a community focusing on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
-The main idea is to represent these different objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
- Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
-  %Thus, instead of working on the manifold $\mathcal{M}$, we can work in the tangent space, which is an Euclidean space at $\mathbf{m}_0$.
+%This type of question has already been addressed in shape analysis \cite{vaillant2004statistics}, a community focusing on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
+%However, time series data with their specific structure has not been properly adressed.
 
- \citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
- However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
- For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
-
- To tackle this issue, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
- Then, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
+ First, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
+ Then, building on the shape analysis litterature, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
   The idea is to represent each element $(\msg(s^j))_{j\in[N]}$ of the dataset as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism.
@@
    The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
-    Finally, these features encoding the shapes can feed any statistical or machine learning model.
-   Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
-   However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g. \Cref{fig:diffeo}.
+    Finally, these features encoding the shapes can feed any statistical or machine learning model as in URL.
+
+   However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
+\Cref{fig:diffeo}, thus a graph time series is more than a simple 1D curve \cite{glaunes2008large}.
+   Our contributions arise from this observation.
+   %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
     We solve this issue by specifying the class of diffeomorphisms to consider and showing how to learn them.
@@
        \item We propose an unsupervised method (TS-LDDMM) to analyze inter-individual variability of shapes in a time series dataset.
-       We especially motivate our extension of LDDMM to time series by introducing a theoretical framework.
+       We especially motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
        \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method with respect to its hyperparameters, also providing guidelines for tuning.
@@
 \section{Related Works}
+The main idea is to represent these different objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
+ Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
+% Depending on the tasks, e.g classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep} or clustering , a meaningful vector representation of time series can be derived.
+%  Initially, in this research, handcrafted features were computed from time series, such as the mean, wavelets \cite{percival2000wavelet}, or Fourier coefficients \cite{bloomfield2004fourier}, depending on expert knowledge.
+%   However, this approach is time-consuming and potentially biased by priors.
+%   Observing the success of neural networks, features are automatically extracted by choosing an objective to optimize and a neural architecture tailored to the problem.
+%    Taking a set of time series as input, features can be identified by solving specific supervised tasks—classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep}— or in an unsupervised manner by deriving labels from the inputs \cite{meng2023unsupervised}.
+%    \thi{Rajouter une image, et une phrase par rapport au temps}
+%    Although the deep learning approach is particularly effective for supervised tasks, annotated datasets are rare, making unsupervised methods more practical.
+%    These methods have proven effective in identifying different patterns in the data and providing a vector representation that segregates well into clusters \cite{meng2023unsupervised}.
+%     However, they may lack precision on small and homogeneous datasets with only one typical shape, often arising from clinical experiments addressing a specific phenomenon (refer to figures).
+%      While current unsupervised methods separate representations into clusters, our focus is on the inter-individual variability within one cluster.
+%       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.
+
+ %\citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
+ %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
+ %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
+ In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
+ Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
 \section{Notations}
@@
     \begin{lemma}
+      \label{lemma:choice_of_kernel_V}
       If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0},

### 2024-05-07 NEURIPS24/neurips_2024_to_submit.tex
update

@@
 \begin{abstract}
-    [TO CHANGE]
-    Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
-    Individuals have specificities depending on their genes or disease
-     while sharing, on average, a similar pattern in the features of their corresponding time series.
+    %TO CHANGE]
+    %Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
+    %Individuals have specificities depending on their genes or disease
+     %while sharing, on average, a similar pattern in the features of their corresponding time series.
 For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
-    In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
-    Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
+    %In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
+    %Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
     % Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
@@
     %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
-    In this paper, we propose an unsupervised representation learning method (URL) for time series \cite{trirat2024universal,meng2023unsupervised} by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
-    In particular, the method handle time series with \textbf{irregular sampling} and \textbf{different sizes}.
+    In this paper, we propose an unsupervised representation learning method (URL) for time series %\cite{trirat2024universal,meng2023unsupervised}
+     by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
+    In particular, the method can handle \textit{irregularly sampled} time series having \textit{different sizes}.
     This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
-     we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) \cite{beg2005computing,vaillant2004statistics} to the case of time series data by applying difformations to their graph.
-    However, time series graph have more structure than 1D curves:
-     a graph time series transformation by a general diffeomorphism is not always a graph time series. Our contribution arise from this observation.
+     we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) %\cite{beg2005computing,vaillant2004statistics}
+      to the case of time series data by applying deformations to their graph.
+    However, the method is not a simple application of LDDMM since time series graph have more structure than 1D curves.
+     Indeed, a graph time series transformed by a general diffeomorphism is not always a graph time series.
+      We solve this issue by establishing a representation theorem on time series graph and derive its consequences on LDDMM.
     This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.

### 2024-05-07 NEURIPS24/paper.tex
update

@@
 
-In this paper, we propose an unsupervised representation learning method (URL) for time series \cite{trirat2024universal,meng2023unsupervised} by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
-In particular, the method handle time series with \textbf{irregular sampling} and \textbf{different sizes}.
-This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
- we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) \cite{beg2005computing,vaillant2004statistics} to the case of time series data by applying difformations to their graph.
-However, time series graph have more structure than 1D curves:
- a graph time series transformation by a general diffeomorphism is not always a graph time series.
-This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.
+% In this paper, we propose an unsupervised representation learning method (URL) for time series \cite{trirat2024universal,meng2023unsupervised} by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
+% In particular, the method handle time series with \textbf{irregular sampling} and \textbf{different sizes}.
+% This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
+%  we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) \cite{beg2005computing,vaillant2004statistics} to the case of time series data by applying difformations to their graph.
+% However, time series graph have more structure than 1D curves:
+%  a graph time series transformation by a general diffeomorphism is not always a graph time series.
+% This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.
 %A time series is a set of values indexed by time \cite{trirat2024universal}.
@@
 
-  Although a community structure with representatives can be learned in an unsupervised way using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}
+  Although a community structure with representatives can be learned in an unsupervised way \cite{trirat2024universal,meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}
   or similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},
@@
  First, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
- Then, building on the shape analysis litterature, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
+ Then, building on the shape analysis litterature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
   The idea is to represent each element $(\msg(s^j))_{j\in[N]}$ of the dataset as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism.
@@
      This change is fruitful in representing time transformation as illustrated in \Cref{fig:transport}.
+     In particular, the method can handle \textit{irregularly sampled} time series having \textit{different sizes}.
 

### 2024-05-07 NEURIPS24/neurips_2024_to_submit.tex
update

@@
      by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
-    In particular, the method can handle \textit{irregularly sampled} time series having \textit{different sizes}.
+    In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
     This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
@@
       to the case of time series data by applying deformations to their graph.
-    However, the method is not a simple application of LDDMM since time series graph have more structure than 1D curves.
+    However, the method is not a simple application of LDDMM since time series graph have more structure than curves.
      Indeed, a graph time series transformed by a general diffeomorphism is not always a graph time series.

### 2024-05-07 NEURIPS24/paper.tex
update

@@
    However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
-\Cref{fig:diffeo}, thus a graph time series is more than a simple 1D curve \cite{glaunes2008large}.
+\Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
    Our contributions arise from this observation.
@@
      This change is fruitful in representing time transformation as illustrated in \Cref{fig:transport}.
-     In particular, the method can handle \textit{irregularly sampled} time series having \textit{different sizes}.
+     %In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
 
@@
        \item We propose an unsupervised method (TS-LDDMM) to analyze inter-individual variability of shapes in a time series dataset.
-       We especially motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
+In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
+       \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
        \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method with respect to its hyperparameters, also providing guidelines for tuning.

### 2024-05-13 NEURIPS24/paper.tex
up

@@
 %\paragraph*{Regularizing the problem to derive uniqueness.}
-It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the space of deformations.
-Following this geometric point of view, geodesics can be constructed on the space of deformations by using the norm of $\msv$. 
-More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
- the following squared norm can be defined
+It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$.
+Following this geometric point of view, geodesics can be constructed on $\msh$ by using the following squared norm 
  \begin{equation}
@@
  \end{equation}
-     as the minimal "energy" needed to perform the deformation $g$.
-By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
-However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
-To circumvent this issue, another characterization of geodesics can be considered.
- As in Riemannian geometry, 
- instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
- \sam{To CHANGE}
- In other words, an Exponential map on deformations is wanted:
-given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
+By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy lipshcitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$, as straight lines in Euclidean space.
+Denoting by $w(v_0)$ the geodesic starting from the identity with inital velocity $v_0$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \phi^v$ and the previous matching problem becomes a \textit{geodesic shooting problem}:
+ \begin{equation}
+  \label{eq:geodesics_shooting}
+  \inf_{v_0 \in \msv} \mathscr{L}(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}).
+ \end{equation}
+ Using $\varphi^{\{v_0\}}$ instead of $\phi^v$ for any $v\in \rml^2([0,1],\msv)$ regularizes the problem and induces a sparse representation for the learning diffeomorphisms.
+ Moreover, by setting $\msv$ as an RKHS, the geodesic shooting problem has a unique solution and becomes tractable, as described in the next section.
+
+% More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
+%  the following squared norm can be defined
 %  \begin{equation}
-%   \label{eq:geodesics_shooting}
-%   v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
+%   \label{eq:geodesics_original}
+%     \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
 %  \end{equation}
-% since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
- By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
+%      as the minimal "energy" needed to perform the deformation $g$.
+% By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
+% However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
+% To circumvent this issue, another characterization of geodesics can be considered.
+%  As in Riemannian geometry, 
+%  instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
+%  \sam{To CHANGE}
+%  %In other words, an Exponential map on deformations is wanted:
+% given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
+% %  \begin{equation}
+% %   \label{eq:geodesics_shooting}
+% %   v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
+% %  \end{equation}
+% % since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
+%  By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
 % Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.

### 2024-05-13 NEURIPS24/paper.tex
update

@@
 \section{Related Works}
+Shape analysis focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
  The main idea is to represent these different objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
  Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
+ The goal is then to find a well suited Riemannian struvture according to the studied object.
+
+ A time series graph can be seen a curve and LDDMM structure is relevant to tackle curves as presented in \cite{glaunes2008large}. %, LDDMM structure is relevant to tackle curves and
+  However, time series graph has more structure as depicted in \Cref{fig:diffeo} due to the temporal evolution. 
+  \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM, but distinguish from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
+
+  Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} address the representation of curves having unitary velocity by using the Square-Root Velocity (SRV) representation.
+ However, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
+  Again, the time series graph structure is not respected.
+   Very recently in a functionnal data analysis framework \cite{wu2024shape} (Shape-FPCA) improved by giving also a representation for the original time evolution.
+    Nevertheless, this methods applies only on time series of \textit{same size} and is made for \textit{continuous objects}, making the estimation more sensitive to noise.
+  %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
+  %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
 % Depending on the tasks, e.g classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep} or clustering , a meaningful vector representation of time series can be derived.
@@
 %       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.
-
+[Ajouter littérature de URL demander à Sisi ?]
  %\citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
@@
  %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
- In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
- Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
+ %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
+ %Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
 \section{Notations}
@@
   \label{eq:geodesics_original}
-    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
+    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv\dd \tau
  \end{equation}
-By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy lipshcitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$, as straight lines in Euclidean space.
+By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy lipshcitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$ \cite{miller2006geodesic}, as straight lines in Euclidean space.
 Denoting by $w(v_0)$ the geodesic starting from the identity with inital velocity $v_0$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \phi^v$ and the previous matching problem becomes a \textit{geodesic shooting problem}:
@@
 \label{section:experiments}
+[L'intro est ce necessaire ?]
 

### 2024-05-14 NEURIPS24/neurips_2024_to_submit.tex
update

@@
     %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
-    In this paper, we propose an unsupervised representation learning method (URL) for time series %\cite{trirat2024universal,meng2023unsupervised}
-     by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
-    In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
-    This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
-     we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) %\cite{beg2005computing,vaillant2004statistics}
-      to the case of time series data by applying deformations to their graph.
-    However, the method is not a simple application of LDDMM since time series graph have more structure than curves.
-     Indeed, a graph time series transformed by a general diffeomorphism is not always a graph time series.
-      We solve this issue by establishing a representation theorem on time series graph and derive its consequences on LDDMM.
-    This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.
-    % These features can be post-processed with any statistical or machine learning tools. 
-     We demonstrate the advantages of our representation compared to existing methods using synthetic data and real-world examples, motivated by applications in medicine.
+    This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies.
+The proposed algorithm manages irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
+    At the crossroads between URL for time series and shape analysis, the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series.
+This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension.
+    In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework.
+We showcase the advantages of our representation compared to existing methods using synthetic data and real-world examples motivated by biomedical applications.
 \end{abstract}

### 2024-05-14 classification/msvc_exp.py
add code

@@ -1,0 +1,230 @@
+import sys
+sys.path.insert(0,"../")
+
+import pandas as pd
+import numpy as np 
+import time
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+from sklearn.model_selection import train_test_split
+from sklearn.preprocessing import LabelEncoder
+from sklearn.metrics import accuracy_score,f1_score
+
+from aeon.datasets import load_classification
+
+from src.utils import from_timeseries_to_dataset
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel,GaussKernel,MyTSGaussKernel
+from src.loss import VarifoldLoss,MyLoss,SumVarifoldLoss
+from src.classifier import MomentaSVC
+from src.lddmm import batch_time_initializer
+from src.barycenter import batch_barycenter_initializer
+from src.utils import batch_dataset,from_timeseries_to_dataset
+
+
+def mouse_loader(seed=0):
+    y =pd.read_csv("../mouse/dataset/y.csv",index_col=0)
+    X = np.load("../mouse/dataset/X.npy")
+    X_mask = np.load("../mouse/dataset/X_mask.npy")
+    X = X[:,::2,:]
+    X_mask = X_mask[:,::2,:]
+
+    lst = []
+    for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+        idx = y[(y.filename == filename)*(y.before == "Y")].sample(20,random_state=seed).index
+        lst.append(idx)
+    idxs = np.concatenate(lst)
+
+    X,X_mask = X[idxs],X_mask[idxs]
+    y = y.iloc[idxs]
+    y = y.genotype.values.astype(str)
+    train_idx,test_idx = train_test_split(np.arange(X.shape[0]),test_size=0.33,shuffle=True,random_state=seed)
+    return X[train_idx],X_mask[train_idx],X[test_idx],X_mask[test_idx],y[train_idx],y[test_idx]
+
+def gait_loader(seed=0): 
+    y = np.load("../gait_exp/dataset/y.npy")
+    X = np.load("../gait_exp/dataset/X.npy")
+    X_mask = np.load("../gait_exp/dataset/X_mask.npy")
+    train_idx,test_idx = train_test_split(np.arange(X.shape[0]),test_size=0.33,shuffle=True,random_state=seed)
[…]

### 2024-05-14 gait_exp/gait_experiment.py

@@ -1,0 +1,67 @@
+import sys 
+sys.path.insert(0,"../")
+
+import numpy as np 
+import jax.numpy as jnp
+import json 
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+
+
+
+from src.utils import batch_dataset
+from src.kernel import VFTSGaussKernel, TSGaussGaussKernel,GaussKernel
+from src.barycenter import batch_varifold_barycenter_registration, batch_barycenter_registration
+from src.loss import SumVarifoldLoss
+
+
+BATCH_SIZE = 75
+NITER = 400
+
+
+scheduler = warmup_cosine_decay_schedule(0,0.3,30,400,0.)
+optimizer = optax.adam(scheduler)
+
+
+
+Kv = VFTSGaussKernel(1,0.1,40,1,1)
+Kl1 = TSGaussGaussKernel(2,2,2,2)
+Kl2 = TSGaussGaussKernel(1,1,1,1)
+Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
+Kls=[Kl1,Kl2,Kl3]
+dataloss = SumVarifoldLoss(Kls)
+
+
+
+
+
+SAVE_PATH = "./results/SS_2.json"
+
+
+if __name__ == "__main__":
+
+    # Import data 
+    X = np.load("./dataset/X.npy")
+    X_mask = np.load("./dataset/X_mask.npy")
+    batched_dataset, batched_masks = batch_dataset(X,BATCH_SIZE,X_mask)
+
+    print("Running experiment")
+    ps ,qb, qb_mask= batch_barycenter_registration(batched_dataset,batched_masks,Kv,dataloss,niter=NITER,optimizer=optimizer,gamma_loss=0.00)
+    
+    print("Exporting results")
+    dct =dict(
+        ps = json.dumps(ps.tolist()),
+        qb = json.dumps(qb.tolist()),
+        qb_mask = json.dumps(qb_mask.tolist())
+    )
+    
+    with open(SAVE_PATH, "w") as f: 
+        f.write(json.dumps(dct))
+
+    print("Done")
+
+
+
+
+
+

### 2024-05-14 mouse/experiment_1.py

@@ -1,0 +1,50 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel
+from src.loss import VarifoldLoss,SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("./dataset/y.csv",index_col=0)
+X = np.load("./dataset/X.npy")
+X_mask = np.load("./dataset/X_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.filename == filename)*(y.before == "Y")].sample(50).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+Kv = VFTSGaussKernel(1,0.1,100,1,1)
+Kl1 = TSGaussGaussKernel(2,2,2,2)
+Kl2 = TSGaussGaussKernel(1,1,1,1)
+Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
+Kls=[Kl1,Kl2,Kl3]
+dataloss = SumVarifoldLoss(Kls)
+schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
+optimizer = optax.adabelief(schedule)
+
+
+if __name__ == "__main__": 
+
+  
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
+
+    np.save("./results/exp_1_0/idxs.npy",idxs)
+    np.save("./results/exp_1_0/p0s.npy",p0s)
+    np.save("./results/exp_1_0/q0.npy",q0)
+    np.save("./results/exp_1_0/q0_mask.npy",q0_mask)

### 2024-05-14 mouse/experiment_1_bis.py

@@ -1,0 +1,56 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel
+from src.loss import VarifoldLoss,SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("./dataset/y.csv",index_col=0)
+X = np.load("./dataset/X.npy")
+X_mask = np.load("./dataset/X_mask.npy")
+idxs = np.load("./results/exp_1_0/idxs.npy")
+
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+print(y.before.unique())
+print(y.genotype.unique())
+
+
+if __name__ == "__main__": 
+
+    Kv = VFTSGaussKernel(1,0.1,100,1,1)
+    Kv = VFTSGaussKernel(1,0.1,100,1,1)
+    Kl1 = TSGaussGaussKernel(2,2,2,2)
+    Kl2 = TSGaussGaussKernel(1,1,1,1)
+    Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
+    Kls=[Kl1,Kl2,Kl3]
+    dataloss = SumVarifoldLoss(Kls)
+    schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
+    optimizer = optax.adabelief(schedule)
+
+
+
+    for i,filename in enumerate(y.filename.unique()): 
+       
+        print(f"Mouse: {i+1}/{y.filename.unique().shape[0]} -- {filename}")
+        tidxs = y[y.filename == filename].index
+        bX,bX_mask = batch_dataset(X[tidxs],1,X_mask[tidxs])
+        p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
+
+        np.save("./results/exp_1_0/"+f"{filename[:-4]}_p0s.npy",p0s)
[…]

### 2024-05-14 mouse/experiment_1_lddmm.py

@@ -1,0 +1,48 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+from src.masked.lddmm import registration,Shooting,Flowing
+from src.masked.kernel import VFTSGaussKernel,TSGaussGaussKernel, TSGaussKernel
+from src.masked.loss import VarifoldLoss
+from src.plotting import plot2Dfigure
+from src.masked.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("./dataset/inv_y.csv",index_col=0)
+X = np.load("./dataset/inv_X.npy")
+X_mask = np.load("./dataset/inv_X_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.filename == filename)*(y.before == "Y")].sample(50).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+Kv = TSGaussKernel(100,1)
+Kl = TSGaussGaussKernel(2,1,2,0.6)
+dataloss = VarifoldLoss(Kl)
+schedule = warmup_cosine_decay_schedule(0,0.01,40,400,0)
+optimizer = optax.adabelief(schedule)
+
+
+if __name__ == "__main__": 
+
+  
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
+
+    np.save("./results/exp_1_lddmm_2/idxs.npy",idxs)
+    np.save("./results/exp_1_lddmm_2/p0s.npy",p0s)
+    np.save("./results/exp_1_lddmm_2/q0.npy",q0)
+    np.save("./results/exp_1_lddmm_2/q0_mask.npy",q0_mask)

### 2024-05-14 mouse/experiment_2.py

@@ -1,0 +1,46 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+from src.masked.lddmm import registration,Shooting,Flowing
+from src.masked.kernel import VFTSGaussKernel,TSGaussGaussKernel
+from src.masked.loss import VarifoldLoss
+from src.plotting import plot2Dfigure
+from src.masked.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+y =pd.read_csv("./dataset/inv_y.csv",index_col=0)
+X = np.load("./dataset/inv_X.npy")
+X_mask = np.load("./dataset/inv_X_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+Kv = VFTSGaussKernel(1,0.1,150,1,1)
+Kl = TSGaussGaussKernel(5,1,5,0.6)
+dataloss = VarifoldLoss(Kl)
+schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
+optimizer = optax.adabelief(schedule)
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.before == "Y")*(y.filename == filename)].sample(25).index
+    lst.append(idx)
+    idx = y[(y.before == "N")*(y.filename == filename)].sample(50).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+if __name__ == "__main__": 
+
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=800,optimizer=optimizer,gamma_loss=1e-4)
+
+    np.save("./results/exp_2_3/idxs.npy",idxs)
+    np.save("./results/exp_2_3/p0s.npy",p0s)
+    np.save("./results/exp_2_3/q0.npy",q0)
+    np.save("./results/exp_2_3/q0_mask.npy",q0_mask)

### 2024-05-14 mouse/experiment_2_lddmm.py

@@ -1,0 +1,51 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+from src.masked.lddmm import registration,Shooting,Flowing
+from src.masked.kernel import VFTSGaussKernel,TSGaussGaussKernel, TSGaussKernel
+from src.masked.loss import VarifoldLoss
+from src.plotting import plot2Dfigure
+from src.masked.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("./dataset/inv_y.csv",index_col=0)
+X = np.load("./dataset/inv_X.npy")
+X_mask = np.load("./dataset/inv_X_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.before == "Y")*(y.filename == filename)].sample(25).index
+    lst.append(idx)
+    idx = y[(y.before == "N")*(y.filename == filename)].sample(50).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+
+Kv = TSGaussKernel(100,1)
+Kl = TSGaussGaussKernel(2,1,2,0.6)
+dataloss = VarifoldLoss(Kl)
+schedule = warmup_cosine_decay_schedule(0,0.01,40,400,0)
+optimizer = optax.adabelief(schedule)
+
+
+if __name__ == "__main__": 
+
+  
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
+
+    np.save("./results/exp_2_lddmm_1/idxs.npy",idxs)
+    np.save("./results/exp_2_lddmm_1/p0s.npy",p0s)
+    np.save("./results/exp_2_lddmm_1/q0.npy",q0)
+    np.save("./results/exp_2_lddmm_1/q0_mask.npy",q0_mask)

### 2024-05-14 mouse/segmentation.py

@@ -1,0 +1,181 @@
+import ruptures as rpt
+from scipy import signal
+import numpy as np
+from sklearn.base import BaseEstimator,TransformerMixin
+from itertools import groupby 
+
+class PhaseSeg(BaseEstimator,TransformerMixin):
+
+    def __init__(self, sampfreq:int, prominence:float, wlen:float, cycle_minimum_duration:float,cycle_maximum_duration:float,trainig_size_per_interval=None,interval=None) -> None:
+        """Initialization 
+
+        Args: 
+            sampfreq (int): Sampling frequency 
+            prominence (float): Peak prominence for volume, in ml.
+            wlen (float): Peak prominence search window, in seconds. 
+            cycle_minimum_duration (float): minimum cycle duration, in seconds. 
+            cycle_maximum_duration (float): maximum cycle duration, in seconds.  
+            trainig_size_per_interval (int): number of cycles slected randomly on each interval. 
+            interval (float): interval duration, in seconds. 
+        """
+        super().__init__()
+        self.sampfreq = sampfreq
+        self.prominence = prominence
+        self.wlen = wlen 
+        self.cycle_minimum_duration = cycle_minimum_duration
+        self.cycle_maximum_duration = cycle_maximum_duration
+        self.trainig_size_per_interval = trainig_size_per_interval
+        self.interval = interval
+
+    def fit(self,X): 
+        """Compute detrended volume
+
+        Args:
+            X (np.ndarray): original sequence, shape (N,).
+
+        """
+        self.flow_ = X
+        self._event_detection()
+        self._filtration()
+        if (self.trainig_size_per_interval is not None) & (self.interval is not None): 
+            self._training_selection()
+
+    def _volume(self,X:np.ndarray)->np.ndarray:
+        """Compute detrended volume
+
+        Args:
+            X (np.ndarray): original sequence, shape (N,).
+
+        Returns:
+            np.ndarray: Detrended volume, shape(N,)
+        """
+        arr = signal.detrend(np.cumsum(X*1/self.sampfreq))
+        arr = arr.astype(float)
+        return arr
+
+    def _event_detection(self)->None:
+        """
+        Detect inhalation and exhalation start_time
+        """
+        #removing unwanted time
+        self.flow_ = self.flow_.astype(float)
+
+        #getting volume
+        self.volume_ = self._volume(self.flow_)
+
+        #Getting start inspiration
[…]

### 2024-05-14 src/barycenter.py

@@ -1,0 +1,118 @@
+import numpy as np
+import jax.numpy as jnp
+from jax import vmap,grad,jit
+from jax.tree_util import Partial
+import optax
+
+from src.lddmm import LDDMMLoss,TimeLDDMMLoss
+from src.loss import VarifoldLoss
+from src.optimizer import Optimizer, BatchBarycenterOptimizer,BatchBarycenterTimeOptimizer
+
+####################################################################################################################################
+####################################################################################################################################
+### GENERAL ###
+####################################################################################################################################
+####################################################################################################################################
+
+def argmedian_length_from_mask(x):
+    return np.argpartition(x, np.sum(x) // 2)[len(x) // 2]
+
+def BarycenterLDDMMLoss(Kv,dataloss:callable,gamma_loss =0.,nt=10,deltat=1.0): 
+    unit_loss = LDDMMLoss(Kv,dataloss,gamma_loss,nt,deltat)
+    f = vmap(unit_loss,(0,None,None,0,0),0)
+    def bloss(p0,q0,q0_mask,q1,q1_mask): 
+        return jnp.sum(f(p0,q0,q0_mask,q1,q1_mask))
+    return bloss
+
+def barycenter_registration(sigs,sigs_masks,Kv,dataloss:callable,init=None,init_mask=None,niter=100, optimizer = optax.adam(learning_rate=0.1),gamma_loss =0.,nt=10,deltat=1.0,verbose=True):
+    if init is None: 
+        median_length_idx = np.argsort(np.sum(sigs_masks,axis=1).flatten(),kind='stable')[sigs_masks.shape[0]//2]
+        q0 = sigs[median_length_idx]
+        q0_mask = sigs_masks[median_length_idx]
+    else: 
+        q0 = init
+        q0_mask = init_mask
+    p0 = jnp.zeros_like(sigs,dtype = jnp.float32)
+    bloss = BarycenterLDDMMLoss(Kv,dataloss,gamma_loss,nt,deltat)
+    opt = Optimizer(bloss,niter,optimizer,static_p0=False,static_q0=False,verbose=verbose)
+    return *opt(p0,q0,q0_mask,sigs,sigs_masks),q0_mask
+
[…]

### 2024-05-14 src/classifier.py

@@ -1,0 +1,114 @@
+import numpy as np
+import optax
+
+from sklearn.svm import SVC
+from sklearn.metrics import accuracy_score
+from src.utils import MaskMomentaGramMatrix,batch_dataset,unbatch_dataset
+from src.lddmm import batch_one_to_many_registration
+from src.barycenter import batch_barycenter_registration
+
+
+class MomentaSVC(SVC):
+
+    def __init__(self,Kv,dataloss,barycenter_initializer = None,registration_initializer=None,C=1,max_per_batch =10,gamma_loss=0.0,niter=200,optimizer = optax.adam(learning_rate=0.1),nt=10,deltat=1.0, shrinking = True, probability = False, tol = 0.001, cache_size = 200, class_weight = None, verbose = True, max_iter =-1, decision_function_shape = "ovr", break_ties = False, random_state = None) -> None:
+        super().__init__(C=C, kernel="precomputed",shrinking=shrinking, probability=probability, tol=tol, cache_size=cache_size, class_weight=class_weight, verbose=False, max_iter=max_iter, decision_function_shape=decision_function_shape, break_ties=break_ties, random_state=random_state)
+        
+        self.Kv = Kv
+        self.dataloss = dataloss
+        self.barycenter_initializer = barycenter_initializer
+        self.registration_initializer = registration_initializer
+        self.max_per_batch = max_per_batch
+        self.gamma_loss = gamma_loss
+        self.niter = niter
+        self.optimizer = optimizer
+        self.nt = nt
+        self.deltat = deltat
+        self.verbose = verbose
+
+    def _batch(self,X,X_mask): 
+        n_samples = X.shape[0]
+        for i in np.arange(self.max_per_batch,0,-1): 
+            if n_samples % i ==0: 
+                self.batch_size_ = i
+                batched_X, batched_X_mask = batch_dataset(X,self.batch_size_,X_mask)
+                break
+        return batched_X, batched_X_mask
+    
+    def _unbached(self,batched_X, batched_X_mask=None): 
+        return unbatch_dataset(batched_X,batched_X_mask)
+    
+    def fit(self,X,X_mask,y,init=None,init_mask=None): 
+        batched_X,batched_X_mask = self._batch(X,X_mask)
[…]

### 2024-05-14 src/kernel.py

@@ -1,0 +1,172 @@
+import numpy as np
+import jax.numpy as jnp
+
+
+def GaussKernel(sigma,mu=1): 
+    oos2 = 1/sigma**2
+    def K(x,mask_x,y,mask_y,b):
+        res = jnp.exp(-oos2*jnp.sum((x[:,None,:]-y[None,:,:])**2,axis=2))* (mask_x*mask_y.T)
+        return mu*res@b
+    return K
+
+
+def ExpKernel(sigma,mu=1): 
+    oos2 = 1/sigma
+    def K(x,mask_x,y,mask_y,b):
+        m = jnp.sum((x[:,None,:]-y[None,:,:])**2,axis=2)
+        m = jnp.sqrt(jnp.where(m>0,m,0))
+        res = jnp.exp(-oos2*m)* (mask_x*mask_y.T)
+        return mu*res@b
+    return K
+
+def TSGaussKernel(t_sigma:float,s_sigma:float,np_dtype = np.float32): 
+    t_oos, s_oos  = 1/t_sigma, 1/s_sigma
+    def K(x,mask_x,y,mask_y,b): 
+        n_d = x.shape[1]
+        oos = np.ones(n_d,dtype=np_dtype).reshape(1,-1)
+        oos[0,0], oos[0,1:] = t_oos, s_oos
+        res = jnp.exp(-jnp.sum(((x*oos)[:,None,:]-(y*oos)[None,:,:])**2,axis=2))
+        mask = (mask_x*mask_y.T)
+        return (res*mask)@b
+    return K
+
+def TSGaussGaussKernel(t_sigma_1:float,s_sigma_1:float,t_sigma_2:float,s_sigma_2:float,np_dtype = np.float32): 
+    t_oos_1, s_oos_1  = 1/t_sigma_1, 1/s_sigma_1
+    t_oos_2, s_oos_2 = 1/t_sigma_2, 1/s_sigma_2
+    def K(x,y,u,v,mask_xu,mask_yv,b): 
+        n_d = x.shape[1]
+        oos_1 = np.ones(n_d,dtype=np_dtype).reshape(1,-1)
+        oos_1[0,0], oos_1[0,1:] = t_oos_1, s_oos_1
[…]

### 2024-05-14 src/lddmm.py

@@ -1,0 +1,371 @@
+import numpy as np 
+import jax.numpy as jnp
+import jax.lax as lax
+from jax import jit, grad, vmap, jacfwd
+import optax
+from src.loss import VarifoldLoss
+from src.optimizer import Optimizer, RegistrationOptimizer, BatchOneToManyRegistrationOptimizer
+####################################################################################################################################
+####################################################################################################################################
+### GENERAL ###
+####################################################################################################################################
+####################################################################################################################################
+
+def Hamiltonian(K:callable): 
+    def H(p,q,q_mask): 
+       return 0.5*jnp.sum(p*K(q,q_mask,q,q_mask,p))
+    return H 
+
+def HamiltonianSystem(K:callable):
+    H = Hamiltonian(K)
+    def HS(p,q,q_mask):
+        Gp,Gq = grad(H,(0,1))(p,q,q_mask)
+        return -Gq,Gp
+    return HS
+
+def HamiltonianRalstonIntegrator(K:callable,nt=10,deltat=1.): 
+    HSystem = HamiltonianSystem(K)
+    dt = deltat/nt
+    def body_function(i,arg): 
+        p,q,q_mask = arg
+        dp,dq = HSystem(p,q,q_mask)
+        pi,qi = p+(2*dt/3)*dp,q+(2*dt/3)*dq
+        dpi,dqi = HSystem(pi,qi,q_mask)
+        p,q = p+0.25*dt*(dp+3*dpi),q+0.25*dt*(dq+3*dqi)
+        return p,q,q_mask
+    def f(p,q,q_mask):
+        p,q,q_mask = lax.fori_loop(0,nt,body_function,(p,q,q_mask))
+        return p,q 
+    return f
+
+def Shooting(K:callable,nt=10,deltat=1.0): 
+    Hri = HamiltonianRalstonIntegrator(K,nt,deltat)
+    def shoot(p0,q0,q0_mask): 
+        return Hri(p0,q0,q0_mask)
+    return shoot
+
[…]

### 2024-05-14 src/loss.py

@@ -1,0 +1,87 @@
+import jax.numpy as jnp
+import numpy as np
+import jax
+
+
+def compCLN(x:jnp.ndarray): 
+    x0, x1 = x[:-1,:], x[1:,:]
+    p = .5*(x0+x1)
+    v = (x1-x0)
+    m = jnp.sum(v**2,axis=1).reshape(-1,1)
+    m = jnp.sqrt(jnp.where(m>0,m,0))
+    v = v*jnp.where(m==0.,0.,1./m)
+    return m,p,v
+
+
+def compELT(x:jnp.ndarray): 
+    t_x0, s_x0, t_x1, s_x1 = x[:-1,:1], x[:-1,1:], x[1:,:1],x[1:,1:]
+    t_p, s_p = .5*(t_x0+t_x1), .5*(s_x0+s_x1)
+    t_v = t_x1 - t_x0
+    s_v = (s_x1-s_x0)*jnp.where(t_v==0.,0.,1/t_v)
+    return t_p,s_p,t_v,s_v
+
+
+def MyLoss(K:callable): 
+    def loss(s_x,mask_s_x,t_x,mask_t_x):
+        t_t_p,t_s_p,t_t_v,t_s_v = compELT(t_x)
+        s_t_p,s_s_p,s_t_v,s_s_v = compELT(s_x)
+        c0 = jnp.sum(t_t_v * K(t_t_p,t_s_p,t_s_v,mask_t_x[1:,:],t_t_p,t_s_p,t_s_v,mask_t_x[1:,:],t_t_v))
+        c1 = jnp.sum(s_t_v * K(s_t_p,s_s_p,s_s_v,mask_s_x[1:,:],t_t_p,t_s_p,t_s_v,mask_t_x[1:,:],t_t_v))
+        c2 = jnp.sum(s_t_v * K(s_t_p,s_s_p,s_s_v,mask_s_x[1:,:],s_t_p,s_s_p,s_s_v,mask_s_x[1:,:],s_t_v))
+        return c0 -2*c1 +c2
+    return loss
+
+
+
+def VarifoldLoss(K:callable): 
+    def loss(s_x,mask_s_x,t_x,mask_t_x): 
+        t_m,t_p,t_v = compCLN(t_x)
+        s_m,s_p,s_v = compCLN(s_x)
+        c0 = jnp.sum(t_m * K(t_p,t_p,t_v,t_v,mask_t_x[1:,:],mask_t_x[1:,:],t_m))
[…]

### 2024-05-14 src/optimizer.py

@@ -1,0 +1,195 @@
+import optax
+from jax import jit,grad,value_and_grad,vmap
+import jax.numpy as jnp
+from jax import lax
+import numpy as np
+from functools import partial
+
+
+####################################################################################################################################
+####################################################################################################################################
+### GENERAL ###
+####################################################################################################################################
+####################################################################################################################################
+
+def Optimizer(loss:callable,niter=100,optimizer = optax.adam(learning_rate=0.1),static_p0 =False, static_q0=True,verbose=True):
+    def f(p0,q0,q0_mask,q1,q1_mask):
+        
+        if (static_p0==False)*(static_q0==True):
+            arg = p0
+            def map_loss(arg):
+                return loss(arg,q0,q0_mask,q1,q1_mask)
+        elif (static_p0==False)*(static_q0==False): 
+            arg = (p0,q0)
+            def map_loss(arg):
+                return loss(*arg,q0_mask,q1,q1_mask)
+        elif (static_p0==True)*(static_q0==False): 
+            arg = q0
+            def map_loss(arg): 
+                return loss(p0,arg,q0_mask,q1,q1_mask)
+            
+        opt_state = optimizer.init(arg)
+        jit_map_loss = jit(value_and_grad(map_loss))
+
+        def step(arg,opt_state):
+            loss_value,grads = jit_map_loss(arg) 
+            updates, opt_state = optimizer.update(grads, opt_state,arg)
+            arg = optax.apply_updates(arg,updates)
+            return arg, opt_state, loss_value
+        
+        for i in range(niter): 
+            arg,opt_state,loss_value = step(arg,opt_state)
+            if verbose: 
+                if (i==0)+((i+1)%10 == 0):
+                    print("iteration: ",i+1,"/",niter, " -- loss: ", "{:0.2f}".format(loss_value))
+        return arg
+    return f 
+
+####################################################################################################################################
+####################################################################################################################################
+### Registration ###
[…]

### 2024-05-14 src/plotting.py

@@ -1,0 +1,49 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+def plot2Dfigure(s_sig,t_sig,p0,shoot,flow,mask_s_sig=None,mask_t_sig=None,nv=50,nh=50): 
+    if mask_s_sig is None:
+        s_sig = (s_sig,)
+        plot_s_sig = s_sig 
+    else: 
+        plot_s_sig = s_sig[np.where(mask_s_sig==True)[0],:]
+        s_sig = (s_sig,mask_s_sig)
+  
+    if mask_t_sig is None: 
+        t_sig = (t_sig,)
+        plot_t_sig = t_sig
+    else: 
+        plot_t_sig = t_sig[np.where(mask_t_sig==True)[0],:]
+        t_sig = (t_sig,mask_t_sig)
+        
+
+    _,p_sig = shoot(p0,*s_sig)
+    if mask_s_sig is None: 
+        plot_p_sig = p_sig
+    else: 
+        plot_p_sig = p_sig[np.where(mask_s_sig==True)[0],:]
+
+    fig,axs = plt.subplots(2,1,sharex=True,sharey=True, figsize = (10,5))
+    axs[0].plot(*plot_s_sig.T, label = "from")
+    axs[0].legend()
+    axs[1].plot(*plot_t_sig.T, color = "green", label = "to")
+    axs[1].plot(*plot_p_sig.T, color = "r", label = "warping")
+    axs[1].legend()
+    sz = 0.2 
+    a = min(np.min(plot_s_sig[:,0]),np.min(plot_t_sig[:,0]),np.min(plot_p_sig[:,0]))
+    b = max(np.max(plot_s_sig[:,0]),np.max(plot_t_sig[:,0]),np.max(plot_p_sig[:,0]))
+    c = min(np.min(plot_s_sig[:,1]),np.min(plot_t_sig[:,1]),np.min(plot_p_sig[:,1]))
+    d = max(np.max(plot_s_sig[:,1]),np.max(plot_t_sig[:,1]),np.max(plot_p_sig[:,1]))
+    lsp1 = np.linspace(a-sz*(b-a),b+sz*(b-a),nv,dtype=np.float32)
+    lsp2 = np.linspace(c-sz*(d-c),d+sz*(d-c),nh,dtype=np.float32)
+    X1, X2 = np.meshgrid(lsp1,lsp2)
[…]

### 2024-05-14 src/statistic.py

@@ -1,0 +1,130 @@
+import numpy as np
+import jax.numpy as jnp
+from jax import vmap,jit
+from jax.lax import map 
+from sklearn.base import BaseEstimator
+from numpy.linalg import svd
+from scipy.linalg import solve
+from sklearn.discriminant_analysis import LinearDiscriminantAnalysis
+from functools import partial
+from scipy.linalg import sqrtm,inv
+
+def MomentaGramMatrix(K,qbar):
+    f = lambda u,v : jnp.sum(u*K(qbar,qbar,v))
+    f = vmap(f,(0,None),0)
+    f = vmap(f,(None,0),1)
+    return f
+
+def MaskMomentaGramMatrix(K,qbar,qbar_mask):
+    f = lambda u,v : jnp.sum(u*K(qbar,qbar_mask,qbar,qbar_mask,v))
+    f = vmap(f,(0,None),0)
+    f = vmap(f,(None,0),1)
+    return jit(f)
+
+def MapMaskMomentaGramMatrix(K,qbar,qbar_mask):
+    f = lambda u,v : jnp.sum(u*K(qbar,qbar_mask,qbar,qbar_mask,v))
+    def mapf(u,v): 
+        nu = u.shape[0]
+        nv = v.shape[0]
+        arr = np.zeros((nu,nv))
+        for i in range(nu):
+            for j in range(i,nv): 
+                val = f(u[i],v[j])
+                arr[i,j] = val
+                arr[j,i] = val
+        return arr 
+    return mapf
+
+
+
+
+class MomentaPCA(BaseEstimator):
+
+    def __init__(self,n_component,centered=False,type = "cov") -> None:
+        super().__init__()
+        self.n_component = n_component
+        self.centered = centered
+        self.type = type    
+
+
+    def fit(self,Kv,ps,qbar,qbar_mask): 
+
+        self.Kv_ = Kv
+        self.qbar_ = qbar
+        self.ps_ = ps
+
+    
+        self._mgm = MaskMomentaGramMatrix(self.Kv_,qbar,qbar_mask)
+       
+
+        if self.centered:
+            r_ps = ps
+        else:
+            self.m_ps_ = np.mean(ps,axis=0)
+            r_ps = ps - self.m_ps_[None,:,:]
+
+        X = self._mgm(r_ps,r_ps)
+
+        if self.type == "cor": 
+            inv_norm = 1/np.sqrt(np.diag(X))
[…]

### 2024-05-14 src/utils.py

@@ -1,0 +1,55 @@
+import numpy as np
+import jax.numpy as jnp
+from jax import vmap 
+
+def ComputeKernelGramMatrix(K):
+    f = lambda x,y,u,v : jnp.sum(u*K(x,y,v))
+    f = vmap(f,(None,None,0,None),0)
+    f = vmap(f,(None,None,None,0),1)
+    return f
+
+def time_shape_embedding(x:np.ndarray,sampfreq=1,max_length=None,dtype = jnp.float32): 
+    n_s = x.shape[0]
+    time = np.arange(n_s,dtype=float).reshape(-1,1)/sampfreq
+    t_x = np.hstack((time,x),dtype=dtype)
+    if max_length is None:
+        return t_x
+    else:
+        t_x = np.pad(t_x,((0,max_length-n_s),(0,0))).astype(dtype)
+        mask = np.full_like(t_x[:,:1],True,dtype=bool)
+        mask[n_s:,:] = False
+        return t_x, mask
+
+def from_timeseries_to_dataset(X:list,sampfreq=1,dtype=jnp.float32): 
+    lengths = np.array([x.shape[0] for x in X])
+    max_length = np.max(lengths)
+    mask_lst = []
+    ts_lst = []
+    for ts in X: 
+        t_ts, t_mask = time_shape_embedding(ts,sampfreq,max_length,dtype=dtype)
+        ts_lst.append(t_ts)
+        mask_lst.append(t_mask)
+    return jnp.array(ts_lst,dtype=dtype), jnp.array(mask_lst,dtype=jnp.bool_)
+    
+def batch_dataset(dataset,batch_size,masks=None,dtype=jnp.float32): 
+    if dataset.shape[0]%batch_size!=0: 
+        raise ValueError("dataset size is not a multiple of batch_size")
+    n_ts,n_s,n_d = dataset.shape
+    n_batches = n_ts//batch_size
+    if masks is None: 
+        return dataset.reshape(n_batches,batch_size,n_s,n_d).astype(dtype)
+    else: 
+        return dataset.reshape(n_batches,batch_size,n_s,n_d).astype(dtype), masks.reshape(n_batches,batch_size,n_s,1)
+    
+def unbatch_dataset(batched_dataset,batched_masks=None,dtype=jnp.float32): 
+    _,_,n_s,n_d = batched_dataset.shape
+    if batched_masks is None: 
[…]

### 2024-05-14 NEURIPS24/neurips_2024_to_submit.tex
update

@@
     %TO CHANGE]
-    %Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
-    %Individuals have specificities depending on their genes or disease
-     %while sharing, on average, a similar pattern in the features of their corresponding time series.
+    Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
+    Individuals have specificities depending on their genes or disease
+     while sharing, on average, a similar pattern in the features of their corresponding time series.
 For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
@@
     This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies.
-The proposed algorithm manages irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
-    At the crossroads between URL for time series and shape analysis, the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series.
+    The idea is to represent time series as the deformation of a time series of reference using Large Deformation Diffeomorphic Metric Mapping (LDDMM).
+     Once the deformations and time series of reference are learned, the vector representations of time series are then given by the parametrization of the deformation.
+     At the crossroads between URL for time series and shape analysis, the proposed algorithm manages irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
+    %At the crossroads between URL for time series and shape analysis.
+%the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series.
 This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension.

### 2024-05-14 NEURIPS24/paper.tex
update
up

@@
 %       experiments on synthetic data are presented to show that the parameter of representation are identifiable when the time seriess are not too sharp and the kernels parameter well tuned. 
-%       Then, we show how to apply the method on an unsupervised case of moouse respeitaory dataset and a supervised one on ... .
+%       Then, we sh-ow how to apply the method on an unsupervised case of moouse respeitaory dataset and a supervised one on ... .
 
@@
 
- Our goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology \cite{wang2016research,bar2012studying,germain2023unsupervised}.
+ Our goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
   More specifically, we aim to find an unsupervised features representation method which encode the specificity of an individual compared to another. 

### 2024-05-14 NEURIPS24/neurips_2024_to_submit.tex
up
update

@@
 \maketitle
-
-\begin{abstract}
-    %TO CHANGE]
-    Analyzing the inter-individual variability within a cluster of time series is particularly appealing in medicine and biology.
-    Individuals have specificities depending on their genes or disease
-     while sharing, on average, a similar pattern in the features of their corresponding time series.
-For instance, an electrocardiogram presents a typical shape of the heartbeat, which varies depending on the individual.
 %In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
@@
     %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
-    This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies.
-    The idea is to represent time series as the deformation of a time series of reference using Large Deformation Diffeomorphic Metric Mapping (LDDMM).
-     Once the deformations and time series of reference are learned, the vector representations of time series are then given by the parametrization of the deformation.
-     At the crossroads between URL for time series and shape analysis, the proposed algorithm manages irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
     %At the crossroads between URL for time series and shape analysis.
@@
 This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension.
+
+\begin{abstract}
+    %TO CHANGE]
+    Analyzing inter-individual variability of physiological functions is particularly appealing in medicine and biology to describe or quantify health conditions.
+     Such analysis can be done by comparing individuals to a reference one using their time series as biomedical data.
+     % In biomedical research, time series data is a prevalent source of information, necessitating dedicated methods comparing them and averaging them to derive a valuable reference time series. 
+ This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies.
+ The idea is to represent time series as deformations of a reference time series. The deformations are diffeomorphisms parameterized and learned by our method called TS-LDDMM.
+  Once the deformations and the reference time series are learned, the vector representations of individual time series are given by the parametrization of their corresponding deformation.
+   At the crossroads between URL for time series and shape analysis, the proposed algorithm handles irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
  In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework.

### 2024-05-15 NEURIPS24/neurips_2024_to_submit.tex
update

@@
     %TO CHANGE]
-    Analyzing inter-individual variability of physiological functions is particularly appealing in medicine and biology to describe or quantify health conditions.
-     Such analysis can be done by comparing individuals to a reference one using their time series as biomedical data.
-     % In biomedical research, time series data is a prevalent source of information, necessitating dedicated methods comparing them and averaging them to derive a valuable reference time series. 
- This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to biomedical inter-individual studies.
- The idea is to represent time series as deformations of a reference time series. The deformations are diffeomorphisms parameterized and learned by our method called TS-LDDMM.
+    Analyzing inter-individual variability of physiological functions is particularly appealing in medical and biological contexts to describe or quantify health conditions.
+Such analysis can be done by comparing individuals to a reference one with time series as biomedical data.
+This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to inter-individual studies.
+The idea is to represent time series as deformations of a reference time series.
+The deformations are diffeomorphisms parameterized and learned by our method called TS-LDDMM.
 Once the deformations and the reference time series are learned, the vector representations of individual time series are given by the parametrization of their corresponding deformation.

### 2024-05-15 NEURIPS24/paper.tex
update

@@
 
- A time series graph can be seen a curve and LDDMM structure is relevant to tackle curves as presented in \cite{glaunes2008large}. %, LDDMM structure is relevant to tackle curves and
-  However, time series graph has more structure as depicted in \Cref{fig:diffeo} due to the temporal evolution. 
+ A time series graph can be seen as a curve and LDDMM structure is relevant to tackle curves as presented in \cite{glaunes2008large}. %, LDDMM structure is relevant to tackle curves and
+  However, time series graph has more structure than curves as depicted in \Cref{fig:diffeo} due to the temporal evolution. 
   \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM, but distinguish from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
@@
   Again, the time series graph structure is not respected.
-   Very recently in a functionnal data analysis framework \cite{wu2024shape} (Shape-FPCA) improved by giving also a representation for the original time evolution.
+   Very recently in a functionnal data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by giving a representation for the original time evolution.
     Nevertheless, this methods applies only on time series of \textit{same size} and is made for \textit{continuous objects}, making the estimation more sensitive to noise.

### 2024-05-15 NEURIPS24/neurips_2024_to_submit.tex
polish intro

@@
 \begin{abstract}
-    %TO CHANGE]
 Analyzing inter-individual variability of physiological functions is particularly appealing in medical and biological contexts to describe or quantify health conditions.

### 2024-05-15 NEURIPS24/paper.tex
polish intro

@@
 
-  Although a community structure with representatives can be learned in an unsupervised way \cite{trirat2024universal,meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}
-  or similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},
+Although a community structure with representatives can be learned in an unsupervised way \cite{trirat2024universal,meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised} or similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},
 studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem in URL.
@@
 in order to apply classical statiscal or machine learning tools depending on the needs such as Principal Component Analysis (PCA), logistic regression or deep learning methods for classification or anomaly detection.
+
+\thi{
+  Our goal is to analyze the inter-individual variability within a time series dataset, an approach of prime interest in physiological contexts \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
+More specifically, we aim to find an unsupervised features representation method that encodes individual time series specificities compared to a reference one.
+}
+
+\thi{
+  Studying shape differences between time series related to biological mechanisms is a common practice in physiology to characterize healthy and pathological functioning CITE.
+For instance, the shapes of heartbeats in electrocardiograms are discriminant for some cardiovascular pathologies CITE.
+Several approaches have been proposed for such comparison.
+Some employ shape-based similarity measures between time series \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},  others embed time series as vectors of predefined features CITE, and, with the rise of deep neural networks, unsupervised learning representation of time series~\cite{trirat2024universal,meng2023unsupervised} has shown to be a valuable approach CITE notably with contrastive learning~\cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
+However, shape-based representation of time series within cohorts~\cite{niennattrakul2007inaccuracies,shirato2023identifying} remains an open problem in URL.
+}
+
+\thi{
+  Our work focuses explicitly on learning shape-based representation of time series.
+First, we propose not to see the shape of a time series through its curve $\{s_t:\eqsp t\in\msi\}$, but rather through its graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\}$.
+Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
+The idea is to represent each element $\msg(s^j)$ of a dataset $(s^j)_{j\in[N]}$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism $\phi_j$ (ie $\msg(s^j) \sim \phi_j .
+\msg(\mathbf{s}_0)$ ).
+The diffeomorphism $\phi_j$ is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
+The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
+Finally, these shape-encoding features can feed any statistical or machine-learning model.
+  }
+
+\thi{
+  However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
+\Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
+Our contributions arise from this observation: we specify the class of diffeomorphisms to consider and show how to learn them.
+This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
+}
+
+\thi{
+  Our contributions can be summarized as follows:
+    \begin{itemize}
+       \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset.
+In particular, the method can handle multivariate time series \textit{irregularly sampled} and with \textit{variable sizes}.
+       \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
+       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters, also providing guidelines for tuning.
+       \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
+        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets.
+ \end{itemize}
+}
 

### 2024-05-15 NEURIPS24/paper.tex
update

@@
 
-Our goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
-More specifically, we aim to find an unsupervised features representation method which encode the specificity of an individual compared to another. 
+ Our goal is to analyze the inter-individual variability within a time series dataset, an approach of prime interest in physiological contexts \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
+  More specifically, we aim to find an unsupervised features representation method that encodes individual time series specificities compared to a reference one.
 In physiology, studying the different \textit{shapes} in a time series related to biological phenomena and their variations according to individual or pathology is common.
@@
 %However, time series data with their specific structure has not been properly adressed.
-
- First, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
- Then, building on the shape analysis litterature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
-  The idea is to represent each element $(\msg(s^j))_{j\in[N]}$ of the dataset as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism.
-  Then, the diffeomorphism is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
+Our work focuses explicitly on learning shape-based representation of time series.
+First, we propose not to see the shape of a time series through its curve $\{s_t:\eqsp t\in\msi\}$, but rather through its graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\}$.
+   Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
+The idea is to represent each element $\msg(s^j)$ of a dataset $(s^j)_{j\in[N]}$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism $\phi_j$, i.e.
+$\msg(s^j) \sim \phi_j .
+\msg(\mathbf{s}_0)$ .
+    The diffeomorphism $\phi_j$ is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
      The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
-    Finally, these features encoding the shapes can feed any statistical or machine learning model as in URL.
+Finally, these shape-encoding features can feed any statistical or machine-learning model.
+%  First, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
+%  Then, building on the shape analysis litterature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
+%   The idea is to represent each element $(\msg(s^j))_{j\in[N]}$ of the dataset as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism.
+%   Then, the diffeomorphism is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
+%    The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
+%     Finally, these features encoding the shapes can feed any statistical or machine learning model as in URL.
 
@@
 \Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
-   Our contributions arise from this observation.
+ Our contributions arise from this observation: we specify the class of diffeomorphisms to consider and show how to learn them.
+  This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
    %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
-    We solve this issue by specifying the class of diffeomorphisms to consider and showing how to learn them.
-     This change is fruitful in representing time transformation as illustrated in \Cref{fig:transport}.
+
      %In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
 
+    %  Our contributions can be summarized as follows:
+    %  \begin{itemize}
+    %    \item We propose an unsupervised method (TS-LDDMM) to analyze inter-individual variability of shapes in a time series dataset.
+In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
+    %    \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
+    %    \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method with respect to its hyperparameters, also providing guidelines for tuning.
+    %    We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
+    %     We illustrate the quantitative interest of the representation on classification tasks on real shape-based datasets.
+
+    %  \end{itemize}
     Our contributions can be summarized as follows:
     \begin{itemize}
-       \item We propose an unsupervised method (TS-LDDMM) to analyze inter-individual variability of shapes in a time series dataset.
-In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
+       \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset.
+In particular, the method can handle multivariate time series \textit{irregularly sampled} and with \textit{variable sizes}.
        \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method with respect to its hyperparameters, also providing guidelines for tuning.
-       We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-        We illustrate the quantitative interest of the representation on classification tasks on real shape-based datasets.
-
+       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters, also providing guidelines for tuning.
+       \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
+        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets.
  \end{itemize}
@@
 
-\thi{
-  Our goal is to analyze the inter-individual variability within a time series dataset, an approach of prime interest in physiological contexts \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
+% \thi{
+%   Our goal is to analyze the inter-individual variability within a time series dataset, an approach of prime interest in physiological contexts \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
 More specifically, we aim to find an unsupervised features representation method that encodes individual time series specificities compared to a reference one.
-}
+% }
 
@@
 
-\thi{
-  Our work focuses explicitly on learning shape-based representation of time series.
+% \thi{
+%   Our work focuses explicitly on learning shape-based representation of time series.
 First, we propose not to see the shape of a time series through its curve $\{s_t:\eqsp t\in\msi\}$, but rather through its graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\}$.
-Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
+%    Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
 The idea is to represent each element $\msg(s^j)$ of a dataset $(s^j)_{j\in[N]}$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism $\phi_j$ (ie $\msg(s^j) \sim \phi_j .
 \msg(\mathbf{s}_0)$ ).
-The diffeomorphism $\phi_j$ is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
-The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
+%     The diffeomorphism $\phi_j$ is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
+%      The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
 Finally, these shape-encoding features can feed any statistical or machine-learning model.
-  }
+%   }
 
@@
 
-\thi{
-  Our contributions can be summarized as follows:
-    \begin{itemize}
-       \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset.
+% \thi{
+%   Our contributions can be summarized as follows:
+%     \begin{itemize}
+%        \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset.
 In particular, the method can handle multivariate time series \textit{irregularly sampled} and with \textit{variable sizes}.
-       \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters, also providing guidelines for tuning.
-       \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets.
- \end{itemize}
-}
+%        \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
+%        \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters, also providing guidelines for tuning.
+%        \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
+%         \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets.
+%  \end{itemize}
+% }
 

### 2024-05-20 NEURIPS24/appendix.tex
first results

@@
 
-\section{Classification datasets}
-\label{appendix:classification_dataset}
+\section{Classification: Comparison with shape analysis methods}
 
-All datasets were taken from UCR/UEA archives \cite{dau2019ucr,bagnall2018uea}. Among all available 
-datasets\footnote{https://timeseriesclassification.com}, we selected 4 datasets related to time series 
-shape comparison. All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/} 
-which already includes the train test split. Essential dataset information is summarized in \Cref{appendix:table:datasets}.
+In this section, we compare classification performances of TS-LDDMM with other state-of-the-art methods coming from shape analysis on 15 shape-based datasets of time-series.
+
+\paragraph{Methods} We compare TS-LDDMM with a method from function~\cite{wu2024shape}
+
+\paragraph{Protocole}
 
@@
   \centering
-  \caption{Time series datasets summary for shape based classification.}
+  \resizebox{\columnwidth}{!}{%
+  \begin{tabular}{llrrrr}
+    \toprule
+     & \textbf{Dataset} & \textbf{Shape-FPCA (2024)} & \textbf{TCLR (2024)} & \textbf{LDDMM (2008)} & \textbf{TS-LDDMM (ours)} \\
+    \midrule
+    \multirow[c]{7}{*}{Univariate} & ArrowHead & 0.18 & 0.75 & \underline{0.84} & \textbf{0.91} \\
+     & BME & 0.16 & \underline{1.00} & 0.82 & \textbf{1.00} \\
+     & ECG200 & 0.40 & 0.67 & \textbf{0.81} & \underline{0.79} \\
+     & FacesUCR & 0.08 & \underline{0.73} & 0.69 & \textbf{0.86} \\
+     & GunPoint & 0.93 & \underline{0.97} & 0.83 & \textbf{1.00} \\
+     & PhalangesOutlinesCorrect & 0.39 & \textbf{0.63} & \underline{0.53} & 0.52 \\
+     & Trace & 0.55 & \underline{1.00} & 0.46 & \textbf{1.00} \\
+    \cline{1-6}
+    \multirow[c]{8}{*}{Multivariate} & ArticularyWordRecognition & -- & -- & \underline{0.98} & \textbf{1.00} \\
+     & Cricket & -- & -- & \underline{0.77} & \textbf{0.93} \\
+     & ERing & -- & -- & \underline{0.95} & \textbf{0.98} \\
+     & Handwriting & -- & -- & \underline{0.22} & \textbf{0.44} \\
+     & Libras & -- & -- & \underline{0.56} & \textbf{0.60} \\
+     & NATOPS & -- & -- & \underline{0.82} & \textbf{0.82} \\
+     & RacketSports & -- & -- & \textbf{0.83} & \underline{0.79} \\
+     & UWaveGestureLibrary & -- & -- & \underline{0.72} & \textbf{0.81} \\
+    \bottomrule
+    \end{tabular}
+    %
+  }
+
+\end{table}
+
+\section{Classification datasets}
+\label{appendix:classification_dataset}
+We selected 15 shape-based datasets (7 univariates and 8 multivariates) from the from the University of East Anglia (UEA) and the University of California Riverside (UCR) Time Series Classification Repository\footnote{https://timeseriesclassification.com} \cite{dau2019ucr,bagnall2018uea}.
+All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/}.
+Essential datasets information are summarized in \Cref{appendix:table:datasets} and further can be found in \cite{dau2019ucr,bagnall2018uea}.
+
+\begin{table}[hbt!]
+  \centering
+  \caption{UCR/UEA shape-based time series datasets for classification.}
+  \resizebox{\columnwidth}{!}{%
   \begin{tabular}{lllllll}
     \toprule
-    Dataset & Train size & test size & Lenghth & Number of classes & Number of dimensions & Type \\
+    & \textbf{Dataset} &  \textbf{Size} & \textbf{Lengh} & \textbf{Number of classes} & \textbf{Number of dimensions} & \textbf{Type} \\
     \midrule
-    \href{https://timeseriesclassification.com/description.php?Dataset=ArrowHead}{ArrowHead} & 36 & 175 & 251 & 3 & 1 & IMAGE \\
-    \href{https://timeseriesclassification.com/description.php?Dataset=ECG200}{ECG200} & 100 & 100 & 96 & 2 & 1 & ECG \\
-    \href{https://timeseriesclassification.com/description.php?Dataset=GunPoint}{GunPoint} & 50 & 150 & 150 & 2 & 1 & MOTION \\
-    \href{https://timeseriesclassification.com/description.php?Dataset=NATOPS}{NATOPS} & 180 & 180 & 51 & 6 & 24 & MOTION \\
+    \multirow[c]{7}{*}{Univariate} &  ArrowHead & 211 & 251 & 3 & 1 & IMAGE \\
+    & BME & 180 & 128 & 3 & 1 & SIMULATED \\
+    & ECG200 & 200 & 96 & 2 & 1 & ECG \\
+    & FacesUCR & 2250 & 131 & 14 & 1 & IMAGE \\
+    & GunPoint & 200 & 150 & 2 & 1 & MOTION \\
+    & PhalangesOutlinesCorrect & 2658 & 80 & 2 & 1 & IMAGE \\
+    & Trace & 200 & 275 & 4 & 1 & SENSOR \\
+    \cline{1-7}
+    \multirow[c]{8}{*}{Multivariate}& ArticularyWordRecognition & 575 & 144 & 25 & 9 & SENSOR \\
+    & Cricket & 180 & 1197 & 12 & 6 & MOTION \\
+    & ERing & 60 & 65 & 6 & 4 & SENSOR \\
+    & Handwriting & 1000 & 152 & 26 & 3 & MOTION \\
+    & Libras & 360 & 45 & 15 & 2 & VIDEO \\ 
+    & NATOPS & 360 & 51 & 6 & 24 & MOTION \\
+    & RacketSports & 303 & 30 & 4 & 6 & SENSOR \\
+    & UWaveGestureLibrary & 240 & 315 & 8 &3 & SENSOR \\
     \bottomrule
   \end{tabular}
+  %
+  }
   \label{appendix:table:datasets}

### 2024-05-20 NEURIPS24/paper.tex
update
first results

@@
 
-\thi{
-  However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
+% \thi{
+%   However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
 \Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
@@
 This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
-}
+% }
 
@@
  %Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
+
+\thi{
+  Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
+The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
+Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
+The goal is to find a well-suited Riemannian structure according to the nature of the studied object.
+LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
+However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
+Closely related, \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM but distinguishes from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
+Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
+However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment.
+Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation.
+Very recently, in a functional data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution.
+Nevertheless, this method is made for \textit{continuous objects} and only applies to time series of \textit{same length}, making the estimation more sensitive to noise.
+}
 \section{Notations}

### 2024-05-21 NEURIPS24/appendix.tex
separate exp

@@
 
+\section{Robustness to missing data}
+
+\begin{table}[hbt!]
+  \centering
+  \begin{tabular}{lcccc}
+    \toprule
+    \multirow[c]{2}{*}{\textbf{Methods}} & \multicolumn{4}{c}{\textbf{Test F1-score}} \\
+    \cline{2-5}
+     &  \textbf{Regular} & \textbf{30 \% dropped} &  \textbf{50 \% dropped} & \textbf{70 \% dropped} \\
+    \midrule
+    RNN (1999) & $0.64 \pm 0.21$ & $0.53 \pm 0.23$ & $0.48 \pm 0.21$ & $0.44 \pm 0.21$ \\
+    LSTM (1997) & $0.61 \pm 0.29$ & $0.57 \pm 0.29$ & $0.53 \pm 0.25$ & $0.51 \pm 0.29$ \\
+    GRU (2014) & $0.71 \pm 0.26$ & $0.68 \pm 0.28$ & $0.66 \pm 0.28$ & $\underline{0.59 \pm 0.28}$ \\
+    MTAN (2021) & $0.59 \pm 0.28$ & $0.58 \pm 0.28$ & $0.54 \pm 0.29$ & $0.51 \pm 0.28$ \\
+    MIAM (2022) & $0.48 \pm 0.35$ & $0.42 \pm 0.33$ & $0.47 \pm 0.31$ & $0.35 \pm 0.31$ \\
+    ODE-LSTM (2020) & $0.63 \pm 0.24$ & $0.57 \pm 0.25$ & $0.51 \pm 0.24$ & $0.45 \pm 0.23$ \\
+    Neural SDE (2019) & $0.48 \pm 0.28$ & $0.47 \pm 0.26$ & $0.45 \pm 0.27$ & $0.45 \pm 0.25$ \\
+    Neural LNSDE (2024) & $0.7 \pm 0.27$ & $0.68 \pm 0.29$ & $\underline{0.67 \pm 0.25}$ & $\mathbf{0.66 \pm 0.23}$ \\
+    LDDMM (2008) & $\underline{0.72 \pm 0.2}$ & $\underline{0.7 \pm 0.21}$ & $0.57 \pm 0.25$ & $0.4 \pm 0.25$ \\
+    %\cline{1-5}
+    TS-LDDMM (ours) & $\mathbf{0.83 \pm 0.18}$ & $\mathbf{0.8 \pm 0.18}$ & $\mathbf{0.7 \pm 0.26}$ & $0.51 \pm 0.27$ \\
+    \bottomrule
+  \end{tabular}  
+\end{table}
+
 \section{Classification datasets}

### 2024-05-21 NEURIPS24/experiment.tex
separate exp

--- 2024-05-15 NEURIPS24/paper.tex
@@
-% citep
-% \section{TODO}
-% %De ce que je comprends , les gens font surtout des variational auto encoder, ou du masking
-% %Mentionner que la méthode proposé peut être utilisé avec des réseaux
-% %\sam{Approche NTK, fédéré personalisation, projet avec Axel, Lien entre continuous normalizing flow}
-% Plan intro: 
-% \begin{itemize}
-%     \item Representation of patterns in time series has several applications in biology, but halso for classification, deep learning finds features automaticly.
-%      Taking or not taking into account time can be done with DTW. However, either the representation is the result of a black box and need post analysis to be understand,
-%       which is difficult to apply statiscal tools on it, or the representation is desgned by hand to select relevant features, but the capacity of analysis is limited by the knowledge of the designer.
-%      \item In shape analysis, representation of shapes for statistical analysis in a long time issue, which has been solved mainly by the LDDMM framework.
-Representing shapes as diffeomorpshims of a common shape.
-%      \item However, this tool has not been yet applied on time series, even if there is already some tools developped for oriented curves. 
-%      \item In this work, we propose to apply LDDMM to time series and to show its interest to derive an interpretable and unsupervised representation of patterns in time series.
-%      We proposed some tools developped in the shape analysis community to the researcher interested in machine learning and time series ones.
-%      \item First, we introduce our assumptions regarding the dataset and the main problem of diffeomorpshim learning. Then, we show how to solve it with LDDMM .
-%       We expose how to apply it to time series by selecting wisely a RKHS kernel, motivating it with a representation theorem.
-%       experiments on synthetic data are presented to show that the parameter of representation are identifiable when the time seriess are not too sharp and the kernels parameter well tuned. 
-%       Then, we sh-ow how to apply the method on an unsupervised case of moouse respeitaory dataset and a supervised one on ... .
-
-% \end{itemize}
-
-% One option to tackle this issue is to derive ...
-% feature representation of time series which depends on the problem at hand.
-% which is parti... -> which is of prime interest in ...
-
-\section{Introduction}
-%https://arxiv.org/abs/1811.00075
-%Appliqué sur les "UWaveGestureLibrary","ArticularyWordRecognition","Cricket" pour l'archive UEA
-%"ArrowHead", "BME", "ECG200","FacesUCR","GunPoint" dans le cas univarié
-
-%Idée de parler d'inter individualité, spécialiser davantage
-%k shape, shapelets , ye2009time
-%\thi{à reformuler pour faire moins stupide}
-
-%\sam{Aller droit au but, se servir de la section related works pour détailler, notamment enlever le premier paragraphe dire directement que l'on veut faire du unsupervised represnetation learning for time series mais en s'intéressant par ticulièrement à la notion de shape.
-%Parler du contexte biomédical, définir la shape. Parler de shape analysis, introduire notre problème de représentation, dire que nous on propose une structure adapté aux time series.}
-
-% In this paper, we propose an unsupervised representation learning method (URL) for time series \cite{trirat2024universal,meng2023unsupervised} by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
-% In particular, the method handle time series with \textbf{irregular sampling} and \textbf{different sizes}.
-% This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
-%  we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) \cite{beg2005computing,vaillant2004statistics} to the case of time series data by applying difformations to their graph.
-% However, time series graph have more structure than 1D curves:
-%  a graph time series transformation by a general diffeomorphism is not always a graph time series.
-% This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.
-%A time series is a set of values indexed by time \cite{trirat2024universal}.
-%This type of data is prevalent in biology and medicine for recording physiological time series , as well as in the broader context of the Internet of Things sensors \cite{fathy2018large} and measurements in physical systems \cite{luo2021deep}.
- %However, due to the nature of such data, which have different sample sizes and incorporate a temporal evolution, classical statistical analysis, and machine learning methods need refinement \cite{anderson2011statistical}.
- %One option to tackle this issue is to derive a feature representation of time series, which depends on the problem at hand \cite{trirat2024universal}.
-
-Our goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
-More specifically, we aim to find an unsupervised features representation method which encode the specificity of an individual compared to another. 
-In physiology, studying the different \textit{shapes} in a time series related to biological phenomena and their variations according to individual or pathology is common.
-However, a \textit{shape} has no clear definition; it is more an intuitive way to speak about the silhouette of a pattern in a time series.
-In this paper, we refer to as the shape of a time series, the graph of this signal.
-
-Although a community structure with representatives can be learned in an unsupervised way \cite{trirat2024universal,meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised} or similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},
-studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem in URL.
-     %In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.
-  %For instance, the shape of a heartbeat in an Electrocardiogram (ECG) -the pattern- has been well studied \cite{madona2021pqrst} with the PQRST decomposition (see figure\sam{fill}).
-  %There exist several methods to detect and segment patterns in a time series (\sam{mettre ref thibuat} ).
-   %The goal is then to study their shape according to the property of the individual.
-
-  %  In most clinical studies, depending on expert knowledge, people inspect hand-crafted features such as the mean, wavelets decomposition, or temporal correlation \cite{anderson2011statistical} to keep interpretability.
-  %  However, this approach is time-consuming and potentially biased by priors, especially in exploratory research.
-  %  Given the success of neural networks \cite{bishop1994neural}, features can now be automatically extracted, depending only on the data, by choosing an objective to optimize and a neural architecture tailored to the problem \cite{ismail2019deep,blazquez2021review,torres2021deep}, even in an unsupervised way \cite{meng2023unsupervised}.
-  %  Moreover, depending on the method, it is still possible to have an interpretability of the features.
-  %   For instance, among individuals, a contrastive clustering method can be applied to spot community structures \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
-  %   More generally, interpretability \cite{yuan2019wave2vec,zhang2018visual,ismail2020benchmarking} is possible by visualizing the shapes of representative patterns to which the model is sensitive.
-  %   Other methods, based on similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k}, also propose to perform clustering on shapes in time series or to find some representatives \cite{ye2009time}.
-
-    %Although a community structure with representatives can be learned in an unsupervised way,
-    % studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem.
-     %In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.% \thi{In an individual cluster ?}
-
-% Alternatively, one can represent time series with an embedding that can be processed using conventional tools.
-% or we can represent the time series with an embedding which can be treated by classical tools.
-
-% Depending on the tasks, e.g classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep} or clustering , a meaningful vector representation of time series can be derived.
-%  Initially, in this research, handcrafted features were computed from time series, such as the mean, wavelets \cite{percival2000wavelet}, or Fourier coefficients \cite{bloomfield2004fourier}, depending on expert knowledge.
-%   However, this approach is time-consuming and potentially biased by priors.
-%   Observing the success of neural networks, features are automatically extracted by choosing an objective to optimize and a neural architecture tailored to the problem.
-%    Taking a set of time series as input, features can be identified by solving specific supervised tasks—classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep}— or in an unsupervised manner by deriving labels from the inputs \cite{meng2023unsupervised}.
-%    \thi{Rajouter une image, et une phrase par rapport au temps}
-%    Although the deep learning approach is particularly effective for supervised tasks, annotated datasets are rare, making unsupervised methods more practical.
-%    These methods have proven effective in identifying different patterns in the data and providing a vector representation that segregates well into clusters \cite{meng2023unsupervised}.
-%     However, they may lack precision on small and homogeneous datasets with only one typical shape, often arising from clinical experiments addressing a specific phenomenon (refer to figures).
-%      While current unsupervised methods separate representations into clusters, our focus is on the inter-individual variability within one cluster.
-%       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.
-
-%This type of question has already been addressed in shape analysis \cite{vaillant2004statistics}, a community focusing on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
-%However, time series data with their specific structure has not been properly adressed.
-
- First, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
- Then, building on the shape analysis litterature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
-  The idea is to represent each element $(\msg(s^j))_{j\in[N]}$ of the dataset as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism.
-  Then, the diffeomorphism is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
-   The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
-    Finally, these features encoding the shapes can feed any statistical or machine learning model as in URL.
-
-   However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
-\Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
-   Our contributions arise from this observation.
-   %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
-    We solve this issue by specifying the class of diffeomorphisms to consider and showing how to learn them.
-     This change is fruitful in representing time transformation as illustrated in \Cref{fig:transport}.
-     %In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
-
-     Our contributions can be summarized as follows:
-     \begin{itemize}
-       \item We propose an unsupervised method (TS-LDDMM) to analyze inter-individual variability of shapes in a time series dataset.
-In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
-       \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method with respect to its hyperparameters, also providing guidelines for tuning.
-       We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-        We illustrate the quantitative interest of the representation on classification tasks on real shape-based datasets.
-
-     \end{itemize}
-
-    %To the best of our knowledge, these theoretical and practical works have never been done before.
-      % On synthetic data, we show the sensibility of the results according to the kernel parameters and give guidelines to tune it.
-      % In an unsupervised framework, we vectorize respitory patterns of a mouse dataset and analyze the variations depending on the drug taken by the mouse.
-      % Finally, we compare to the state-of-the-art methods in classification an ... classifier taking as inputs the representation given by LDDMM.
-
-  %    \paragraph*{Related papers}
-  %    %https://arxiv.org/pdf/2106.00750.pdf look to the related papers
-  %  In the shape analysis community, a famous paper \cite{srivastava2010shape} addresses the problem of representation of curves $C:\msi \to \Rset^d$ by using the Square-Root Velocity (SRV) representation.
-  %      This structure is particularly relevant to compute geodesics and distances between curves when the space is quotiented by time reparametrization.
-  %       However, in our case the time axis deserves a particular attention, since the curves are derived from time-series. 
-  %       %While \cite{srivastava2010shape} tackles the computations of geodesics and distances on quotient curves space, our aim is to get a vector representation of each pattern. 
-  %       Very recently, in \cite{heo2024logistic}, this representation was used to compute PCA and to perform classification on quotient curves space, we follow the same ideas, but we focus on time series and we use LDDMM instead of SRV.
- % optimization framework with a well chosen neural architecture in order to learn automatically the features, which has been very 
-% To understand these complex data, we are looking to simpler representations toperform classical statiscal analysis or to perform efficient downstream tasks of machine learning, the time series representation res time series to a real valued vector representation.
-in order to apply classical statiscal or machine learning tools depending on the needs such as Principal Component Analysis (PCA), logistic regression or deep learning methods for classification or anomaly detection.
-
-\thi{
-  Our goal is to analyze the inter-individual variability within a time series dataset, an approach of prime interest in physiological contexts \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
-More specifically, we aim to find an unsupervised features representation method that encodes individual time series specificities compared to a reference one.
-}
-
-\thi{
-  Studying shape differences between time series related to biological mechanisms is a common practice in physiology to characterize healthy and pathological functioning CITE.
-For instance, the shapes of heartbeats in electrocardiograms are discriminant for some cardiovascular pathologies CITE.
-Several approaches have been proposed for such comparison.
-Some employ shape-based similarity measures between time series \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},  others embed time series as vectors of predefined features CITE, and, with the rise of deep neural networks, unsupervised learning representation of time series~\cite{trirat2024universal,meng2023unsupervised} has shown to be a valuable approach CITE notably with contrastive learning~\cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
-However, shape-based representation of time series within cohorts~\cite{niennattrakul2007inaccuracies,shirato2023identifying} remains an open problem in URL.
-}
-
-\thi{
-  Our work focuses explicitly on learning shape-based representation of time series.
-First, we propose not to see the shape of a time series through its curve $\{s_t:\eqsp t\in\msi\}$, but rather through its graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\}$.
-Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
-The idea is to represent each element $\msg(s^j)$ of a dataset $(s^j)_{j\in[N]}$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism $\phi_j$ (ie $\msg(s^j) \sim \phi_j .
-\msg(\mathbf{s}_0)$ ).
-The diffeomorphism $\phi_j$ is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
-The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
-Finally, these shape-encoding features can feed any statistical or machine-learning model.
-  }
-
-\thi{
-  However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
-\Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
-Our contributions arise from this observation: we specify the class of diffeomorphisms to consider and show how to learn them.
-This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
-}
-
-\thi{
-  Our contributions can be summarized as follows:
-    \begin{itemize}
-       \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset.
-In particular, the method can handle multivariate time series \textit{irregularly sampled} and with \textit{variable sizes}.
-       \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters, also providing guidelines for tuning.
-       \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets.
- \end{itemize}
-}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.7\linewidth]{"./pictures/diffeo.jpeg"}
-
-  \caption{A time series' graph $\msg=\{(t,s(t)): \eqsp t\in\msi\} $ can lose its structure after applying a general diffeomorphism $\phi.\msg$: a time value can be related to two values on the space axis.}
-  \label{fig:diffeo}
-
-\end{figure}
-
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=\linewidth]{"./pictures/transport.jpeg"}
-
-  \caption{LDDMM and TS-LDDMM are applied to ECG data.
-  We observe that LDDMM, using a general Gaussian kernel, does not learn the time translation of the first spike but changes the space values, i.e., one spike disappears before emerging at a translated position.
-At the same time, TS-LDDMM handles the time change in the shape.
-  This difference of \textit{deformations} implies differences in features \textit{representations}.   }
-  \label{fig:transport}
-
-\end{figure*}
-
-\section{Related Works}
-Shape analysis focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
- The main idea is to represent these different objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
- Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
- The goal is then to find a well suited Riemannian struvture according to the studied object.
-
- A time series graph can be seen as a curve and LDDMM structure is relevant to tackle curves as presented in \cite{glaunes2008large}. %, LDDMM structure is relevant to tackle curves and
-  However, time series graph has more structure than curves as depicted in \Cref{fig:diffeo} due to the temporal evolution. 
-  \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM, but distinguish from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
-
-  Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} address the representation of curves having unitary velocity by using the Square-Root Velocity (SRV) representation.
- However, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
-  Again, the time series graph structure is not respected.
-   Very recently in a functionnal data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by giving a representation for the original time evolution.
-    Nevertheless, this methods applies only on time series of \textit{same size} and is made for \textit{continuous objects}, making the estimation more sensitive to noise.
-  %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
-  %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
-% Depending on the tasks, e.g classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep} or clustering , a meaningful vector representation of time series can be derived.
-%  Initially, in this research, handcrafted features were computed from time series, such as the mean, wavelets \cite{percival2000wavelet}, or Fourier coefficients \cite{bloomfield2004fourier}, depending on expert knowledge.
-%   However, this approach is time-consuming and potentially biased by priors.
-%   Observing the success of neural networks, features are automatically extracted by choosing an objective to optimize and a neural architecture tailored to the problem.
-%    Taking a set of time series as input, features can be identified by solving specific supervised tasks—classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep}— or in an unsupervised manner by deriving labels from the inputs \cite{meng2023unsupervised}.
-%    \thi{Rajouter une image, et une phrase par rapport au temps}
-%    Although the deep learning approach is particularly effective for supervised tasks, annotated datasets are rare, making unsupervised methods more practical.
-%    These methods have proven effective in identifying different patterns in the data and providing a vector representation that segregates well into clusters \cite{meng2023unsupervised}.
-%     However, they may lack precision on small and homogeneous datasets with only one typical shape, often arising from clinical experiments addressing a specific phenomenon (refer to figures).
-%      While current unsupervised methods separate representations into clusters, our focus is on the inter-individual variability within one cluster.
-%       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.
-[Ajouter littérature de URL demander à Sisi ?]
- %\citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
- %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
- %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
- %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
- %Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
-\section{Notations}
-We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$,
-by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
- by $||u||_\infty=\inf_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
-and by $\Nset_{>0}$ is the set of positive integers. 
-%We are in a sub class of curve quotiented by temporal reparametrisation since a time series' graph can be understood as the equivalent class of t ->(t,s(t))
-
-%Reproducing Kernel Hilbert Space (RKHS)
-\section{Background on LDDMM}
-\label{section:LDDMM}
-
-In this part, there is no novelty, we simply expose how to
- learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
-  In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
-
-  In the next section, the time series are going to be represented by diffeomorphism parameters $(\alpha_j)_{j\in[N]}$.
-  That's why LDDMM is chosen since it offers a parametrization for diffemorphisms which is sparse and interpretable, two features particularly relevant in the biomedical context.
-
-The basic problem that we consider in this section is the following.
-Given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$\footnote{Note that we denote by $d'\in\nset$ the ambient space}, a set of starting points $\mathbf{x}=(x_{i})_{i\in[T_1]}$ in $\Rset^{d'}$, we aim to find a diffeomorphism $\phi$ such that the finite set of points $\mathbf{y}$ is similar in a certain sense to the set of finite sets of transformed points $\phi \cdot \mathbf{x} =(\phi(x_i))_{i\in[T_1]} $.
- The function $\phi$ is occasionally referred to as a \textit{deformation}. In general, these sets $\mathbf{x},\mathbf{y}$ are meshes of continous objects, e.g surfaces, curves, images and so on.
-
-% LDDMM is designed to analyzed an existing dataset, while NF and CNF are made to generalize a dataset for data augmentation.
-\paragraph{Representing diffeomorpshims as deformations.}
-Such \textit{deformations} $\phi$ are constructed via differential flow equations, for any $x_0\in \Rset^{d'} $ and $\tau\in[0,1]$:
-\begin{equation}
-  \label{eq:LDDMM_dynamic}
-    \frac{\dd X(\tau)}{\dd \tau}= v_\tau(X(\tau)), \quad X(0)=x_0\eqsp ,
-    \phi^v_\tau(x_0)=X(\tau), \quad \phi^v=\phi^v_1  \eqsp ,
-\end{equation}
-where the velocity field is $v:\tau\in [0,1]\mapsto v_\tau\in \msv $
-and $\msv$ is a Hilbert space of continuously differentiable function
-on $\Rset^{d'}$.  If
-$||\dd u ||_{\infty}+|| u ||_{\infty}\leq ||u ||_\msv $ for any
-$u\in \msv$ and
-$v\in \rml^2([0,1],\msv)=\{v\in \rmC^0([0,1],\msv): \int_0^1 ||
-v_\tau||^2_\msv \dd \tau<\infty \} $, by \citep[Theorem 5]{glaunes2005transport}
-$\phi^v$ exists and belongs to $\mcd(\Rset^{d'})$, where we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
- Therefore, for any choice of $v$, $\phi^v$ defines a valid deformation. 
-This offers a general recipe to construct diffeomorphism given a functional space $\msv$.
-
-With this in mind, the velocity field $v$ fitting the data can be
-estimated by minimizing 
-$v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
- However, two computational challenges arise.
-  First, this optimization problem is ill-posed, and a penalty term is needed to obtain a unique solution.
-   In addition, we have to find a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, which allows us to solve this minimization problem efficiently.
-
-%\paragraph*{Regularizing the problem to derive uniqueness.}
-It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$.
-Following this geometric point of view, geodesics can be constructed on $\msh$ by using the following squared norm 
- \begin{equation}
-  \label{eq:geodesics_original}
-    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv\dd \tau
- \end{equation}
-By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy lipshcitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$ \cite{miller2006geodesic}, as straight lines in Euclidean space.
-Denoting by $w(v_0)$ the geodesic starting from the identity with inital velocity $v_0$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \phi^v$ and the previous matching problem becomes a \textit{geodesic shooting problem}:
- \begin{equation}
-  \label{eq:geodesics_shooting}
-  \inf_{v_0 \in \msv} \mathscr{L}(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}).
- \end{equation}
- Using $\varphi^{\{v_0\}}$ instead of $\phi^v$ for any $v\in \rml^2([0,1],\msv)$ regularizes the problem and induces a sparse representation for the learning diffeomorphisms.
- Moreover, by setting $\msv$ as an RKHS, the geodesic shooting problem has a unique solution and becomes tractable, as described in the next section.
-
-% More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
-%  the following squared norm can be defined
-%  \begin{equation}
-%   \label{eq:geodesics_original}
-%     \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
-%  \end{equation}
-%      as the minimal "energy" needed to perform the deformation $g$.
-% By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
-% However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
-% To circumvent this issue, another characterization of geodesics can be considered.
-%  As in Riemannian geometry, 
-%  instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
-%  \sam{To CHANGE}
-%  %In other words, an Exponential map on deformations is wanted:
-% given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
-% %  \begin{equation}
-% %   \label{eq:geodesics_shooting}
-% %   v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
-% %  \end{equation}
-% % since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
-%  By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
-% Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.
-%  Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
-%  Moreover, we have $\mathscr{R}(\phi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in L^2([0,1],\msv)$.
-%  %Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
-%  This is presented with more details in the next paragraph.
-
-\paragraph{Discrete parametrization of diffeomorpshim.}
-%\alain{adapt this without mentioning applicaiton to time seriess}
-%\sam{Est ce qu'on devrait pas introduire la dicretization du graphe ici et dire que l'on peut utiliser le théoreme de représentation}
-In this part, $\msv$ is chosen as an RKHS \cite{berlinet2011reproducing} generated by a smooth kernel $K$ (e.g., Gaussian).
-% $K(x,y)\triangleq \exp(-|x-y|^2/\sigma^2)\operatorname{Id}_{d'}$ for any $(x,y)\in(\Rset^{d'})^2$, where $\sigma^2>0$ is the kernel width and $\operatorname{Id}_{d'}$ is the identity matrix.
-We follow \cite{durrleman2013sparse} and define a 
- discrete parameterization of the velocity fields to perform geodesics shooting \eqref{eq:geodesics_shooting}.  
-  The initial velocity field $v_0$ is chosen as a finite linear combination of the RKHS basis vector fields, 
-$\mathbf{n}_0$ control points $\msx_0=(x_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0}$ and momentum vectors $\alpha_0=(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d'})^{\mathbf{n}_0} $ are defined such that for any $x\in \Rset^{d'}$,
-  \begin{equation}
-    \label{eq:def_v0}
-    v_0\left(\alpha_0,\msx_0\right)(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_{k,0})\alpha_{k,0} \eqsp .
-  \end{equation}
-   In our applications, the control points $(x_{k,0})_{k\in[\mathbf{n}_0]}$ can be understood as the discretized graph $(t_k,\mathbf{s}_0(t_k))_{k\in[\mathbf{n}_0]}$ of a starting time series $\mathbf{s}_0$.
-  With this parametrization of $v_0$, \citet{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same
-  structure along time, such that for any $x\in\Rset^{d'}$ and $\tau\in[0,1]$, 
-  \begin{equation}
-    \label{eq:specific_form}
-    v_\tau(x)=\sum_{k=1}^{\mathbf{n}_0} K(x,x_k(\tau))\alpha_{k}(\tau) \eqsp ,
-  \end{equation}
-  \begin{equation} 
-    \label{eq:integration}
-      \left\{
-        \begin{aligned}
-        & \frac{\dd x_k(\tau)}{\dd \tau}=v_\tau(x_k(\tau)) \eqsp, \quad
-        \frac{\dd \alpha_k(\tau)}{\dd \tau}=-\sum_{k=1}^{\mathbf{n}_0} \dd_{x_k(\tau)} K(x_k(\tau),x_l(\tau))\alpha_{l}(\tau)^\top \alpha_{k}(\tau) \eqsp  \\
-        & \alpha_k(0)=\alpha_{k,0},\quad x_k(0)=x_{k,0} \eqsp , k\in[\mathbf{n}_0] 
-        \end{aligned}
-        \right .
-  \end{equation}
-  These equations are derived from the hamiltonian $H:(\alpha_k,x_k)_{k\in [\mathbf{n}_0]}\mapsto \sum_{k,l=1}^{\mathbf{n}_0} \alpha_{k}^\top K(x_k,x_l)\alpha_{l}  $, such that
-  the velocity norm is preserved $||v_\tau||_{\msv}=||v_0||_\msv $ for any $\tau\in [0,1]$.
-   By \eqref{eq:integration}, the velocity field related to a geodesic $v^*$ is fully parametrized by its initial control points and momentum $(x_{k,0},\alpha_{k,0})_{k\in[\mathbf{n}_0]}$.
-   Thus, given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$, a set of starting points $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$ in $\Rset^{d'}$, a RKHS's kernel $K:\Rset^{d'}\times \Rset^{d'}\to \Rset^{d'\times d'}$, a distance on sets $\mathscr{L}$,
- a numerical integration scheme of ODE and a penalty factor $\lambda>0$, the basic geodesic shooting step minimizes the following function using a gradient descent method:
-   \begin{equation}
-    \label{eq:relaxation}
-    \mathcal{F}_{\mathbf{x},\mathbf{y}}: (\alpha_k)_{k\in [T_1]}\mapsto \mathscr{L}\left(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2 \eqsp,  
-   \end{equation}
-   where $v_0$ is defined by \eqref{eq:def_v0} and $\varphi^{\{v_0\}}.\mathbf{x}$ is the result of the numerical integration of \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_{k\in[T_1]} $.
-
-   %Before to specify the choice of the RKHS kernel $K$ and the loss on sets $\mathscr{L}$ in \Cref{section:time_series_specificity}, we first discuss the connection of LDDMM with normalizing flows.
-% to highlight the interest of the LDDMM approach.
-
-%Conversely to NF or CNF, by \eqref{eq:integration}, different momentums parameters raise different final deformations $\phi^{\{v_0\}}$, this injectivity properties offers interpretability of the representation and gives sens to PCA or LDA post-processing as exposed in \Cref{section:experiments}.
-
-   %If a diffeomorpshim $\Phi$ is given and we want to represent it as a deformation $\phi^v$ with $v$ of minimal ernergy, then
-  %$(\alpha_{k,0})_k$ is the only parameters to learn since $(x_{k,0})_k$ is fixed.
-
-%   \begin{algorithm}[tb]
-%     \caption{Geodesic shooting with LDDMM}
-%     \label{alg:example}
-%  \begin{algorithmic}
-%     \STATE {\bfseries Input:} a RKHS's kernel $K$, a loss on sets $L$, $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$, $\mathbf{y}=(y_i)_{i\in[T_2]}$ two sets of $\Rset^{d'}$, initial momentums 
-%     $(\alpha_{k,0})_{k\in [T_1]} \in (\Rset^{d'})^{T_1}$ and a number of gradient iterations $n_{\text{iter}}$.
-
-%     \STATE $f: (\alpha_k)_k\mapsto L\left(\phi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2  $ with $v_0$ and $\phi^{\{v_0\}}$
-%      respectively defined by \eqref{eq:def_v0} and \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_k $. 
-%      \sam{Peut être à définir dans un autre algo ?}
-%     \STATE $\alpha_{k,0}^*\gets \alpha_{k,0}, \eqsp k\in[T_1]$
-%     \FOR{$i=1$ {\bfseries to} $n_{\text{iter}}$}
-%     \STATE $\alpha_{k,0}^*\gets \text{AdamStep}(f,\alpha_{k,0}^*$) , where $\text{AdamStep}$ is given in appendix (mettre ref)
-%     \ENDFOR
-%     \STATE {\bfseries Return:} $(\alpha_{k,0}^*)_{k\in [T_1]}$
-
-%  \end{algorithmic}
-%  \end{algorithm}
-\paragraph{Relation to Continuous Normalizing Flows.}
-
-One particular popular choice to address the problem of learning a diffeomorphism or a velocity field is Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} (NF) or their continuous counterpart \cite{chen2018neural,grathwohl2019scalable,salman2018deep} (CNF).
-However, we do not rely on this class of learning algorithms for several reasons.
-Indeed, existing and simple normalizing flows are not suitable for the type of data that we are interested in this paper \cite{feng2023multi,deng2020modeling}.
-  In addition,  they are primarily designed to have tractable Jacobian functions, while we do not require such property in our applications. 
-Finally, the use of a differential flow solution of an ODE
-\eqref{eq:LDDMM_dynamic} trick is also at the basis of CNF, which
-then consists of learning a velocity field to address in fitting the data through a loss aiming to address the problem
-at hand. Nevertheless, the main difference between CNF and LDDMM lies in the
-parametrization of the velocity field. LDDMM uses kernels to
-derive closed form formula and enhance interpretability while
-NF and CNF take advantage of deep neural networks to scale with
-large dataset in high dimensions.
-\section{Methodology}
-%\paragraph{Assumptions}
-%Denoting by $C^0(\msi,\msj) $ the space of continous functions between two sets $\msi,\msj$,
-
-We consider in this paper observations which consist in a population of $N$ multivariate time series, for any $j\in[N]$, $s^j \in \rmC^1(\msi_j,\Rset^{d})$. 
-However, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
- Note that \textbf{the number of samples $n_j$ is not necessary the same accross individuals} and the timestamps can be \textbf{irregularly sampled}.
-%To ease the presentation of our methodology, we suppose that we have access to the continuous time series $(s^j)_{j\in[N]}$, while in practice, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
- %However, this difference does not change the overall rationale of our method and its implementation up to minor changes, which we summarize in \eqref{eq:general_optimization_problem}.
- We assume the time series population is globally homogeneous regarding their "shapes" even if inter-individual variability exists.
- Intuitively speaking, the "shape" of a time series $s:\msi\to \Rset^d$ is encoded in its graphs $\msg(s)$ defined as the set $\{(t,s(t)):\eqsp t\in\msi \} $ and not only in its values $s(\msi)=\{s(t):\eqsp t\in\msi \} $ since the time axis is crucial.
-% on the shape as illustrated in Figure\alain{ (mettre ref)}.
- As a motivating use-case, $s^j$ can be the time series of a heartbeat extracted from an individual's electrocardiogram (ECG), see \Cref{fig:transport}.
- The homogeneity in a resulting dataset comes from the fact that humans have similar shapes of heartbeat \cite{ye2012heartbeat,madona2021pqrst}.
- % the shape is also referred as morphology of the time series. 
-  %  [More specifically, the heterogeneity in such dataset is often due to local or glocal dilatation of the time series regarding time, the heartbeat is quicker or slower depending on the individual.
-  %  In maths, denoting by $\mcd(\msi,\msj) $ the set of diffeomorpshim defined on the sets $\msi$ to $\msj$,
-  %   there exist a common pattern $\mathbf{s}_0: \msi \to \Rset^d$ and individual time reparametrisations $ \psi_j\in\mcd(\msi,\msi_j) $ such that,
-  %   for any $j\in[N]$, $s^j=\mathbf{s}_0\circ \psi_j^{-1} $.]
-%  \paragraph{Goals}%\sam{potentiellement changer l'ordre de time and space}
-\paragraph*{The deformation problem.}
-In this paper, we aim to study the inter-individual variability in the dataset by finding a relevant representation of each time series.
-Inspired from the framework of shape analysis \cite{vaillant2004statistics}, addressing similar problems in morphology,
- we suggest to represent each time series' graph $\msg(s^j)$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$, related to a time series $\mathbf{s}_0:\msi \to\Rset^d$, by a diffeomorphism $\phi_j$ on $\Rset^{d+1}$, for any $j\in[N]$,
-\begin{equation}
- \label{eq:transformation}
- \phi_j.\msg(\mathbf{s}_0)=\{\phi_j\left(t,\mathbf{s}_0(t)\right), \eqsp t\in \msi \} \eqsp.
-\end{equation}
-$\bfs_0$ will be understood as the typical representative shape common to the collection of time series $(s^j)_{j\in[N]}$.
-As $\bfs_0$ is supposed to be fixed, then the representation of the time series $(s^j)_{j\in[N]}$ boils down to the one of the transformation $(\phi_j)_{j\in[N]}$.
-We aim to learn $\msg(\bfs_0)$ and $(\phi_j)_{j\in[N]} $.
-
-%First, we introduce the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework.
-% Then, we explain how to learn a discretization of the graph $\msg(\bfs_0)$ and the diffeomorphisms $(\phi_j)_{j\in[N]} $ by using LDDMM and a gradient descent minimization.
-  %Finally, we tackle the specificity of graph time series by deriving a representation theorem on diffeomorphisms, which enables us to select the kernel needed in LDDMM, and thus, we propose TS-LDDMM.
-
-     %\sam{Parler de l'unicté?}
-     \paragraph{Optimization related to \eqref{eq:transformation}.}
-     Defining the \textit{discretized graphs} of the time series $(s^j)_{j\in[N]}$ and a discretization of the reference graph $\msg(\mathbf{s}_0)$  as, for any $j\in[N]$,
-     \begin{equation}
-      \label{eq:descretized_graph}
-      \mathbf{y}_j=\msg(\tsig^j)=(t_i^j,\tsig^j_i)_{ i\in[n_j]}\in (\Rset^{d+1})^{n_j},\quad \tilde{\msg}_0=(t_i^0,\tsig^0_i)_{i\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0} \eqsp ,
-     \end{equation}
-       with $\mathbf{n}_0=\operatorname{median}((n_j)_{j\in[N]})$, the representation problem given in \eqref{eq:transformation} boils down solving:
-      \begin{equation}
-        \label{eq:general_optimization_problem}
-        \argmin_{\tilde{\msg}_0,(\alpha_k^j)_{k\in [\mathbf{n}_0]}^{j\in[N]}} \sum_{j=1}^N \mathcal{F}_{\tilde{\msg}_0,\mathbf{y}_j}\left((\alpha_k^j)_{k\in [\mathbf{n}_0]}\right)\eqsp ,
-      \end{equation}
-      which is carried out by a gradient descent on the control points $\tilde{\msg}_0$ and the momentums $\mathbf{\alpha}_j=(\alpha_k^j)_{k\in [\mathbf{n}_0]}$ for any $j\in[N]$, initialized by a dataset's time series graph of size $\mathbf{n}_0$ and by $0_{(d+1)\mathbf{n}_0}$ respectively.
-      The optimization hyperparameter details are given in \Cref{appendix:optimizers_details}.
-     The result of the minimization $\tilde{\msg}_0$ is then considered as the $\mathbf{n}_0$-samples of a common time series $\mathbf{s}_0$ and the momentums $\mathbf{\alpha}_j$ encoding $\phi_j$ yields a feature vector in $\Rset^{d \mathbf{n}_0} $ of $s^j$ for any $j\in[N]$.
-     %, which will be the representation of the time series when choosing the time series' graph sets as inputs, is dependent on the choice of kernel $K$ and the set distance $L$.
-     Finally, the vectors $(\mathbf{\alpha}_j)_{j\in[N]}$ can be analyzed with any statistical or machine learning tools such as Principal Components Analysis (PCA), Latent Discriminant Analysis (LDA), longitudinal data analysis and so on.
-%As $\bfs_0$ is supposed to be fixed, then the representation of the time series $(s_j)_{j\in[N]}$ boils down to the one of the transformation $(\phi_j)_{j\in[N]}$.
- %In \Cref{section:optimization} the choice of the common time series $\mathbf{s}_0$ will be optimized to have $\phi_j$ of minimal norm. %T is given in the following part.
-
-%  To learn the diffeomorphisms $(\phi_j)_{j\in[N]}$, we propose to use in this paper Large Deformation Diffeomorphic Metric Mapping (LDDMM).
-%   In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
-% We now describe more precisely LDDMM and in a next step the procedure to learn $\mathbf{s}_0$.
-
-% \sam{to improve ?}
-% Compared to the litterature of Unsupervised Representation Learning (URL) \cite{meng2023unsupervised}, the homogeneity assumption can be quite restrictive, however this is quite common in Shape Analysis \cite{vaillant2004statistics} (SA).
-%     While in URL the goal is to find features accounting for the clustering structure in the heterogene data and to transfer this representation for different ML tasks, in SA the goal is to analyze the inter-individual variability with precision to derive clinical conclusions.
-%     The generalization of the method presented in this paper to heterogene dataset is possible, but relagated for future works.
-  %These velocity fields are then learned by minimizing the norm of this RKHS.
-
-%In all the following, we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
-%\sam{Transformer , en : dans les ensembles}
-
-   %\sam{Parler de l'unicté?}
-
-   %Before to specify the choice of the RKHS kernel $K$ and the loss on sets $\mathscr{L}$ in \Cref{section:time_series_specificity}, we first discuss the connection of LDDMM with normalizing flows.
-% to highlight the interest of the LDDMM approach.
-
-%Conversely to NF or CNF, by \eqref{eq:integration}, different momentums parameters raise different final deformations $\phi^{\{v_0\}}$, this injectivity properties offers interpretability of the representation and gives sens to PCA or LDA post-processing as exposed in \Cref{section:experiments}.
-
-   %If a diffeomorpshim $\Phi$ is given and we want to represent it as a deformation $\phi^v$ with $v$ of minimal ernergy, then
-  %$(\alpha_{k,0})_k$ is the only parameters to learn since $(x_{k,0})_k$ is fixed.
-
-%   \begin{algorithm}[tb]
-%     \caption{Geodesic shooting with LDDMM}
-%     \label{alg:example}
-%  \begin{algorithmic}
-%     \STATE {\bfseries Input:} a RKHS's kernel $K$, a loss on sets $L$, $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$, $\mathbf{y}=(y_i)_{i\in[T_2]}$ two sets of $\Rset^{d'}$, initial momentums 
-%     $(\alpha_{k,0})_{k\in [T_1]} \in (\Rset^{d'})^{T_1}$ and a number of gradient iterations $n_{\text{iter}}$.
-
-%     \STATE $f: (\alpha_k)_k\mapsto L\left(\phi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2  $ with $v_0$ and $\phi^{\{v_0\}}$
-%      respectively defined by \eqref{eq:def_v0} and \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_k $. 
-%      \sam{Peut être à définir dans un autre algo ?}
-%     \STATE $\alpha_{k,0}^*\gets \alpha_{k,0}, \eqsp k\in[T_1]$
-%     \FOR{$i=1$ {\bfseries to} $n_{\text{iter}}$}
-%     \STATE $\alpha_{k,0}^*\gets \text{AdamStep}(f,\alpha_{k,0}^*$) , where $\text{AdamStep}$ is given in appendix (mettre ref)
-%     \ENDFOR
-%     \STATE {\bfseries Return:} $(\alpha_{k,0}^*)_{k\in [T_1]}$
-
-%  \end{algorithmic}
-%  \end{algorithm}
-
-Nevertheless, \eqref{eq:general_optimization_problem} ask to define a kernel and a loss in order to perform geodesic shooting \ref{eq:relaxation}, which is the purpose of the next subsection.
-   \subsection{Application of LDDMM to time series analysis: TS-LDDMM}
-%\tcr{Mettre l'applicaiton clair pour les times series}
-
-        \label{section:time_series_specificity}
-        In this section, we present our theoretical contribution: we tailor the LDDMM framework to handle time series data.
-          The reason is that applying a general diffeomorphism $\phi$ from $\Rset^{d+1}$ to a time series' graph $\msg(s)$ can result in a set $\phi.\msg(s)$ that does not correspond to the graph of any time series, as illustrated in the \Cref{fig:diffeo}.
-           Thus, Time series graph have more structure than a simple 1D curve \cite{glaunes2008large} and deserve their special analysis which will prove fruitful as demonstrated in \ref{section:experiments}.% thus this case is not an application of \cite{glaunes2004diffeomorphic}
-
-To address this challenge, we need to identify an RKHS kernel $K:\Rset^{d+1}\times \Rset^{d+1}\to \Rset^{(d+1)^2}$ that generates deformations preserving the structure of the time series graph.
-This goal motivates us to clarify, in \Cref{theorem:representation}, the specific representation of diffeomorphisms we require before presenting a class of kernels that produce deformations with this representation.
-
-Similarly, selecting a loss function on sets $\mathscr{L}$ that considers the temporal evolution in a time series' graph is crucial for meaningful comparisons with time series data.
-Consequently, we introduce the oriented Varifold distance.
-
-       \paragraph{A representation separating space and time.}
-       We prove that two time series graphs can always be linked by a time transformation composed of a space transformation.
-Moreover, a time series graph transformed by this kind of transformation is always a time series graph.
-        We define $\Psi_\gamma\in \mcd(\Rset^{d+1}) : (t,x)\in\Rset^{d+1}\to (\gamma(t),x)$ for any $\gamma\in \mcd(\Rset)$ and $\Phi_f:  (t,x)\in\Rset^{d+1}\to (t,f(t,x)) $ for any $f\in \rmC^1(\Rset^{d+1},\Rset^d)$.
-        We have the following representation theorem.
-        All proofs are given in \Cref{appendix:proofs}.
-
-  Denote by $\msg(s)\triangleq \{ (t,s(t)): \eqsp t\in \msi \} $ the graph of a time series $s: \msi \to \Rset^d$ and $ \phi.\msg(s)\triangleq\{ \phi(t,s(t)): \eqsp t\in \msi\} $ the action of  $\phi\in \mcd(\Rset^{d+1}) $ on $\msg(s)$.
-    \begin{theorem}
-        \label{theorem:representation}
-    Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi,\msj$ two intervals of $\Rset$.
-     There exist $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\gamma\in  \mcd(\Rset) $ such that $\gamma(\msi)=\msj $ and $\Phi_f\in \mcd(\Rset^{d+1})$,
-     \begin{equation}
-        \msg(s)= \Pi_{\gamma,f}.\msg(\mathbf{s}_0),\eqsp \Pi_{\gamma,f}=\Psi_\gamma\circ\Phi_f.
-     \end{equation}
-     Moreover, for any $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\bar{\gamma}\in  \mcd(\Rset) $, there exists a continously differentiable time series $\bar{s}$ such that 
-     $\msg(\bar{s})= \Pi_{\bar{\gamma},\bar{f}}.\msg(\mathbf{s}_0)$
-    \end{theorem}
-    %\sam{preuve todo, classement des variétés à bord, donne un homeomorphisme sur la restriction}
-    % \begin{proof}
-    %   Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi=(a,b),\msj=(\alpha,\beta)$ two intervals of $\Rset$.
-    %   % $\msg(\mathbf{s}_0)$ and $\msg(s)$ are two smooth connected 1-dimensional manifold of $\Rset^{d+1}$, thus, by \cite[Appendix
-    %   % Classfying 1-Manifolds]{milnor1965topology}, there exists $\Pi\in \mcd(\Rset^{d+1})$ such that $\msg(s)=\Pi.\msg(\mathbf{s}_0)$.
-    %   %  Therefore, denoting by $\Pi_1: \Rset^{d+1}\mapsto \Rset$, we define $\gamma(t)\triangleq \Pi_1(t,\mathbf{s}_0(t))$ for any $t\in \msi$, then for any $t\in [b,+\infty)$,
-    %   %   \begin{equation}
-    %   %     \gamma(t)=\\Pi_1(b,\mathbf{s}_0(b))+(t-b)\frac{\dd\gamma}{\dd^- t}(b) \eqsp,
-    %   %   \end{equation}
-    %   %   and for any $t\in (-\infty,a]$,
-    %   %   \begin{equation}
-    %   %      \gamma(t)=\Pi_1(a,\mathbf{s}_0(a))+(t-a)\frac{\dd\gamma}{\dd^+ t}(a)\eqsp ,
-    %   %   \end{equation}
-    %   %    where $ \frac{\dd\gamma}{\dd^- t},\frac{\dd\gamma}{\dd^+ t}$ are respectively the left and right derivative.
-
-    %   By setting $\gamma: t\in \Rset \mapsto (\beta-\alpha)(t-a)/(b-a)+\alpha\in \Rset $, we have $ \gamma(\msi)=\msj$ and $\gamma \in \mcd(\Rset) $.
-    %    By defining $f:(t,x)\in\Rset^{d+1}\mapsto x-\mathbf{s}_0(t)+s\circ \gamma(t) $, the map $\Phi_f\in \mcd(\Rset^{d+1})$,
-    %     indeed, its inverse is $\Phi_f^{-1}:(t,x)\in\Rset^{d+1}\mapsto (t,x+\mathbf{s}_0(t)-s(t)) $ and is continuously differentiable.
-    %      Moreover, we have $\Pi_{\gamma,f}.\msg(\mathbf{s}_0)=\{(\gamma(t),s\circ \gamma(t)):\eqsp t\in\msi \}=\msg(s) $.
-
-    %     Let $\bar{f}\in \rmC^0(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^0(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
-    %     We have :
-    %     \begin{align}
-    %       \Pi_{\gamma,f}.\msg(\mathbf{s}_0)&=\{(\gamma(t),f(t,\mathbf{s}_0(t))),\eqsp t\in \msi \} \\
-    %       &\label{eq:proof1_last_eq}=\{(t,f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right),\eqsp t\in \gamma(\msi) \} \eqsp .
-    %     \end{align}
-    %     By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^0(\gamma(\msi), \Rset^d) $ by composition of continuous functions
-    %     and $ \msg(\bar{s})= \Pi_{\gamma,f}.\msg(\mathbf{s}_0)$ by \eqref{eq:proof1_last_eq}, which concludes the proof.
-    % \end{proof}
-    \begin{remark}
-      that for any $\gamma \in \mcd(\Rset) $ and $s\in \rmC^0(\msi,\Rset^d)$,
-   \begin{equation}
-    \{(\gamma(t),s(t)),\eqsp t\in \msi \}=\{(t,s\circ \gamma^{-1}(t)):\eqsp t\in\gamma(\msi) \}\eqsp .
-   \end{equation}
-   As a result, $\Psi_\gamma $ can be understood as a temporal reparametrization and $\Phi_f$ encodes the transformation about the space.
- \end{remark}
-
-    \paragraph{Choice for the kernel associated with the RKHS $\msv$}
-    \label{paragraph:kernel_V}
-    As depicted on \Cref{fig:diffeo}-\ref{fig:transport}, we can not use any kernel $K$ to apply the previous methodology to learn deformations on time series' graphs.
-    We describe and motivate our choice in this paragraph.
-     Denote the one-dimensional Gaussian kernel by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
-    To solve the geodesic shooting problem \eqref{eq:relaxation} on $\Rset^{d+1}$, we consider for $\msv$ the RKHS associated with the kernel defined for any $(t,x),(t',x')\in (\Rset^{d+1})^2$:
-    \begin{align}
-      \label{eq:kernel_TAS}
-      &K_{\msg}((t,x),(t',x'))=\begin{pmatrix}
-        c_0K_{\text{time}} & 0 \\
-        0 & c_1 K_{\text{space}} 
-        \end{pmatrix} \eqsp ,\\
-       & K_{\text{space}}=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K_{\text{time}}=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
-    \end{align}
-    parametrized by the widths $\sigma_{T,0},\sigma_{T,1},\sigma_x>0$ and the constants $c_0,c_1>0$.
-This choice for $K_\msg$ is motivated by the representation \Cref{theorem:representation} and the following result. 
-    \begin{lemma}
-      \label{lemma:choice_of_kernel_V}
-      If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0},
-       there exist $\gamma \in \msd(\Rset) $ and $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ such that $\phi^v=\Psi_\gamma\circ\Phi_f $.
-    \end{lemma}
-    \sam{Parler des Cauchy kernel en appndice et du choix de la loss}
-    % \begin{proof}
-    %   Let $v$ be a vector field generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0}.
-    %  We remark that the first coordinate of the velocity field $v_\tau$ denoted by $v_\tau^{\text{time}}$ only depends on the time variable $t$ for any $\tau\in[0,1]$.
-    %  Thus, when computing the first coordinate of the deformation $\phi^v$, denoted by $\gamma$, we integrate \eqref{eq:LDDMM_dynamic} with $v_\tau$ replaced by $v_\tau^{\text{time}}$,
-    %   thus $\gamma$ is independant of the variable $x$.
-Moreover, $\gamma\in \mcd(\Rset)$ since a Gaussian kernel induced an Hilbert space $\msv$ satisfying $|f|_V\leq |f|_\infty+ |\dd f|_\infty  $ for any $f\in \msv$ by \citep[Theorem 9]{glaunes2005transport}.
-    %   For the same reason, we have $\phi^v\in \mcd(\Rset^{d+1})$, and thus its last coordinates denoted by $f$ belongs to $\rmC^1(\Rset^{d+1},\Rset^d)$, and by construction $\phi^v=\Psi_\gamma\circ\Phi_f $.
-    % \end{proof
-    \begin{remark}
-      With this choice of kernel, the features associated to the time transformation can be extracted from the momentums $(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}$ in \eqref{eq:def_v0} by taking the coordinates related to time.
-      However, the features related to the space transformation are not only in the space coordinates since the related kernel $K_{\text{space}}$ depends on time as well.
-     \end{remark}
-    In \Cref{appendix:kernel_TS_LDDMM}, we give guidelines for selecting the hyperparameters $(\sigma_{T,0},\sigma_{T,1},\sigma_x,c_0,c_1)$.
-
-    %  and then we can perform a statistical analysis on the representations $(\psi_j,\phi^j)_{j\in[N]}$,
-    % such as applying PCA decomposition \cite{vaillant2004statistics}, carring out longitudinal analysis \cite{guigui2022parallel} and so on (\sam{mettre ref}).
-    % % Cette hypothèse intéressante car ça permet de gérer des signaux de taille variables, tout les signaux se repréentent comme ça 
-    % % Espace homogène, injectivité de t ,s(t), permet de garder la structure temporel en compte
-    % \sam{Can we show unicity by assuming that $\phi,\psi$ minimize a norm ?}
- % KEOPS tp pour lddmm
-
-%and their empiric samples version $\tilde{G}_j=(\{(t_i,s^j_i), \eqsp i\in [n_i]\})_j $
-
-%  We translate the notation from the continous to discrete by adding a tilde :
-%   $\tilde{G}^j=\triangleq \{ (t_i,\tilde{s}_i^j),\eqsp i\in[n_i] \}$.
-
-% \paragraph{A problem of minimization}
-% In the next section, we will present that diffeomorpshim on space an time can be parametrized as $\phi^{v^S},\phi^{v^T}$
-%  with parameters $(v^S,v^T) \in \Theta_S\times \Theta_T$. Then, pattern will be represented by the parameters $v^S,v^T$ solutions of the following minimization :
-% \begin{align}
-%   \label{eq:minimization}
-%   &\argmin_{v^S \in \Theta_S,\eqsp v^T\in \Theta_T } \mathbf{D}(v^S,v^T)+\lambda \mathbf{N}(v^S,v^T)\eqsp , \\
-%   & \mathbf{D}(v^S,v^T)=L(\phi^{v^S}.G^\msj(\mathbf{s_0}\circ (\psi^{v^T})^{-1}),G^\msj(s)) \eqsp , 
-% \end{align} 
-% where $L$ is a loss on graphs and $\mathbf{N}$ is a term of regularization.
-
-% $s_i: t\in[0,T_0]\mapsto s(t)\in \Rset^d $ is only known through its samples $(S_i^j(t_i)=s_i)_{i\in[N]} $ at timestamps $(t_i)_{i\in[T]}$.
-
-%\sam{améliorer transition}
-\paragraph{Loss}
-
-This section specifies the distance function $\scrl$ introduced in the loss function defined in \eqref{eq:relaxation}.
-
-In practice, we can only access discretized graphs of time series, $(t_i^j,\tsig^j_i)_{i\in[n_j]}$ for any $j\in[N]$, that are potentially of 
-different sizes $n_j$ and sampled at different timestamps $(t_i^j)_{i\in[n_j]}$ for any $j\in[N]$.
- Usual metrics, such as the Euclidean distance, are not appealing as they 
-make the underlying assumptions of equal size sets and the existence of a pairing between points.
- Distances between measures on 
-sets (taking the empirical distribution), such as Maximum Mean Discaprency (MMD) \cite{dziugaite2015training,borgwardt2006integrating}, alleviate those issues; however, MMD only accounts for positional information
-and lacks information about the time evolution between sampled points.
- A classical data fidelity metric from shape analysis 
-corresponding to the distance between \textit{oriented varifolds} associated with curves alleviates this last issue \cite{kaltenmark2017general}.  
-Intuitively, an oriented varifold is a measure that accounts for positional and tangential information about the underlying 
-curves at sample points. More details and information about \textit{oriented varifolds} can be found in \Cref{appendix:varifold}.
-
-More precisely, given two sets $\msg_0=(g_i^0)_{i\in[T_0]},\msg_1=(g_i^1)_{i\in[T_1]}\in (\Rset^{d+1})^{T_1}$ and a kernel\footnote{$\mathbb{S}^d=\{x\in\Rset^{d+1}:\eqsp |x|=1\}$} $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$
-verifying \citep[Proposition 2 \& 4]{kaltenmark2017general}, for any $\xi\in\{0,1\}$ and $i\in[T_\xi-1]$, denoting the center and length of the $i^{th}$ segment $[g_i^\xi,g_{i+1}^\xi]$ by
-$c_i^\xi = (g_i^\xi + g_{i+1}^\xi)/2$, $l_i^\xi = \| g_{i+1}^\xi-g_{i}^\xi\|$, and 
-$\overrightarrow{v_i}^\xi = (g_{i+1}^\xi-g_{i}^\xi)/l_i^\xi$, the varifold distance between $\msg_0$ and $\msg_1$  is defined as,
-\begin{align}
-  &d_{\msw^*}^2(\msg_0,\msg_1) = \sum_{i,j = 1}^{T_0-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^0_j,\overrightarrow{v_j}^0))l^0_j
-  - 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l^0_i k((c^0_i,\overrightarrow{v_i}^0),(c^1_j,\overrightarrow{v_j}^1))l^1_j \\
-  &+ \sum_{i,j = 1}^{T_1-1}l^1_i k((c^1_i,\overrightarrow{v_i}^1),(c^1_j,\overrightarrow{v_j}^1))l^1_j 
-\end{align}
-
-In practice, we set the kernel $k$ as the product of two anisotropic Gaussian kernels, $k_{\pos}$ and $k_{\dir}$, 
-such that for any $(x,\overrightarrow{u}),(y,\overrightarrow{v}) \in (\Rset^{d+1} \times \mathbb{S}^d)^2$
-\begin{equation}
-    k((x,\overrightarrow{u}),(y,\overrightarrow{v})) = k_{\pos}(x,y)k_{\dir}(\overrightarrow{u},\overrightarrow{v}) \eqsp.
-  \end{equation}
-  The specific kernels $k_{\pos},k_{\dir}$ that we use in our experiments are given \Cref{appendix:kernel_implementation}.
-   Note that the loss kernel $k$ has nothing to do with the velocity field kernel denoted by $K_\msg$ or $K$ specified in \Cref{paragraph:kernel_V}.
-Finally, we define the data fidelity loss function, $\scrl$, as $ d_{\msw^*}^2$, which is differentiable with regards to its first variable.
-For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}.
-
-%\sam{TO improve by Thibaut
-%\thi{parler du fait qu'on a pas pairing, grace au difféo on conserve la structure}
-%To relax the geodesic problem \Cref{eq:geodesics_original}, the end point of the geodesics is compared to the target through a loss $\mathscr{L}$ as proposed in \eqref{eq:relaxation}.
-%Since we are dealing with discrete time series $\tsig^j$ of different samples size $n_j$, we need a loss on set measures and not only on vectors.
-% A natural choice would be the Maximum Mean Discrepency (MMD) \cite{dziugaite2015training,borgwardt2006integrating}, for any $X=(x_i)_{i\in[T_1]},Y=(y_i)_{i\in[T_2]}\subset Rset^{d'}$, given a kernel $k:\Rset^{d'}\times \Rset^{d'} \to \Rset$,
-% \begin{align}
-%  &\operatorname{MMD}^k(\mu_X,\mu_Y)=\frac{1}{T_1^2}\sum_{i,j=1}^{T_1} k(x_i,x_j)+\frac{1}{T_2^2}\sum_{i,j=1}^{T_2} k(y_i,y_j) \\
-%  &-\frac{2}{T_1T_2}\sum_{i=1}^{T_1}\sum_{j=1}^{T_2} k(x_i,y_j) \eqsp , 
-% \end{align}
-% where $\mu_X=\sum_{i=1}^{T_1} \delta_{x_i}/T_1$, $\mu_Y=\sum_{j=1}^{T_2} \delta_{y_j}/T_2$.
-%However, this distance is agnostic of the time evolutions in time series' graph $(t_i^j,\tsig^j_i)_{i\in[n_j]}$, there is no orientation of the set taken into account.
-% That is why, we prefer to choose the Varifold distance $d_V(\cdot,\cdot)$ \cite{kaltenmark2017general} which was specifically designed for oriented curves.
-%  The Varifold distance extends the MMD by adding tangent vectors to the input:
-%   instead of considering a set $X=(x_i)_{i\in[T_1]}$, we extend the input as $\overrightarrow{X}=(x_i,v_i)_{i\in[T_1]}$ where $v_i$ is the tangent vector at $x_i$ such that
-%    \begin{equation}
-%      d_V(X,Y)=\operatorname{MMD}^{K_p}(\mu_{\overrightarrow{X}},\mu_{\overrightarrow{X}}) \eqsp ,
-%    \end{equation}
-%    where in practice the tangent vectors are taken as discrete derivative and $K_p$ is a kernel on $\Rset^{2d'}$ such that for any $(x,v),(y,w)\in (\Rset^{2d})^2$, 
-%    we have, 
-%    \begin{equation}
-%      K_p((x,v),(y,w))=k(x,v)k_T(v,w) \eqsp ,
-%    \end{equation}
-%   where $k_T $ is a kernel on the tangent vector space.
-
-\sam{Parler de méthode adaptatif ici}
-
-%Wavelet-Based Multiscale Flow For Realistic Image Deformation in the Large Diffeomorphic Deformation Model Framework
-% 3.1 +3.2 + specialité de la décomposition (t,f(t,x))
-% \section{Specificity to time series}
-% We are looking to time seriess 
-% Each time series $s: t\in[0,T]\mapsto s(t) $ is only known through its samples $(s(t_i)=s_i)_{i\in[N]} $ at timestamps $(t_i)_{i\in[N]}$.
-%is described by its graph $G(s)=\{(y,s(t)), \} $
-
-% \section{optimzation
-% % We are not doing Pre-ESPAce car l'on s'intéresse à la reparamétrisation en temps, au lieu de calculer la transfo en temps après avoir
-% % push en espace, on le fait avant pour que tout soit défini depuis s_0, méthode itérative car ça marche mieux, moins de calcul de gradient
-% %Natural smoothing with TTS
-% \label{section:optimization}
-%     \begin{center}
-%         \begin{tikzpicture}
-%             \node at (0,0) {$G_f^I$};
-%             \draw [-stealth](0.3,0) -- (1.3,0);
-%             \node at (1.6,0) {$G_f^J$};
-%             \draw [-stealth](1.6,-0.3) -- (1.6,-1.3);
-%             \node at (1.6,-1.6) {$G_g^J$};
-%             \draw [-stealth](1.3,-1.6) -- (0.3,-1.6);
-%             \node at (0,-1.6) {$G_g^I$};
-%             \draw [-stealth](0,-1.3) -- (0,-0.3);
-
-%             \node at (0.8,0.3) {$T$};
-%             \node at (0.8,-1.9) {$T^{-1}$};
-%             \node at (1.9,-0.8) {$S_J$};
-%             \node at (-0.4,-0.8) {$S_I^{-1}$};
-%         \end{tikzpicture}
-%         \end{center}
-
-%         where $S_I^{-1} = T^{-1} \circ S_J^{-1} \circ T$
-
 \section{Experiments}
@@
   \end{table}
-\section{Conclusion}
-In this paper, we propose a feature representation method, TS-LDDMM, designed for 
-shape comparison in homogeneous time series datasets. We show on a real dataset 
-its ability to study, with high interpretability, the inter-individual shape 
-variability. As an unsupervised approach, it is user-friendly and enables knowledge 
-transfer for different supervised tasks such as classification. Although TS-LDDMM 
-is already competitive for classification, its performances can be leveraged on 
-more heterogeneous datasets using a hierarchical clustering extension, which is relagated for future work.
-
-%\section{Societal impact}
-%This paper presents work whose goal is to advance the field of Machine Learning. There are many potential societal consequences of our work, none which we feel must be specifically highlighted here.
-% \section{Acknowledgements}
-
-%The TS-LDDMM feature representation proposed in this paper enables us to study the inter-individual variability of shapes in a time series dataset with high interpretability.
-%Its unsupervised aspect is user-friendly and enables knowledge transfer for different supervised tasks such as classification.
-%Although the method is already competitive,
-% its capacity can be leveraged on more heterogeneous datasets using a hierarchical clustering extension.
-% This is relegated to future works.
-
-%Applying NF or CNF to our
- %  framework with high dimensional time series is relagated for future works.
-
-% \begin{figure}[ht]
-% \vskip 0.2in
-% \begin{center}
-% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
-% \caption{Historical locations and number of accepted papers for International
-% Machine Learning Conferences (ICML 1993 -- ICML 2008) and International
-% Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
-% produced, the number of accepted papers for ICML 2008 was unknown and instead
-% estimated.}
-% \label{icml-historical}
-% \end{center}
-% \vskip -0.2in
-% \end{figure}
-
-% \begin{algorithm}[tb]
-%    \caption{Bubble Sort}
-%    \label{alg:example}
-% \begin{algorithmic}
-%    \STATE {\bfseries Input:} data $x_i$, size $m$
-%    \REPEAT
-%    \STATE Initialize $noChange = true$.
-%    \FOR{$i=1$ {\bfseries to} $m-1$}
-%    \IF{$x_i > x_{i+1}$}
-%    \STATE Swap $x_i$ and $x_{i+1}$
-%    \STATE $noChange = false$
-%    \ENDIF
-%    \ENDFOR
-%    \UNTIL{$noChange$ is $true$}
-% \end{algorithmic}
-% \end{algorithm}

### 2024-05-21 NEURIPS24/paper.tex
separate exp

@@
 
-\section{Experiments}
-\label{section:experiments}
-[L'intro est ce necessaire ?]
+\include{experiment}
 
-First, we show on synthetic data that the proposed representation is identifiable provided that the hyperparameters and the reference graph are wisely selected, i.e.,
- the parameter $v_0^*$ generating a deformation $\varphi^{\{v_0^*\}}$ of a time series graph $\msg$ can be estimated from the data $\msg,\varphi^{\{v_0^*\}}.\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation}.
- Secondly, we illustrate the qualitative interest of TS-LDDMM in studying inter-individual variability on a clinical dataset.
-  Thirdly, we demonstrate the quantitative performance of our representation by performing classification on shape-based datasets.
-  The method is implemented on Python using the library JAX\footnote{https://github.com/google/jax}.
-The code was compiled on a server with NVIDIA RTX A2000 12GB GPU, Intel(R) Xeon(R) Gold 5220R CPU @ 2.20GHz, and 250 GB of RAM.
-The code will be available on Github.
-\subsection{Synthetic experiments}
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
-    \vspace{-2.5em}
-    \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
-    \label{fig:exemple_synthetic}
-    \vspace{-1em}
-\end{figure}
-
-\begin{table}
-    \caption{Values of $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ as $\mathbf{\alpha}^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
-      \centering
-         \begin{tabular}{lrrrrrrr}
-         \toprule
-         $\sigma_{T,0} \backslash \sigma_x$  & 1 & 10 & 50 & 100 & 200 & 300 \\
-         \midrule
-         0.1 & 2e+0 & 3e-4  & 1e-5&4e-6&7e-4&4e-3 \\
-        1 & 4e-2 & 1e-4  & 1e-5&4e-6&7e-4 &4e-3  \\
-         100 & 4e-2 & 2e-4  & 1e-5&4e-6&7e-4&4e-3  \\
-         \bottomrule
-         \end{tabular}
-      \label{table:synthetic2}
-      \vspace{-1em}
-  \end{table}
-
-% \begin{subfigure}{0.7\linewidth}
-%   \centering
-%   \resizebox{\columnwidth}{!}{%
-%      \begin{tabular}{lrrrr}
-%      \toprule
-%      $\sigma_{T,0} \backslash \sigma_x$  & 0.1 & 1 & 100 \\
-%      \midrule
-%      1 & 2e+0 & 4e-2 & 4e-2 \\
-%      10 & 3e-4 & 1e-4  & 2e-4 \\
-%      50 & 1e-5 & 1e-5  & 1e-5 \\
-%      100 & 4e-6 & 4e-6 & 4e-6 \\
-%      200 & 7e-4 & 7e-4 & 7e-4 \\
-%      300 & 4e-3 & 4e-3 & 4e-3 \\
-%      \bottomrule
-%      \end{tabular}
-%      %
-%      }
-%  \caption{Values of $\scrl(y^*,\varphi^{\{\hat{v}_0\}}.\msx)$ where $y^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
-% \end{subfigure}
-First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
-All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
-We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
-Given $m_{s}\in \Nset_{>0}$ and $t_{a},s_{a}>0$, we randomly generate initial momentums $\mathbf{\alpha}^*=(\alpha_k^*)_{k\in[\mathbf{n}_0]}$ with the following sampling, called Gen($m_s,t_a,s_a$):
- For any $k\in[\mathbf{n}_0]$, $\mathbf{\alpha}_k'$ is sampled according to a Gaussian normal distribution $\mathcal{N}(0_{d+1},I_{d+1})$.
-Then, $(\alpha_k')_{k\in[\mathbf{n}_0]}$ is regularized by a rolling average of size $m_{s}$, we get $\bar{\mathbf{\alpha}}'=(\bar{\alpha}_k')_{k\in[\mathbf{n}_0]}$.
-Finally, we normalize $\bar{\mathbf{\alpha}}'$ to derive $\mathbf{\alpha}^*$ such that $|([\alpha_k^*]_t)_{k\in[\mathbf{n}_0]}|=t_{\text{amp}}$ and $|([\alpha_k^*]_s)_{k\in[\mathbf{n}_0]}|=s_{\text{amp}}$ for any $k\in[\mathbf{n}_0]$, denoting by $[\alpha_k^*]_t,[\alpha_k^*]_s$ the time and space coordinates of $\alpha_k^*$ respectively.
-Note that the regularizing step $(\mathbf{\alpha}_k')_{k\in[\mathbf{n}_0]}\to \bar{\mathbf{\alpha}}' $ is necessary to obtain realistic deformations which take into account the regularity induced by the RKHS $\msv$.
-%  uniformly on the set $\mss_M=\{ (\alpha_k)_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}:\eqsp |\alpha_k|=M\}  $ with $M>0,\mathbf{n}_0\in \Nset_{>0}$,
-%  then we regularize $\mathbf{\alpha}$ by convolving
-%   with ... to get the generation parameter $\mathbf{\alpha}^*$.
-Then, using $v_0(\mathbf{\alpha}^*,\msx)$ as defined in \eqref{eq:def_v0} with initial momentums $\mathbf{\alpha}^*$ and control points $\msx$, we apply the induced deformation $\varphi^{\{v_0\}} $ by \eqref{eq:integration} to $\msx$ and obtain $\varphi^{\{v_0\}}.\msx$.
-Finally, we solve \eqref{eq:relaxation} to recover an estimation $\hat{\mathbf{\alpha}}$ of $\mathbf{\alpha}^*$ and report the average relative error (ARE) $|v_0(\hat{\mathbf{\alpha}},\msx)-v_0(\mathbf{\alpha}^*,\msx)|_\msv/|v_0(\mathbf{\alpha}^*,\msx)|_\msv$ on 50 repetitions.
-This procedure is performed for any $m_{s},t_{a},s_{a}\in \{10,50,100\}\times \{5,10,15,20\}^2 $.
- Mean, standard deviation, and maximum of the ARE on all these hyperparameters choices are respectively $\mathbf{0.10, 0.03, 0.17}$.
- Therefore, the estimation procedure \eqref{eq:relaxation} offers a good approximation of the true parameter when the kernel $K_\msg$ is well specified.
- We observe that the estimation is difficult when $t_a\ll s_a$ because the time series can be very noisy as illustrated in \Cref{fig:exemple_synthetic}: this impacts the Varifold loss which is sensitive to tangents.
-
- Secondly, we demonstrate a weak identifiability when the kernel $K_\msg$ is misspecified: we can reconstruct the graph time series' after deformations even if the hyperparameters of $K_\msg$
-  are different during the generation and the estimation.
-   The hyperparameters of $K_\msg$ during generation are $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)=(1,0.1,100,1,1)$ and we fix $\sigma_{T,1},c_0,c_1=(1,1,0.1) $ for $K_\msg$ during estimation.
-   We aim to understand the impact of $\sigma_{T,1},\sigma_x$ on the reconstruction since they are encoding the smoothness of the transformation according to time and space.
-
-    For any choice of the hyperparameters $\sigma_{T,1},\sigma_x\in \{1,10,50,100,200,300 \}\times \{0.1,1,100\}$ related to $K_\msg$ in the estimation,
-     we average $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ on 50 repetitions when $\mathbf{\alpha}^*$ is sampled according to Gen$(10,10,50)$ and $\hat{v}_0=v_0(\hat{\alpha},\msx)$ denoting by $\hat{\alpha}$ the result of the minimization \eqref{eq:relaxation}.
-  We observe in \Cref{table:synthetic2} that the reconstruction is almost perfect except in the case when $\sigma_{t,0}=1$ during estimation, while $ \sigma_{t,0}=100$ during generation.
-   Compared to $\sigma_{T,0}$, $\sigma_x$ has nearly no impact on the reconstruction.
-   In \Cref{appendix:kernel_implementation}-\ref{appendix:kernel_TS_LDDMM}, we propose guidelines to drive future hyperparameters tuning and further discussions related to $\sigma_{T,1},c_0,c_1$. 
-\subsection{Qualitative analysis of respiratory behavior in mice}
-%We consider a dataset composed of mouse respiratory time series before and after a drug injection.
-% A complete presentation of this dataset is given in \Cref{appendix:mouse_dataset}.
-%  The mouse are divided in two groups depending on their genotypes : \textit{colq} and \textit{wt}.
-%  We aim to study the difference of respiratory shapes according to the genotype.
-%   That is why, TS-LDDMM \eqref{eq:general_optimization_problem} is applied to derive the features representations $(\mathbf{\alpha}_j)_{j\in[N]}\in (\Rset^2)^N$ related to $N$=\sam{fill} respiratory time series coming from 14 mouse (7 \textit{colq} and \textit{wt}) before drug injection.
-%   Then, a Principal Components Analysis (PCA) is performed on $(\mathbf{\alpha}_j)_{j\in[N]}$
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=0.95\linewidth]{"./pictures/exp_1_bis.pdf"}
-  \vspace{-1.5em}
-  \caption{Analysis of the three principal components (PC) related to the respiratory cycles of the mouse before exposure.
-  In Figure A), the densities of each genotype according to each PC are displayed.
-In Figure B), the deformations of the reference graph $S_0$ along each PC are given.
-In Figure D), the graph of reference $S^j$, also called barycenter, related to each mouse, is displayed according to their coordinates on PC1 and PC3.
-In Figure C) et E), illustrations of respiratory cycles related to mice coming from the \textbf{wt} and \textbf{colq} group are displayed.
-}
-  \label{fig:exp_1_PCA}
-  %\vspace{-1em}
-\end{figure*}
-
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=0.95\linewidth]{"./pictures/exp_2.pdf"}
-  \vspace{-1em}
-  \caption{Analysis of the first Principal Component (PC1) related to the respiratory cycles of the mouse 
-  before and after exposure.
-In Figure A), the densities of each genotype according are displayed.
-In Figure B), the deformations of the reference graph $S_0$ PC1 is given.
-In Figure C), respiratory cycles displayed with respect to time and according to their coordinates on PC1 and PC2}
-  \label{fig:exp_2_PCA}
-  \vspace{-1.5em}
-\end{figure*}
-
-This experiment highlights the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-We consider a time series dataset recording the evolution of the respiratory airflow of mice exposed 
-to an irritant molecule altering respiratory functions \cite{nervo2019respiratory}. The dataset is divided into two groups, one 
-composed of 7 control mice (\textbf{wt}) and the other of 7 mice (\textbf{colq}) deficient in an enzyme 
-involved in the control of respiration. For each mouse, the respiratory airflow was recorded for 
-15 to 20 minutes before exposure to the irritant molecule and then for 35 to 40 minutes. A complete 
-description of the dataset is given in the \Cref{appendix:mouse_dataset}.
-By comparing the shape of individual respiratory cycles (inspiration + expiration, see \Cref{fig:exp_1_PCA}-C)), 
-we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution 
-after exposure to the irritant molecule.
-
-We first compare breathing behaviors before exposure.
-Solving \eqref{eq:general_optimization_problem}, we derive the reference respiratory cycle's graph $S_0$ and the TS-LDDMM features representations 
-$(\alpha_j)_{j\in[N_1]}$ related to $N_1=700$ respiratory cycles extracted according 
-to the procedure \cite{germain2023unsupervised}.
-Then, we perform a kernel PCA on the initial velocity field $\left(v_0(\alpha_j,S_0)\right)_{j\in[N_1]}\in \msv^{N_1}$ defined in \eqref{eq:def_v0}.
-In \Cref{fig:exp_1_PCA}, we focus on the analysis of the three Principal Components (PC).
-
-As observable from \Cref{fig:exp_1_PCA}-B), principal components refer to different types of deformations. 
-By interpreting \Cref{fig:exp_1_PCA}-B): Only PC1 accounts for time warping, PC2 expresses the trade-off between inspiration and expiration duration, and PC3 corresponds to a change in signal amplitude.
- Compared to \textbf{wt} mice, the distribution of \textbf{colq} mice 
-TS-LDMMM feature representation along the PC1 axis has a heavy tail and the associated deformation (+3 $\sigma_{\text{PC}}$) 
-shows an inspiration with two peaks. As illustrated in \Cref{fig:exp_1_PCA}-A), such respiratory cycles are preponderant
- with \textbf{colq} mice and may be caused by motor impairment due to their enzyme deficiency, \cite{germain2023unsupervised}.
-  In addition, 
- the \textbf{colq} mice were smaller than the \textbf{wt} mice due to a delay in growth caused by their lack of an enzyme. 
- This difference can be seen on PC3 since the volumes of air (area under the curve) inspired and exhaled are
-  smaller for the smaller mice. In correlation, the distribution of \textbf{wt} mice TS-LDDMM feature representations 
-  along the PC3 axis have a heavy tail corresponding to large air volume as depicted by the deformation 
-  (+3 $\sigma_{\text{PC}}$) in \Cref{fig:exp_1_PCA}-B). Finally, \Cref{fig:exp_1_PCA}-D) shows that PC1 and PC3 capture the main differences between
-   the two groups as their respective reference graphs $S^j$ are located in different parts of the space. 
-%In Figure \Cref{fig:exp_1_PCA}, we display coordinates densities per genotype and PC.
-
-We perform a second experiment to analyze the evolution of breathing behaviors when mice are exposed to the irritant 
-molecule. We follow the same procedure as before. However, we take $N_2=1400$ with 25\% (resp. 75\%) 
-before (resp. after) exposure. In \Cref{fig:exp_2_PCA}, we focus on the first principal component PC since
-it encodes the effect of the irritant molecule as depicted in \Cref{fig:exp_2_PCA}-C) (the exposure occurs at 20 minutes).
-\Cref{fig:exp_2_PCA}-B) shows that the deformation (+3 $\sigma_{\text{PC}}$) leads to longer respiratory cycles that include pauses,
-as observed in \cite{germain2023unsupervised}. As well, \Cref{fig:exp_2_PCA}-A) shows that TS-LDDMM features distributions are less spread 
-out for \textbf{colq} mice compared to \textbf{wt} mice. Indeed, the irritant molecule inhibits the action of the 
-deficient enzyme, \textbf{wt} mice strongly react to the irritant molecule, whereas \textbf{colq} mice are better 
-adapted due to their deficiency.
-%Secondly, we compare breathing behaviors before and after exposure to observe the impact of the irritant molecule.
-% We follow the same procedure as for before exposure, but we take $N_2=1400$ respiratory cycles extracted according 
-% to the procedure CITE....
-% In \Cref{fig:exp_2_PCA}, we focus on the first Principal Component (PC) since it encodes the effect of the irritant molecule as demonstrated on Figure \ref{fig:exp_2_PCA}-C) (exposure at time 20).
-% We observe on \Cref{fig:exp_2_PCA}-B) that after exposure the mouse have longuer breath such that the expiration is longuer than inspiration.
-%  Moreover, we deduce from Figure \ref{fig:exp_2_PCA}-A) that \textbf{colq} are more constant in their breath compared to \textbf{wt} after exposure.
-%
-% In Figure REF, we 
-% also display the reference respiratory cycle S0 and its deformations in the principal component directions.
-%  Additionally, we learn each mouse's reference respiratory cycle and represent them in the first and third PC coordinates system in Figure REF.
-
-\subsection{Quantitative performances of the TS-LDDMM representation in classification}
-Combined with a Support Vector Classifier (SVC) \cite{hsu2003practical}, TS-LDDMM representation can be used for 
-classification tasks using the kernel associated with the initial velocity space $\msv$.
-We compare TS-LDDMM-SVC classification performances with another SVC using representation 
-learned with T-loss \cite{franceschi2019unsupervised}, an unsupervised deep learning feature 
-representation method for time series. We also include fully supervised methods in deep learning 
--ResNet, CNN \cite{ismail2019deep}- and machine learning: Catch22 \cite{lubba2019catch22}, Rocket
-\cite{dempster2020rocket}, Dynamic Time Wrapping k-Nearest Neighbors (DTW-kNN) 
-\cite{muller2007dynamic}. Methods are compared using f1-score on several shape-based UCR/UEA datasets 
-\cite{dau2019ucr,bagnall2018uea} introduced in \Cref{appendix:classification_dataset}. All implementation details are given 
-in \Cref{appendix:classification_implementation}.
-\Cref{table:classification} presents the reuslts. TS-LDDMM-SVC consistently outperforms the other unsupervised methods. It is ranked 1,3,4,3 for all methods combined, 
-demonstrating its competitiveness as an unsupervised method on time series dataset homogeneous regarding shape.
-%TS-LDDMM representation can be used in a Support Vector Classifier (SVC) \cite{hsu2003practical} using the kernel of the space of initial velocity $\msv$.
-%We compare its capacity of representation to an SVC using T-Loss representation \cite{franceschi2019unsupervised}, which is an unsupervised representation deep learning method for time series,
-%and to others fully supervised methods using deep learning -ResNet, CNN \cite{ismail2019deep}- or not: Catch22 \cite{lubba2019catch22}, Rocket \cite{dempster2020rocket}, Dynamic Time Wrapping \cite{muller2007dynamic} k-Nearest Neighbors (DTW-kNN).
-%These methods are compared using f1-score on several datasets introduced in \Cref{appendix:classification_dataset}.
-All details of implementations are given in \Cref{appendix:classification_implementation}.
-%TS-LDDMM-SVC is always ranked 2 or 3 behind a fully supervised method, demonstrating its competitiveness as an unsupervised method.
-%The more homogeneous the shapes in the dataset are, the better TS-LDDMM-SVC performs.
-
-\begin{table}[t]
-  \vspace{-0.5em}
-  \caption{Classification results in f1-score (U: unsupervised, S: supervised, DL: deep learning, ML: machine learning). \textbf{x} best unsupervised method, \underline{x} best supervised method.}
-  \resizebox{\linewidth}{!}{%
-  \begin{tabular}{lllrrrr}
-    \toprule
-     &  &  & ArrowHead & ECG200 & GunPoint & NATOPS \\
-    \midrule
-    \multirow[m]{3}{*}{U} & \multirow[t]{3}{*}{} & TS-LDDMM-SVC & \textbf{0.84} & \textbf{0.82} & \textbf{0.94} & \textbf{0.93} \\
-     &  & T-loss-SVC & 0.57 & 0.76 & 0.82 & 0.88 \\
-     &  & DTW-kNN & 0.70 & 0.75 & 0.91 & 0.88 \\
-    \cline{1-7} 
-    \multirow[m]{5}{*}{S} & \multirow[m]{2}{*}{DL} & CNN & 0.70 & 0.79 & 0.85 & \underline{0.96} \\
-     &  & ResNet & 0.77 & 0.87 & 0.97 & 0.95 \\
-    \cline{2-7}
-     & \multirow[m]{2}{*}{ML} & Catch22 & 0.73 & 0.81 & 0.96 & 0.89 \\
-     &  & Rocket & \underline{0.81} & \underline{0.91} & \underline{1.00} & 0.88 \\
-    \bottomrule
-    \end{tabular}
-    \label{table:classification}
-  %
-}
-\vspace{-2em}
-  \end{table}
 \section{Conclusion}

### 2024-05-21 NEURIPS24/paper.tex
update related works

@@
 Although a community structure with representatives can be learned in an unsupervised way \cite{trirat2024universal,meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised} or similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},
-studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem in URL.
+studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem in URL and even more for \textit{irregularly sampled} time series with \textit{variable lengths}.
      %In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.
@@
 
-\thi{
-  Studying shape differences between time series related to biological mechanisms is a common practice in physiology to characterize healthy and pathological functioning CITE.
+% \thi{
+%   Studying shape differences between time series related to biological mechanisms is a common practice in physiology to characterize healthy and pathological functioning CITE.
 For instance, the shapes of heartbeats in electrocardiograms are discriminant for some cardiovascular pathologies CITE.
@@
 However, shape-based representation of time series within cohorts~\cite{niennattrakul2007inaccuracies,shirato2023identifying} remains an open problem in URL.
-}
+% }
 
@@
 \section{Related Works}
-Shape analysis focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
- The main idea is to represent these different objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
+Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
+   The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
     Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
- The goal is then to find a well suited Riemannian struvture according to the studied object.
+     The goal is to find a well-suited Riemannian structure according to the nature of the studied object.
 
- A time series graph can be seen as a curve and LDDMM structure is relevant to tackle curves as presented in \cite{glaunes2008large}. %, LDDMM structure is relevant to tackle curves and
-  However, time series graph has more structure than curves as depicted in \Cref{fig:diffeo} due to the temporal evolution. 
-  \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM, but distinguish from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
+     LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
+However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
+      Closely related, \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM but distinguishes from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
 
-  Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} address the representation of curves having unitary velocity by using the Square-Root Velocity (SRV) representation.
- However, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
-  Again, the time series graph structure is not respected.
-   Very recently in a functionnal data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by giving a representation for the original time evolution.
-    Nevertheless, this methods applies only on time series of \textit{same size} and is made for \textit{continuous objects}, making the estimation more sensitive to noise.
+       Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
+        However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment.
+         Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation.
+          Very recently, in a functional data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution.
+           Nevertheless, this method is made for \textit{continuous objects} and only applies to time series of \textit{same length}, making the estimation more sensitive to noise.
   %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
@@
 %       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.
-[Ajouter littérature de URL demander à Sisi ?]
+
  %\citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
@@
  %Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
+ Balancing between discrete and continuous elements is a challenging task.
+  In the deep learning literature \cite{chen2018neural, kidger2020neural, tzen2019neural, jia2019neural, liu2019neural, ansari2023neural}, Neural Ordinary Differential Equations (Neural ODEs) \cite{chen2018neural} learn continuous latent representations using a vector field parameterized by a neural network, serving as a continuous analogue to Residual Networks \cite{zagoruyko2016wide}.
+   This approach was further enhanced by Neural Controlled Differential Equations (Neural CDEs) \cite{kidger2020neural} for handling irregular time series, functioning as continuous-time analogs of RNNs \cite{schuster1997bidirectional}.
+ Extending Neural ODEs, Neural Stochastic Differential Equations (Neural SDEs) introduce regularization effects \cite{liu2019neural}, although optimization remains challenging.
+  Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied succesfully Neural SDEs to irregular time series.
+  Oh et al. \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
 
-\thi{
-  Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
-The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
-Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
-The goal is to find a well-suited Riemannian structure according to the nature of the studied object.
-LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
-However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
-Closely related, \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM but distinguishes from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
-Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
-However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment.
-Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation.
-Very recently, in a functional data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution.
-Nevertheless, this method is made for \textit{continuous objects} and only applies to time series of \textit{same length}, making the estimation more sensitive to noise.
-}
 \section{Notations}

### 2024-05-21 NEURIPS24/appendix.tex
draft exp

@@
 
-\section{Numerical details}
-A report of all the hyperparameters selected is given in \Cref{appendix:table:set_up_exp_practical}.
-\subsection{Optimization details of \eqref{eq:general_optimization_problem}}
+\section{Experimental settings}
+
+All experiments were performed on a Debian 6.1.69-1 server with NVIDIA RTX A2000 12GB GPU, Intel(R) Xeon(R) Gold 5220R CPU @ 2.20GHz, and 250 GB of RAM. The source code will be available on Github.
+
+\subsection{Optimization details of TS-LDDMM~\eqref{eq:general_optimization_problem}}
 \label{appendix:optimizers_details}
-\paragraph{Initialization}
-At the initialization of \eqref{eq:general_optimization_problem}, all the momentums parameter are set to $0$ and the graph of reference is set to the graph of a time series in the dataset having a mediane samples size.
+
+We implemented TS-LDDMM in Python with the JAX library \footnote{https://github.com/google/jax}.
+
+\paragraph{Initialization.}
+As initialization of \eqref{eq:general_optimization_problem}, all momentum parameters are set to $0$, and the initial graph of reference is picked from the dataset such that its length is equal to the median length observed in the dataset.
 
 \paragraph{Gradient descent.}
-The chosen gradient descent method is "adabelief" \cite{zhuang2020adabelief} implemented in the library OPTAX\footnote{https://optax.readthedocs.io/en/latest/}.
-There are two main parameters in the gradient descent: the number of steps nb\_steps, and the maximum value of step size $\eta_M$.
-The stepsize has a particular scheduling: 
+The chosen gradient descent method is "adabelief" \cite{zhuang2020adabelief} implemented in the OPTAX library~\footnote{https://optax.readthedocs.io/en/latest/}.
+The gradient descent has two main parameters: the number of steps (nb\_steps) and the maximum stepsize value ($\eta_M$).
+The stepsize has a scheduling scheme: 
 \begin{itemize}
-  \item Warmup period on $0.1 \times$ nb\_steps steps: the stepsize increases linearly from $0$ to $\eta_M$. The goal is to learn progressively the parameters.
-   If the stepsize is too large at the start, smaller steps at the end can't make up for the mistakes made at the beginning. 
-  \item Fine tuning periode on $ 0.9  \times$ nb\_steps : the stepsize decreases from $\eta_M$ to $0$ with a cosine decay implemented in the OPTAX scheduler,
-   i.e. the decreasing factor as the form $0.5  (1 + \cos(\pi  t/T))$. 
+  \item Warmup period on $0.1 \times$ nb\_steps steps: the stepsize increases linearly from $0$ to $\eta_M$.
+The goal is to learn progressively the parameters.
+If the step size is too large at the start, smaller steps at the end cannot make up for the mistakes made at the beginning.
+  \item Fine tuning periode on $ 0.9  \times$ nb\_steps : the stepsize decreases from $\eta_M$ to $0$ with a cosine decay implemented in the OPTAX scheduler, i.e.
+the decreasing factor as the form $0.5  (1 + \cos(\pi  t/T))$.
 \end{itemize}
-The sharper the deformations, the larger the number of steps and the maximum value of step size should be selected.
-We suggest nb\_steps=300, $\eta_M=0.1 $ for small deformations and nb\_steps=800, $\eta_M=0.3 $ for big ones (time dilation with a factor $\lambda\geq 2$).
+By default, we set nb\_steps to 400 and $\eta_M$ to 0.1.
+
+\subsection{Identifiability experiment}
+This experiement only involves the TS-LDDMM method in two different settings: 
+\begin{itemize}
+  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
+  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$.
+The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
+\end{itemize}
+TS-LDDMM parameters of the experiment are summarized in \Cref{appendix:table:set_up_exp_practical}
+
+\subsection{Mice respiratory behaviors experiment}
+
+This experiment involves TS-LDDMM and LDDMM~\cite{glaunes2008large} methods.
+Both methods are run twice, first on respiratory cycles before exposure to the irritant molecule to capture mice breathing behavior at rest and on all respiratory cycles to capture the influence of the irritant molecule.
+Exposure to the irritant molecule leads to significant shape deformation in the respiratory cycles, and the terms must be added to the varifold loss to capture deformations at a large time scale.
+
+\paragraph{TS-LDDMMM parameters.}
+\begin{itemize}
+  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,150,1,2)$.
+The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
+  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,220,1,2)$.
+The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
+\end{itemize}
+
+\paragraph{LDDMMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM.
+Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
+\begin{itemize}
+  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =150$ for the time dimension and $\sigma_x = 2$ for space dimensions. 
+  The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
+  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =220$ for the time dimension and $\sigma_x = 2$ for space dimensions.
+  The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
+\end{itemize}
+\subsection{Benchmark}
+
+A report of all the hyperparameters selected is given in \Cref{appendix:table:set_up_exp_practical}.
 
@@
             \midrule
-            Synthetic data well-specified &(300,0.1) &$(2,1,2,0.6)$& $(1,0.1,100,1,1)$   \\
-            Synthetic data misspecified  & (300,0.1)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
+            Synthetic data well-specified &(400,0.1) &$(2,1,2,0.6)$& $(1,0.1,100,1,1)$   \\
+            Synthetic data misspecified  & (400,0.1)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
             Mouse before injection  & (400,0.3)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
-            Mouse before/after injection  & (800,0.3)  & $ (5,1,5,0.6)$ &$(1,0.1,150,1,1)$  \\
-            classification  & (400,0.1)  & $(\max(2,0.03\bar{l}),N_d,\max(2,0.03\bar{l}),0.6)$  & $(1,0.1,0.33\bar{l},1,N_d)$  \\
+            Mouse before/after injection  & (400,0.3)  & $ (5,1,5,0.6)$ &$(1,0.1,150,1,1)$  \\
+            Classification  & (400,0.1)  & $(2,N_d,2,N_d)$  & $(1,0.1,0.33\bar{l},1,N_d)$  \\
 
@@
     \end{table}
-\section{Mouse respiratory dataset}
+
+\section{Datasets}
+\subsection{Mouse respiratory cycle dataset}
 
@@
 
-\section{Classification: Comparison with shape analysis methods}
+\subsection{Shape-based UCR/UEA time series classification datasets}
+\label{appendix:classification_dataset}
+We selected 15 shape-based datasets (7 univariates and 8 multivariates) from the from the University of East Anglia (UEA) and the University of California Riverside (UCR) Time Series Classification Repository\footnote{https://timeseriesclassification.com} \cite{dau2019ucr,bagnall2018uea}.
+All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/}.
+Essential datasets information are summarized in \Cref{appendix:table:datasets} and further can be found in \cite{dau2019ucr,bagnall2018uea}.
+
+\begin{table}[hbt!]
+  \centering
+  \caption{UCR/UEA shape-based time series datasets for classification.}
+  \resizebox{\columnwidth}{!}{%
+  \begin{tabular}{lllllll}
+    \toprule
+    & \textbf{Dataset} &  \textbf{Size} & \textbf{Lengh} & \textbf{Number of classes} & \textbf{Number of dimensions} & \textbf{Type} \\
+    \midrule
+    \multirow[c]{7}{*}{Univariate} &  ArrowHead & 211 & 251 & 3 & 1 & IMAGE \\
+    & BME & 180 & 128 & 3 & 1 & SIMULATED \\
+    & ECG200 & 200 & 96 & 2 & 1 & ECG \\
+    & FacesUCR & 2250 & 131 & 14 & 1 & IMAGE \\
+    & GunPoint & 200 & 150 & 2 & 1 & MOTION \\
+    & PhalangesOutlinesCorrect & 2658 & 80 & 2 & 1 & IMAGE \\
+    & Trace & 200 & 275 & 4 & 1 & SENSOR \\
+    \cline{1-7}
+    \multirow[c]{8}{*}{Multivariate}& ArticularyWordRecognition & 575 & 144 & 25 & 9 & SENSOR \\
+    & Cricket & 180 & 1197 & 12 & 6 & MOTION \\
+    & ERing & 60 & 65 & 6 & 4 & SENSOR \\
+    & Handwriting & 1000 & 152 & 26 & 3 & MOTION \\
+    & Libras & 360 & 45 & 15 & 2 & VIDEO \\ 
+    & NATOPS & 360 & 51 & 6 & 24 & MOTION \\
+    & RacketSports & 303 & 30 & 4 & 6 & SENSOR \\
+    & UWaveGestureLibrary & 240 & 315 & 8 &3 & SENSOR \\
+    \bottomrule
+  \end{tabular}
+  %
+  }
+  \label{appendix:table:datasets}
+\end{table}
+
+\section{Experiments}
+
+\subsection{Identifiability experiments}
+\label{appendix:identifiability}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
+  \vspace{-2.5em}
+  \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
+  \label{fig:exemple_synthetic}
+  \vspace{-1em}
+\end{figure}
+
+\begin{table}
+  \caption{Values of $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ as $\mathbf{\alpha}^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
+    \centering
+       \begin{tabular}{lrrrrrrr}
+       \toprule
+       $\sigma_{T,0} \backslash \sigma_x$  & 1 & 10 & 50 & 100 & 200 & 300 \\
+       \midrule
+       0.1 & 2e+0 & 3e-4  & 1e-5&4e-6&7e-4&4e-3 \\
+      1 & 4e-2 & 1e-4  & 1e-5&4e-6&7e-4 &4e-3  \\
+       100 & 4e-2 & 2e-4  & 1e-5&4e-6&7e-4&4e-3  \\
+       \bottomrule
+       \end{tabular}
+    \label{table:synthetic2}
+    \vspace{-1em}
+\end{table}
+
+% \begin{subfigure}{0.7\linewidth}
+%   \centering
+%   \resizebox{\columnwidth}{!}{%
+%      \begin{tabular}{lrrrr}
+%      \toprule
+%      $\sigma_{T,0} \backslash \sigma_x$  & 0.1 & 1 & 100 \\
+%      \midrule
+%      1 & 2e+0 & 4e-2 & 4e-2 \\
+%      10 & 3e-4 & 1e-4  & 2e-4 \\
+%      50 & 1e-5 & 1e-5  & 1e-5 \\
+%      100 & 4e-6 & 4e-6 & 4e-6 \\
+%      200 & 7e-4 & 7e-4 & 7e-4 \\
+%      300 & 4e-3 & 4e-3 & 4e-3 \\
+%      \bottomrule
+%      \end{tabular}
+%      %
+%      }
+%  \caption{Values of $\scrl(y^*,\varphi^{\{\hat{v}_0\}}.\msx)$ where $y^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
+% \end{subfigure}
+
+provided that the hyperparameters and the reference graph are wisely selected, i.e., the parameter $v_0^*$ generating a deformation $\varphi^{\{v_0^*\}}$ of a time series graph $\msg$ can be estimated from the data $\msg,\varphi^{\{v_0^*\}}.\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation}.
+
+First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
+All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
+We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
+Given $m_{s}\in \Nset_{>0}$ and $t_{a},s_{a}>0$, we randomly generate initial momentums $\mathbf{\alpha}^*=(\alpha_k^*)_{k\in[\mathbf{n}_0]}$ with the following sampling, called Gen($m_s,t_a,s_a$):
+For any $k\in[\mathbf{n}_0]$, $\mathbf{\alpha}_k'$ is sampled according to a Gaussian normal distribution $\mathcal{N}(0_{d+1},I_{d+1})$.
+Then, $(\alpha_k')_{k\in[\mathbf{n}_0]}$ is regularized by a rolling average of size $m_{s}$, we get $\bar{\mathbf{\alpha}}'=(\bar{\alpha}_k')_{k\in[\mathbf{n}_0]}$.
+Finally, we normalize $\bar{\mathbf{\alpha}}'$ to derive $\mathbf{\alpha}^*$ such that $|([\alpha_k^*]_t)_{k\in[\mathbf{n}_0]}|=t_{\text{amp}}$ and $|([\alpha_k^*]_s)_{k\in[\mathbf{n}_0]}|=s_{\text{amp}}$ for any $k\in[\mathbf{n}_0]$, denoting by $[\alpha_k^*]_t,[\alpha_k^*]_s$ the time and space coordinates of $\alpha_k^*$ respectively.
+Note that the regularizing step $(\mathbf{\alpha}_k')_{k\in[\mathbf{n}_0]}\to \bar{\mathbf{\alpha}}' $ is necessary to obtain realistic deformations which take into account the regularity induced by the RKHS $\msv$.
+%  uniformly on the set $\mss_M=\{ (\alpha_k)_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}:\eqsp |\alpha_k|=M\}  $ with $M>0,\mathbf{n}_0\in \Nset_{>0}$,
+%  then we regularize $\mathbf{\alpha}$ by convolving
+%   with ... to get the generation parameter $\mathbf{\alpha}^*$.
+Then, using $v_0(\mathbf{\alpha}^*,\msx)$ as defined in \eqref{eq:def_v0} with initial momentums $\mathbf{\alpha}^*$ and control points $\msx$, we apply the induced deformation $\varphi^{\{v_0\}} $ by \eqref{eq:integration} to $\msx$ and obtain $\varphi^{\{v_0\}}.\msx$.
+Finally, we solve \eqref{eq:relaxation} to recover an estimation $\hat{\mathbf{\alpha}}$ of $\mathbf{\alpha}^*$ and report the average relative error (ARE) $|v_0(\hat{\mathbf{\alpha}},\msx)-v_0(\mathbf{\alpha}^*,\msx)|_\msv/|v_0(\mathbf{\alpha}^*,\msx)|_\msv$ on 50 repetitions.
+This procedure is performed for any $m_{s},t_{a},s_{a}\in \{10,50,100\}\times \{5,10,15,20\}^2 $.
+Mean, standard deviation, and maximum of the ARE on all these hyperparameters choices are respectively $\mathbf{0.10, 0.03, 0.17}$.
+Therefore, the estimation procedure \eqref{eq:relaxation} offers a good approximation of the true parameter when the kernel $K_\msg$ is well specified.
+We observe that the estimation is difficult when $t_a\ll s_a$ because the time series can be very noisy as illustrated in \Cref{fig:exemple_synthetic}: this impacts the Varifold loss which is sensitive to tangents.
+
+Secondly, we demonstrate a weak identifiability when the kernel $K_\msg$ is misspecified: we can reconstruct the graph time series' after deformations even if the hyperparameters of $K_\msg$
+are different during the generation and the estimation.
+ The hyperparameters of $K_\msg$ during generation are $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)=(1,0.1,100,1,1)$ and we fix $\sigma_{T,1},c_0,c_1=(1,1,0.1) $ for $K_\msg$ during estimation.
+ We aim to understand the impact of $\sigma_{T,1},\sigma_x$ on the reconstruction since they are encoding the smoothness of the transformation according to time and space.
+
+  For any choice of the hyperparameters $\sigma_{T,1},\sigma_x\in \{1,10,50,100,200,300 \}\times \{0.1,1,100\}$ related to $K_\msg$ in the estimation,
+   we average $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ on 50 repetitions when $\mathbf{\alpha}^*$ is sampled according to Gen$(10,10,50)$ and $\hat{v}_0=v_0(\hat{\alpha},\msx)$ denoting by $\hat{\alpha}$ the result of the minimization \eqref{eq:relaxation}.
+We observe in \Cref{table:synthetic2} that the reconstruction is almost perfect except in the case when $\sigma_{t,0}=1$ during estimation, while $ \sigma_{t,0}=100$ during generation.
+ Compared to $\sigma_{T,0}$, $\sigma_x$ has nearly no impact on the reconstruction.
+ In \Cref{appendix:kernel_implementation}-\ref{appendix:kernel_TS_LDDMM}, we propose guidelines to drive future hyperparameters tuning and further discussions related to $\sigma_{T,1},c_0,c_1$.
+
+\subsection{Classification: Comparison with shape analysis methods}
+\label{appendix:classification_shape_analysis}
 
@@
   \end{tabular}  
-\end{table}
-
-\section{Classification datasets}
-\label{appendix:classification_dataset}
-We selected 15 shape-based datasets (7 univariates and 8 multivariates) from the from the University of East Anglia (UEA) and the University of California Riverside (UCR) Time Series Classification Repository\footnote{https://timeseriesclassification.com} \cite{dau2019ucr,bagnall2018uea}.
-All datasets were downloaded with the python package aeon\footnote{https://www.aeon-toolkit.org/en/stable/}.
-Essential datasets information are summarized in \Cref{appendix:table:datasets} and further can be found in \cite{dau2019ucr,bagnall2018uea}.
-
-\begin{table}[hbt!]
-  \centering
-  \caption{UCR/UEA shape-based time series datasets for classification.}
-  \resizebox{\columnwidth}{!}{%
-  \begin{tabular}{lllllll}
-    \toprule
-    & \textbf{Dataset} &  \textbf{Size} & \textbf{Lengh} & \textbf{Number of classes} & \textbf{Number of dimensions} & \textbf{Type} \\
-    \midrule
-    \multirow[c]{7}{*}{Univariate} &  ArrowHead & 211 & 251 & 3 & 1 & IMAGE \\
-    & BME & 180 & 128 & 3 & 1 & SIMULATED \\
-    & ECG200 & 200 & 96 & 2 & 1 & ECG \\
-    & FacesUCR & 2250 & 131 & 14 & 1 & IMAGE \\
-    & GunPoint & 200 & 150 & 2 & 1 & MOTION \\
-    & PhalangesOutlinesCorrect & 2658 & 80 & 2 & 1 & IMAGE \\
-    & Trace & 200 & 275 & 4 & 1 & SENSOR \\
-    \cline{1-7}
-    \multirow[c]{8}{*}{Multivariate}& ArticularyWordRecognition & 575 & 144 & 25 & 9 & SENSOR \\
-    & Cricket & 180 & 1197 & 12 & 6 & MOTION \\
-    & ERing & 60 & 65 & 6 & 4 & SENSOR \\
-    & Handwriting & 1000 & 152 & 26 & 3 & MOTION \\
-    & Libras & 360 & 45 & 15 & 2 & VIDEO \\ 
-    & NATOPS & 360 & 51 & 6 & 24 & MOTION \\
-    & RacketSports & 303 & 30 & 4 & 6 & SENSOR \\
-    & UWaveGestureLibrary & 240 & 315 & 8 &3 & SENSOR \\
-    \bottomrule
-  \end{tabular}
-  %
-  }
-  \label{appendix:table:datasets}
 \end{table}

### 2024-05-21 NEURIPS24/experiment.tex
draft exp

@@
 \label{section:experiments}
-[L'intro est ce necessaire ?]
 
-First, we show on synthetic data that the proposed representation is identifiable provided that the hyperparameters and the reference graph are wisely selected, i.e.,
- the parameter $v_0^*$ generating a deformation $\varphi^{\{v_0^*\}}$ of a time series graph $\msg$ can be estimated from the data $\msg,\varphi^{\{v_0^*\}}.\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation}.
+For consiceness, several experiments are relagated in Appendix:
+
+\begin{enumerate}
+    \item \textbf{TS-LDDMM representation identifiability, \Cref{appendix:identifiability}:} Results show that TS-LDDMM representations are identifiable or weakly identifiable depending on the specification of the velocity field kernel $K_G$.
+    \item \textbf{Classification benchmark on regulary sampled datasets, \Cref{appendix:classification_shape_analysis}:} Results shows that methods based on TS-LDDMM representation outperform other methods from shape analysis on classification of 15 shape-based datasets (7 univariates \& 8 multivariates).
+    \item \textbf{Robustness to irregular sampling} We compared the robustness of TS-LDDMM representation with 9 URL methods handling irregularly sampled multivariate time series on 15 shape-based datasets (7 univariates \& 8 multivariates).
+We assess methods' classification performances under regular sampling (0\% missing rate) and three irregular sampling regimes (30\%,50\%, and 70\% missing rates), according to the protocol depicted in \cite{kidger2020neural}.
+\textbf{Results} show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%,30\%, and 50\%.
+The performance decrease of the Support Vector Machine classifier (SVC) based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.
+\end{enumerate}
+
+experiment on r.  As well, comparison of TS-LDDMM representation with other shape-analysis methods on a classification task is relagated in \Cref{appendix:classification_shape_analysis}.
+
+First, we show on synthetic data that the proposed representation is identifiable
  Secondly, we illustrate the qualitative interest of TS-LDDMM in studying inter-individual variability on a clinical dataset.
@@
 The code will be available on Github.
-\subsection{Synthetic experiments}
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
-    \vspace{-2.5em}
-    \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
-    \label{fig:exemple_synthetic}
-    \vspace{-1em}
-\end{figure}
 
-\begin{table}
-    \caption{Values of $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ as $\mathbf{\alpha}^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
-      \centering
-         \begin{tabular}{lrrrrrrr}
-         \toprule
-         $\sigma_{T,0} \backslash \sigma_x$  & 1 & 10 & 50 & 100 & 200 & 300 \\
-         \midrule
-         0.1 & 2e+0 & 3e-4  & 1e-5&4e-6&7e-4&4e-3 \\
-        1 & 4e-2 & 1e-4  & 1e-5&4e-6&7e-4 &4e-3  \\
-         100 & 4e-2 & 2e-4  & 1e-5&4e-6&7e-4&4e-3  \\
-         \bottomrule
-         \end{tabular}
-      \label{table:synthetic2}
-      \vspace{-1em}
-  \end{table}
-
-% \begin{subfigure}{0.7\linewidth}
-%   \centering
-%   \resizebox{\columnwidth}{!}{%
-%      \begin{tabular}{lrrrr}
-%      \toprule
-%      $\sigma_{T,0} \backslash \sigma_x$  & 0.1 & 1 & 100 \\
-%      \midrule
-%      1 & 2e+0 & 4e-2 & 4e-2 \\
-%      10 & 3e-4 & 1e-4  & 2e-4 \\
-%      50 & 1e-5 & 1e-5  & 1e-5 \\
-%      100 & 4e-6 & 4e-6 & 4e-6 \\
-%      200 & 7e-4 & 7e-4 & 7e-4 \\
-%      300 & 4e-3 & 4e-3 & 4e-3 \\
-%      \bottomrule
-%      \end{tabular}
-%      %
-%      }
-%  \caption{Values of $\scrl(y^*,\varphi^{\{\hat{v}_0\}}.\msx)$ where $y^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
-% \end{subfigure}
-First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
-All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
-We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
-Given $m_{s}\in \Nset_{>0}$ and $t_{a},s_{a}>0$, we randomly generate initial momentums $\mathbf{\alpha}^*=(\alpha_k^*)_{k\in[\mathbf{n}_0]}$ with the following sampling, called Gen($m_s,t_a,s_a$):
- For any $k\in[\mathbf{n}_0]$, $\mathbf{\alpha}_k'$ is sampled according to a Gaussian normal distribution $\mathcal{N}(0_{d+1},I_{d+1})$.
-Then, $(\alpha_k')_{k\in[\mathbf{n}_0]}$ is regularized by a rolling average of size $m_{s}$, we get $\bar{\mathbf{\alpha}}'=(\bar{\alpha}_k')_{k\in[\mathbf{n}_0]}$.
-Finally, we normalize $\bar{\mathbf{\alpha}}'$ to derive $\mathbf{\alpha}^*$ such that $|([\alpha_k^*]_t)_{k\in[\mathbf{n}_0]}|=t_{\text{amp}}$ and $|([\alpha_k^*]_s)_{k\in[\mathbf{n}_0]}|=s_{\text{amp}}$ for any $k\in[\mathbf{n}_0]$, denoting by $[\alpha_k^*]_t,[\alpha_k^*]_s$ the time and space coordinates of $\alpha_k^*$ respectively.
-Note that the regularizing step $(\mathbf{\alpha}_k')_{k\in[\mathbf{n}_0]}\to \bar{\mathbf{\alpha}}' $ is necessary to obtain realistic deformations which take into account the regularity induced by the RKHS $\msv$.
-%  uniformly on the set $\mss_M=\{ (\alpha_k)_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}:\eqsp |\alpha_k|=M\}  $ with $M>0,\mathbf{n}_0\in \Nset_{>0}$,
-%  then we regularize $\mathbf{\alpha}$ by convolving
-%   with ... to get the generation parameter $\mathbf{\alpha}^*$.
-Then, using $v_0(\mathbf{\alpha}^*,\msx)$ as defined in \eqref{eq:def_v0} with initial momentums $\mathbf{\alpha}^*$ and control points $\msx$, we apply the induced deformation $\varphi^{\{v_0\}} $ by \eqref{eq:integration} to $\msx$ and obtain $\varphi^{\{v_0\}}.\msx$.
-Finally, we solve \eqref{eq:relaxation} to recover an estimation $\hat{\mathbf{\alpha}}$ of $\mathbf{\alpha}^*$ and report the average relative error (ARE) $|v_0(\hat{\mathbf{\alpha}},\msx)-v_0(\mathbf{\alpha}^*,\msx)|_\msv/|v_0(\mathbf{\alpha}^*,\msx)|_\msv$ on 50 repetitions.
-This procedure is performed for any $m_{s},t_{a},s_{a}\in \{10,50,100\}\times \{5,10,15,20\}^2 $.
- Mean, standard deviation, and maximum of the ARE on all these hyperparameters choices are respectively $\mathbf{0.10, 0.03, 0.17}$.
- Therefore, the estimation procedure \eqref{eq:relaxation} offers a good approximation of the true parameter when the kernel $K_\msg$ is well specified.
- We observe that the estimation is difficult when $t_a\ll s_a$ because the time series can be very noisy as illustrated in \Cref{fig:exemple_synthetic}: this impacts the Varifold loss which is sensitive to tangents.
-
- Secondly, we demonstrate a weak identifiability when the kernel $K_\msg$ is misspecified: we can reconstruct the graph time series' after deformations even if the hyperparameters of $K_\msg$
-  are different during the generation and the estimation.
-   The hyperparameters of $K_\msg$ during generation are $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)=(1,0.1,100,1,1)$ and we fix $\sigma_{T,1},c_0,c_1=(1,1,0.1) $ for $K_\msg$ during estimation.
-   We aim to understand the impact of $\sigma_{T,1},\sigma_x$ on the reconstruction since they are encoding the smoothness of the transformation according to time and space.
-
-    For any choice of the hyperparameters $\sigma_{T,1},\sigma_x\in \{1,10,50,100,200,300 \}\times \{0.1,1,100\}$ related to $K_\msg$ in the estimation,
-     we average $\scrl(\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx,\varphi^{\{\hat{v}_0\}}.\msx)$ on 50 repetitions when $\mathbf{\alpha}^*$ is sampled according to Gen$(10,10,50)$ and $\hat{v}_0=v_0(\hat{\alpha},\msx)$ denoting by $\hat{\alpha}$ the result of the minimization \eqref{eq:relaxation}.
-  We observe in \Cref{table:synthetic2} that the reconstruction is almost perfect except in the case when $\sigma_{t,0}=1$ during estimation, while $ \sigma_{t,0}=100$ during generation.
-   Compared to $\sigma_{T,0}$, $\sigma_x$ has nearly no impact on the reconstruction.
-   In \Cref{appendix:kernel_implementation}-\ref{appendix:kernel_TS_LDDMM}, we propose guidelines to drive future hyperparameters tuning and further discussions related to $\sigma_{T,1},c_0,c_1$. 
-\subsection{Qualitative analysis of respiratory behavior in mice}
+\subsection{Analysis of respiratory behavior in mice}
 %We consider a dataset composed of mouse respiratory time series before and after a drug injection.

### 2024-05-21 NEURIPS24/paper.tex
update remarque laurent

@@
 %k shape, shapelets , ye2009time
-%\thi{à reformuler pour faire moins stupide}
 
-%\sam{Aller droit au but, se servir de la section related works pour détailler, notamment enlever le premier paragraphe dire directement que l'on veut faire du unsupervised represnetation learning for time series mais en s'intéressant par ticulièrement à la notion de shape.
-%Parler du contexte biomédical, définir la shape. Parler de shape analysis, introduire notre problème de représentation, dire que nous on propose une structure adapté aux time series.}
+Our goal is to analyze the inter-individual variability within a time series dataset, an approach of significant interest in physiological contexts \cite{guscelli2019importance, wang2016research, bar2012studying, germain2023unsupervised}.
+ Specifically, we aim to develop an unsupervised feature representation method that encodes the specificities of individual time series in comparison to a reference time series.
+In physiology, examining the various "shapes" in a time series related to biological phenomena and their variations due to individual differences or pathological conditions is common.
+ However, the term "shape" lacks a precise definition and is more intuitively understood as the silhouette of a pattern in a time series.
+In this paper, we refer to the shape of a time series as the graph of this signal.
 
-% In this paper, we propose an unsupervised representation learning method (URL) for time series \cite{trirat2024universal,meng2023unsupervised} by giving a special attention to their \textit{shapes} for studying inter-variability in biomedical applications.
-% In particular, the method handle time series with \textbf{irregular sampling} and \textbf{different sizes}.
-% This work belongs to the shape analysis literrature: %which focus on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
-%  we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) \cite{beg2005computing,vaillant2004statistics} to the case of time series data by applying difformations to their graph.
-% However, time series graph have more structure than 1D curves:
-%  a graph time series transformation by a general diffeomorphism is not always a graph time series.
-% This work is at the crossroad of URL for time series and shape analysis; we hope it will benefit to both communities.
-%A time series is a set of values indexed by time \cite{trirat2024universal}.
-%This type of data is prevalent in biology and medicine for recording physiological time series , as well as in the broader context of the Internet of Things sensors \cite{fathy2018large} and measurements in physical systems \cite{luo2021deep}.
- %However, due to the nature of such data, which have different sample sizes and incorporate a temporal evolution, classical statistical analysis, and machine learning methods need refinement \cite{anderson2011statistical}.
- %One option to tackle this issue is to derive a feature representation of time series, which depends on the problem at hand \cite{trirat2024universal}.
+ Although community structures with representatives can be learned in an unsupervised manner \cite{trirat2024universal, meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised, tonekaboni2021unsupervised, meng2023unsupervised} or similarity measures \cite{asgari2023clustering, germain2023unsupervised, paparrizos2015k, ye2009time},
+  the study of inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies, shirato2023identifying} remains an open problem in unsupervised representation learning (URL), particularly for \textit{irregularly sampled} time series with \textit{variable lengths}.
 
- Our goal is to analyze the inter-individual variability within a time series dataset, an approach of prime interest in physiological contexts \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
-  More specifically, we aim to find an unsupervised features representation method that encodes individual time series specificities compared to a reference one.
-In physiology, studying the different \textit{shapes} in a time series related to biological phenomena and their variations according to individual or pathology is common.
-However, a \textit{shape} has no clear definition; it is more an intuitive way to speak about the silhouette of a pattern in a time series.
-In this paper, we refer to as the shape of a time series, the graph of this signal.
-
-Although a community structure with representatives can be learned in an unsupervised way \cite{trirat2024universal,meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised} or similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},
-studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem in URL and even more for \textit{irregularly sampled} time series with \textit{variable lengths}.
-     %In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.
-  %For instance, the shape of a heartbeat in an Electrocardiogram (ECG) -the pattern- has been well studied \cite{madona2021pqrst} with the PQRST decomposition (see figure\sam{fill}).
-  %There exist several methods to detect and segment patterns in a time series (\sam{mettre ref thibuat} ).
-   %The goal is then to study their shape according to the property of the individual.
-
-  %  In most clinical studies, depending on expert knowledge, people inspect hand-crafted features such as the mean, wavelets decomposition, or temporal correlation \cite{anderson2011statistical} to keep interpretability.
-  %  However, this approach is time-consuming and potentially biased by priors, especially in exploratory research.
-  %  Given the success of neural networks \cite{bishop1994neural}, features can now be automatically extracted, depending only on the data, by choosing an objective to optimize and a neural architecture tailored to the problem \cite{ismail2019deep,blazquez2021review,torres2021deep}, even in an unsupervised way \cite{meng2023unsupervised}.
-  %  Moreover, depending on the method, it is still possible to have an interpretability of the features.
-  %   For instance, among individuals, a contrastive clustering method can be applied to spot community structures \cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
-  %   More generally, interpretability \cite{yuan2019wave2vec,zhang2018visual,ismail2020benchmarking} is possible by visualizing the shapes of representative patterns to which the model is sensitive.
-  %   Other methods, based on similarity measures \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k}, also propose to perform clustering on shapes in time series or to find some representatives \cite{ye2009time}.
-
-    %Although a community structure with representatives can be learned in an unsupervised way,
-    % studying the inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies,shirato2023identifying} is still an open problem.
-     %In this paper, we aim to go beyond group structure by giving an unsupervised data-driven method that represents the specificity of each individual's shape \cite{vidal2021perceptual}.% \thi{In an individual cluster ?}
-
-% Alternatively, one can represent time series with an embedding that can be processed using conventional tools.
-% or we can represent the time series with an embedding which can be treated by classical tools.
-
-% Depending on the tasks, e.g classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep} or clustering , a meaningful vector representation of time series can be derived.
-%  Initially, in this research, handcrafted features were computed from time series, such as the mean, wavelets \cite{percival2000wavelet}, or Fourier coefficients \cite{bloomfield2004fourier}, depending on expert knowledge.
-%   However, this approach is time-consuming and potentially biased by priors.
-%   Observing the success of neural networks, features are automatically extracted by choosing an objective to optimize and a neural architecture tailored to the problem.
-%    Taking a set of time series as input, features can be identified by solving specific supervised tasks—classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep}— or in an unsupervised manner by deriving labels from the inputs \cite{meng2023unsupervised}.
-%    \thi{Rajouter une image, et une phrase par rapport au temps}
-%    Although the deep learning approach is particularly effective for supervised tasks, annotated datasets are rare, making unsupervised methods more practical.
-%    These methods have proven effective in identifying different patterns in the data and providing a vector representation that segregates well into clusters \cite{meng2023unsupervised}.
-%     However, they may lack precision on small and homogeneous datasets with only one typical shape, often arising from clinical experiments addressing a specific phenomenon (refer to figures).
-%      While current unsupervised methods separate representations into clusters, our focus is on the inter-individual variability within one cluster.
-%       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.
-
-%This type of question has already been addressed in shape analysis \cite{vaillant2004statistics}, a community focusing on the statistical analysis of various mathematical objects invariant under rotations, dilations, or time parameterization.
-%However, time series data with their specific structure has not been properly adressed.
-Our work focuses explicitly on learning shape-based representation of time series.
-First, we propose not to see the shape of a time series through its curve $\{s_t:\eqsp t\in\msi\}$, but rather through its graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\}$.
-   Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
-The idea is to represent each element $\msg(s^j)$ of a dataset $(s^j)_{j\in[N]}$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism $\phi_j$, i.e.
+Our work explicitly focuses on learning shape-based representation of time series.
+First, we propose to view the shape of a time series not merely as its curve $\{s_t:\eqsp t\in\msi\}$, but as its graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\}$.
+   Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we adopt the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
+    The core idea is to represent each element $\msg(s^j)$ of a dataset $(s^j)_{j\in[N]}$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism $\phi_j$, i.e.
 $\msg(s^j) \sim \phi_j .
@@
      The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
-Finally, these shape-encoding features can feed any statistical or machine-learning model.
-%  First, we propose not to see time series through their curve $\{s_t:\eqsp t\in\msi\}$, but through their graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\} $. 
-%  Then, building on the shape analysis litterature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
-%   The idea is to represent each element $(\msg(s^j))_{j\in[N]}$ of the dataset as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism.
-%   Then, the diffeomorphism is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
-%    The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
-%     Finally, these features encoding the shapes can feed any statistical or machine learning model as in URL.
+     Finally, these shape-encoding features can be used as inputs to any statistical or machine-learning model.
 
@@
 However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
-      Closely related, \cite{qiu2009time} tracks anatomical shape changes in serial images using LDDMM but distinguishes from us by including the temporal evolution at a higher level: the goal is to perform longitudinal data modeling.
+     In a similar vein, Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
+They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
 
-       Leaving the LDDMM representation, \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
+       Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
         However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment.
@@
   Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied succesfully Neural SDEs to irregular time series.
-  Oh et al. \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
+  Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
+  With TS-LDDMM, the representation is also driven by an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm.
 

### 2024-05-21 NEURIPS24/paper.tex
update

@@
   Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
-  With TS-LDDMM, the representation is also driven by an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm.
+  With TS-LDDMM, the representation is also derived from an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm, which enhance interpretability.
 

### 2024-05-21 NEURIPS24/experiment.tex
restore bis

@@
 
-For consiceness, several experiments are relagated in Appendix:
+For consiceness, several experiments are relagated in appendix:
 
 \begin{enumerate}
-    \item \textbf{TS-LDDMM representation identifiability, \Cref{appendix:identifiability}:} Results show that TS-LDDMM representations are identifiable or weakly identifiable depending on the specification of the velocity field kernel $K_G$.
-    \item \textbf{Classification benchmark on regulary sampled datasets, \Cref{appendix:classification_shape_analysis}:} Results shows that methods based on TS-LDDMM representation outperform other methods from shape analysis on classification of 15 shape-based datasets (7 univariates \& 8 multivariates).
-    \item \textbf{Robustness to irregular sampling} We compared the robustness of TS-LDDMM representation with 9 URL methods handling irregularly sampled multivariate time series on 15 shape-based datasets (7 univariates \& 8 multivariates).
-We assess methods' classification performances under regular sampling (0\% missing rate) and three irregular sampling regimes (30\%,50\%, and 70\% missing rates), according to the protocol depicted in \cite{kidger2020neural}.
-\textbf{Results} show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%,30\%, and 50\%.
-The performance decrease of the Support Vector Machine classifier (SVC) based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.
+    \item \textbf{TS-LDDMM representation identifiability, \Cref{appendix:identifiability}:} On synthetic data, we evaluate the ability of our method to retrieve the parameter $v_0^*$ that encodes the deformation $\varphi^{\{v_0^*\}}$ acting on a time series graph $\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation} between $\msg$ and $\varphi^{\{v_0^*\}}.\msg$.
+\textbf{Results} show that TS-LDDMM representations are identifiable or weakly identifiable depending on the velocity field kernel $K_G$ specification.
+
+    \item \textbf{Robustness to irregular sampling} We compare the robustness of TS-LDDMM representation with 9 URL methods handling irregularly sampled multivariate time series on 15 shape-based datasets (7 univariates \& 8 multivariates).
+We assess methods' classification performances under regular sampling (0\% missing rate) and three irregular sampling regimes (30\%, 50\%, and 70\% missing rates), according to the protocol depicted in \cite{kidger2020neural}.
+\textbf{Results} show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%, 30\%, and 50\%.
+The performance decrease of the Kernel-SVC based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.
+
+    \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix:classification_shape_analysis}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates).
+\textbf{Results} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.
 \end{enumerate}
 
-experiment on r.  As well, comparison of TS-LDDMM representation with other shape-analysis methods on a classification task is relagated in \Cref{appendix:classification_shape_analysis}.
-
-First, we show on synthetic data that the proposed representation is identifiable
- Secondly, we illustrate the qualitative interest of TS-LDDMM in studying inter-individual variability on a clinical dataset.
-  Thirdly, we demonstrate the quantitative performance of our representation by performing classification on shape-based datasets.
-  The method is implemented on Python using the library JAX\footnote{https://github.com/google/jax}.
-The code was compiled on a server with NVIDIA RTX A2000 12GB GPU, Intel(R) Xeon(R) Gold 5220R CPU @ 2.20GHz, and 250 GB of RAM.
-The code will be available on Github.
-
-\subsection{Analysis of respiratory behavior in mice}
+\subsection{Interpretability: analysis of respiratory behavior in mice}
 %We consider a dataset composed of mouse respiratory time series before and after a drug injection.
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{"./pictures/exp_1_bis.pdf"}
-  \vspace{-1.5em}
-  \caption{Analysis of the three principal components (PC) related to the respiratory cycles of the mouse before exposure.
-  In Figure A), the densities of each genotype according to each PC are displayed.
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/ts_ddmm_shooting.pdf}
+    \caption{TS-LDDMM}
+    \label{fig:ts-lddmm shooting}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/lddmm_shooting.pdf}
+    \caption{LDDMM}
+    \label{fig:lddmm shooting}
+  \end{subfigure}
+  \caption{Analysis of the two principal components (PC) related to mice' respiratory cycles before exposure for TS-LDDMM \Cref{fig:ts-lddmm shooting}, and LDDMM \Cref{fig:lddmm shooting}.
+  In both cases, I) displays PC densities according to mice genotype and II) displays  deformations of the reference graph $S_0$ along each PC.}
+\end{figure*}
+
+\begin{wrapfigure}{r}{0.50\textwidth}
+  \centering
+  \begin{subfigure}[b]{0.16\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/exp_1_cycle_example.jpeg}
+    \caption{Colq cycle}
+    \label{fig:colq-cycle}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.16\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/exp_1_scatter.jpeg}
+    \caption{PC1 vs PC2}
+    \label{fig: pcs-scatter}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.16\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/exp_1_barycenter.jpeg}
+    \caption{WT reference}
+    \label{fig:wt-reference}
+  \end{subfigure}
+  \caption{\Cref{fig:colq-cycle} is a \textbf{colq} respiratory cycle.
+\Cref*{fig: pcs-scatter} displays individual mouse reference respiratory cycle in the TS-LDDMM PC1-PC2 coordinates.
+\Cref{fig:wt-reference} is a referent respiratory cycle of a \textbf{wt} mouse learned with TS-LDDMM.
+}
+\end{wrapfigure}
+
+%the densities of each genotype according to each PC are displayed.
+In Figure B), the deformations of the reference graph $S_0$ along each PC are given.
+In Figure D), the graph of reference $S^j$, also called barycenter, related to each mouse, is displayed according to their coordinates on PC1 and PC3.
+In Figure C) et E), illustrations of respiratory cycles related to mice coming from the \textbf{wt} and \textbf{colq} group are displayed.
+
+%\begin{figure*}[t]
+%  \centering
+%  \includegraphics[width=0.95\linewidth]{"./pictures/exp_1_bis.pdf"}
+%  \vspace{-1.5em}
+%  \caption{Analysis of the two principal components (PC) related to the mice respiratory cycles before exposure for TS-LDDMM \Cref{fig:ts-lddmm shooting}.
+%  In Figure A), the densities of each genotype according to each PC are displayed.
 In Figure B), the deformations of the reference graph $S_0$ along each PC are given.
@@
 }
-  \label{fig:exp_1_PCA}
-  %\vspace{-1em}
-\end{figure*}
+%  \label{fig:exp_1_PCA}
+%  %\vspace{-1em}
+%\end{figure*}
 
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=0.95\linewidth]{"./pictures/exp_2.pdf"}
-  \vspace{-1em}
-  \caption{Analysis of the first Principal Component (PC1) related to the respiratory cycles of the mouse 
-  before and after exposure.
-In Figure A), the densities of each genotype according are displayed.
-In Figure B), the deformations of the reference graph $S_0$ PC1 is given.
-In Figure C), respiratory cycles displayed with respect to time and according to their coordinates on PC1 and PC2}
-  \label{fig:exp_2_PCA}
-  \vspace{-1.5em}
-\end{figure*}
+This experiment highlights the \textit{interpretability} of TS-LDDMM representation for studying the inter-individual variability in time series biomedical datasets.
+We consider a multivariate time series dataset that accounts for mice's nasal and thoracic airflow evolution when exposed to an irritant molecule altering respiratory functions \cite{nervo2019respiratory}.
+The dataset is divided into two groups, one composed of 7 control mice (\textbf{wt}) and the other of 7 mice (\textbf{colq}) deficient in an enzyme involved in the control of respiration.
+For each mouse, airflows were recorded for 15 to 20 minutes before exposure to the irritant molecule and then for 35 to 40 minutes.
+A complete
+description of the dataset is given in the \Cref{appendix:mouse_dataset}.
 
-This experiment highlights the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-We consider a time series dataset recording the evolution of the respiratory airflow of mice exposed 
-to an irritant molecule altering respiratory functions \cite{nervo2019respiratory}. The dataset is divided into two groups, one 
-composed of 7 control mice (\textbf{wt}) and the other of 7 mice (\textbf{colq}) deficient in an enzyme 
-involved in the control of respiration. For each mouse, the respiratory airflow was recorded for 
-15 to 20 minutes before exposure to the irritant molecule and then for 35 to 40 minutes. A complete 
-description of the dataset is given in the \Cref{appendix:mouse_dataset}.
-By comparing the shape of individual respiratory cycles (inspiration + expiration, see \Cref{fig:exp_1_PCA}-C)), 
-we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution 
-after exposure to the irritant molecule.
+\paragraph{Result summary:} By comparing the shape of individual respiratory cycles (inspiration + expiration, see \Cref{fig:exp_1_PCA}-C), we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule.
 
-We first compare breathing behaviors before exposure.
+\paragraph{Breathing behaviors before exposure} We first compare breathing behaviors before exposure.
 Solving \eqref{eq:general_optimization_problem}, we derive the reference respiratory cycle's graph $S_0$ and the TS-LDDMM features representations 
@@
 %In Figure \Cref{fig:exp_1_PCA}, we display coordinates densities per genotype and PC.
-
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=0.95\linewidth]{"./pictures/exp2.pdf"}
+  \caption{Analysis of the first Principal Component (PC1) related to mice' respiratory cycles before and after exposure for TS-LDDMM.
+Figure A) displays PC densities according to mice genotype, Figure B) displays  deformations of the reference graph $S_0$ along PC1, and Figure C) displays respiratory cycles with respect to time in PC1 and PC3 coordinates}
+  \label{fig:exp_2_PCA}
+  \vspace{-1.5em}
+\end{figure*}
 We perform a second experiment to analyze the evolution of breathing behaviors when mice are exposed to the irritant 

### 2024-05-22 NEURIPS24/paper.tex
restore bis
add Samuel update
update

@@
 
-In this part, there is no novelty, we simply expose how to
+In this part, we expose how to
  learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.

### 2024-05-22 NEURIPS24/experiment.tex
experiment all

@@
   In both cases, I) displays PC densities according to mice genotype and II) displays  deformations of the reference graph $S_0$ along each PC.}
+  \label{fig:exp1}
 \end{figure*}
@@
     \includegraphics[width = \textwidth]{pictures/exp_1_cycle_example.jpeg}
-    \caption{Colq cycle}
+    \caption{colq cycle}
     \label{fig:colq-cycle}
@@
     \centering
-    \includegraphics[width = \textwidth]{pictures/exp_1_barycenter.jpeg}
-    \caption{WT reference}
+    \includegraphics[width = \textwidth]{pictures/exp_1_wt.jpeg}
+    \caption{wt cycle}
     \label{fig:wt-reference}
@@
 }
+  \vspace{-1em}
 \end{wrapfigure}
@@
 
-\paragraph{Result summary:} By comparing the shape of individual respiratory cycles (inspiration + expiration, see \Cref{fig:exp_1_PCA}-C), we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule.
+\paragraph{Result summary.} By comparing the shape of individual respiratory cycles, we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule.
 
-\paragraph{Breathing behaviors before exposure} We first compare breathing behaviors before exposure.
-Solving \eqref{eq:general_optimization_problem}, we derive the reference respiratory cycle's graph $S_0$ and the TS-LDDMM features representations 
-$(\alpha_j)_{j\in[N_1]}$ related to $N_1=700$ respiratory cycles extracted according 
-to the procedure \cite{germain2023unsupervised}.
-Then, we perform a kernel PCA on the initial velocity field $\left(v_0(\alpha_j,S_0)\right)_{j\in[N_1]}\in \msv^{N_1}$ defined in \eqref{eq:def_v0}.
-In \Cref{fig:exp_1_PCA}, we focus on the analysis of the three Principal Components (PC).
+\paragraph{Protocol.} For both experiments and representations (TS-LDDMM and LDDMM), we derive the reference respiratory cycle's graph $S_0$ and the representations $(\alpha_j)_{j\in[N]}$ related to $N$ respiratory cycles extracted according to the procedure \cite{germain2023unsupervised} by solving  \eqref{eq:general_optimization_problem}.
+Then, we perform a kernel PCA on the initial velocity field $\left(v_0(\alpha_j,S_0)\right)_{j\in[N]}\in \msv^{N_1}$ defined in \eqref{eq:def_v0} to breathing behavior.
+The first experiment includes $N_1 = 700$ respiratory cycles collected before exposure.
+The second experiment includes $N_2 = 1400$ respiratory cycles with 25\% (resp.
+75\%) before (resp.
+after) exposure.
+\Cref{appendix: mice_exp_setting} describes methods settings.
+Essentially, varifold losses are identical for both methods, and the velocity field kernels are set to encompass time and space scales.
 
-As observable from \Cref{fig:exp_1_PCA}-B), principal components refer to different types of deformations. 
-By interpreting \Cref{fig:exp_1_PCA}-B): Only PC1 accounts for time warping, PC2 expresses the trade-off between inspiration and expiration duration, and PC3 corresponds to a change in signal amplitude.
- Compared to \textbf{wt} mice, the distribution of \textbf{colq} mice 
-TS-LDMMM feature representation along the PC1 axis has a heavy tail and the associated deformation (+3 $\sigma_{\text{PC}}$) 
-shows an inspiration with two peaks. As illustrated in \Cref{fig:exp_1_PCA}-A), such respiratory cycles are preponderant
- with \textbf{colq} mice and may be caused by motor impairment due to their enzyme deficiency, \cite{germain2023unsupervised}.
-  In addition, 
- the \textbf{colq} mice were smaller than the \textbf{wt} mice due to a delay in growth caused by their lack of an enzyme. 
- This difference can be seen on PC3 since the volumes of air (area under the curve) inspired and exhaled are
-  smaller for the smaller mice. In correlation, the distribution of \textbf{wt} mice TS-LDDMM feature representations 
-  along the PC3 axis have a heavy tail corresponding to large air volume as depicted by the deformation 
-  (+3 $\sigma_{\text{PC}}$) in \Cref{fig:exp_1_PCA}-B). Finally, \Cref{fig:exp_1_PCA}-D) shows that PC1 and PC3 capture the main differences between
-   the two groups as their respective reference graphs $S^j$ are located in different parts of the space. 
-%In Figure \Cref{fig:exp_1_PCA}, we display coordinates densities per genotype and PC.
+\paragraph{Breathing behaviors before exposure.} We focus on the analysis of the two first Principal Components (PC) for TS-LDDMM (\Cref{fig:ts-lddmm shooting}), and LDDMM (\Cref{fig:lddmm shooting}).
+A respiratory cycle is an inspiration (positive flow) followed by an expiration (negative flow); \Cref{fig:wt-reference} displays an example of \textbf{wt} mouse respiratory cycle.
+\Cref{fig:exp1} shows that principal components learned with TS-LDMM lead to deformations that remain respiratory cycles, while deformations learned with LDDMM are challenging to interpret as respiratory cycles.
+The LDDMM velocity field kernel is a Gaussian anisotropic kernel that accounts for time and space scales; however, the entanglement of time and space dimensions in the kernel does not guarantee the graph structure, and it makes the convergence of the method difficult (relative varifold loss error: TS-LDDMM: 0.06, LDDMM: 0.11).
+Regarding TS-LDDMM, \Cref{fig:ts-lddmm shooting} shows its principal components refer to different types of deformations.
+By interpreting \Cref{fig:ts-lddmm shooting}: PC1 accounts for time warping, PC2 expresses the trade-off between inspiration and expiration duration.
+Compared to \textbf{wt} mice, the distribution of \textbf{colq} mice feature along the PC1 axis has a heavy left tail, and the associated deformation (-2 $\sigma_{\text{PC}}$) shows an inspiration with two peaks.
+\Cref{fig:colq-cycle} shows an example of such respiratory cycles, which may be caused by motor impairment due to enzyme deficiency, \cite{germain2023unsupervised}.
+Finally, \Cref{fig: pcs-scatter} shows that PC1 and PC2 capture the main differences between the two groups as their respective reference graphs $S^j$ are located in different parts of the space.
+
 \begin{figure*}[t]
@@
 \end{figure*}
-We perform a second experiment to analyze the evolution of breathing behaviors when mice are exposed to the irritant 
-molecule. We follow the same procedure as before. However, we take $N_2=1400$ with 25\% (resp. 75\%) 
-before (resp. after) exposure. In \Cref{fig:exp_2_PCA}, we focus on the first principal component PC since
-it encodes the effect of the irritant molecule as depicted in \Cref{fig:exp_2_PCA}-C) (the exposure occurs at 20 minutes).
-\Cref{fig:exp_2_PCA}-B) shows that the deformation (+3 $\sigma_{\text{PC}}$) leads to longer respiratory cycles that include pauses,
-as observed in \cite{germain2023unsupervised}. As well, \Cref{fig:exp_2_PCA}-A) shows that TS-LDDMM features distributions are less spread 
-out for \textbf{colq} mice compared to \textbf{wt} mice. Indeed, the irritant molecule inhibits the action of the 
-deficient enzyme, \textbf{wt} mice strongly react to the irritant molecule, whereas \textbf{colq} mice are better 
-adapted due to their deficiency.
+
+\paragraph{Breathing behaviors' evolution after exposure to the irritant molecule} 
+TS-LDDMM representation features are learned on a dataset that includes respiratory cycles before and after exposure.
+\Cref{fig:exp_2_PCA} displays the first principal component PC since it encodes the effect of the irritant molecule as depicted in \Cref{fig:exp_2_PCA}-C) (the exposure occurs at 20 minutes).
+\Cref{fig:exp_2_PCA}-B) shows that the deformation (-1.5 $\sigma_{\text{PC}}$) leads to longer respiratory cycles that include a pause between inspiration and expiration as observed in \cite{germain2023unsupervised}.
+Conjointly, \Cref{fig:exp_2_PCA}-A) shows a bimodal distribution for \textbf{wt} mice whereas \textbf{wt}' distributions before exposure were unimodal (\Cref{fig:ts-lddmm shooting}). Indeed, the irritant molecule inhibits the action of the deficient enzyme, \textbf{wt} mice strongly react to the irritant molecule, whereas \textbf{colq} mice better adapt due to their deficiency \cite{germain2023unsupervised}.
 %Secondly, we compare breathing behaviors before and after exposure to observe the impact of the irritant molecule.
@@
 %  Additionally, we learn each mouse's reference respiratory cycle and represent them in the first and third PC coordinates system in Figure REF.
-
-\subsection{Quantitative performances of the TS-LDDMM representation in classification}
-Combined with a Support Vector Classifier (SVC) \cite{hsu2003practical}, TS-LDDMM representation can be used for 
-classification tasks using the kernel associated with the initial velocity space $\msv$.
-We compare TS-LDDMM-SVC classification performances with another SVC using representation 
-learned with T-loss \cite{franceschi2019unsupervised}, an unsupervised deep learning feature 
-representation method for time series. We also include fully supervised methods in deep learning 
--ResNet, CNN \cite{ismail2019deep}- and machine learning: Catch22 \cite{lubba2019catch22}, Rocket
-\cite{dempster2020rocket}, Dynamic Time Wrapping k-Nearest Neighbors (DTW-kNN) 
-\cite{muller2007dynamic}. Methods are compared using f1-score on several shape-based UCR/UEA datasets 
-\cite{dau2019ucr,bagnall2018uea} introduced in \Cref{appendix:classification_dataset}. All implementation details are given 
-in \Cref{appendix:classification_implementation}.
-\Cref{table:classification} presents the reuslts. TS-LDDMM-SVC consistently outperforms the other unsupervised methods. It is ranked 1,3,4,3 for all methods combined, 
-demonstrating its competitiveness as an unsupervised method on time series dataset homogeneous regarding shape.
-%TS-LDDMM representation can be used in a Support Vector Classifier (SVC) \cite{hsu2003practical} using the kernel of the space of initial velocity $\msv$.
-%We compare its capacity of representation to an SVC using T-Loss representation \cite{franceschi2019unsupervised}, which is an unsupervised representation deep learning method for time series,
-%and to others fully supervised methods using deep learning -ResNet, CNN \cite{ismail2019deep}- or not: Catch22 \cite{lubba2019catch22}, Rocket \cite{dempster2020rocket}, Dynamic Time Wrapping \cite{muller2007dynamic} k-Nearest Neighbors (DTW-kNN).
-%These methods are compared using f1-score on several datasets introduced in \Cref{appendix:classification_dataset}.
-All details of implementations are given in \Cref{appendix:classification_implementation}.
-%TS-LDDMM-SVC is always ranked 2 or 3 behind a fully supervised method, demonstrating its competitiveness as an unsupervised method.
-%The more homogeneous the shapes in the dataset are, the better TS-LDDMM-SVC performs.
-
-\begin{table}[t]
-  \vspace{-0.5em}
-  \caption{Classification results in f1-score (U: unsupervised, S: supervised, DL: deep learning, ML: machine learning). \textbf{x} best unsupervised method, \underline{x} best supervised method.}
-  \resizebox{\linewidth}{!}{%
-  \begin{tabular}{lllrrrr}
-    \toprule
-     &  &  & ArrowHead & ECG200 & GunPoint & NATOPS \\
-    \midrule
-    \multirow[m]{3}{*}{U} & \multirow[t]{3}{*}{} & TS-LDDMM-SVC & \textbf{0.84} & \textbf{0.82} & \textbf{0.94} & \textbf{0.93} \\
-     &  & T-loss-SVC & 0.57 & 0.76 & 0.82 & 0.88 \\
-     &  & DTW-kNN & 0.70 & 0.75 & 0.91 & 0.88 \\
-    \cline{1-7} 
-    \multirow[m]{5}{*}{S} & \multirow[m]{2}{*}{DL} & CNN & 0.70 & 0.79 & 0.85 & \underline{0.96} \\
-     &  & ResNet & 0.77 & 0.87 & 0.97 & 0.95 \\
-    \cline{2-7}
-     & \multirow[m]{2}{*}{ML} & Catch22 & 0.73 & 0.81 & 0.96 & 0.89 \\
-     &  & Rocket & \underline{0.81} & \underline{0.91} & \underline{1.00} & 0.88 \\
-    \bottomrule
-    \end{tabular}
-    \label{table:classification}
-  %
-}
-\vspace{-2em}
-  \end{table}

### 2024-05-22 NEURIPS24/experiment.tex
update sam

@@
 \section{Experiments}
+\vspace{-1ex}
 \label{section:experiments}
@@
       \textbf{Results} show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%, 30\%, and 50\%.
-The performance decrease of the Kernel-SVC based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.
+      {\color{red} J'enlèverai cette phrase ou je développerai en Appendix : The performance decrease of the Kernel-SVC based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.}
 
@@
 
+\vspace{-1ex}
 \subsection{Interpretability: analysis of respiratory behavior in mice}
+\label{section:interpretability}
+\vspace{-1ex}
 %We consider a dataset composed of mouse respiratory time series before and after a drug injection.
@@
 description of the dataset is given in the \Cref{appendix:mouse_dataset}.
-
+\vspace{-1ex}
 \paragraph{Result summary.} By comparing the shape of individual respiratory cycles, we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule.
-
+\vspace{-1ex}
 \paragraph{Protocol.} For both experiments and representations (TS-LDDMM and LDDMM), we derive the reference respiratory cycle's graph $S_0$ and the representations $(\alpha_j)_{j\in[N]}$ related to $N$ respiratory cycles extracted according to the procedure \cite{germain2023unsupervised} by solving  \eqref{eq:general_optimization_problem}.
@@
 
-\paragraph{Breathing behaviors before exposure.} We focus on the analysis of the two first Principal Components (PC) for TS-LDDMM (\Cref{fig:ts-lddmm shooting}), and LDDMM (\Cref{fig:lddmm shooting}).
+    \vspace{-1ex}
+\paragraph{Breathing behaviors before exposure.}
+ We focus on the analysis of the two first Principal Components (PC) for TS-LDDMM (\Cref{fig:ts-lddmm shooting}), and LDDMM (\Cref{fig:lddmm shooting}).
   A respiratory cycle is an inspiration (positive flow) followed by an expiration (negative flow); \Cref{fig:wt-reference} displays an example of \textbf{wt} mouse respiratory cycle.
@@
       Compared to \textbf{wt} mice, the distribution of \textbf{colq} mice feature along the PC1 axis has a heavy left tail, and the associated deformation (-2 $\sigma_{\text{PC}}$) shows an inspiration with two peaks.
-\Cref{fig:colq-cycle} shows an example of such respiratory cycles, which may be caused by motor impairment due to enzyme deficiency, \cite{germain2023unsupervised}.
+       \Cref{fig:colq-cycle} shows an example of such respiratory cycles, which may be caused by motor impairment due to enzyme deficiency \cite{germain2023unsupervised}.
         Finally, \Cref{fig: pcs-scatter} shows that PC1 and PC2 capture the main differences between the two groups as their respective reference graphs $S^j$ are located in different parts of the space.
@@
 \end{figure*}
-
+\vspace{-1ex}
 \paragraph{Breathing behaviors' evolution after exposure to the irritant molecule} 

### 2024-05-22 NEURIPS24/paper.tex
update sam

@@
 % which is parti... -> which is of prime interest in ...
-
+\vspace{-1ex}
 \section{Introduction}
+\vspace{-1ex}
 %https://arxiv.org/abs/1811.00075
@@
     \begin{itemize}
-       \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset.
+       \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset (\Cref{section:methodology}).
 In particular, the method can handle multivariate time series \textit{irregularly sampled} and with \textit{variable sizes}.
        \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters, also providing guidelines for tuning.
-       \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets.
+       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters (\Cref{appendix:identifiability}), also providing guidelines for tuning (\Cref{appendix:kernel_TS_LDDMM}).
+       \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset (\Cref{section:experiments}).
+        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets with regular and irregular sampling (\Cref{appendix:classification_shape_analysis}).
  \end{itemize}
@@
 
-\section{Related Works}
-Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
-   The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
-    Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
-     The goal is to find a well-suited Riemannian structure according to the nature of the studied object.
-
-     LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
-However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
-     In a similar vein, Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
-They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
-
-       Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
-        However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment.
-         Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation.
-          Very recently, in a functional data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution.
-           Nevertheless, this method is made for \textit{continuous objects} and only applies to time series of \textit{same length}, making the estimation more sensitive to noise.
-  %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
-  %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
-% Depending on the tasks, e.g classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep} or clustering , a meaningful vector representation of time series can be derived.
-%  Initially, in this research, handcrafted features were computed from time series, such as the mean, wavelets \cite{percival2000wavelet}, or Fourier coefficients \cite{bloomfield2004fourier}, depending on expert knowledge.
-%   However, this approach is time-consuming and potentially biased by priors.
-%   Observing the success of neural networks, features are automatically extracted by choosing an objective to optimize and a neural architecture tailored to the problem.
-%    Taking a set of time series as input, features can be identified by solving specific supervised tasks—classification \cite{ismail2019deep}, anomaly detection \cite{blazquez2021review}, forecasting \cite{torres2021deep}— or in an unsupervised manner by deriving labels from the inputs \cite{meng2023unsupervised}.
-%    \thi{Rajouter une image, et une phrase par rapport au temps}
-%    Although the deep learning approach is particularly effective for supervised tasks, annotated datasets are rare, making unsupervised methods more practical.
-%    These methods have proven effective in identifying different patterns in the data and providing a vector representation that segregates well into clusters \cite{meng2023unsupervised}.
-%     However, they may lack precision on small and homogeneous datasets with only one typical shape, often arising from clinical experiments addressing a specific phenomenon (refer to figures).
-%      While current unsupervised methods separate representations into clusters, our focus is on the inter-individual variability within one cluster.
-%       Typically, we aim to perform Principal Components Analysis (PCA) \cite{bro2014principal} on individual patterns with similar shapes. The representation is dependant on the input dataset.
-
- %\citet{srivastava2010shape,heo2024logistic} address the representation of curves $C:\msi \to \Rset^d$, $\msi\subset \Rset$, having unitary velocity by using the Square-Root Velocity (SRV) representation.
- %However, the case of time series has not been adequately addressed, considering the special attention the time axis requires.
- %For instance, the SRV representation applies after a reparametrization in time, such that the original time evolution of the time series is not represented in the final features.
- %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
- %Compared to Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} or Continuous Normalizing Flows \cite{chen2018neural,grathwohl2019scalable,salman2018deep} for diffeomorphisms learning, the number of hyperparameters to tune is minimal, and the related optimization problem is well-posed.
- Balancing between discrete and continuous elements is a challenging task.
-  In the deep learning literature \cite{chen2018neural, kidger2020neural, tzen2019neural, jia2019neural, liu2019neural, ansari2023neural}, Neural Ordinary Differential Equations (Neural ODEs) \cite{chen2018neural} learn continuous latent representations using a vector field parameterized by a neural network, serving as a continuous analogue to Residual Networks \cite{zagoruyko2016wide}.
-   This approach was further enhanced by Neural Controlled Differential Equations (Neural CDEs) \cite{kidger2020neural} for handling irregular time series, functioning as continuous-time analogs of RNNs \cite{schuster1997bidirectional}.
- Extending Neural ODEs, Neural Stochastic Differential Equations (Neural SDEs) introduce regularization effects \cite{liu2019neural}, although optimization remains challenging.
-  Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied succesfully Neural SDEs to irregular time series.
-  Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
-  With TS-LDDMM, the representation is also derived from an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm, which enhance interpretability.
-
+  \vspace{-1ex}
 \section{Notations}
@@
 %Reproducing Kernel Hilbert Space (RKHS)
+\vspace{-1ex}
 \section{Background on LDDMM}
+\vspace{-1ex}
 \label{section:LDDMM}
@@
 
+ \vspace{-1ex}
 % LDDMM is designed to analyzed an existing dataset, while NF and CNF are made to generalize a dataset for data augmentation.
@@
 
+\vspace{-1ex}
 \paragraph{Discrete parametrization of diffeomorpshim.}
@@
 %  \end{algorithm}
+\vspace{-1ex}
 \paragraph{Relation to Continuous Normalizing Flows.}
@@
 large dataset in high dimensions.
+\vspace{-1ex}
 \section{Methodology}
+\label{section:methodology}
 %\paragraph{Assumptions}
 %Denoting by $C^0(\msi,\msj) $ the space of continous functions between two sets $\msi,\msj$,
-
+\vspace{-1ex}
 We consider in this paper observations which consist in a population of $N$ multivariate time series, for any $j\in[N]$, $s^j \in \rmC^1(\msi_j,\Rset^{d})$. 
@@
 %  \paragraph{Goals}%\sam{potentiellement changer l'ordre de time and space}
+\vspace{-1ex}
 \paragraph*{The deformation problem.}
@@
      %\sam{Parler de l'unicté?}
+     \vspace{-1ex}
      \paragraph{Optimization related to \eqref{eq:transformation}.}
@@
 Nevertheless, \eqref{eq:general_optimization_problem} ask to define a kernel and a loss in order to perform geodesic shooting \ref{eq:relaxation}, which is the purpose of the next subsection.
+\vspace{-1ex}
    \subsection{Application of LDDMM to time series analysis: TS-LDDMM}
+   \vspace{-1ex}
 %\tcr{Mettre l'applicaiton clair pour les times series}
@@
 Consequently, we introduce the oriented Varifold distance.
-
+\vspace{-1ex}
        \paragraph{A representation separating space and time.}
@@
 
+ \vspace{-1ex}
     \paragraph{Choice for the kernel associated with the RKHS $\msv$}
@@
 %\sam{améliorer transition}
+\vspace{-1ex}
 \paragraph{Loss}
@@
 For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}. 
+{\color{red} mettre une phrase sur la loss multiscale}
 
@@
 
-\sam{Parler de méthode adaptatif ici}
+%\sam{Parler de méthode adaptatif ici}
 
@@
 
-\include{experiment}
+\vspace{-1ex}
+\input{experiment}
+\vspace{-1ex}
 
+\section{Related Works}
+\vspace{-1ex}
+Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
+   The main idea is to represent these objects in a complete Riemannian manifold $(\mathcal{M},\mathbf{g})$ with a metric $\mathbf{g}$ adapted to the geometry of the problem \cite{miller2006geodesic}.
+    Then, any set of points in $\mathcal{M}$ can be represented as points in the tangent space of their Frechet mean $\mathbf{m}_0$ \cite{pal2017riemannian,le2001locating} by considering their logarithms.
+     The goal is to find a well-suited Riemannian structure according to the nature of the studied object.
+
+     LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
+However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
+     In a similar vein, Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
+They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
+
+       Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
+        However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment.
+         Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation.
+          Very recently, in a functional data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution.
+           Nevertheless, this method is made for \textit{continuous objects} and only applies to time series of \textit{same length}, making the estimation more sensitive to noise.
+
+ Balancing between discrete and continuous elements is a challenging task.
+  In the deep learning literature \cite{chen2018neural, kidger2020neural, tzen2019neural, jia2019neural, liu2019neural, ansari2023neural}, Neural Ordinary Differential Equations (Neural ODEs) \cite{chen2018neural} learn continuous latent representations using a vector field parameterized by a neural network, serving as a continuous analogue to Residual Networks \cite{zagoruyko2016wide}.
+   This approach was further enhanced by Neural Controlled Differential Equations (Neural CDEs) \cite{kidger2020neural} for handling irregular time series, functioning as continuous-time analogs of RNNs \cite{schuster1997bidirectional}.
+ Extending Neural ODEs, Neural Stochastic Differential Equations (Neural SDEs) introduce regularization effects \cite{liu2019neural}, although optimization remains challenging.
+  Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied succesfully Neural SDEs to irregular time series.
+  Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
+  With TS-LDDMM, the representation is also derived from an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm, which enhance interpretability.
+
+  All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are comapred to TS-LDDMM in \Cref{appendix:classification_shape_analysis}.
+  \section{Limitations}
+  {\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
 \section{Conclusion}
+\vspace{-1ex}
 In this paper, we propose a feature representation method, TS-LDDMM, designed for 

### 2024-05-22 NEURIPS24/appendix.tex
experiment all
appendix 0

@@
 
-\section{Tuning the hyperparameters of the TS-LDDMM kernel given in \eqref{eq:kernel_TAS}}
+\section{Tuning the hyperparameters of the TS-LDDMM velocity field kernel}
 \label{appendix:kernel_TS_LDDMM}
@@
 
-\subsection{Optimization details of TS-LDDMM~\eqref{eq:general_optimization_problem}}
+\subsection{Optimization details of TS-LDDMM \& LDDMM}
 \label{appendix:optimizers_details}
@@
 
-\subsection{Identifiability experiment}
-This experiement only involves the TS-LDDMM method in two different settings: 
-\begin{itemize}
-  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
-  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$.
-The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
-\end{itemize}
-TS-LDDMM parameters of the experiment are summarized in \Cref{appendix:table:set_up_exp_practical}
-
-\subsection{Mice respiratory behaviors experiment}
+\subsection{Settings for experiment: analysis of respiratory behavior in mice}
+\label{appendix: mice_exp_setting}
 
@@
 \end{itemize}
-\subsection{Benchmark}
 
-A report of all the hyperparameters selected is given in \Cref{appendix:table:set_up_exp_practical}.
+\subsection{Settings for experiment: TS-LDDMM representation identifiability}
+\label{appendix: settings_identifiability}
+This experiment only involves the TS-LDDMM method in two different settings: 
+\begin{itemize}
+  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
+  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$.
+The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
+\end{itemize}
+
+\subsection{Settings for experiment: Robustness to irregular sampling}
+This experiment is inspired from \cite{oh2024stable} where authors performs an extensive comparison of Neural Ordinary Differential Equations (Neural ODEs) methods~\cite{kidger2020neural}.
+Methods, network architectures, and training strategy are taken from its associated Github\footnote{\url{https://github.com/yongkyung-oh/Stable-Neural-SDEs}} and describe in what follows.
+\label{appendix: settings_robustness}
+\subsubsection{Benchmark methods}
+In related work, we give an overview of Neurals ODEs methods and its relation with TS-LDDMM.  
+\begin{itemize}
+  \item RNN-based methods: Baseline reccurent neural networks including RNN~\cite{medsker1999recurrent}, LSTM~\cite{hochreiter1997long}, and GRU~\cite{chung2014empirical}.
+
+  \item Attention-based methods: Multi-Time Attention Networks (MTAN)~\cite{shukla2021multi} and Multi-Integration Attention Module (MIAM)~\cite{lee2022multi}.
+Both handles multivariate time series irregularly sampled with attention mechanisms.
+  \item Neural ODEs: ODE-LSTM~\cite{lechner2020learning} a form of Neural-ODEs used to learned continous latent representations. 
+  \item Neural SDEs:  Neural SDE~\cite{liu2019neural} and Neural LNSDE~\cite{oh2024stable} have been proposed to model randomness in time-series using drift and diffusion terms as an extension of Neural-ODEs.
+  \item Shape-Analysis methods: TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}.
+From shape analysis, both methods learn representations by solving ODEs parametrized with Kernels.
+While both methods hamdle multivariate signal irregularly sampled, TS-LDDMM is specifically desgined for time series.
+\end{itemize}
+
+\subsection{Settings for experiment: Classification benchmark on regularly sampled datasets}
+\label{appendix: settings_classification}
 

### 2024-05-22 NEURIPS24/paper.tex
update

@@
     \end{lemma}
-    \sam{Parler des Cauchy kernel en appndice et du choix de la loss}
+    Instead of Gaussian kernels, other types of smooth kernel can be selected as long as the structure \eqref{eq:kernel_TAS} is respected.
     % \begin{proof}
@@
   \end{equation}
+   Note that the loss kernel $k$ has nothing to do with the velocity field kernel denoted by $K_\msg$ or $K$ specified in \Cref{paragraph:kernel_V}.
+Finally, we define the data fidelity loss function, $\scrl$, as a sum of $ d_{\msw^*}^2$ using different kernel's width parameters $\sigma$ to incorporate multiscale information such that.
+$\scrl$ is indeed differentiable with regards to its first variable.
 The specific kernels $k_{\pos},k_{\dir}$ that we use in our experiments are given \Cref{appendix:kernel_implementation}.
-   Note that the loss kernel $k$ has nothing to do with the velocity field kernel denoted by $K_\msg$ or $K$ specified in \Cref{paragraph:kernel_V}.
-Finally, we define the data fidelity loss function, $\scrl$, as $ d_{\msw^*}^2$, which is differentiable with regards to its first variable.
 For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}.
-{\color{red} mettre une phrase sur la loss multiscale}
 
@@
   \section{Limitations}
-  {\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
+  While TS-LDDMM performs well on shape-based datasets, it assumes a certain degree of homogeneity with the existence of a reference time series graph$\msg_0$.
+  Extending TS-LDDMM to more heterogeneous datasets using clustering is ongoing work.
+   TS-LDDMM employs kernel computations, which require specific libraries (e.g., KeOps \cite{charlier2021kernel}) to be efficient and scalable.
+   However, in our experiments, the time complexity of TS-LDDMM is comparable to that of competitors.
+   It is clear that TS-LDDMM needs to be extended to handle very large datasets with high-dimensional time series (such as videos).
+    Additionally, TS-LDDMM requires tuning several hyperparameters, though this is a common requirement among competitors \cite{glaunes2008large, oh2024stable, wu2024shape, heo2024logistic}.
+    In future work, adaptive methods are expected to be developed to provide a more user-friendly interface.
+  {%\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
 \section{Conclusion}

### 2024-05-22 NEURIPS24/appendix.tex
appendix add

@@
 
-\subsection{Settings for experiment: analysis of respiratory behavior in mice}
-\label{appendix: mice_exp_setting}
-
-This experiment involves TS-LDDMM and LDDMM~\cite{glaunes2008large} methods.
-Both methods are run twice, first on respiratory cycles before exposure to the irritant molecule to capture mice breathing behavior at rest and on all respiratory cycles to capture the influence of the irritant molecule.
-Exposure to the irritant molecule leads to significant shape deformation in the respiratory cycles, and the terms must be added to the varifold loss to capture deformations at a large time scale.
-
-\paragraph{TS-LDDMMM parameters.}
-\begin{itemize}
-  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,150,1,2)$.
-The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
-  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,220,1,2)$.
-The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
-\end{itemize}
-
-\paragraph{LDDMMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM.
-Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
-\begin{itemize}
-  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =150$ for the time dimension and $\sigma_x = 2$ for space dimensions. 
-  The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
-  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =220$ for the time dimension and $\sigma_x = 2$ for space dimensions.
-  The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
-\end{itemize}
-
-\subsection{Settings for experiment: TS-LDDMM representation identifiability}
-\label{appendix: settings_identifiability}
-This experiment only involves the TS-LDDMM method in two different settings: 
-\begin{itemize}
-  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
-  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$.
-The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
-\end{itemize}
-
-\subsection{Settings for experiment: Robustness to irregular sampling}
-This experiment is inspired from \cite{oh2024stable} where authors performs an extensive comparison of Neural Ordinary Differential Equations (Neural ODEs) methods~\cite{kidger2020neural}.
-Methods, network architectures, and training strategy are taken from its associated Github\footnote{\url{https://github.com/yongkyung-oh/Stable-Neural-SDEs}} and describe in what follows.
-\label{appendix: settings_robustness}
-\subsubsection{Benchmark methods}
-In related work, we give an overview of Neurals ODEs methods and its relation with TS-LDDMM.  
-\begin{itemize}
-  \item RNN-based methods: Baseline reccurent neural networks including RNN~\cite{medsker1999recurrent}, LSTM~\cite{hochreiter1997long}, and GRU~\cite{chung2014empirical}.
-
-  \item Attention-based methods: Multi-Time Attention Networks (MTAN)~\cite{shukla2021multi} and Multi-Integration Attention Module (MIAM)~\cite{lee2022multi}.
-Both handles multivariate time series irregularly sampled with attention mechanisms.
-  \item Neural ODEs: ODE-LSTM~\cite{lechner2020learning} a form of Neural-ODEs used to learned continous latent representations. 
-  \item Neural SDEs:  Neural SDE~\cite{liu2019neural} and Neural LNSDE~\cite{oh2024stable} have been proposed to model randomness in time-series using drift and diffusion terms as an extension of Neural-ODEs.
-  \item Shape-Analysis methods: TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}.
-From shape analysis, both methods learn representations by solving ODEs parametrized with Kernels.
-While both methods hamdle multivariate signal irregularly sampled, TS-LDDMM is specifically desgined for time series.
-\end{itemize}
-
-\subsection{Settings for experiment: Classification benchmark on regularly sampled datasets}
-\label{appendix: settings_classification}
-
-\subsection{Synthetic experiments}
-\label{appendix:numerics_synthetic}
-For any deformations generation in both experiments (well-specified and misspecified), we take $\sigma_{T,0},\sigma_{T,1},\sigma_x=(100,1,1)$ and $c_0,c_1=(1,0.1)$ for the kernel $K_\msg$
-and $\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}=(2,1,2,0.6) $ for the varifold kernels $k_\pos,k_\dir$ related to the loss $\scrl$.
-
-In both experiments, we have nb\_steps=300 abd $\eta_M=0.1$.
-
-\subsection{Mouse experiments}
-The number of steps is larger in the second experiment (before/after injection) because the deformations are sharper.
-
-\subsection{Classification experiments}
-\label{appendix:classification_implementation}
-
-We defined a default parametrization for all classifiers.
-
-For classifiers: \href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.deep_learning.CNNClassifier.html#aeon.classification.deep_learning.CNNClassifier}{CNN}, 
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.networks.ResNetNetwork.html#aeon.networks.ResNetNetwork}{ResNet}, 
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.feature_based.Catch22Classifier.html#aeon.classification.feature_based.Catch22Classifier}{Catch22}, 
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.distance_based.KNeighborsTimeSeriesClassifier.html#aeon.classification.distance_based.KNeighborsTimeSeriesClassifier}{DTW-KNN},
-\href{https://www.aeon-toolkit.org/en/stable/api_reference/auto_generated/aeon.classification.convolution_based.RocketClassifier.html#aeon.classification.convolution_based.RocketClassifier}{Rocket}
-we used the aeon\footnote{https://www.aeon-toolkit.org/en/stable/index.html} implementations with their default settings.
-
-For Tloss-SVC we used the implementation provided on github\footnote{https://github.com/mqwfrog/ULTS} 
-with the following parameters for learning representations: batch\_size: 10, channels: 40, depth: 10, nb\_steps: 200, in\_channels: 1, 
-kernel\_size: 3, lr: 0.001, nb\_random\_samples: 10, negative\_penalty: 1, out\_channels: 320, reduced\_size: 160. 
-We used the Support Vector Classifier (\href{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html#sklearn.svm.SVC}{SVC})
-from scikit-learn with thee regularization term C: 1. Others parameters are set to default.
-
-For TS-LDDMMM-SVC, all kernels' parameters et optimizer parameter are presented in \Cref{appendix:table:set_up_exp_practical}.
-As well, we used the Support Vector Classifier from scikit-learn with thee regularization term C: 1. Others parameters are set to default.
-
- \begin{table}[hbt!]
-  \caption{Parameters used in all the experiments.
-   For synthetic data, $K_\msg$ refers to the kernel used in the generation, which is the same for the estimation only in the well-specified case.
-$\bar{l}$ refers to the average time series length and $N_d$ refers to the number of dimensions.}
-  \centering
-  \begin{tabular}{lllll}
-
-            \toprule 
-            objects & Optimizer &$k_\pos,k_\dir$ & $K_\msg $ \\
-            \midrule
-            Parameter & (nb\_steps,$\eta_M$) &$(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$ & $(c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x)$ \\
-            \midrule
-            Synthetic data well-specified &(400,0.1) &$(2,1,2,0.6)$& $(1,0.1,100,1,1)$   \\
-            Synthetic data misspecified  & (400,0.1)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
-            Mouse before injection  & (400,0.3)  & $ (2,1,2,0.6)$ &$(1,0.1,100,1,1)$  \\
-            Mouse before/after injection  & (400,0.3)  & $ (5,1,5,0.6)$ &$(1,0.1,150,1,1)$  \\
-            Classification  & (400,0.1)  & $(2,N_d,2,N_d)$  & $(1,0.1,0.33\bar{l},1,N_d)$  \\
-
-            \bottomrule
-          \end{tabular}
-        \label{appendix:table:set_up_exp_practical}
-    \end{table}
-
 \section{Datasets}
@@
 
-\section{Experiments}
+\section{Appendix for experiment: Analysis of respiratory behavior in mice}
+\subsection{Settings}
+\label{appendix: mice_exp_setting}
 
-\subsection{Identifiability experiments}
-\label{appendix:identifiability}
+This experiment involves TS-LDDMM and LDDMM~\cite{glaunes2008large} methods.
+Both methods are run twice, first on respiratory cycles before exposure to the irritant molecule to capture mice breathing behavior at rest and on all respiratory cycles to capture the influence of the irritant molecule.
+Exposure to the irritant molecule leads to significant shape deformation in the respiratory cycles, and the terms must be added to the varifold loss to capture deformations at a large time scale.
+
+\paragraph{TS-LDDMMM parameters.}
+\begin{itemize}
+  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,150,1,2)$.
+The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
+  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,220,1,2)$.
+The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
+\end{itemize}
+
+\paragraph{LDDMMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM.
+Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
+\begin{itemize}
+  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =150$ for the time dimension and $\sigma_x = 2$ for space dimensions. 
+  The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
+  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =220$ for the time dimension and $\sigma_x = 2$ for space dimensions.
+  The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
+\end{itemize}
+
+\section{Appendix for experiment: TS-LDDMM representation identifiability}
+\subsection{Settings}
+\label{appendix: settings_identifiability}
+This experiment only involves the TS-LDDMM method in two different settings: 
+\begin{itemize}
+  \item \textbf{The velocity field kernel $K_G$ is well-specified:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,100,1,1)$, the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
+  \item \textbf{The velocity field kernel $K_G$ is missspecified:} The velocity field kernel $K_G$ is set with  $(c_0,c_1,\sigma_{T,1}) = (1,0.1,1)$, $\sigma_{T,0}$ ranging in $(1,5,10,50,100,200,300)$, and $\sigma_x$  ranging in $(0.1,1,10,100)$.
+The varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,1,2,0.6)$, and the optimizer has 400 steps with a maximum stepsize $\eta_M$ of 0.05.
+\end{itemize}
 
@@
   \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
-  \vspace{-2.5em}
+  \vspace{-1em}
   \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
@@
 
-% \begin{subfigure}{0.7\linewidth}
-%   \centering
-%   \resizebox{\columnwidth}{!}{%
-%      \begin{tabular}{lrrrr}
-%      \toprule
-%      $\sigma_{T,0} \backslash \sigma_x$  & 0.1 & 1 & 100 \\
-%      \midrule
-%      1 & 2e+0 & 4e-2 & 4e-2 \\
-%      10 & 3e-4 & 1e-4  & 2e-4 \\
-%      50 & 1e-5 & 1e-5  & 1e-5 \\
-%      100 & 4e-6 & 4e-6 & 4e-6 \\
-%      200 & 7e-4 & 7e-4 & 7e-4 \\
-%      300 & 4e-3 & 4e-3 & 4e-3 \\
-%      \bottomrule
-%      \end{tabular}
-%      %
-%      }
-%  \caption{Values of $\scrl(y^*,\varphi^{\{\hat{v}_0\}}.\msx)$ where $y^*$ is sampled according to Gen(10,10,50) and $\hat{v}_0$ is estimated using $K_\msg$ with varying parameters $\sigma_{T,1},\sigma_x$.}
-% \end{subfigure}
-
 provided that the hyperparameters and the reference graph are wisely selected, i.e., the parameter $v_0^*$ generating a deformation $\varphi^{\{v_0^*\}}$ of a time series graph $\msg$ can be estimated from the data $\msg,\varphi^{\{v_0^*\}}.\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation}.
@@
 Note that the regularizing step $(\mathbf{\alpha}_k')_{k\in[\mathbf{n}_0]}\to \bar{\mathbf{\alpha}}' $ is necessary to obtain realistic deformations which take into account the regularity induced by the RKHS $\msv$.
-%  uniformly on the set $\mss_M=\{ (\alpha_k)_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}:\eqsp |\alpha_k|=M\}  $ with $M>0,\mathbf{n}_0\in \Nset_{>0}$,
-%  then we regularize $\mathbf{\alpha}$ by convolving
-%   with ... to get the generation parameter $\mathbf{\alpha}^*$.
+
 Then, using $v_0(\mathbf{\alpha}^*,\msx)$ as defined in \eqref{eq:def_v0} with initial momentums $\mathbf{\alpha}^*$ and control points $\msx$, we apply the induced deformation $\varphi^{\{v_0\}} $ by \eqref{eq:integration} to $\msx$ and obtain $\varphi^{\{v_0\}}.\msx$.
@@
 
-\subsection{Classification: Comparison with shape analysis methods}
-\label{appendix:classification_shape_analysis}
+\section{Appendix for experiment: Robustness to irregular sampling}
+\label{appendix: robustness}
+
+This experiment is inspired by \cite{oh2024stable} where the authors perform an extensive comparison of Neural Ordinary Differential Equations (Neural ODEs) methods~\cite{kidger2020neural}.
+We assess the classification
+performances of several methods under regular sampling (0\% missing rate) and three irregular sampling regimes on 15 shape-based datasets (7 univariate \& 8 multivariate).
+Methods and training strategy are taken from its associated Github\footnote{\url{https://github.com/yongkyung-oh/Stable-Neural-SDEs}} and described in what follows.
+We conclude with the results, which show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%, 30\%, and 50\%.
+
+\subsection{Benchmark methods}
+\label{appendix: benchmark robustness}
+In related work, we give an overview of Neurals ODEs methods and their relation with TS-LDDMM.  
+\begin{itemize}
+  \item RNN-based methods: Baseline reccurent neural networks including RNN~\cite{medsker1999recurrent}, LSTM~\cite{hochreiter1997long}, and GRU~\cite{chung2014empirical}.
+
+  \item Attention-based methods: Multi-Time Attention Networks (MTAN)~\cite{shukla2021multi} and Multi-Integration Attention Module (MIAM)~\cite{lee2022multi}.
+Both handle multivariate time series irregularly sampled with attention mechanisms.
+  \item Neural ODEs: ODE-LSTM~\cite{lechner2020learning} a form of Neural-ODEs used to learn continuous latent representations. 
+  \item Neural SDEs:  Neural SDE~\cite{liu2019neural} and Neural LNSDE~\cite{oh2024stable} have been proposed to model randomness in time-series using drift and diffusion terms as an extension of Neural-ODEs.
+  \item Shape-Analysis methods: TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}.
+From shape analysis, both methods learn representations by solving ODEs parametrized with Kernels.
+While both methods handle multivariate signals irregularly sampled, TS-LDDMM is specifically designed for time series.
+\end{itemize}
+
+\subsection{Model architecture}
+
+\paragraph{Neural ODEs methods}
+As depicted in~\cite{oh2024stable}, any Neural ODEs layer in \Cref{appendix: benchmark robustness} is followed by an MLP with two fully connected layers with \texttt{ReLU} activations.
+The risk of overfitting and the model regularization are handled with a dropout rate of 10\% and an early-stopping mechanism, ceasing the training when the validation loss does not improve for 10 successive epochs.
+
+\paragraph{TS-LDDMM and LDDMM}
+
+\subsection{Protocol}
+
+\subsection{Results}
+
+\begin{table}[hbt!]
+  \centering
+  \begin{tabular}{lcccc}
+    \toprule
+    \multirow[c]{2}{*}{\textbf{Methods}} & \multicolumn{4}{c}{\textbf{Test F1-score}} \\
+    \cline{2-5}
+     &  \textbf{Regular} & \textbf{30 \% dropped} &  \textbf{50 \% dropped} & \textbf{70 \% dropped} \\
+    \midrule
+    RNN (1999) & $0.64 \pm 0.21$ & $0.53 \pm 0.23$ & $0.48 \pm 0.21$ & $0.44 \pm 0.21$ \\
+    LSTM (1997) & $0.61 \pm 0.29$ & $0.57 \pm 0.29$ & $0.53 \pm 0.25$ & $0.51 \pm 0.29$ \\
+    GRU (2014) & $0.71 \pm 0.26$ & $0.68 \pm 0.28$ & $0.66 \pm 0.28$ & $\underline{0.59 \pm 0.28}$ \\
+    MTAN (2021) & $0.59 \pm 0.28$ & $0.58 \pm 0.28$ & $0.54 \pm 0.29$ & $0.51 \pm 0.28$ \\
+    MIAM (2022) & $0.48 \pm 0.35$ & $0.42 \pm 0.33$ & $0.47 \pm 0.31$ & $0.35 \pm 0.31$ \\
+    ODE-LSTM (2020) & $0.63 \pm 0.24$ & $0.57 \pm 0.25$ & $0.51 \pm 0.24$ & $0.45 \pm 0.23$ \\
+    Neural SDE (2019) & $0.48 \pm 0.28$ & $0.47 \pm 0.26$ & $0.45 \pm 0.27$ & $0.45 \pm 0.25$ \\
+    Neural LNSDE (2024) & $0.7 \pm 0.27$ & $0.68 \pm 0.29$ & $\underline{0.67 \pm 0.25}$ & $\mathbf{0.66 \pm 0.23}$ \\
+    LDDMM (2008) & $\underline{0.72 \pm 0.2}$ & $\underline{0.7 \pm 0.21}$ & $0.57 \pm 0.25$ & $0.4 \pm 0.25$ \\
+    %\cline{1-5}
+    TS-LDDMM (ours) & $\mathbf{0.83 \pm 0.18}$ & $\mathbf{0.8 \pm 0.18}$ & $\mathbf{0.7 \pm 0.26}$ & $0.51 \pm 0.27$ \\
+    \bottomrule
+  \end{tabular}  
+\end{table}
+
+\subsection{Appendix for experiment: Classification benchmark on regularly sampled datasets}
+\subsection{Settings}
+\label{appendix: settings_classification}
 
@@
 \end{table}
-
-\section{Robustness to missing data}
-
-\begin{table}[hbt!]
-  \centering
-  \begin{tabular}{lcccc}
-    \toprule
-    \multirow[c]{2}{*}{\textbf{Methods}} & \multicolumn{4}{c}{\textbf{Test F1-score}} \\
-    \cline{2-5}
-     &  \textbf{Regular} & \textbf{30 \% dropped} &  \textbf{50 \% dropped} & \textbf{70 \% dropped} \\
-    \midrule
-    RNN (1999) & $0.64 \pm 0.21$ & $0.53 \pm 0.23$ & $0.48 \pm 0.21$ & $0.44 \pm 0.21$ \\
-    LSTM (1997) & $0.61 \pm 0.29$ & $0.57 \pm 0.29$ & $0.53 \pm 0.25$ & $0.51 \pm 0.29$ \\
-    GRU (2014) & $0.71 \pm 0.26$ & $0.68 \pm 0.28$ & $0.66 \pm 0.28$ & $\underline{0.59 \pm 0.28}$ \\
-    MTAN (2021) & $0.59 \pm 0.28$ & $0.58 \pm 0.28$ & $0.54 \pm 0.29$ & $0.51 \pm 0.28$ \\
-    MIAM (2022) & $0.48 \pm 0.35$ & $0.42 \pm 0.33$ & $0.47 \pm 0.31$ & $0.35 \pm 0.31$ \\
-    ODE-LSTM (2020) & $0.63 \pm 0.24$ & $0.57 \pm 0.25$ & $0.51 \pm 0.24$ & $0.45 \pm 0.23$ \\
-    Neural SDE (2019) & $0.48 \pm 0.28$ & $0.47 \pm 0.26$ & $0.45 \pm 0.27$ & $0.45 \pm 0.25$ \\
-    Neural LNSDE (2024) & $0.7 \pm 0.27$ & $0.68 \pm 0.29$ & $\underline{0.67 \pm 0.25}$ & $\mathbf{0.66 \pm 0.23}$ \\
-    LDDMM (2008) & $\underline{0.72 \pm 0.2}$ & $\underline{0.7 \pm 0.21}$ & $0.57 \pm 0.25$ & $0.4 \pm 0.25$ \\
-    %\cline{1-5}
-    TS-LDDMM (ours) & $\mathbf{0.83 \pm 0.18}$ & $\mathbf{0.8 \pm 0.18}$ & $\mathbf{0.7 \pm 0.26}$ & $0.51 \pm 0.27$ \\
-    \bottomrule
-  \end{tabular}  
-\end{table}
-
-%  [11:35, 31/01/2024] Thibaut Germain: avant/après :
-%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,150,1,1)
-%  Kl = TSGaussGaussKernel(5,1,5,0.6)
-%  dataloss = VarifoldLoss(Kl)
-%  schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
-%  [11:35, 31/01/2024] Thibaut Germain: avant :
-%  [11:35, 31/01/2024] Thibaut Germain: Kv = VFTSGaussKernel(1,0.1,100,1,1)
-%  Kl = TSGaussGaussKernel(2,1,2,0.6)
-%  dataloss = VarifoldLoss(Kl)
-%  schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
-% Let $s$ be a time series and $\msg(\tilde{s}) = (g_i)_{i \in [n]}$ be its discretized graph.
-%  The graph $\msg(s)$ can be approximated as 
-% the union of piecewise linear segments between time-consecutive samples, 
-% and we associate to the approximated graph the oriented varifold:
-
-% \begin{equation}
-%   \mu_{\msg(\tilde{s})} : w \in \msw \mapsto \sum_{i=1}^{n-1} l_i\delta_{(c_i,\overrightarrow{v_i})}(w)
-% \end{equation}
-% where $\msw$ is a set of real valued test functions defined on $\Rset^{d+1} \times \mathbb{S}^d$, 
-% $\delta_{(c_i,\overrightarrow{v_i})}(w) = w(c_i,\overrightarrow{v_i})$ is a Dirac delta function, 
-% $c_i = (g_i + g_{i+1})/2$ (resp. $v_i = \| g_{i+1}-g_{i}\|$) is the center (resp. length) of the $i^{th}$ segment, 
-% and $\overrightarrow{v_i} = (g_{i+1}-g_{i})/l_i$ is the unit norm vector of direction $\overrightarrow{g_i g_{i+1}}$.
-
-% Assuming that the test functions' space $\msw$ is the RKHS associated with a $C^1$ positive definite kernel $k$ 
-% that verifies \citep[Proposition 2 \& 4]{kaltenmark2017general}, the oriented varifold belongs to the dual space $\msw^*$. Additionally, 
-% we can define a distance $d_{\msw^*}$ between signals' graph sample sets, which, thanks to the kernel reproducing property,
-% has an explicit formulation:
-
-% \section{Continuous normalizing flows}
-% %Modeling Continuous Stochastic Processes with Dynamic Normalizing Flows time series generative
-% CONTUNYIYS NORMALIZING FLOW bofbof
-% Here, we present the framework of Continuous Normalizing Flows (CNF) and show how LDDMM can be seen as a particular case.
-% \cite{salman2018deep} was already inspired from the work of shape analysis to present deep diffeomorphic normalizing flows.
-% Discussion with Alain :
-% Given a time-dependant vector field $v:[0,1]\times \Rset^{d'} \to \Rset^{d'} $, the integration of this vector field gives a flow $ \phi:[0,1]\times \Rset^{d'} \to \Rset^{d'}$, defined via the ordinary differential equation (ODE),
-% \begin{equation}
-%   \frac{\dd }{\dd \tau}\phi_\tau(x)=v(\tau,\phi_\tau(x)), \phi_0(x)=x \eqsp.
-%  \end{equation}
-% In \cite{chen2018neural}, they introduce Continous Normalizing flows for generative modeling.
-%  By modeling $v=f_\theta$ is a neural networks and choosing a prior $X_0\sim\rho_0$ on the inital value, the law of $\phi_1(X_0)$ has the pullback density $\phi_1\# \rho_0$ and the parameter $\theta$ is optimized to maximize the log-likelihood of a given dataset $(y_i)_{i\in[N]}\in (\Rset^{d'})^N$ which is assumed to follow the law of $\phi_1(X_0)$.
-% Instead of modeling the law of the end point $\phi_1(X_0) $, we can try to model the law of the whole path $(\phi_\tau(X_0))_{\tau\in[0,1]} $ having a density $p:[0,1]\times \Rset^d\to \Rset_{>0}$ such that $p_1=q, p_0=\rho_0$ where $q$ is the empirical distribution in the dataset, as proposed in \cite{lipman2022flow}.
-% In so doing, the problem is ill-posed since $p_t$ can be whatever as soon as $0<t<1$, that is why we try to constraint the learned velocity field $v=f_\theta$ such that the path law $p$ is tractable.
-% One way of doing it is to look for geodesics \eqref{eq:geodesics_original} as proposed in LDDMM, it corresponds to the case where $q=\sum_{i=1}^N \delta_{y_i}/N$ and $\rho_0=\sum_{i=1}^N \delta_{x_i}/N$ are empirical distribution and $y_i=\phi^v_1(x_i) $ for a given velocity field $v$.
-
-% The goal is to target $\nu\in \mathcal{P}(\rmC^0([0,1]\times \Rset^d, \Rset^d)) $ by generating randomness on a initial value $x \sim \delta_x$
-%  and by integrating a velocity field $v_\theta \in \msw \subset \rmC^0([0,1]\times \Rset^d, \Rset^d)$: 
-%  \begin{equation}
-%   \frac{\dd h}{\dd \tau}=v_\theta(\tau,h(\tau,x)), h(0,x)=x \eqsp,
-%  \end{equation}
-%  denoting by $\mu$ the law of $(h(\tau,x))_{\tau \in [0,1]}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \subsection{TTS}
-% To circumvant this issue, we propose two methods :
-%       \begin{enumerate}
-%         \item \label{enum:first_method} [Time And Space (TAS)] Using a specific form of the RKHS's kernel in LDDMM, to stick to the classical shape analysis framework.
-%         \item \label{enum:second_method} [Time Then Space (TTS)] Tackling the time axis and space axis separetely in the transformation.
-%          This distangles time and space variability in the patterns, increasing the interpretability of the representation.
-%           Although the related optimization is more complex than method \ref{enum:first_method}, the results is more robust to the choice of hyper-parameter related to the RKHS's kernel.
-%       \end{enumerate}
-%        In the following, we present both methods to highlights their pro and cons.
-
-% As described in \Cref{section:optimization}, the method (TAS) learns $\Pi_{\gamma,f}$ directly, while (TTS) find the temporal reparametrisation $\Phi_f$ then the space transformation $\Psi_\gamma$.
-%     % If we define a diffeomorpshim $\Phi_\gamma$ as $(\gamma,\Id_d) $, we have $\Phi_\gamma.G(s) =G(s).\gamma^{-1}$.
-%     %  To increase the variability, we allow some deformations on the graph of the time series after being reparametrized in time, which allow us to recover every possible time seriess.
-%     % Remark that, $G(\mathbf{s}_0\circ \psi^{-1})=\{(t,\mathbf{s}_0\circ \psi^{-1}(t)), t\in \msj \}= \{(\psi(t),\mathbf{s}_0(t)) ,\eqsp t\in \msi \}  $.
-%     % [The action of the diffeomorpshim group 
-%     % \begin{align}
-%     %     &\mcd_*=\mcd(\Rset,\Rset)\times \{\phi\in \mcd(\Rset^{d+1},\Rset^{d+1}): \\
-%     %     &\phi(t,x)=(t,f(t,x)),
-%     %      (t,x)\in \Rset \times \Rset^{d}, f \in C^0(\Rset^{d+1}, \Rset^d) \} 
-%     % \end{align}
-%     % on the space of time series graph $\{G^{(a,b)}(s), s\in C^0((a,b), \Rset), (a,b)\in \Rset_+^2, a<b  \}$
-%     % is homogene.]
-%      %In the next section we define a metric and a norm on this space in order to find a representation $(\phi,\psi)$ with a minimal norm, and we show in (...) how to recover a proxy.
-%      \paragraph{The representation as the solution of an optimization}
-%      By \Cref{theorem:representation}, for any $\mathbf{s}_0\in C^0(\msi,\Rset^d)$, we can find $ F=(\gamma_j,f_j)_{j\in[N]}$ such that $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$ for any $j\in[N]$.
-%      However, to get a meaningfull representation and to tackle the unicity problem, we should optimize the choice of $\mathbf{s}_0 $ and $F$, formally speaking, denoting by $R(\cdot) $ a regularization norm on $\mcd(\Rset^{d+1})$ and $L$ a loss on sets which will be specify later, we aim to solve
-%      \begin{equation}
-%       \label{eq:minimization}
-%       \underset{\mathbf{s_0},(\gamma_j,f_j)_{j\in[N]}}{ \argmin} \sum_{j=1}^N \lambda R(\Pi_{\gamma_j,f_j})+ L(\msg(s^j),\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)) \eqsp ,
-%     \end{equation} 
-%     where $\lambda>0$ is a penalization factor related to $L$ to relax the condition $\msg(s^j)=\Pi_{\gamma_j,f_j}.\msg(\mathbf{s}_0)$.
-
-%     In practice, as already proposed in shape analysis, we perform an alternative minimization between $\mathbf{s_0}$ and $(\gamma_j,f_j)_j$.
-%     The optimization on $(\gamma_j,f_j)_j$ is parrallelisable, for each couple $(\mathbf{s}_0,s^j)$, we should learn diffemorphisms of minimal norm satisfying a condition. 
-%     To this end, we adapt the geodesic shooting method using LDDMM presented in \cite{durrleman2013sparse} to our special case.
-%       Before to describe how to optimize the whole in \Cref{section:optimization}, we introduce LDDMM in the next section to specify the norm on diffemorphisms.
-
-% \paragraph{Kernel choices}
-% As depicted on figure (mettre ref), we can not use any kernel $K$ to apply the previous methodology to learn deformations on time series' graphs.
-
-% In the following, we expose two choice of RKHS's kernel which answers this issue : (TAS) and (TTS).
-%  We rely on the Gaussian kernel, denoting by $K_\sigma^{(a)}(x,y)=\exp(-|x-y|^2/\sigma)$ for any $(x,y)\in (\Rset^a)^2$, $a\in \Nset$ and $\sigma>0$.
-% To apply (TAS) in LDDMM with $d'=d+1$, we use the following anisotropic Gaussian kernel  :
-% \begin{align}
-%   \label{eq:kernel_TAS}
-%   &K_{\text{TAS}}((t,x),(t',x'))=\begin{pmatrix}
-%     c_0K & 0 \\
-%     0 & c_1 D 
-%     \end{pmatrix} \eqsp , \\
-%     &D=K_{\sigma_{T,1}}^{(1)}(t,t')K_{\sigma_x}^{(d)}(x,x') \Idd\eqsp,K=K_{\sigma_{T,0}}^{(1)}(t,t') \eqsp,
-% \end{align}
-% parametrized by the widths $\sigma_{T,0},\sigma_{T,1},\sigma_x>0$ and the constants $c_0,c_1>0$.
-%  If we use this kernel, by considering \eqref{eq:specific_form}, we remark that the first coordinate of the velocity field $v_\tau$ only depends on the time variable $t$.
-%   It implies that the final deformation will have the form $\phi^v=(\gamma,f) $ with  $f\in C^1(\Rset^{d+1},\Rset^d) $, and $\gamma\in \mcd(\Rset)$
-%   since the first coordinate follows a LDDMM-ODE \eqref{eq:LDDMM_dynamic}.
-
-% To apply (TTS), we use LDDMM on the time axis with $d'=1$ by taking $K_{\text{TTS}}^{\text{time}}=K_{\sigma_{T,0}}^{(1)}$, then on the space axis with $d'=d$ by choosing
-% $K_{\text{TTS}}^{\text{space}}=K_{\sigma_{T,1}}^{(1)}\times K_{\sigma_x}^{(d)} $ where the widths are the same than in \eqref{eq:kernel_TAS}.
-The alternate optimization will be more detailed in \Cref{section:optimization}.
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
-% by Iain Murray in 2018, and modified by Alexandre Bouchard in
-% 2019 and 2021 and by Csaba Szepesvari, Gang Niu and Sivan Sabato in 2022.
-% Modified again in 2023 by Sivan Sabato and Jonathan Scarlett.
-% Previous contributors include Dan Roy, Lise Getoor and Tobias
-% Scheffer, which was slightly modified from the 2010 version by
-% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
-% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
-% slightly modified from Prasad Tadepalli's 2007 version which is a
-% lightly changed version of the previous year's version by Andrew
-% Moore, which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2024-05-22 NEURIPS24/appendix.tex
update

--- 2024-05-22 NEURIPS24/appendix.tex
@@
 \section{Experimental settings}
-
+\label{appendix:settings}
 All experiments were performed on a Debian 6.1.69-1 server with NVIDIA RTX A2000 12GB GPU, Intel(R) Xeon(R) Gold 5220R CPU @ 2.20GHz, and 250 GB of RAM. The source code will be available on Github.

### 2024-05-22 NEURIPS24/checklist.tex
update

--- 2024-05-15 NEURIPS24/neurips_2024_to_submit.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2024
-
-% ready for submission
-\usepackage{neurips_2024}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2024}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2024}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2024}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{mathtools}
-\usepackage{amsthm}
-\usepackage[capitalize,noabbrev]{cleveref}
-
-\title{Shapes analysis for time series.}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{%
-  David S.~Hippocampus\thanks{Use footnote for providing further information
-    about author (webpage, alternative address)---\emph{not} for acknowledging
-    funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  % examples of more authors
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-}
-\input{header.tex}
-\input{def.tex}
-\newcommand\sam[1]{\textcolor{blue}{#1}}
-\newcommand\thi[1]{\textcolor{orange}{#1}}
-\begin{document}
-
-\maketitle
-%In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
-    %Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
-    % Given a dataset of $N$ individual's time series $(s^j)_{j\in[N]}$,
-    %  we learn a time series of reference $\mathbf{s}_0$ encoding the common part of the individuals as well as diffeomorphisms $(\phi^j)_{j\in[N]}$ encoding the specificities of each individual:
-    %  $s^j$ is seen as the transformation of $\mathbf{s}_0$ by $\phi_j$.
-    %  Then, the parameters $(\alpha_j)_{j\in[N]}$ encoding the diffeomorphisms $(\phi^j)_{j\in[N]}$ are used as the features representation related to each time series $s^j$.
-    %At the crossroads between URL for time series and shape analysis.
-%the algorithm extends the Large Deformation Diffeomorphic Metric Mapping (LDDMM) to the case of time series.
-This extension is a nontrivial application of the LDDMM framework since it involves applying deformations to the graph of time series, which have more structure than curves due to the inclusion of the temporal dimension.
-
-\begin{abstract}
-Analyzing inter-individual variability of physiological functions is particularly appealing in medical and biological contexts to describe or quantify health conditions.
-Such analysis can be done by comparing individuals to a reference one with time series as biomedical data.
-This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to inter-individual studies.
-The idea is to represent time series as deformations of a reference time series.
-The deformations are diffeomorphisms parameterized and learned by our method called TS-LDDMM.
-Once the deformations and the reference time series are learned, the vector representations of individual time series are given by the parametrization of their corresponding deformation.
-At the crossroads between URL for time series and shape analysis, the proposed algorithm handles irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
-In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework.
-We showcase the advantages of our representation compared to existing methods using synthetic data and real-world examples motivated by biomedical applications.
-\end{abstract}
-
-\input{paper.tex}
-
-\bibliographystyle{plain}
-\bibliography{ref.bib}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\input{appendix.tex}
 \section*{NeurIPS Paper Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist is designed to encourage best practices for responsible machine learning research, addressing issues of reproducibility, transparency, research ethics, and societal impact.
-Do not remove the checklist: {\bf The papers not including the checklist will be desk rejected.} The checklist should follow the references and precede the (optional) supplemental material.
-The checklist does NOT count towards the page
-limit.
-
-Please read the checklist guidelines carefully for information on how to answer these questions. For each question in the checklist:
-\begin{itemize}
-    \item You should answer \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item \answerNA{} means either that the question is Not Applicable for that particular paper or the relevant information is Not Available.
-    \item Please provide a short (1–2 sentence) justification right after your answer (even for NA). 
-   % \item {\bf The papers not including the checklist will be desk rejected.}
-\end{itemize}
-
-{\bf The checklist answers are an integral part of your paper submission.} They are visible to the reviewers, area chairs, senior area chairs, and ethics reviewers.
-You will be asked to also include it (after eventual revisions) with the final version of your paper, and its final version will be published with the paper.
-
-The reviewers of your paper will be asked to use the checklist as one of the factors in their evaluation.
-While "\answerYes{}" is generally preferable to "\answerNo{}", it is perfectly acceptable to answer "\answerNo{}" provided a proper justification is given (e.g., "error bars are not reported because it would be too computationally expensive" or "we were unable to find the license for the dataset we used").
-In general, answering "\answerNo{}" or "\answerNA{}" is not grounds for rejection.
-While the questions are phrased in a binary way, we acknowledge that the true answer is often more nuanced, so please just use your best judgment and write a justification to elaborate.
-All supporting evidence can appear either in the main paper or the supplemental material, provided in appendix.
-If you answer \answerYes{} to a question, in the justification please point to the section(s) where related material for the question can be found.
-
-IMPORTANT, please:
-\begin{itemize}
-    \item {\bf Delete this instruction block, but keep the section heading ``NeurIPS paper checklist"},
-    \item  {\bf Keep the checklist subsection headings, questions/answers and guidelines below.}
-    \item {\bf Do not modify the questions and only use the provided macros for your answers}.
-\end{itemize}
-
-%%% END INSTRUCTIONS %%%
 
@@
     \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: Each claim in the introduction is referring to the part where it is tackled.
     \item[] Guidelines:
@@
     \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: We have provided a special section for this purpose \Cref{sec:limitations}.
     \item[] Guidelines:
@@
     \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (and correct) proof?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: All the proofs are given in \Cref{appendix:proofs} and each proof state all its arguments in a logical order.
     \item[] Guidelines:
@@
     \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper (regardless of whether the code and data are provided or not)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: The optimization methodology is described in \Cref{section:methodology} and all numerical details and protocol are given in \Cref{section:experiments} and in \Cref{appendix:classification_implementation}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix:identifiability}.
     \item[] Guidelines:
@@
     \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results, as described in supplemental material?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: Synthetic data can be generated, benchmark data are publicly available (\Cref{appendix:classification_dataset}), but the mouse dataset is not.
+     The code is provided as supplementary material and will be made publicly available if the paper is accepted.
     \item[] Guidelines:
@@
     \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerNo{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: Error bars are given for the synthetic experiments, but not for the real experiments due to limited computational budget.
+     There is no randomness in the train/test split but there is randomness in the optimization method (Adabelief), but its impact is not significant on the results.
     \item[] Guidelines:
@@
     \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources (type of compute workers, memory, time of execution) needed to reproduce the experiments?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: This is given at the beginning of \Cref{appendix:settings}.
     \item[] Guidelines:
@@
     \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics \url{https://neurips.cc/public/EthicsGuidelines}?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: Every authors are funded for their work and preserve research integrity.
     \item[] Guidelines:
@@
     \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work performed?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: A special section is given on this subject .
     \item[] Guidelines:
@@
     \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
- .     \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: \justificationTODO{}
     \item[] Guidelines:
@@
 \end{enumerate}
-
-\end{document}

### 2024-05-22 NEURIPS24/paper.tex
update

@@
   \section{Limitations}
+  \label{sec:limitations}
   While TS-LDDMM performs well on shape-based datasets, it assumes a certain degree of homogeneity with the existence of a reference time series graph$\msg_0$.

### 2024-05-22 NEURIPS24/checklist.tex
update
without societal impact

@@
     \item[] Question: Does the paper specify all the training and test details (e.g., data splits, hyperparameters, how they were chosen, type of optimizer, etc.) necessary to understand the results?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: The experiments setting is clearly presented in \Cref{section:experiments} for the experiments on the mouse dataset.
+The full details of all experiments are presented in \Cref{appendix:classification_implementation}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix:identifiability}.
     \item[] Guidelines:
@@
     \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: A special section is given on this subject .
+    \item[] Justification: A special section is given on this subject \Cref{appendix:societal_impact}.
     \item[] Guidelines:
@@
     \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models that have a high risk for misuse (e.g., pretrained language models, image generators, or scraped datasets)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerNA{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: There is no risk.
     \item[] Guidelines:
@@
     \item[] Question: Are the creators or original owners of assets (e.g., code, data, models), used in the paper, properly credited and are the license and terms of use explicitly mentioned and properly respected?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: All dataset and python library were cited when necessary.
     \item[] Guidelines:
@@
     \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: The mouse dataset is described in \Cref{appendix:mouse_dataset}.
     \item[] Guidelines:
@@
     \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants and screenshots, if applicable, as well as details about compensation (if any)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerNA{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: The paper does not involve crowdsourcing nor research with human subjects.
     \item[] Guidelines:
@@
     \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
+    \item[] Answer: \answerNA{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
+    \item[] Justification: The paper does not involve crowdsourcing nor research with human subjects.
     \item[] Guidelines:

### 2024-05-22 NEURIPS24/experiment.tex
update
update
without societal impact

@@
       \textbf{Results} show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%, 30\%, and 50\%.
-      {\color{red} J'enlèverai cette phrase ou je développerai en Appendix : The performance decrease of the Kernel-SVC based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.}
+      %{\color{red} J'enlèverai cette phrase ou je développerai en Appendix : The performance decrease of the Kernel-SVC based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.}
 
@@
 \vspace{-1ex}
-\paragraph{Result summary.} By comparing the shape of individual respiratory cycles, we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule.
+\paragraph{Result summary.} By comparing the shape of individual respiratory cycles, we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule contrary to LDDMM \cite{glaunes2008large}.
 \vspace{-1ex}

### 2024-05-22 NEURIPS24/neurips_2024_to_submit.tex
restore bis
update
without societal impact

@@
 \usepackage{amsthm}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{wrapfig}
 \usepackage[capitalize,noabbrev]{cleveref}
@@
 \input{appendix.tex}
-\section*{NeurIPS Paper Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist is designed to encourage best practices for responsible machine learning research, addressing issues of reproducibility, transparency, research ethics, and societal impact.
-Do not remove the checklist: {\bf The papers not including the checklist will be desk rejected.} The checklist should follow the references and precede the (optional) supplemental material.
-The checklist does NOT count towards the page
-limit.
-
-Please read the checklist guidelines carefully for information on how to answer these questions. For each question in the checklist:
-\begin{itemize}
-    \item You should answer \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item \answerNA{} means either that the question is Not Applicable for that particular paper or the relevant information is Not Available.
-    \item Please provide a short (1–2 sentence) justification right after your answer (even for NA). 
-   % \item {\bf The papers not including the checklist will be desk rejected.}
-\end{itemize}
-
-{\bf The checklist answers are an integral part of your paper submission.} They are visible to the reviewers, area chairs, senior area chairs, and ethics reviewers.
-You will be asked to also include it (after eventual revisions) with the final version of your paper, and its final version will be published with the paper.
-
-The reviewers of your paper will be asked to use the checklist as one of the factors in their evaluation.
-While "\answerYes{}" is generally preferable to "\answerNo{}", it is perfectly acceptable to answer "\answerNo{}" provided a proper justification is given (e.g., "error bars are not reported because it would be too computationally expensive" or "we were unable to find the license for the dataset we used").
-In general, answering "\answerNo{}" or "\answerNA{}" is not grounds for rejection.
-While the questions are phrased in a binary way, we acknowledge that the true answer is often more nuanced, so please just use your best judgment and write a justification to elaborate.
-All supporting evidence can appear either in the main paper or the supplemental material, provided in appendix.
-If you answer \answerYes{} to a question, in the justification please point to the section(s) where related material for the question can be found.
-
-IMPORTANT, please:
-\begin{itemize}
-    \item {\bf Delete this instruction block, but keep the section heading ``NeurIPS paper checklist"},
-    \item  {\bf Keep the checklist subsection headings, questions/answers and guidelines below.}
-    \item {\bf Do not modify the questions and only use the provided macros for your answers}.
-\end{itemize}
-
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item {\bf Claims}
-    \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the abstract and introduction do not include the claims made in the paper.
-        \item The abstract and/or introduction should clearly state the claims made, including the contributions made in the paper and important assumptions and limitations.
-A No or NA answer to this question will not be perceived well by the reviewers.
-        \item The claims made should match theoretical and experimental results, and reflect how much the results can be expected to generalize to other settings. 
-        \item It is fine to include aspirational goals as motivation as long as it is clear that these goals are not attained by the paper. 
-    \end{itemize}
-
-\item {\bf Limitations}
-    \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper has no limitation while the answer No means that the paper has limitations, but those are not discussed in the paper. 
-        \item The authors are encouraged to create a separate "Limitations" section in their paper.
-        \item The paper should point out any strong assumptions and how robust the results are to violations of these assumptions (e.g., independence assumptions, noiseless settings, model well-specification, asymptotic approximations only holding locally).
-The authors should reflect on how these assumptions might be violated in practice and what the implications would be.
-        \item The authors should reflect on the scope of the claims made, e.g., if the approach was only tested on a few datasets or with a few runs.
-In general, empirical results often depend on implicit assumptions, which should be articulated.
-        \item The authors should reflect on the factors that influence the performance of the approach.
-For example, a facial recognition algorithm may perform poorly when image resolution is low or images are taken in low lighting.
-Or a speech-to-text system might not be used reliably to provide closed captions for online lectures because it fails to handle technical jargon.
-        \item The authors should discuss the computational efficiency of the proposed algorithms and how they scale with dataset size.
-        \item If applicable, the authors should discuss possible limitations of their approach to address problems of privacy and fairness.
-        \item While the authors might fear that complete honesty about limitations might be used by reviewers as grounds for rejection, a worse outcome might be that reviewers discover limitations that aren't acknowledged in the paper.
-The authors should use their best judgment and recognize that individual actions in favor of transparency play an important role in developing norms that preserve the integrity of the community.
-Reviewers will be specifically instructed to not penalize honesty concerning limitations.
-    \end{itemize}
-
-\item {\bf Theory Assumptions and Proofs}
-    \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (and correct) proof?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include theoretical results. 
-        \item All the theorems, formulas, and proofs in the paper should be numbered and cross-referenced.
-        \item All assumptions should be clearly stated or referenced in the statement of any theorems.
-        \item The proofs can either appear in the main paper or the supplemental material, but if they appear in the supplemental material, the authors are encouraged to provide a short proof sketch to provide intuition.
-        \item Inversely, any informal proof provided in the core of the paper should be complemented by formal proofs provided in appendix or supplemental material.
-        \item Theorems and Lemmas that the proof relies upon should be properly referenced. 
-    \end{itemize}
-
-    \item {\bf Experimental Result Reproducibility}
-    \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper (regardless of whether the code and data are provided or not)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item If the paper includes experiments, a No answer to this question will not be perceived well by the reviewers: Making the paper reproducible is important, regardless of whether the code and data are provided or not.
-        \item If the contribution is a dataset and/or model, the authors should describe the steps taken to make their results reproducible or verifiable. 
-        \item Depending on the contribution, reproducibility can be accomplished in various ways.
-For example, if the contribution is a novel architecture, describing the architecture fully might suffice, or if the contribution is a specific model and empirical evaluation, it may be necessary to either make it possible for others to replicate the model with the same dataset, or provide access to the model.
-In general.
-releasing code and data is often one good way to accomplish this, but reproducibility can also be provided via detailed instructions for how to replicate the results, access to a hosted model (e.g., in the case of a large language model), releasing of a model checkpoint, or other means that are appropriate to the research performed.
-        \item While NeurIPS does not require releasing code, the conference does require all submissions to provide some reasonable avenue for reproducibility, which may depend on the nature of the contribution.
-For example
-        \begin{enumerate}
-            \item If the contribution is primarily a new algorithm, the paper should make it clear how to reproduce that algorithm.
-            \item If the contribution is primarily a new model architecture, the paper should describe the architecture clearly and fully.
-            \item If the contribution is a new model (e.g., a large language model), then there should either be a way to access this model for reproducing the results or a way to reproduce the model (e.g., with an open-source dataset or instructions for how to construct the dataset).
-            \item We recognize that reproducibility may be tricky in some cases, in which case authors are welcome to describe the particular way they provide for reproducibility.
-In the case of closed-source models, it may be that access to the model is limited in some way (e.g., to registered users), but it should be possible for other researchers to have some path to reproducing or verifying the results.
-        \end{enumerate}
-    \end{itemize}
-
-\item {\bf Open access to data and code}
-    \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results, as described in supplemental material?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that paper does not include experiments requiring code.
-        \item Please see the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
-        \item While we encourage the release of code and data, we understand that this might not be possible, so “No” is an acceptable answer.
-Papers cannot be rejected simply for not including code, unless this is central to the contribution (e.g., for a new open-source benchmark).
-        \item The instructions should contain the exact command and environment needed to run to reproduce the results.
-See the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
-        \item The authors should provide instructions on data access and preparation, including how to access the raw data, preprocessed data, intermediate data, and generated data, etc.
-        \item The authors should provide scripts to reproduce all experimental results for the new proposed method and baselines.
-If only a subset of experiments are reproducible, they should state which ones are omitted from the script and why.
-        \item At submission time, to preserve anonymity, the authors should release anonymized versions (if applicable).
-        \item Providing as much information as possible in supplemental material (appended to the paper) is recommended, but including URLs to data and code is permitted.
-    \end{itemize}
-
-\item {\bf Experimental Setting/Details}
-    \item[] Question: Does the paper specify all the training and test details (e.g., data splits, hyperparameters, how they were chosen, type of optimizer, etc.) necessary to understand the results?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item The experimental setting should be presented in the core of the paper to a level of detail that is necessary to appreciate the results and make sense of them.
-        \item The full details can be provided either with the code, in appendix, or as supplemental material.
-    \end{itemize}
-
-\item {\bf Experiment Statistical Significance}
-    \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item The authors should answer "Yes" if the results are accompanied by error bars, confidence intervals, or statistical significance tests, at least for the experiments that support the main claims of the paper.
-        \item The factors of variability that the error bars are capturing should be clearly stated (for example, train/test split, initialization, random drawing of some parameter, or overall run with given experimental conditions).
-        \item The method for calculating the error bars should be explained (closed form formula, call to a library function, bootstrap, etc.)
-        \item The assumptions made should be given (e.g., Normally distributed errors).
-        \item It should be clear whether the error bar is the standard deviation or the standard error of the mean.
-        \item It is OK to report 1-sigma error bars, but one should state it.
-The authors should preferably report a 2-sigma error bar than state that they have a 96\% CI, if the hypothesis of Normality of errors is not verified.
-        \item For asymmetric distributions, the authors should be careful not to show in tables or figures symmetric error bars that would yield results that are out of range (e.g.
-negative error rates).
-        \item If error bars are reported in tables or plots, The authors should explain in the text how they were calculated and reference the corresponding figures or tables in the text.
-    \end{itemize}
-
-\item {\bf Experiments Compute Resources}
-    \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources (type of compute workers, memory, time of execution) needed to reproduce the experiments?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item The paper should indicate the type of compute workers CPU or GPU, internal cluster, or cloud provider, including relevant memory and storage.
-        \item The paper should provide the amount of compute required for each of the individual experimental runs as well as estimate the total compute. 
-        \item The paper should disclose whether the full research project required more compute than the experiments reported in the paper (e.g., preliminary or failed experiments that didn't make it into the paper).
-    \end{itemize}
-
-\item {\bf Code Of Ethics}
-    \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics \url{https://neurips.cc/public/EthicsGuidelines}?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the authors have not reviewed the NeurIPS Code of Ethics.
-        \item If the authors answer No, they should explain the special circumstances that require a deviation from the Code of Ethics.
-        \item The authors should make sure to preserve anonymity (e.g., if there is a special consideration due to laws or regulations in their jurisdiction).
-    \end{itemize}
-
-\item {\bf Broader Impacts}
-    \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work performed?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that there is no societal impact of the work performed.
-        \item If the authors answer NA or No, they should explain why their work has no societal impact or why the paper does not address societal impact.
-        \item Examples of negative societal impacts include potential malicious or unintended uses (e.g., disinformation, generating fake profiles, surveillance), fairness considerations (e.g., deployment of technologies that could make decisions that unfairly impact specific groups), privacy considerations, and security considerations.
-        \item The conference expects that many papers will be foundational research and not tied to particular applications, let alone deployments.
-However, if there is a direct path to any negative applications, the authors should point it out.
-For example, it is legitimate to point out that an improvement in the quality of generative models could be used to generate deepfakes for disinformation.
-On the other hand, it is not needed to point out that a generic algorithm for optimizing neural networks could enable people to train models that generate Deepfakes faster.
-        \item The authors should consider possible harms that could arise when the technology is being used as intended and functioning correctly, harms that could arise when the technology is being used as intended but gives incorrect results, and harms following from (intentional or unintentional) misuse of the technology.
-        \item If there are negative societal impacts, the authors could also discuss possible mitigation strategies (e.g., gated release of models, providing defenses in addition to attacks, mechanisms for monitoring misuse, mechanisms to monitor how a system learns from feedback over time, improving the efficiency and accessibility of ML).
-    \end{itemize}
-
-\item {\bf Safeguards}
-    \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models that have a high risk for misuse (e.g., pretrained language models, image generators, or scraped datasets)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper poses no such risks.
-        \item Released models that have a high risk for misuse or dual-use should be released with necessary safeguards to allow for controlled use of the model, for example by requiring that users adhere to usage guidelines or restrictions to access the model or implementing safety filters.
-        \item Datasets that have been scraped from the Internet could pose safety risks. The authors should describe how they avoided releasing unsafe images.
-        \item We recognize that providing effective safeguards is challenging, and many papers do not require this, but we encourage authors to take this into account and make a best faith effort.
-    \end{itemize}
-
-\item {\bf Licenses for existing assets}
-    \item[] Question: Are the creators or original owners of assets (e.g., code, data, models), used in the paper, properly credited and are the license and terms of use explicitly mentioned and properly respected?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not use existing assets.
-        \item The authors should cite the original paper that produced the code package or dataset.
-        \item The authors should state which version of the asset is used and, if possible, include a URL.
-        \item The name of the license (e.g., CC-BY 4.0) should be included for each asset.
-        \item For scraped data from a particular source (e.g., website), the copyright and terms of service of that source should be provided.
-        \item If assets are released, the license, copyright information, and terms of use in the package should be provided.
-For popular datasets, \url{paperswithcode.com/datasets} has curated licenses for some datasets.
-Their licensing guide can help determine the license of a dataset.
-        \item For existing datasets that are re-packaged, both the original license and the license of the derived asset (if it has changed) should be provided.
-        \item If this information is not available online, the authors are encouraged to reach out to the asset's creators.
-    \end{itemize}
-
-\item {\bf New Assets}
-    \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not release new assets.
-        \item Researchers should communicate the details of the dataset/code/model as part of their submissions via structured templates.
-This includes details about training, license, limitations, etc.
-        \item The paper should discuss whether and how consent was obtained from people whose asset is used.
-        \item At submission time, remember to anonymize your assets (if applicable). You can either create an anonymized URL or include an anonymized zip file.
-    \end{itemize}
-
-\item {\bf Crowdsourcing and Research with Human Subjects}
-    \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants and screenshots, if applicable, as well as details about compensation (if any)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
-        \item Including this information in the supplemental material is fine, but if the main contribution of the paper involves human subjects, then as much detail as possible should be included in the main paper.
-        \item According to the NeurIPS Code of Ethics, workers involved in data collection, curation, or other labor should be paid at least the minimum wage in the country of the data collector. 
-    \end{itemize}
-
-\item {\bf Institutional Review Board (IRB) Approvals or Equivalent for Research with Human Subjects}
-    \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
- .     \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
-        \item Depending on the country in which research is conducted, IRB approval (or equivalent) may be required for any human subjects research.
-If you obtained IRB approval, you should clearly state this in the paper.
-        \item We recognize that the procedures for this may vary significantly between institutions and locations, and we expect authors to adhere to the NeurIPS Code of Ethics and the guidelines for their institution.
-        \item For initial submissions, do not include any information that would break anonymity (if applicable), such as the institution conducting the review.
-    \end{itemize}
-
-\end{enumerate}
+\input{checklist.tex}
 

### 2024-05-22 NEURIPS24/appendix.tex
update
without societal impact
update

@@
 
+\section{Societal impact}
+\label{appendix:societal_impact}
+We believe that the paper has a positive societal impact for the following reasons:
+\begin{itemize}
+\item TS-LDDMM is an interpretable method for understanding inter-individual variability in biomedical datasets, potentially offering new insights in medicine.
+\item TS-LDDMM bridges the gap between the shape analysis community and the unsupervised representation learning (URL) community, fostering potential future collaborations between these fields.
+\end{itemize}
+However, the computational cost of the method may raise environmental concerns similar to those associated with deep learning \cite{oh2024stable}.
+ Additionally, while TS-LDDMM has promising biomedical applications, it could also be misused for creating poison.
 \section{Proofs}

### 2024-05-22 NEURIPS24/paper.tex
update

@@
 
-  In the next section, the time series are going to be represented by diffeomorphism parameters $(\alpha_j)_{j\in[N]}$.
-  That's why LDDMM is chosen since it offers a parametrization for diffemorphisms which is sparse and interpretable, two features particularly relevant in the biomedical context.
+ In the next section, time series are going to be represented by diffeomorphism parameters $(\alpha_j)_{j\in[N]}$.
+ That is why LDDMM is chosen since it offers a parametrization for diffeomorphisms that is sparse and interpretable, two features particularly relevant in the biomedical context.
 
@@
 Given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$\footnote{Note that we denote by $d'\in\nset$ the ambient space}, a set of starting points $\mathbf{x}=(x_{i})_{i\in[T_1]}$ in $\Rset^{d'}$, we aim to find a diffeomorphism $\phi$ such that the finite set of points $\mathbf{y}$ is similar in a certain sense to the set of finite sets of transformed points $\phi \cdot \mathbf{x} =(\phi(x_i))_{i\in[T_1]} $.
- The function $\phi$ is occasionally referred to as a \textit{deformation}. In general, these sets $\mathbf{x},\mathbf{y}$ are meshes of continous objects, e.g surfaces, curves, images and so on.
+ The function $\phi$ is occasionally referred to as a \textit{deformation}. In general, these sets $\mathbf{x},\mathbf{y}$ are meshes of continuous objects, e.g., surfaces, curves, images, and so on.
 
@@
  \end{equation}
-By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy lipshcitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$ \cite{miller2006geodesic}, as straight lines in Euclidean space.
+By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy-Lipschitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$ \cite{miller2006geodesic}, as straight lines in Euclidean space.
 Denoting by $w(v_0)$ the geodesic starting from the identity with inital velocity $v_0$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \phi^v$ and the previous matching problem becomes a \textit{geodesic shooting problem}:
@@
 \paragraph{Discrete parametrization of diffeomorpshim.}
-%\alain{adapt this without mentioning applicaiton to time seriess}
-%\sam{Est ce qu'on devrait pas introduire la dicretization du graphe ici et dire que l'on peut utiliser le théoreme de représentation}
+
 In this part, $\msv$ is chosen as an RKHS \cite{berlinet2011reproducing} generated by a smooth kernel $K$ (e.g., Gaussian). 
-% $K(x,y)\triangleq \exp(-|x-y|^2/\sigma^2)\operatorname{Id}_{d'}$ for any $(x,y)\in(\Rset^{d'})^2$, where $\sigma^2>0$ is the kernel width and $\operatorname{Id}_{d'}$ is the identity matrix.
 We follow \cite{durrleman2013sparse} and define a 
@@
 However, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
- Note that \textbf{the number of samples $n_j$ is not necessary the same accross individuals} and the timestamps can be \textbf{irregularly sampled}.
-%To ease the presentation of our methodology, we suppose that we have access to the continuous time series $(s^j)_{j\in[N]}$, while in practice, we can only access a $n_j$-samples $\tsig^j=(\tsig_i^j=s^j(t^j_i))_{i\in[n_j]}$ collected at timestamps $(t^j_i)_{i\in[n_j]}$ for any $j \in [N]$.
- %However, this difference does not change the overall rationale of our method and its implementation up to minor changes, which we summarize in \eqref{eq:general_optimization_problem}.
+ Note that \textbf{the number of samples $n_j$ is not necessarily the same across individuals} and the timestamps can be \textbf{irregularly sampled}.
  We assume the time series population is globally homogeneous regarding their "shapes" even if inter-individual variability exists.
@@
  \end{equation}
-      which is carried out by a gradient descent on the control points $\tilde{\msg}_0$ and the momentums $\mathbf{\alpha}_j=(\alpha_k^j)_{k\in [\mathbf{n}_0]}$ for any $j\in[N]$, initialized by a dataset's time series graph of size $\mathbf{n}_0$ and by $0_{(d+1)\mathbf{n}_0}$ respectively.
+ which is carried out by gradient descent on the control points $\tilde{\msg}_0$ and the momentums $\mathbf{\alpha}_j=(\alpha_k^j)_{k\in [\mathbf{n}_0]}$ for any $j\in[N]$, initialized by a dataset's time series graph of size $\mathbf{n}_0$ and by $0_{(d+1)\mathbf{n}_0}$ respectively.
  The optimization hyperparameter details are given in \Cref{appendix:optimizers_details}.
  The result of the minimization $\tilde{\msg}_0$ is then considered as the $\mathbf{n}_0$-samples of a common time series $\mathbf{s}_0$ and the momentums $\mathbf{\alpha}_j$ encoding $\phi_j$ yields a feature vector in $\Rset^{d \mathbf{n}_0} $ of $s^j$ for any $j\in[N]$.
-     %, which will be the representation of the time series when choosing the time series' graph sets as inputs, is dependent on the choice of kernel $K$ and the set distance $L$.
  Finally, the vectors $(\mathbf{\alpha}_j)_{j\in[N]}$ can be analyzed with any statistical or machine learning tools such as Principal Components Analysis (PCA), Latent Discriminant Analysis (LDA), longitudinal data analysis and so on.
@@
 
-Nevertheless, \eqref{eq:general_optimization_problem} ask to define a kernel and a loss in order to perform geodesic shooting \ref{eq:relaxation}, which is the purpose of the next subsection.
+Nevertheless, \eqref{eq:general_optimization_problem} asks to define a kernel and a loss in order to perform geodesic shooting \ref{eq:relaxation}, which is the purpose of the following subsection.
 \vspace{-1ex}
@@
         \label{section:time_series_specificity}
-        In this section, we present our theoretical contribution: we tailor the LDDMM framework to handle time series data.
+        This section presents our theoretical contribution: we tailor the LDDMM framework to handle time series data.
         The reason is that applying a general diffeomorphism $\phi$ from $\Rset^{d+1}$ to a time series' graph $\msg(s)$ can result in a set $\phi.\msg(s)$ that does not correspond to the graph of any time series, as illustrated in the \Cref{fig:diffeo}.
-           Thus, Time series graph have more structure than a simple 1D curve \cite{glaunes2008large} and deserve their special analysis which will prove fruitful as demonstrated in \ref{section:experiments}.% thus this case is not an application of \cite{glaunes2004diffeomorphic}
+        Thus, time series graphs have more structure than a simple 1D curve \cite{glaunes2008large} and deserve their unique analysis, which will prove fruitful as demonstrated in \ref{section:experiments}.
 
@@
     \begin{remark}
-      that for any $\gamma \in \mcd(\Rset) $ and $s\in \rmC^0(\msi,\Rset^d)$,
+      Note that for any $\gamma \in \mcd(\Rset) $ and $s\in \rmC^0(\msi,\Rset^d)$,
    \begin{equation}
@@
     \end{lemma}
-    Instead of Gaussian kernels, other types of smooth kernel can be selected as long as the structure \eqref{eq:kernel_TAS} is respected.
+    Instead of Gaussian kernels, other types of smooth kernels can be selected as long as the structure \eqref{eq:kernel_TAS} is respected.
     % \begin{proof}
@@
     \begin{remark}
-      With this choice of kernel, the features associated to the time transformation can be extracted from the momentums $(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}$ in \eqref{eq:def_v0} by taking the coordinates related to time.
+      With this choice of kernel, the features associated with the time transformation can be extracted from the momentums $(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}$ in \eqref{eq:def_v0} by taking the coordinates related to time.
       However, the features related to the space transformation are not only in the space coordinates since the related kernel $K_{\text{space}}$ depends on time as well.
@@
  Note that the loss kernel $k$ has nothing to do with the velocity field kernel denoted by $K_\msg$ or $K$ specified in \Cref{paragraph:kernel_V}.
-Finally, we define the data fidelity loss function, $\scrl$, as a sum of $ d_{\msw^*}^2$ using different kernel's width parameters $\sigma$ to incorporate multiscale information such that.
-$\scrl$ is indeed differentiable with regards to its first variable.
+Finally, we define the data fidelity loss function, $\scrl$, as a sum of $ d_{\msw^*}^2$ using different kernel's width parameters $\sigma$ to incorporate multiscale information.
+$\scrl$ is indeed differentiable with respect to its first variable.
 The specific kernels $k_{\pos},k_{\dir}$ that we use in our experiments are given \Cref{appendix:kernel_implementation}.
-For further readings on curves and surfaces representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}.
+For further readings on curves and surface representation as varifolds, readers can refer to \cite{kaltenmark2017general,charon2013varifold}.
 
@@
 
-       Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} addresses the representation of curves with the Square-Root Velocity (SRV) representation.
-        However, the SRV representation is applied after a reparametrization of the temporal dimension on the unit length segment.
+ Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} address the representation of curves with the Square-Root Velocity (SRV) representation.
+ However, the SRV representation is applied after reparametrization of the temporal dimension of the unit length segment.
  Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation.
@@
  Balancing between discrete and continuous elements is a challenging task.
-  In the deep learning literature \cite{chen2018neural, kidger2020neural, tzen2019neural, jia2019neural, liu2019neural, ansari2023neural}, Neural Ordinary Differential Equations (Neural ODEs) \cite{chen2018neural} learn continuous latent representations using a vector field parameterized by a neural network, serving as a continuous analogue to Residual Networks \cite{zagoruyko2016wide}.
+ In the deep learning literature \cite{chen2018neural, kidger2020neural, tzen2019neural, jia2019neural, liu2019neural, ansari2023neural}, Neural Ordinary Differential Equations (Neural ODEs) \cite{chen2018neural} learn continuous latent representations using a vector field parameterized by a neural network, serving as a continuous analog to Residual Networks \cite{zagoruyko2016wide}.
  This approach was further enhanced by Neural Controlled Differential Equations (Neural CDEs) \cite{kidger2020neural} for handling irregular time series, functioning as continuous-time analogs of RNNs \cite{schuster1997bidirectional}.
  Extending Neural ODEs, Neural Stochastic Differential Equations (Neural SDEs) introduce regularization effects \cite{liu2019neural}, although optimization remains challenging.
-  Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied succesfully Neural SDEs to irregular time series.
+ Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied successfully Neural SDEs to irregular time series.
  Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
-  With TS-LDDMM, the representation is also derived from an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm, which enhance interpretability.
+ With TS-LDDMM, the representation is also derived from an ODE, but the velocity field is parameterized with kernels and optimized to have a minimal norm, which enhances interpretability.
 
@@
  In future work, adaptive methods are expected to be developed to provide a more user-friendly interface.
-  {%\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
+  %\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
 \section{Conclusion}
 \vspace{-1ex}
-In this paper, we propose a feature representation method, TS-LDDMM, designed for 
-shape comparison in homogeneous time series datasets. We show on a real dataset 
+This paper proposes a feature representation method, TS-LDDMM, designed for 
+shape comparison on homogeneous time series datasets. We show on a real dataset 
 its ability to study, with high interpretability, the inter-individual shape 
@@
 is already competitive for classification, its performances can be leveraged on 
-more heterogeneous datasets using a hierarchical clustering extension, which is relagated for future work.
+more heterogeneous datasets using a hierarchical clustering extension, which is relegated for future work.
 

### 2024-05-22 NEURIPS24/paper.tex
up grammarly

@@
 
-Nevertheless, \eqref{eq:general_optimization_problem} asks to define a kernel and a loss in order to perform geodesic shooting \ref{eq:relaxation}, which is the purpose of the following subsection.
+Nevertheless, \eqref{eq:general_optimization_problem} asks to define a kernel and a loss in order to perform geodesic shooting \eqref{eq:relaxation}, which is the purpose of the following subsection.
 \vspace{-1ex}
@@
         The reason is that applying a general diffeomorphism $\phi$ from $\Rset^{d+1}$ to a time series' graph $\msg(s)$ can result in a set $\phi.\msg(s)$ that does not correspond to the graph of any time series, as illustrated in the \Cref{fig:diffeo}.
-        Thus, time series graphs have more structure than a simple 1D curve \cite{glaunes2008large} and deserve their unique analysis, which will prove fruitful as demonstrated in \ref{section:experiments}.
+        Thus, time series graphs have more structure than a simple 1D curve \cite{glaunes2008large} and deserve their unique analysis, which will prove fruitful as demonstrated in \Cref{section:experiments}.
 
@@
        \paragraph{A representation separating space and time.}
-       We prove that two time series graphs can always be linked by a time transformation composed of a space transformation.
+       We prove that two time series graphs can always be linked by a time transformation composed with a space transformation.
 Moreover, a time series graph transformed by this kind of transformation is always a time series graph.

### 2024-05-22 NEURIPS24/paper.tex
update

@@
 
- All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are comapred to TS-LDDMM in \Cref{appendix:classification_shape_analysis}.
-  \section{Limitations}
+ All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix:classification_shape_analysis}.
+  \section{Limitations and conclusion}
   \label{sec:limitations}
+  \paragraph*{Limitations}
   While TS-LDDMM performs well on shape-based datasets, it assumes a certain degree of homogeneity with the existence of a reference time series graph$\msg_0$.
@@
   %\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
-\section{Conclusion}
+\paragraph*{Conclusion}
 \vspace{-1ex}

### 2024-05-22 NEURIPS24/appendix.tex
Changes

@@
 
-\paragraph{TS-LDDMMM parameters.}
+\paragraph{TS-LDDMM parameters.}
 \begin{itemize}
@@
 
-\paragraph{LDDMMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM.
+\paragraph{LDDMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM.
 Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
@@
 
+For each method and dataset, the learning rate, the hidden vector dimensions, and the number of layers are optimized to minimize the CrossEntropy loss on a validation set using the \texttt{Ray}~\footnote{\url{https://github.com/ray- project/ray}} Python library.
+The learning rate varies from $10^{-4}$ to $10^{-1}$ using log uniform search, the hidden vector dimension ranges from ${16, 32, 64, 128}$ using grid search, and the number of layers ranges from ${1, 2, 3, 4}$ using grid search.
+The batch size was selected from ${16, 32, 64, 128}$ according to the size of the dataset.
+All methods were trained for 100 epochs, and the best method was selected based on the lowest validation loss.
+
 \paragraph{TS-LDDMM and LDDMM}
+Representations learned with TS-LDDMM or LDDMM are fed to a Support Vector Classifier (\href{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html#sklearn.svm.SVC}{SVC}) from \texttt{scikit-learn}~\footnote{\url{https://scikit-learn.org/stable/}}.
+All SVC's hyperparameters are set to default except the regularization term C, which is set through grid search on a validation set with the macro f1-score~\footnote{\url{https://scikit-learn.org/stable/modules/generated/sklearn.metrics.f1_score.html}}.
+
+To learn TS-LDDMM (resp.
+LDDMM) representations, the velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,0.33\bar{l},1,n_d)$, (resp.
+$ (\sigma_{T},\sigma_x) = (0.33\bar{l},n_d)$) where $\bar{l}$ is the average time series length and $n_d$ the number of dimensions.
+For both methods and all datasets, the varifold loss kernels $(k_{pos},k_{dir})$ are identical and set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,n_d,2,n_d)$.
+For TS-LDDMM (resp.
+LDDMM), the optimizer is set with 400 epochs (resp.
+400) and a maximum learning rate $\eta_M = 0.1$ (resp.
+$\eta_M = 0.01$).
+In all cases, the initial reference graph is selected in the dataset as a time series with the median length.
 
 \subsection{Protocol}
+In this experiment, we investigate the robustness to missing samples and the classification performance of TS-LDDMM compared to Neural ODEs on 15 datasets described in \Cref{appendix:classification_dataset}.
+For fairness between methods of different architectures, the evaluation protocol on each dataset and method is as follows:
+\begin{enumerate}
+  \item The dataset is spilt in train 75\%, validation 15\%, and test 15\%.
+  \item hyperparameters are tuned with train and validation sets and a missing rate of 0\%.
+  \item For each missing rate in [0\%,30\%,50\%,70\%]
+    \begin{itemize}
+      \item Remove samples in time series in the train and test sets according to the missing rate and the drop procedure described in~\cite{kidger2020neural}.
+      \item Train the model on the train set
+      \item Evaluate the macro f1-score on the test set
+    \end{itemize}
+\end{enumerate}
 
 \subsection{Results}
+In this experiment, we investigate the robstuness to missing samples and the classification performance of TS-LDDMM representations.
+We compare TS-LDDMM with LDDMM and 8 neural ODEs networks.
+Performances are evaluated in terms of average macro f1-score and rank on four different regimes of missing rate 0\%,30\%,50\%, and 70\%.
+Results are aggregated in \Cref{table:robstuness}.
+
+On three out of four regimes (0\%,30\%, and 50\%) TS-LDDMM classifier is the best performer in terms of f1-score and rank.
+For missing rates of 0\% and 30\%, the score increases by 10\% compared to the second-best performer, LDDMM.
+However, LDDMM is not the second-best performer in rank (Neural LNSDE), showing its sensitivity to parameterization, unlike TS-LDDMM, which remains consistent.
+Performances of Neural LNSDE remain constant with the increase of the missing rate as observed in~\cite{oh2024stable}, and it becomes the best performer for missing rate 70\%.
+The decrease in TS-LDDMM performances with the increasing missing rate is due to the varifold loss, which poorly approximates the time series shape.
+Other losses might be more relevant for high missing rates.
+
+Overall, TS-LDDMM is a relevant and consistent shape-based representation for irregularly sampled multivariate time series for missing rates up to 50\% .
 
@@
   \centering
-  \begin{tabular}{lcccc}
+  \caption{Comparison of average macro f1-score and rank as the sample dropping rate increases.
+\textbf{First} \& \underline{second} best performers.
+TS-LDDMM is the best performer on three out of four regimes.}
+  \label{table:robstuness}
+  \resizebox{\columnwidth}{!}{%
+  \begin{tabular}{lcccccccc}
     \toprule
-    \multirow[c]{2}{*}{\textbf{Methods}} & \multicolumn{4}{c}{\textbf{Test F1-score}} \\
-    \cline{2-5}
-     &  \textbf{Regular} & \textbf{30 \% dropped} &  \textbf{50 \% dropped} & \textbf{70 \% dropped} \\
+    \multirow[c]{2}{*}{\textbf{Methods}} & \multicolumn{2}{c}{\textbf{Regular}} & \multicolumn{2}{c}{\textbf{30 \% dropped}} &  \multicolumn{2}{c}{\textbf{50 \% dropped}} & \multicolumn{2}{c}{\textbf{70 \% dropped}} \\
+    \cline{2-9}
+     &  \textbf{F1-score} & \textbf{Rank} &  \textbf{F1-score} & \textbf{Rank} &  \textbf{F1-score} & \textbf{Rank} &  \textbf{F1-score} & \textbf{Rank} \\
     \midrule
-    RNN (1999) & $0.64 \pm 0.21$ & $0.53 \pm 0.23$ & $0.48 \pm 0.21$ & $0.44 \pm 0.21$ \\
-    LSTM (1997) & $0.61 \pm 0.29$ & $0.57 \pm 0.29$ & $0.53 \pm 0.25$ & $0.51 \pm 0.29$ \\
-    GRU (2014) & $0.71 \pm 0.26$ & $0.68 \pm 0.28$ & $0.66 \pm 0.28$ & $\underline{0.59 \pm 0.28}$ \\
-    MTAN (2021) & $0.59 \pm 0.28$ & $0.58 \pm 0.28$ & $0.54 \pm 0.29$ & $0.51 \pm 0.28$ \\
-    MIAM (2022) & $0.48 \pm 0.35$ & $0.42 \pm 0.33$ & $0.47 \pm 0.31$ & $0.35 \pm 0.31$ \\
-    ODE-LSTM (2020) & $0.63 \pm 0.24$ & $0.57 \pm 0.25$ & $0.51 \pm 0.24$ & $0.45 \pm 0.23$ \\
-    Neural SDE (2019) & $0.48 \pm 0.28$ & $0.47 \pm 0.26$ & $0.45 \pm 0.27$ & $0.45 \pm 0.25$ \\
-    Neural LNSDE (2024) & $0.7 \pm 0.27$ & $0.68 \pm 0.29$ & $\underline{0.67 \pm 0.25}$ & $\mathbf{0.66 \pm 0.23}$ \\
-    LDDMM (2008) & $\underline{0.72 \pm 0.2}$ & $\underline{0.7 \pm 0.21}$ & $0.57 \pm 0.25$ & $0.4 \pm 0.25$ \\
-    %\cline{1-5}
-    TS-LDDMM (ours) & $\mathbf{0.83 \pm 0.18}$ & $\mathbf{0.8 \pm 0.18}$ & $\mathbf{0.7 \pm 0.26}$ & $0.51 \pm 0.27$ \\
+    RNN (1999) & $0.64 \pm 0.21$  & 6.2 & $0.53 \pm 0.23$  & 6.6 & $0.48 \pm 0.21$  & 7.2 & $0.44 \pm 0.21$  & 6.07 \\
+    LSTM (1997) & $0.61 \pm 0.29$  & 6.0 & $0.57 \pm 0.29$  & 6.27 & $0.53 \pm 0.25$  & 6.07 & $0.51 \pm 0.29$  & 5.27 \\
+    GRU (2014) & $0.71 \pm 0.26$  & 4.2 & $0.68 \pm 0.28$  & 4.27 & $0.66 \pm 0.28$  & 3.73 & $\underline{0.59 \pm 0.28}$  & \underline{3.67} \\
+    MTAN (2021) & $0.59 \pm 0.28$  & 7.13 & $0.58 \pm 0.28$  & 5.8 & $0.54 \pm 0.29$  & 5.33 & $0.51 \pm 0.28$  & 5.0 \\
+    MIAM (2022) & $0.48 \pm 0.35$  & 6.93 & $0.42 \pm 0.33$  & 8.27 & $0.47 \pm 0.31$  & 6.93 & $0.35 \pm 0.31$  & 7.6 \\
+    ODE-LSTM (2020) & $0.63 \pm 0.24$  & 6.0 & $0.57 \pm 0.25$  & 6.53 & $0.51 \pm 0.24$  & 7.27 & $0.45 \pm 0.23$  & 6.73 \\
+    Neural SDE (2019) & $0.48 \pm 0.28$  & 7.67 & $0.47 \pm 0.26$  & 7.47 & $0.45 \pm 0.27$  & 7.13 & $0.45 \pm 0.25$  & 6.0 \\
+    Neural LNSDE (2024) & $0.7 \pm 0.27$  & \underline{3.87} & $0.68 \pm 0.29$  & \underline{4.0} & $\underline{0.67 \pm 0.25}$  & \underline{3.53} & $\mathbf{0.66 \pm 0.23}$  & \textbf{2.47} \\
+    LDDMM (2008) & $\underline{0.72 \pm 0.2}$  & 4.53 & $\underline{0.7 \pm 0.21}$  & 4.2 & $0.57 \pm 0.25$  & 5.0 & $0.4 \pm 0.25$  & 7.13 \\
+    TS-LDDMM (ours) & $\mathbf{0.83 \pm 0.18}$  & \textbf{2.93} & $\mathbf{0.8 \pm 0.18}$  & \textbf{2.07} & $\mathbf{0.7 \pm 0.26}$  & \textbf{3.33} & $0.51 \pm 0.27$  & 5.67 \\
     \bottomrule
   \end{tabular}  
+  %
+  }
 \end{table}
@@
 \subsection{Appendix for experiment: Classification benchmark on regularly sampled datasets}
+
+In this section, we compare the classification performances of TS-LDDMM with other methods from shape analysis on 15 shape-based datasets of time series regularly sampled.
+TS-LDDMM outperforms other methods on 12 out of 15, highlighting its relevance for shape analysis when dealing with time series.
+
+\subsection{Benchmark metohds}
+\begin{itemize}
+  \item SRV-based method: we include TCLR~\cite{heo2024logistic} a logistic regression on the tangent space of the Frechet mean with Square Root Velocity (SRV representation).
+We also include Shape-FPCA~\cite{wu2024shape} that encodes both the time series and its time parameterization.
+Both methods are implemented with the Python library \texttt{FDASRSF}~\footnote{https://fdasrsf-python.readthedocs.io/en/latest/}, which cannot handle multivariate time series.
+  \item LDDMM-Based : TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}.
+Both methods learn representations by solving ODEs parametrized with Kernels.
+While both methods handle multivariate signals, TS-LDDMM is specifically designed for time series.
+ \end{itemize}
 \subsection{Settings}
 \label{appendix: settings_classification}
-
-In this section, we compare classification performances of TS-LDDMM with other state-of-the-art methods coming from shape analysis on 15 shape-based datasets of time-series.
 

### 2024-05-22 NEURIPS24/appendix.tex
update

@@
 First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
-All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
+`%All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
 We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
@@
 
-\subsection{Appendix for experiment: Classification benchmark on regularly sampled datasets}
+\section{Appendix for experiment: Classification benchmark on regularly sampled datasets}
 
@@
 
-\paragraph{Methods} We compare TS-LDDMM with a method from function~\cite{wu2024shape}
+%\paragraph{Methods} We compare TS-LDDMM with a method from function~\cite{wu2024shape}
 
-\paragraph{Protocole}
+%\paragraph{Protocole}
 

### 2024-05-22 NEURIPS24/checklist.tex
update

@@
     \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: The optimization methodology is described in \Cref{section:methodology} and all numerical details and protocol are given in \Cref{section:experiments} and in \Cref{appendix:classification_implementation}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix:identifiability}.
+    \item[] Justification: The optimization methodology is described in \Cref{section:methodology} and all numerical details and protocol are given in \Cref{section:experiments} and in \Cref{appendix: settings_classification}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix: settings_classification}.
     \item[] Guidelines:
@@
     \item[] Justification: The experiments setting is clearly presented in \Cref{section:experiments} for the experiments on the mouse dataset.
-The full details of all experiments are presented in \Cref{appendix:classification_implementation}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix:identifiability}.
+The full details of all experiments are presented in \Cref{appendix: benchmark robustness}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix: settings_identifiability}.
     \item[] Guidelines:

### 2024-05-22 NEURIPS24/experiment.tex
update

@@
 
-For consiceness, several experiments are relagated in appendix:
+For conciseness, several experiments are relegated in appendix:
 
 \begin{enumerate}
-    \item \textbf{TS-LDDMM representation identifiability, \Cref{appendix:identifiability}:} On synthetic data, we evaluate the ability of our method to retrieve the parameter $v_0^*$ that encodes the deformation $\varphi^{\{v_0^*\}}$ acting on a time series graph $\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation} between $\msg$ and $\varphi^{\{v_0^*\}}.\msg$.
+    \item \textbf{TS-LDDMM representation identifiability, \Cref{appendix: settings_identifiability}:} On synthetic data, we evaluate the ability of our method to retrieve the parameter $v_0^*$ that encodes the deformation $\varphi^{\{v_0^*\}}$ acting on a time series graph $\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation} between $\msg$ and $\varphi^{\{v_0^*\}}.\msg$.
      \textbf{Results} show that TS-LDDMM representations are identifiable or weakly identifiable depending on the velocity field kernel $K_G$ specification.
 
-    \item \textbf{Robustness to irregular sampling} We compare the robustness of TS-LDDMM representation with 9 URL methods handling irregularly sampled multivariate time series on 15 shape-based datasets (7 univariates \& 8 multivariates).
+    \item \textbf{Robustness to irregular sampling, \Cref{appendix: benchmark robustness}:} We compare the robustness of TS-LDDMM representation with 9 URL methods handling irregularly sampled multivariate time series on 15 shape-based datasets (7 univariates \& 8 multivariates).
      We assess methods' classification performances under regular sampling (0\% missing rate) and three irregular sampling regimes (30\%, 50\%, and 70\% missing rates), according to the protocol depicted in \cite{kidger2020neural}.
@@
 
-    \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix:classification_shape_analysis}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates).
-\textbf{Results} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.
+    \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix: settings_classification}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates).
+\textbf{Results} in \Cref{table:wewin} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.
 \end{enumerate}
-
-\vspace{-1ex}
-\subsection{Interpretability: analysis of respiratory behavior in mice}
-\label{section:interpretability}
-\vspace{-1ex}
-%We consider a dataset composed of mouse respiratory time series before and after a drug injection.
-% A complete presentation of this dataset is given in \Cref{appendix:mouse_dataset}.
-%  The mouse are divided in two groups depending on their genotypes : \textit{colq} and \textit{wt}.
-%  We aim to study the difference of respiratory shapes according to the genotype.
-%   That is why, TS-LDDMM \eqref{eq:general_optimization_problem} is applied to derive the features representations $(\mathbf{\alpha}_j)_{j\in[N]}\in (\Rset^2)^N$ related to $N$=\sam{fill} respiratory time series coming from 14 mouse (7 \textit{colq} and \textit{wt}) before drug injection.
-%   Then, a Principal Components Analysis (PCA) is performed on $(\mathbf{\alpha}_j)_{j\in[N]}$
 \begin{figure*}[t]
@@
 
-\begin{wrapfigure}{r}{0.50\textwidth}
-  \centering
-  \begin{subfigure}[b]{0.16\textwidth}
-    \centering
-    \includegraphics[width = \textwidth]{pictures/exp_1_cycle_example.jpeg}
-    \caption{colq cycle}
-    \label{fig:colq-cycle}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.16\textwidth}
-    \centering
-    \includegraphics[width = \textwidth]{pictures/exp_1_scatter.jpeg}
-    \caption{PC1 vs PC2}
-    \label{fig: pcs-scatter}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.16\textwidth}
-    \centering
-    \includegraphics[width = \textwidth]{pictures/exp_1_wt.jpeg}
-    \caption{wt cycle}
-    \label{fig:wt-reference}
-  \end{subfigure}
-  \caption{\Cref{fig:colq-cycle} is a \textbf{colq} respiratory cycle.
-\Cref*{fig: pcs-scatter} displays individual mouse reference respiratory cycle in the TS-LDDMM PC1-PC2 coordinates.
-\Cref{fig:wt-reference} is a referent respiratory cycle of a \textbf{wt} mouse learned with TS-LDDMM.
-}
-  \vspace{-1em}
-\end{wrapfigure}
+\vspace{-1ex}
+\subsection{Interpretability: analysis of respiratory behavior in mice}
+\label{section:interpretability}
+\vspace{-1ex}
+%We consider a dataset composed of mouse respiratory time series before and after a drug injection.
+% A complete presentation of this dataset is given in \Cref{appendix:mouse_dataset}.
+%  The mouse are divided in two groups depending on their genotypes : \textit{colq} and \textit{wt}.
+%  We aim to study the difference of respiratory shapes according to the genotype.
+%   That is why, TS-LDDMM \eqref{eq:general_optimization_problem} is applied to derive the features representations $(\mathbf{\alpha}_j)_{j\in[N]}\in (\Rset^2)^N$ related to $N$=\sam{fill} respiratory time series coming from 14 mouse (7 \textit{colq} and \textit{wt}) before drug injection.
+%   Then, a Principal Components Analysis (PCA) is performed on $(\mathbf{\alpha}_j)_{j\in[N]}$
 
@@
 description of the dataset is given in the \Cref{appendix:mouse_dataset}.
-\vspace{-1ex}
-\paragraph{Result summary.} By comparing the shape of individual respiratory cycles, we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule contrary to LDDMM \cite{glaunes2008large}.
-\vspace{-1ex}
+
 \paragraph{Protocol.} For both experiments and representations (TS-LDDMM and LDDMM), we derive the reference respiratory cycle's graph $S_0$ and the representations $(\alpha_j)_{j\in[N]}$ related to $N$ respiratory cycles extracted according to the procedure \cite{germain2023unsupervised} by solving  \eqref{eq:general_optimization_problem}.
@@
     Essentially, varifold losses are identical for both methods, and the velocity field kernels are set to encompass time and space scales. 
+    \begin{table}[t]
+      \centering
+      \resizebox{\columnwidth}{!}{%
+      \begin{tabular}{llrrrr}
+        \toprule
+         & \textbf{Dataset} & \textbf{Shape-FPCA (2024)} & \textbf{TCLR (2024)} & \textbf{LDDMM (2008)} & \textbf{TS-LDDMM (ours)} \\
+        \midrule
+        \multirow[c]{7}{*}{Univariate} & ArrowHead & 0.18 & 0.75 & \underline{0.84} & \textbf{0.91} \\
+         & BME & 0.16 & \underline{1.00} & 0.82 & \textbf{1.00} \\
+         & ECG200 & 0.40 & 0.67 & \textbf{0.81} & \underline{0.79} \\
+         & FacesUCR & 0.08 & \underline{0.73} & 0.69 & \textbf{0.86} \\
+         & GunPoint & 0.93 & \underline{0.97} & 0.83 & \textbf{1.00} \\
+         & PhalangesOutlinesCorrect & 0.39 & \textbf{0.63} & \underline{0.53} & 0.52 \\
+         & Trace & 0.55 & \underline{1.00} & 0.46 & \textbf{1.00} \\
+        \cline{1-6}
+        \multirow[c]{8}{*}{Multivariate} & ArticularyWordRecognition & -- & -- & \underline{0.98} & \textbf{1.00} \\
+         & Cricket & -- & -- & \underline{0.77} & \textbf{0.93} \\
+         & ERing & -- & -- & \underline{0.95} & \textbf{0.98} \\
+         & Handwriting & -- & -- & \underline{0.22} & \textbf{0.44} \\
+         & Libras & -- & -- & \underline{0.56} & \textbf{0.60} \\
+         & NATOPS & -- & -- & \underline{0.82} & \textbf{0.82} \\
+         & RacketSports & -- & -- & \textbf{0.83} & \underline{0.79} \\
+         & UWaveGestureLibrary & -- & -- & \underline{0.72} & \textbf{0.81} \\
+        \bottomrule
+        \end{tabular}
+        %
+      }
+      \label{table:wewin}
+      \caption{All the state-of-the-art methods \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} mentionned in related works are compared to TS-LDDMM in \Cref{appendix: settings_classification}.}
 
+    \end{table}
     \vspace{-1ex}
+    \vspace{-1ex}\begin{wrapfigure}{r}{0.5\textwidth}
+      \centering
+      \begin{subfigure}[b]{0.15\textwidth}
+        \centering
+        \includegraphics[width = \textwidth]{pictures/exp_1_cycle_example.jpeg}
+        \caption{colq cycle}
+        \label{fig:colq-cycle}
+      \end{subfigure}
+      \begin{subfigure}[b]{0.15\textwidth}
+        \centering
+        \includegraphics[width = \textwidth]{pictures/exp_1_scatter.jpeg}
+        \caption{PC1 vs PC2}
+        \label{fig: pcs-scatter}
+      \end{subfigure}
+      \begin{subfigure}[b]{0.15\textwidth}
+        \centering
+        \includegraphics[width = \textwidth]{pictures/exp_1_wt.jpeg}
+        \caption{wt cycle}
+        \label{fig:wt-reference}
+      \end{subfigure}
+      \caption{\Cref{fig:colq-cycle} is a \textbf{colq} respiratory cycle.
+\Cref*{fig: pcs-scatter} displays individual mouse reference respiratory cycle in the TS-LDDMM PC1-PC2 coordinates.
+\Cref{fig:wt-reference} is a referent respiratory cycle of a \textbf{wt} mouse learned with TS-LDDMM.
+}
+      \vspace{-1em}
+    \end{wrapfigure}
 \paragraph{Breathing behaviors before exposure.}
@@
   Conjointly, \Cref{fig:exp_2_PCA}-A) shows a bimodal distribution for \textbf{wt} mice whereas \textbf{wt}' distributions before exposure were unimodal (\Cref{fig:ts-lddmm shooting}). Indeed, the irritant molecule inhibits the action of the deficient enzyme, \textbf{wt} mice strongly react to the irritant molecule, whereas \textbf{colq} mice better adapt due to their deficiency \cite{germain2023unsupervised}.
+%   \vspace{-1ex}
+% \paragraph{Result summary.} By comparing the shape of individual respiratory cycles, we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule contrary to LDDMM \cite{glaunes2008large}.
+% \vspace{-1ex}
 %Secondly, we compare breathing behaviors before and after exposure to observe the impact of the irritant molecule.

### 2024-05-22 NEURIPS24/paper.tex
update

@@
        \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters (\Cref{appendix:identifiability}), also providing guidelines for tuning (\Cref{appendix:kernel_TS_LDDMM}).
+       \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters (\Cref{appendix: settings_identifiability}), also providing guidelines for tuning (\Cref{appendix:kernel_TS_LDDMM}).
        \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset (\Cref{section:experiments}).
-        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets with regular and irregular sampling (\Cref{appendix:classification_shape_analysis}).
+        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets with regular and irregular sampling (\Cref{appendix: robustness}-\ref{appendix: settings_classification}).
  \end{itemize}
@@
 
- LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
+ %LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
 However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
- In a similar vein, Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
+ %In a similar vein, Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
 They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
-
+ Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
+They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
  Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} address the representation of curves with the Square-Root Velocity (SRV) representation.
@@
 
- All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix:classification_shape_analysis}.
+ %All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix: settings_classification}.
+ \vspace{-1ex}
   \section{Limitations and conclusion}
+  \vspace{-1ex}
   \label{sec:limitations}
-  \paragraph*{Limitations}
-  While TS-LDDMM performs well on shape-based datasets, it assumes a certain degree of homogeneity with the existence of a reference time series graph$\msg_0$.
- Extending TS-LDDMM to more heterogeneous datasets using clustering is ongoing work.
- TS-LDDMM employs kernel computations, which require specific libraries (e.g., KeOps \cite{charlier2021kernel}) to be efficient and scalable.
- However, in our experiments, the time complexity of TS-LDDMM is comparable to that of competitors.
- It is clear that TS-LDDMM needs to be extended to handle very large datasets with high-dimensional time series (such as videos).
- Additionally, TS-LDDMM requires tuning several hyperparameters, though this is a common requirement among competitors \cite{glaunes2008large, oh2024stable, wu2024shape, heo2024logistic}.
- In future work, adaptive methods are expected to be developed to provide a more user-friendly interface.
-  %\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
-\paragraph*{Conclusion}
-\vspace{-1ex}
+  %\paragraph*{Limitations}
   This paper proposes a feature representation method, TS-LDDMM, designed for 
@@
 more heterogeneous datasets using a hierarchical clustering extension, which is relegated for future work. 
+%   While TS-LDDMM performs well on shape-based datasets, it assumes a certain degree of homogeneity with the existence of a reference time series graph$\msg_0$.
+%  Extending TS-LDDMM to more heterogeneous datasets using clustering is ongoing work.
+ TS-LDDMM employs kernel computations, which require specific libraries (e.g., KeOps \cite{charlier2021kernel}) to be efficient and scalable.
+ However, in our experiments, the time complexity of TS-LDDMM is comparable to that of competitors.
+ It is clear that TS-LDDMM needs to be extended to handle very large datasets with high-dimensional time series (such as videos).
+ Additionally, TS-LDDMM requires tuning several hyperparameters, though this is a common requirement among competitors \cite{glaunes2008large, oh2024stable, wu2024shape, heo2024logistic}.
+ In future work, adaptive methods are expected to be developed to provide a more user-friendly interface.
+  %\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
+% \paragraph*{Conclusion}
+% \vspace{-1ex}
 

### 2024-05-22 NEURIPS24/experiment.tex
update

@@
     \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix: settings_classification}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates).
-\textbf{Results} in \Cref{table:wewin} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.
+\textbf{Results} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.
 \end{enumerate}
@@
 
+\vspace{-1ex}
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{0.15\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/exp_1_cycle_example.jpeg}
+    \caption{colq cycle}
+    \label{fig:colq-cycle}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.15\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/exp_1_scatter.jpeg}
+    \caption{PC1 vs PC2}
+    \label{fig: pcs-scatter}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.15\textwidth}
+    \centering
+    \includegraphics[width = \textwidth]{pictures/exp_1_wt.jpeg}
+    \caption{wt cycle}
+    \label{fig:wt-reference}
+  \end{subfigure}
+  \caption{\Cref{fig:colq-cycle} is a \textbf{colq} respiratory cycle.
+\Cref*{fig: pcs-scatter} displays individual mouse reference respiratory cycle in the TS-LDDMM PC1-PC2 coordinates.
+\Cref{fig:wt-reference} is a referent respiratory cycle of a \textbf{wt} mouse learned with TS-LDDMM.
+}
+  \vspace{-1em}
+\end{figure*}
 \vspace{-1ex}
@@
     Essentially, varifold losses are identical for both methods, and the velocity field kernels are set to encompass time and space scales.
-    \begin{table}[t]
-      \centering
-      \resizebox{\columnwidth}{!}{%
-      \begin{tabular}{llrrrr}
-        \toprule
-         & \textbf{Dataset} & \textbf{Shape-FPCA (2024)} & \textbf{TCLR (2024)} & \textbf{LDDMM (2008)} & \textbf{TS-LDDMM (ours)} \\
-        \midrule
-        \multirow[c]{7}{*}{Univariate} & ArrowHead & 0.18 & 0.75 & \underline{0.84} & \textbf{0.91} \\
-         & BME & 0.16 & \underline{1.00} & 0.82 & \textbf{1.00} \\
-         & ECG200 & 0.40 & 0.67 & \textbf{0.81} & \underline{0.79} \\
-         & FacesUCR & 0.08 & \underline{0.73} & 0.69 & \textbf{0.86} \\
-         & GunPoint & 0.93 & \underline{0.97} & 0.83 & \textbf{1.00} \\
-         & PhalangesOutlinesCorrect & 0.39 & \textbf{0.63} & \underline{0.53} & 0.52 \\
-         & Trace & 0.55 & \underline{1.00} & 0.46 & \textbf{1.00} \\
-        \cline{1-6}
-        \multirow[c]{8}{*}{Multivariate} & ArticularyWordRecognition & -- & -- & \underline{0.98} & \textbf{1.00} \\
-         & Cricket & -- & -- & \underline{0.77} & \textbf{0.93} \\
-         & ERing & -- & -- & \underline{0.95} & \textbf{0.98} \\
-         & Handwriting & -- & -- & \underline{0.22} & \textbf{0.44} \\
-         & Libras & -- & -- & \underline{0.56} & \textbf{0.60} \\
-         & NATOPS & -- & -- & \underline{0.82} & \textbf{0.82} \\
-         & RacketSports & -- & -- & \textbf{0.83} & \underline{0.79} \\
-         & UWaveGestureLibrary & -- & -- & \underline{0.72} & \textbf{0.81} \\
-        \bottomrule
-        \end{tabular}
-        %
-      }
-      \label{table:wewin}
-      \caption{All the state-of-the-art methods \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} mentionned in related works are compared to TS-LDDMM in \Cref{appendix: settings_classification}.}
 
-    \end{table}
     \vspace{-1ex}
-    \vspace{-1ex}\begin{wrapfigure}{r}{0.5\textwidth}
-      \centering
-      \begin{subfigure}[b]{0.15\textwidth}
-        \centering
-        \includegraphics[width = \textwidth]{pictures/exp_1_cycle_example.jpeg}
-        \caption{colq cycle}
-        \label{fig:colq-cycle}
-      \end{subfigure}
-      \begin{subfigure}[b]{0.15\textwidth}
-        \centering
-        \includegraphics[width = \textwidth]{pictures/exp_1_scatter.jpeg}
-        \caption{PC1 vs PC2}
-        \label{fig: pcs-scatter}
-      \end{subfigure}
-      \begin{subfigure}[b]{0.15\textwidth}
-        \centering
-        \includegraphics[width = \textwidth]{pictures/exp_1_wt.jpeg}
-        \caption{wt cycle}
-        \label{fig:wt-reference}
-      \end{subfigure}
-      \caption{\Cref{fig:colq-cycle} is a \textbf{colq} respiratory cycle.
-\Cref*{fig: pcs-scatter} displays individual mouse reference respiratory cycle in the TS-LDDMM PC1-PC2 coordinates.
-\Cref{fig:wt-reference} is a referent respiratory cycle of a \textbf{wt} mouse learned with TS-LDDMM.
-}
-      \vspace{-1em}
-    \end{wrapfigure}
 \paragraph{Breathing behaviors before exposure.}

### 2024-05-22 NEURIPS24/paper.tex
update

@@
 
- %LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
+ LDDMM framework is a relevant shape analysis tool to represent curves as depicted in \cite{glaunes2008large}.
 However, graphs of time series are a well-structured type of curve due to the inclusion of the temporal dimension that requires specific care (\Cref{fig:diffeo}).
- %In a similar vein, Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
+ In a similar vein, Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
 They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
- Qiu \textit{et al} \cite{qiu2009time} proposes a method for tracking anatomical shape changes in serial images using LDDMM.
-They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
+
  Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} address the representation of curves with the Square-Root Velocity (SRV) representation.
@@
 
- %All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix: settings_classification}.
+ All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix: settings_classification}.
  \vspace{-1ex}
@@
   \label{sec:limitations}
-  %\paragraph*{Limitations}
+
   This paper proposes a feature representation method, TS-LDDMM, designed for 
@@
 variability. As an unsupervised approach, it is user-friendly and enables knowledge 
-transfer for different supervised tasks such as classification. Although TS-LDDMM 
+transfer for different supervised tasks such as classification.
+
+Although TS-LDDMM 
 is already competitive for classification, its performances can be leveraged on 
@@
 %  Extending TS-LDDMM to more heterogeneous datasets using clustering is ongoing work.
+
  TS-LDDMM employs kernel computations, which require specific libraries (e.g., KeOps \cite{charlier2021kernel}) to be efficient and scalable.
@@
  It is clear that TS-LDDMM needs to be extended to handle very large datasets with high-dimensional time series (such as videos).
+
  Additionally, TS-LDDMM requires tuning several hyperparameters, though this is a common requirement among competitors \cite{glaunes2008large, oh2024stable, wu2024shape, heo2024logistic}.

### 2024-05-22 NEURIPS24/appendix.tex
final

@@
 \section{Appendix for experiment: TS-LDDMM representation identifiability}
+In this experiment, we evaluate the ability of TS-LDDMM to retrieve the parameter $v_0^*$ that encodes the deformation $\varphi^{\{v_0^*\}}$ acting on a time series graph $\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation} between $\msg$ and $\varphi^{\{v_0^*\}}.\msg$.
+Parameter identifiability is an important property for subsequent statistical analysis.
+Results show that TS-LDDMM representations are identifiable or weakly identifiable depending on the velocity field kernel $K_G$ specification.
+
 \subsection{Settings}
@@
   \includegraphics[width=0.5\linewidth]{pictures/samples.pdf}
-  \vspace{-1em}
   \caption{Plots of $\varphi^{\{v_0(\mathbf{\alpha}^*,\msx)\}}.\msx$ for different values of $\mathbf{\alpha}^*$ according to its sampling parameter $t_a,s_a,m_s $, taking $\msx=\msg(s_0)$ with $s_0:k\in [300]\to \sin(2\pi k/300) $.}
   \label{fig:exemple_synthetic}
-  \vspace{-1em}
 \end{figure}
@@
     \label{table:synthetic2}
-    \vspace{-1em}
 \end{table}
@@
 
+\paragraph{The velocity field kernel $K_G$ is well specified.}
+
 First, we show the model identifiability when the kernel $K_G$ is well specified: the estimated parameter is a good approximation of the generating parameter when the generation and the estimation procedure use the same hyperparameters for the RKHS kernel $K_\msg$.
-`%All the hyperparameter values for generation and estimation are given in \Cref{appendix:numerics_synthetic}.
+All the hyperparameter values for generation and estimation are given in \Cref{appendix: settings_identifiability}.
+
 We fix the initial control points as $\msx=\left(x_k=(k,\sin(2\pi k/300))\right)_{k\in[300]} $.
@@
 
-Secondly, we demonstrate a weak identifiability when the kernel $K_\msg$ is misspecified: we can reconstruct the graph time series' after deformations even if the hyperparameters of $K_\msg$
+\paragraph{The velocity field kernel $K_G$ is misspecified.}
+
+We demonstrate a weak identifiability when the kernel $K_\msg$ is misspecified: we can reconstruct the graph time series' after deformations even if the hyperparameters of $K_\msg$
 are different during the generation and the estimation.
@@
 \subsection{Protocol}
+
 In this experiment, we investigate the robustness to missing samples and the classification performance of TS-LDDMM compared to Neural ODEs on 15 datasets described in \Cref{appendix:classification_dataset}.
@@
 \begin{enumerate}
-  \item The dataset is spilt in train 75\%, validation 15\%, and test 15\%.
-  \item hyperparameters are tuned with train and validation sets and a missing rate of 0\%.
+  \item Spilt the dataset in train 75\%, validation 15\%, and test 15\%.
+  \item Tune hyperparameters with train and validation sets and a missing rate of 0\%.
   \item For each missing rate in [0\%,30\%,50\%,70\%]
@@
 We also include Shape-FPCA~\cite{wu2024shape} that encodes both the time series and its time parameterization.
-Both methods are implemented with the Python library \texttt{FDASRSF}~\footnote{https://fdasrsf-python.readthedocs.io/en/latest/}, which cannot handle multivariate time series.
   \item LDDMM-Based : TS-LDDMM (ours) and LDDMM~\cite{glaunes2008large}.
@@
  \end{itemize}
-\subsection{Settings}
-\label{appendix: settings_classification}
 
-%\paragraph{Methods} We compare TS-LDDMM with a method from function~\cite{wu2024shape}
+\subsection{Model settings}
 
-%\paragraph{Protocole}
+\paragraph{TCLR \& Shape-FPCA}
+Shape-FPCA is available in the Python library \texttt{FDASRSF}~\footnote{https://fdasrsf-python.readthedocs.io/en/latest/}.
+Once the shape-FPCA representations are learned, they are fed to an \href{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html#sklearn.svm.SVC}{SVC} from \texttt{scikit-learn}.
+\texttt{FDASRSF} provides SRV representation methods that we combined with a \href{https://scikit-learn.org/stable/modules/generated/sklearn.linear_model.LogisticRegressionCV.html}{logistic regression} from \texttt{scikit-learn} to implement TCLR.
+For both methods, the number of steps to learn the Frechet mean is set to 50, and the regularization hyperparameter C is set through grid search on a validation set with the macro f1-score.
+Other parameters are set to default.
+
+\paragraph{TS-LDDMM \& LDDMM}
+Representations learned with TS-LDDMM or LDDMM are fed to an SVC from \texttt{scikit-learn}.
+All SVC's hyperparameters are set to default except the regularization term C, which is set through grid search on a validation set with the macro f1-score.
+
+To learn TS-LDDMM (resp.
+LDDMM) representations, the velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,0.33\bar{l},1,n_d)$, (resp.
+$ (\sigma_{T},\sigma_x) = (0.33\bar{l},n_d)$) where $\bar{l}$ is the average time series length and $n_d$ the number of dimensions.
+For both methods and all datasets, the varifold loss kernels $(k_{pos},k_{dir})$ are identical and set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (2,n_d,2,n_d)$.
+For TS-LDDMM (resp.
+LDDMM), the optimizer is set with 400 epochs (resp.
+400) and a maximum learning rate $\eta_M = 0.1$ (resp.
+$\eta_M = 0.01$).
+In all cases, the initial reference graph is selected in the dataset as a time series with the median length.
+
+\paragraph{Protocole}
+For each dataset and method, the evaluation protocol is a simple train,validation test with hyperparameter tuning:
+\begin{enumerate}
+  \item Split The dataset in train 75\%, validation 15\%, and test 15\%.
+  \item Training and hyperparameters tuning with train and validation sets
+  \item Evaluate the macro f1-score on the test set 
+\end{enumerate}
+
+\subsection{Results}
+In this experiment, we investigate the classification performances of several methods from shape analysis on 15 shape-based time series datasets (7 univariate and 8 multivariate).
+The performances are evaluated in terms of macro f1-score.
+Results are aggregated in \Cref{table: shape classif}.
+
+The TS-LDDMM-based classifier outperforms other methods on 12 out of 15 datasets.
+TCLR is the second-best performer on univariate datasets; however, its current implementation with \texttt{FDASRSF} does not extend to the multivariate case, which limits usage.
+LDDMM performances are lower than TCLR, and Shape-FPCA is the worst performer.
+
+Overall, TS-LDDMM representations are well suited for shape-based time series classification, and its extension to multivariate irregularly sampled time series makes it a relevant option for time series shape analysis.
 
@@
   \centering
+  \caption{F1-score comparison between methods from shape analysis on 15 datasets. \textbf{First} and \underline{second} best performers.}
+  \label{table: shape classif}
   \resizebox{\columnwidth}{!}{%

### 2024-05-22 NEURIPS24/checklist.tex
label

--- 2024-05-22 NEURIPS24/checklist.tex
@@
     \item[] Answer: \answerYes{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: The optimization methodology is described in \Cref{section:methodology} and all numerical details and protocol are given in \Cref{section:experiments} and in \Cref{appendix:classification_implementation}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix:identifiability}.
+    \item[] Justification: The optimization methodology is described in \Cref{section:methodology} and all numerical details and protocol are given in \Cref{section:experiments} and in \Cref{appendix:optimizers_details}-\ref{appendix: identifiability}-\Cref{appendix: robustness}-\Cref{appendix: shape_classification}--\Cref{appendix: mice_exp_setting}.
     \item[] Guidelines:
@@
     \item[] Justification: The experiments setting is clearly presented in \Cref{section:experiments} for the experiments on the mouse dataset.
-The full details of all experiments are presented in \Cref{appendix:classification_implementation}-\ref{appendix:optimizers_details}-\ref{appendix:numerics_synthetic}-\ref{appendix:identifiability}.
+The full details of all experiments are presented in \Cref{appendix: identifiability}-\Cref{appendix: robustness}-\Cref{appendix: shape_classification}-\ref{appendix:optimizers_details}.
     \item[] Guidelines:

### 2024-05-24 Identifiability/misspecified.py
add code

@@ -1,0 +1,30 @@
+import sys
+sys.path.insert(0,"../")
+
+import numpy as np
+import pandas as pd
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+
+from utils import Sampler,TASExperiment
+
+from src.kernel import TSGaussGaussKernel,VFTSGaussKernel
+from src.lddmm import Shooting
+from src.utils import time_shape_embedding
+from src.loss import SumVarifoldLoss
+
+functional_signal = lambda x:  np.sin(2*np.pi*x)
+q0 = time_shape_embedding(functional_signal(np.linspace(0,1,300)).reshape(-1,1),)
+q0_mask = np.full_like(q0[:,:1],True).astype(bool)
+Kv = VFTSGaussKernel(1,0.1,100,1,1)
+shoot = Shooting(Kv)
+
+spl = Sampler(q0,q0_mask,shoot,[10],[10],[50],0)
+ps,qs,df= spl.rvs(50)
+qs_mask = np.full_like(qs[:,:,:1],True).astype(bool)
+Kls = [TSGaussGaussKernel(2,1,2,0.6)]
+dataloss = SumVarifoldLoss(Kls)
+
+tasexp = TASExperiment(1,0.1,[1,5,10,50,100,200,300],1,[0.1,1,10,100],q0,q0_mask,dataloss,20,gamma_loss=0,niter=400,optimizer=optax.adabelief(warmup_cosine_decay_schedule(0,0.05,40,400,0)))
+tasexp.fit(ps,qs,qs_mask)
+tasexp.df_.to_csv("./results/misspecified.csv")

### 2024-05-24 Identifiability/utils.py

@@ -1,0 +1,110 @@
+import sys
+sys.path.insert(0,"../")
+
+import numpy as np
+import pandas as pd
+import jax.numpy as jnp
+from jax.lax import map
+from jax import vmap
+from sklearn.utils import check_random_state
+from functools import partial
+from scipy.ndimage import convolve1d
+import optax
+
+
+
+from src.lddmm import Shooting,batch_one_to_many_registration
+from src.utils import batch_dataset
+from src.loss import MomentaLoss
+
+
+
+class Sampler: 
+
+    def __init__(self,q0,q0_mask,shoot,t_amp,s_amp,smoothness,random_state=None) -> None:
+        self.q0 = q0 
+        self.q0_mask = q0_mask
+        self.shoot = shoot
+        self.t_amp =t_amp
+        self.s_amp = s_amp
+        self.smoothness = smoothness
+        self.random_state = check_random_state(random_state)
+
+    def rvs(self,n_samples): 
+        p_lst = []
+        i_lst = []
+        for smoothness in self.smoothness: 
+            for t_amp in self.t_amp: 
+                for s_amp in self.s_amp: 
+                    p = self.random_state.randn(n_samples,*self.q0.shape)
+                    p = convolve1d(p,np.ones(smoothness)/smoothness,axis=1,mode="constant")
+                    t_norms = np.linalg.norm(p[:,:,0],axis=(1))
+                    s_norms = np.linalg.norm(p[:,:,1:],axis=(1,2))
+                    p[:,:,:1]*=t_amp/t_norms[:,None,None]
+                    p[:,:,1:]*=s_amp/s_norms[:,None,None]
+                    p_lst.append(p)
+                    i_lst += [[smoothness,t_amp,s_amp]]*n_samples
+
+        self.df_ = pd.DataFrame(i_lst,columns=["smothness", "t_amp", "s_amp"])
+        self.p_ = np.concatenate(p_lst)
+        self.q_ = map(partial(self.shoot,q0=self.q0,q0_mask=self.q0_mask),self.p_)[1]
+        return self.p_, self.q_, self.df_
+
+class TASExperiment: 
+
+    def __init__(self,Kv,q0,q0_mask,dataloss,batch_size = 6,gamma_loss=0.001,niter=100,optimizer = optax.adam(learning_rate=0.1),nt=10,deltat=1.0,verbose=True) -> None:
[…]

### 2024-05-24 Identifiability/wellspecified.py

@@ -1,0 +1,33 @@
+import sys
+sys.path.insert(0,"../")
+
+import numpy as np
+import pandas as pd
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+
+from utils import Sampler,TASExperiment
+
+from src.kernel import TSGaussGaussKernel,VFTSGaussKernel
+from src.lddmm import Shooting
+from src.utils import time_shape_embedding
+from src.loss import SumVarifoldLoss
+
+functional_signal = lambda x:  np.sin(2*np.pi*x)
+q0 = time_shape_embedding(functional_signal(np.linspace(0,1,300)).reshape(-1,1),)
+q0_mask = np.full_like(q0[:,:1],True).astype(bool)
+Kv = VFTSGaussKernel(1,0.1,100,1,1)
+shoot = Shooting(Kv)
+
+spl = Sampler(q0,q0_mask,shoot,[5,10,15,20],[5,10,15,20],[10,50,100],0)
+ps,qs,df= spl.rvs(50)
+
+qs_mask = np.full_like(qs[:,:,:1],True).astype(bool)
+Kls = [TSGaussGaussKernel(2,1,2,0.6)]
+dataloss = SumVarifoldLoss(Kls)
+tasexp = TASExperiment(Kv,q0,q0_mask,dataloss,20,gamma_loss=0,niter=400,optimizer=optax.adabelief(warmup_cosine_decay_schedule(0,0.05,40,400,0)))
+tasexp.fit(ps,qs,qs_mask)
+
+edf = tasexp.error_df_
+fdf = pd.concat((df,edf),axis=1)
+fdf.to_csv("./results/wellspecified.csv")

### 2024-05-24 Robustness--Classification/lddmm_methods/script_missing_data.py

@@ -1,0 +1,100 @@
+import pandas as pd
+import numpy as np 
+import jax.numpy as jnp
+import time 
+import os
+import random
+import sys
+sys.path.insert(0,"../../")
+
+
+from sklearn.metrics import f1_score,accuracy_score
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel,TSGaussKernel
+from src.loss import VarifoldLoss
+from src.classifier import MomentaSVC
+from utils import get_data, from_mask_timeseries_to_dataset
+
+
+def seed_everything(seed):
+    os.environ['PYTHONHASHSEED'] = str(seed)
+    np.random.seed(seed)
+    random.seed(seed)
+
+SEED = 0
+seed_everything(SEED) 
+
+
+valid_datasets = [
+    "ArrowHead",
+    "BME",
+    "ECG200",
+    "FacesUCR",
+    "GunPoint",
+    "PhalangesOutlinesCorrect",
+    "Trace",
+    "Cricket",
+    "ERing",
+    "Handwriting",
+    "Libras",
+    "NATOPS",
+    "RacketSports",
+    "UWaveGestureLibrary",
+    "ArticularyWordRecognition"
+]
+
+estimator_parameters = dict(
+    C=1,
+    max_per_batch = 1,
+    gamma_loss = 0.0,
+    niter = 400,
+    optimizer = optax.adabelief(warmup_cosine_decay_schedule(0,0.1,40,400,0)),
+    nt = 10,
+    deltat = 1.,
+    verbose = True
+)
+
+settings_df = pd.read_csv("settings.csv",index_col=0)
+
+
+for missing_rate in [0.3,0.5,0.7]: 
+    lst = []
+    for model in ['ts-lddmm','lddmm']:
+        for dataset in valid_datasets: 
+
+            #load data
+            X_missing, X_raw_mask, X_delta,Y,train_idx,valid_idx,test_idx = get_data(dataset,missing_rate=missing_rate)
+            X,X_mask = from_mask_timeseries_to_dataset(X_missing,X_raw_mask)
+            X_train,X_train_mask,y_train = X[train_idx],X_mask[train_idx],Y[train_idx]
+            X_val,X_val_mask,y_val = X[valid_idx],X_mask[valid_idx],Y[valid_idx]
+            X_test,X_test_mask,y_test = X[test_idx],X_mask[test_idx],Y[test_idx]
+
+            #set estimator
[…]

### 2024-05-24 Robustness--Classification/lddmm_methods/script_no_missing_data.py

@@ -1,0 +1,131 @@
+import pandas as pd
+import numpy as np 
+import time 
+import os
+import random
+import sys
+sys.path.insert(0,"../../")
+
+from sklearn.metrics import f1_score,accuracy_score
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel,TSGaussKernel
+from src.loss import VarifoldLoss
+from src.classifier import MomentaSVC
+from utils import get_data, from_mask_timeseries_to_dataset
+
+
+def seed_everything(seed):
+    os.environ['PYTHONHASHSEED'] = str(seed)
+    np.random.seed(seed)
+    random.seed(seed)
+
+SEED = 0
+seed_everything(SEED)
+
+
+valid_datasets = [
+    "ArrowHead",
+    "BME",
+    "ECG200",
+    "FacesUCR",
+    "GunPoint",
+    "PhalangesOutlinesCorrect",
+    "Trace",
+    "Cricket",
+    "ERing",
+    "Handwriting",
+    "Libras",
+    "NATOPS",
+    "RacketSports",
+    "UWaveGestureLibrary",
+    "ArticularyWordRecognition"
+]
+
+estimator_parameters = dict(
+    C=1,
+    max_per_batch = 1,
+    gamma_loss = 0.0,
+    niter = 400,
+    optimizer = optax.adabelief(warmup_cosine_decay_schedule(0,0.1,40,400,0)),
+    nt = 10,
+    deltat = 1.,
+    verbose = True
+)
+
+
+def set_VFTSGausGaussKernel(X,X_mask,prop=0.33,sampfreq =1.): 
+    nd = X.shape[2]-1
+    ns = np.mean(np.sum(X_mask.squeeze(),axis=-1))
+    t_sig = max(np.around(np.mean(ns)*prop,0)*sampfreq,2*sampfreq)
+    return VFTSGaussKernel(1,0.1,t_sig,1,nd)
+
+def set_TSGaussKernel(X,X_mask,prop=0.33,sampfreq =1.): 
+    nd = X.shape[2]-1
+    ns = np.mean(np.sum(X_mask.squeeze(),axis=-1))
+    t_sig = max(np.around(np.mean(ns)*prop,0)*sampfreq,2*sampfreq)
+    return TSGaussKernel(t_sig,nd)
+
[…]

### 2024-05-24 Robustness--Classification/lddmm_methods/utils.py

@@ -1,0 +1,62 @@
+import numpy as np
+import pathlib
+import jax.numpy as jnp
+
+valid_datasets = [
+    "ArrowHead",
+    "BME",
+    "ECG200",
+    "FacesUCR",
+    "GunPoint",
+    "PhalangesOutlinesCorrect",
+    "Trace",
+    "Cricket",
+    "ERing",
+    "Handwriting",
+    "Libras",
+    "NATOPS",
+    "RacketSports",
+    "UWaveGestureLibrary",
+    "ArticularyWordRecognition"
+]
+
+here = pathlib.Path(__file__).resolve().parent.parent.parent
+
+def get_data(dataname,missing_rate=0.0):
+    '''
+    get data as tensor
+    '''
+    assert dataname in valid_datasets, "Must specify a valid dataset name."
+    base_filename = here /"datasets"/dataname
+    train_idx = np.load(base_filename/"train_idx.npy")
+    val_idx = np.load(base_filename/"val_idx.npy")
+    test_idx = np.load(base_filename/"test_idx.npy")
+    y = np.load(base_filename/"y.npy")
+    base_filename = here /"datasets"/dataname/f"0.{np.round(missing_rate*10,0).astype(int)}"
+    X = np.load(base_filename/"X.npy")
+    X_mask = np.load(base_filename/"X_mask.npy")
+    X_delta = np.load(base_filename/"X_delta.npy")    
+    return X, X_mask,X_delta,y ,train_idx,val_idx,test_idx
+
+
+def time_shape_mask_embedding(x,x_mask,max_length,sampfreq=1,dtype = jnp.float32):
+    n_s = x.shape[0]
+    time = np.arange(n_s,dtype=float).reshape(-1,1)/sampfreq
+    t_x = np.hstack((time,x),dtype=dtype)
+    t_x = t_x[x_mask.astype(bool)]
+    n_s = t_x.shape[0]
+    t_x = np.pad(t_x,((0,max_length-n_s),(0,0))).astype(dtype)
+    mask = np.full_like(t_x[:,:1],True,dtype=bool)
+    mask[n_s:,:] = False
+    return t_x, mask
+
+def from_mask_timeseries_to_dataset(X:list,X_mask:list,sampfreq=1,dtype=jnp.float32): 
+    lengths = np.array([np.sum(x_mask) for x_mask in X_mask],dtype=int)
+    max_length = np.max(lengths)
[…]

### 2024-05-24 Robustness--Classification/srv_methods/fda.py

@@ -1,0 +1,126 @@
+import numpy as np
+import fdasrsf as fs
+import fdasrsf.utility_functions as uf 
+from sklearn.decomposition import PCA
+from sklearn.linear_model import LogisticRegression
+from sklearn.svm import SVC
+from joblib import Parallel, delayed
+
+def dataset_to_normalized_srvf(dataset,time,smooth=False): 
+    # dataset : (n_ts, n_samples)
+    lst = []
+    for f in dataset: 
+        q = uf.f_to_srsf(f,time,smooth)
+        nq = q/np.sqrt(uf.innerprod_q(time,q,q))
+        lst.append(nq)
+    return np.array(lst)
+
+def shoot_exp(time,qs,v): 
+    norm = np.sqrt(uf.innerprod_q(time,v,v))
+    if norm>0:
+        return np.cos(norm)*qs + np.sin(norm)*(v/norm)
+    else: 
+        return qs
+
+def inv_shoot_exp(time,qs,qt): 
+    inn = uf.innerprod_q(time,qs,qt)
+    if (inn<1.0)*(inn>-1.0):
+        theta = np.arccos(inn)
+        return theta/np.sin(theta)*(qt - inn *qs)
+    else: 
+        return np.zeros_like(qs,dtype=float)
+    
+def compute_intermediate(qbar,time,qt):
+    gam = uf.optimum_reparam(qbar,time,qt)
+    qt_aligned = uf.warp_q_gamma(time,qt,gam)
+    qt_aligned = qt_aligned/np.sqrt(uf.innerprod_q(time,qt_aligned,qt_aligned))
+    vt = inv_shoot_exp(time,qbar,qt_aligned)
+    return gam,qt_aligned,vt
+
+def compute_karcher_mean(time,qs,init=None,niter=20,lr=0.1,random_seed=None,verbose = False,n_jobs=1):
+    #qs shape (m_ts, n_samples)
+    if not random_seed is None: 
+        np.random.seed(random_seed)
+    if init is None: 
+        idx = np.random.choice(qs.shape[0])
+        init = qs[idx]
+    qbar = init.copy()
+    for i in range(niter): 
+        results = Parallel(n_jobs=n_jobs)(delayed(compute_intermediate)(qbar,time,qt) for qt in qs)
+        gams,qs_aligned,vs_aligned = list(zip(*results))
+        vbar = np.mean(vs_aligned,axis=0)
+        qbar = shoot_exp(time,qbar,lr*vbar)
[…]

### 2024-05-24 Robustness--Classification/srv_methods/script_srv.py

@@ -1,0 +1,123 @@
+import pandas as pd
+import numpy as np 
+import time 
+import os
+import random
+import copy
+
+
+from utils import get_data
+from sklearn.model_selection import GridSearchCV,train_test_split
+from sklearn.base import clone
+from sklearn.metrics import f1_score,accuracy_score
+
+from fda import TangentCurveLogisticRegression, FpcaSVM
+
+valid_datasets = [
+    "ArrowHead",
+    "BME",
+    "ECG200",
+    "FacesUCR",
+    "GunPoint",
+    "PhalangesOutlinesCorrect",
+    "Trace",
+    "Cricket",
+    "ERing",
+    "Handwriting",
+    "Libras",
+    "NATOPS",
+    "RacketSports",
+    "UWaveGestureLibrary",
+    "ArticularyWordRecognition"
+]
+
+univariate_datasets = [
+    "ArrowHead", 
+    "BME", 
+    "ECG200",
+    "FacesUCR",
+    "GunPoint", 
+    "PhalangesOutlinesCorrect",
+    "Trace",
+]
+
+def seed_everything(seed):
+    os.environ['PYTHONHASHSEED'] = str(seed)
+    np.random.seed(seed)
+    random.seed(seed)
+
+SEED = 0
+seed_everything(SEED)
+Cs = [100.,10.,1.,0.1]
+
+
+lst = []
+for dataset in univariate_datasets[:1]:
+
+    #load data 
+    X, X_raw_mask, X_delta,Y,train_idx,valid_idx,test_idx = get_data(dataset,missing_rate=0.0)
+    X_train, y_train = X[train_idx].squeeze(),Y[train_idx]
+    X_val, y_val = X[valid_idx].squeeze(), Y[valid_idx]
+    X_test, y_test = X[test_idx].squeeze(), Y[test_idx]
+
+    best_performer = None
+    best_score = -np.inf
+
+    for C in Cs:
+            tclr = TangentCurveLogisticRegression("l2",C=C,karcher_niter=50,n_jobs=30)
+            start_time = time.time()
+            tclr.fit(X_train.astype(float),y_train)
+            ex_time = time.time()-start_time
+            y_pred = tclr.predict(X_val)
+            score = f1_score(y_val,y_pred,average="macro")
+            print(f"TCLR -- Valdidation: dataset: {dataset} -- C: {C} -- score: {np.around(score,2)}")
+            if score > best_score:
+                    best_performer = copy.deepcopy(tclr)
[…]

### 2024-05-24 Robustness--Classification/srv_methods/utils.py

@@ -1,0 +1,38 @@
+import numpy as np
+import pathlib
+
+valid_datasets = [
+    "ArrowHead",
+    "BME",
+    "ECG200",
+    "FacesUCR",
+    "GunPoint",
+    "PhalangesOutlinesCorrect",
+    "Trace",
+    "Cricket",
+    "ERing",
+    "Handwriting",
+    "Libras",
+    "NATOPS",
+    "RacketSports",
+    "UWaveGestureLibrary",
+    "ArticularyWordRecognition"
+]
+
+here = pathlib.Path(__file__).resolve().parent.parent.parent
+
+def get_data(dataname,missing_rate=0.0):
+    '''
+    get data as tensor
+    '''
+    assert dataname in valid_datasets, "Must specify a valid dataset name."
+    base_filename = here /"datasets"/dataname
+    train_idx = np.load(base_filename/"train_idx.npy")
+    val_idx = np.load(base_filename/"val_idx.npy")
+    test_idx = np.load(base_filename/"test_idx.npy")
+    y = np.load(base_filename/"y.npy")
+    base_filename = here /"datasets"/dataname/f"0.{np.round(missing_rate*10,0).astype(int)}"
+    X = np.load(base_filename/"X.npy")
+    X_mask = np.load(base_filename/"X_mask.npy")
+    X_delta = np.load(base_filename/"X_delta.npy")    
+    return X, X_mask,X_delta,y ,train_idx,val_idx,test_idx

### 2024-05-24 Robustness--Classification/url/layer.py

@@ -1,0 +1,431 @@
+import os
+import math
+import numpy as np
+import torch
+import torchcde
+import torchsde
+
+from torch import nn
+from signatory import logsignature_channels
+
+from module import GRUD, TLSTM, PLSTM, TGLSTM, ODELSTM, GRU_dt, GRU_D, ODERNN
+from attn_module import SAnD_layer, mTAN_layer, MIAM_layer
+from diff_module.NCDE import NeuralCDE, ContinuousRNNConverter, SingleHiddenLayer, FinalTanh, GRU_ODE #NeuralRDE, FinalTanhT
+from diff_module.ANCDE import ANCDE, FinalTanh, FinalTanh_ff6
+from diff_module.EXIT import NeuralCDE_IDEA4, FinalTanh, FinalTanh_g, ODEFunc_f
+#from diff_module.LEAP import NeuralCDE_Learnable, FinalTanh, ODEFunc_ff, Mapping_f
+from diff_module.NFE import NeuralFlow, NeuralFlowCDE, NeuralControlledFlow, NeuralMixture #, FinalTanhT
+from diff_module.NSDE import LatentSDE, NeuralSDE, NN_model, Diffusion_model
+
+## list seq base for ists
+model_name_list = [
+    'cnn', 'cnn-3', 'cnn-5', 'cnn-7', 
+    'rnn', 'lstm', 'gru', 'gru-simple', 'grud',
+    'bilstm', 'tlstm', 'plstm', 'tglstm',
+    'transformer', 'sand', 'mtan', 'miam',
+    'gru-dt', 'gru-d', 'gru-ode', 'ode-rnn', 'ode-lstm',
+    'neuralcde', 'neuralcde-l', 'neuralcde-r', 'neuralcde-c', 'neuralcde-h', 
+    'neuralrde-1', 'neuralrde-2', 'neuralrde-3', 
+    'ancde', 'exit', 'leap',
+    'latentsde', 'latentsde-kl', 'neuralsde-x', 'neuralsde-y', 'neuralsde-z', 
+]
+
+model_name_list = [
+    'rnn', 'lstm', 'gru',
+    'mtan', 'miam',
+    'gru-dt', 'gru-d', 'gru-ode', 'ode-rnn', 'ode-lstm',
+    'neuralcde', 'neuralcde-l', 'neuralcde-r', 'neuralcde-c', 'neuralcde-h', 
[…]

### 2024-05-24 Robustness--Classification/url/model.py

@@ -1,0 +1,115 @@
+import torch
+from torch import nn
+from torch.utils.data import Dataset
+
+# from .module import *
+from layer import *
+
+
+class ists_classifier(nn.Module):
+    def __init__(self, model_name='lstm', input_dim=3, seq_len=128, num_class=2, 
+                 hidden_dim=32, hidden_hidden_dim=32, num_layers=1, num_hidden_layers=1, 
+                 bidirectional=False, dropout=0.1, use_intensity=True, method=None, file=None, device='cuda'):
+        super().__init__()
+
+        self.seq_layer = ists_layer(model_name=model_name, input_dim=input_dim, seq_len=seq_len,
+                                    hidden_dim=hidden_dim, hidden_hidden_dim=hidden_hidden_dim, 
+                                    num_layers=num_layers, num_hidden_layers=num_hidden_layers, 
+                                    bidirectional=bidirectional, dropout=dropout, use_intensity=use_intensity, 
+                                    method=method, file=file, device=device).to(device) 
+
+        # classifier
+        self.fc = nn.Sequential(nn.Linear(hidden_dim, hidden_dim),
+                                nn.BatchNorm1d(hidden_dim), nn.ReLU(), nn.Dropout(dropout),
+                                nn.Linear(hidden_dim, num_class)).to(device) 
+
+        self.fc.apply(self.init_weights)
+        self.fc[-1].weight.register_hook(lambda grad: 100 * grad)
+        self.fc[-1].bias.register_hook(lambda grad: 100 * grad)
+                
+    def init_weights(self, m):
+        if type(m) == nn.Linear:
+            torch.nn.init.xavier_uniform_(m.weight)
+            m.bias.data.fill_(0.01)
+
+    def forward(self, seq, coeffs):
+        out, hn = self.seq_layer(seq, coeffs)
+        assert out.shape[0] == seq.shape[0]
+        seq_hidden = out[:, -1, :]
+        x = self.fc(seq_hidden)
+        return x
+
+
+# Define dataset
+class ists_dataset(Dataset):
+    def __init__(self, y, X_missing, X_mask, X_delta, coeffs, split):
+        self.Y = y[split]
+        self.X_missing = X_missing[split, ...]
+        self.X_mask = X_mask[split, ...]
+        self.X_delta = X_delta[split, ...]
+        self.Coeffs = coeffs[split, ...]
+
+    def __len__(self):
+        return len(self.Y)
+
[…]

### 2024-05-24 Robustness--Classification/url/model_run.py

@@ -1,0 +1,269 @@
+#-*- coding:utf-8 -*-
+
+import os
+import sys
+import copy
+import pickle
+import random
+import numpy as np
+import pandas as pd
+import sklearn
+from sklearn.metrics import *
+import time
+
+
+from tqdm import tqdm
+
+import torch
+import torch.optim as optim
+from torch import nn
+from torch.utils.data import  DataLoader
+
+# setup seed
+def seed_everything(seed):
+    os.environ['PYTHONHASHSEED'] = str(seed)
+    np.random.seed(seed)
+    random.seed(seed)
+    torch.manual_seed(seed)
+    torch.random.manual_seed(seed)
+    torch.cuda.manual_seed(seed)
+    torch.cuda.manual_seed_all(seed)
+    torch.backends.cudnn.deterministic = True
+    torch.backends.cudnn.benchmark = True
+
+SEED = 0
+seed_everything(SEED)
+
+use_cuda = torch.cuda.is_available()
+device = torch.device("cuda" if use_cuda else "cpu")
+
+
+from utils import get_data, get_coeffs
+from model import ists_dataset, ists_classifier, train, evaluate 
+
+
+model_name_list = [
+    'rnn', 'lstm', 'gru',
+    'mtan', 'miam',
+    'ode-lstm',
+    'ancde', 'exit',
+    "neuralsde_1_18", "neuralsde_4_17"
+]
+
+
+if not os.path.exists('../results'):
+    os.mkdir('../results')
+    
+    
+def run_model(data_name, missing_rate, model_name, model_config, EPOCHS=100, SEED=SEED, CHECK=False):
+    print(data_name, missing_rate, model_name)
+
+    # check exist
+    out_name = '_'.join([data_name, str(missing_rate), model_name, str(SEED)])
+    out_path = '../results/{}/{}/{}'.format(data_name, str(missing_rate), out_name)
+
+    if CHECK & os.path.exists(out_path):
+        return None
+    
+    # setup out path
+    if not os.path.exists('../results/{}'.format(data_name)):
+        os.mkdir('../results/{}'.format(data_name))
+    
+    # load data
+    X_missing, X_mask, X_delta,Y,train_idx,valid_idx,test_idx = get_data(data_name,missing_rate=missing_rate)
+    X_missing,X_mask,X_delta = torch.tensor(X_missing),torch.tensor(X_mask),torch.tensor(X_delta)
[…]

### 2024-05-24 Robustness--Classification/url/param_search.py

@@ -1,0 +1,253 @@
+#-*- coding:utf-8 -*-
+
+import os
+import copy
+import pickle
+import random
+
+import numpy as np
+import pandas as pd
+import sklearn.model_selection
+from sklearn.metrics import *
+
+from tqdm import tqdm
+
+import torch
+import torch.optim as optim
+from torch import nn
+from torch.utils.data import DataLoader
+
+# setup seed
+def seed_everything(seed):
+    os.environ['PYTHONHASHSEED'] = str(seed)
+    np.random.seed(seed)
+    random.seed(seed)
+    torch.manual_seed(seed)
+    torch.random.manual_seed(seed)
+    torch.cuda.manual_seed(seed)
+    torch.cuda.manual_seed_all(seed)
+    
+SEED = 0
+seed_everything(SEED)
+
+# CUDA for PyTorch
+use_cuda = torch.cuda.is_available()
+device = torch.device("cuda" if use_cuda else "cpu")
+
+
+from utils import get_data, get_coeffs
+from model import ists_dataset, ists_classifier, train, evaluate 
+
+import ray
+from ray import tune
+from ray.tune.schedulers import ASHAScheduler
+
+
+model_name_list = [
+    'rnn', 'lstm', 'gru',
+    'mtan', 'miam',
+    'ode-lstm',
+    'ancde', 'exit',
+    "neuralsde_1_18", "neuralsde_4_17"
+]
+
+
+if not os.path.exists('params'):
+    os.mkdir('params')
+
+# set model
+def tune_model(data_name, missing_rate, model_name, model_config, EPOCHS=100, SEED=SEED):
+    print(data_name)
+    
+    # load data
+    X_missing, X_mask, X_delta,Y,train_idx,valid_idx,test_idx = get_data(data_name,missing_rate=missing_rate)
+    X_missing,X_mask,X_delta = torch.tensor(X_missing),torch.tensor(X_mask),torch.tensor(X_delta)
+    num_data = X_missing.shape[0]
+    seq_len = X_missing.shape[1]
+    num_dim = X_missing.shape[2]
+    num_class = len(np.unique(Y))
+    
+    # set batch_size by the number of data
+    batch_size = 2**4
+    for i in range(4,5): #8
+        if 2**i > num_data/2:
+            break
+        batch_size = 2**i
+
+    # set learning params
+    if model_config['lr'] is None:
+        lr = 1e-3 * (batch_size / 2**4)
+    else:
+        lr = model_config['lr']
+    
+    # check model_name and settings
[…]

### 2024-05-24 Robustness--Classification/url/utils.py

@@ -1,0 +1,73 @@
+import numpy as np
+import torch
+import torchcde
+import pathlib
+
+from diff_module.NCDE.controldiffeq import natural_cubic_spline_coeffs
+
+def _pad(channel, max_len):
+    channel = torch.tensor(channel)
+    out = torch.full((max_len,), channel[-1])
+    out[:channel.size(0)] = channel
+    return out
+
+valid_datasets = [
+    "ArrowHead",
+    "BME",
+    "ECG200",
+    "FacesUCR",
+    "GunPoint",
+    "PhalangesOutlinesCorrect",
+    "Trace",
+    "Cricket",
+    "ERing",
+    "Handwriting",
+    "Libras",
+    "NATOPS",
+    "RacketSports",
+    "UWaveGestureLibrary",
+    "ArticularyWordRecognition"
+]
+
+here = pathlib.Path(__file__).resolve().parent.parent.parent
+
+def get_data(dataname,missing_rate=0.0):
+    '''
+    get data as tensor
+    '''
+    assert dataname in valid_datasets, "Must specify a valid dataset name."
+    base_filename = here /"datasets"/dataname
+    train_idx = np.load(base_filename/"train_idx.npy")
+    val_idx = np.load(base_filename/"val_idx.npy")
+    test_idx = np.load(base_filename/"test_idx.npy")
+    y = np.load(base_filename/"y.npy")
+    base_filename = here /"datasets"/dataname/f"0.{np.round(missing_rate*10,0).astype(int)}"
+    X = np.load(base_filename/"X.npy")
+    X_mask = np.load(base_filename/"X_mask.npy")
+    X_delta = np.load(base_filename/"X_delta.npy")    
+    return X, X_mask,X_delta,y ,train_idx,val_idx,test_idx
+
+
+def get_coeffs(X_missing,X_mask,X_delta,interpolate='natural', use_intensity=True):
+    max_len = torch.tensor([len(x) for x in X_missing]).max()
+    times = torch.linspace(0, 1, max_len)
+    intensity = ~torch.isnan(X_missing)
+    intensity = intensity.to(X_missing.dtype).cumsum(dim=1)
+
+    values_T = torch.cat([times.repeat((X_missing.shape[0],1)).unsqueeze(-1), X_missing], dim=-1)
+    values_TI = torch.cat([times.repeat((X_missing.shape[0],1)).unsqueeze(-1), intensity, X_missing], dim=-1) 
+
+    if interpolate == 'natural':
+        if use_intensity:
[…]

### 2024-05-24 classification/msvc_exp.py

@@ -1,230 +1,0 @@
-import sys
-sys.path.insert(0,"../")
-
-import pandas as pd
-import numpy as np 
-import time
-import optax
-from optax.schedules import warmup_cosine_decay_schedule
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import LabelEncoder
-from sklearn.metrics import accuracy_score,f1_score
-
-from aeon.datasets import load_classification
-
-from src.utils import from_timeseries_to_dataset
-from src.kernel import VFTSGaussKernel,TSGaussGaussKernel,GaussKernel,MyTSGaussKernel
-from src.loss import VarifoldLoss,MyLoss,SumVarifoldLoss
-from src.classifier import MomentaSVC
-from src.lddmm import batch_time_initializer
-from src.barycenter import batch_barycenter_initializer
-from src.utils import batch_dataset,from_timeseries_to_dataset
-
-
-def mouse_loader(seed=0):
-    y =pd.read_csv("../mouse/dataset/y.csv",index_col=0)
-    X = np.load("../mouse/dataset/X.npy")
-    X_mask = np.load("../mouse/dataset/X_mask.npy")
-    X = X[:,::2,:]
-    X_mask = X_mask[:,::2,:]
-
-    lst = []
-    for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
-        idx = y[(y.filename == filename)*(y.before == "Y")].sample(20,random_state=seed).index
-        lst.append(idx)
-    idxs = np.concatenate(lst)
-
-    X,X_mask = X[idxs],X_mask[idxs]
-    y = y.iloc[idxs]
-    y = y.genotype.values.astype(str)
-    train_idx,test_idx = train_test_split(np.arange(X.shape[0]),test_size=0.33,shuffle=True,random_state=seed)
-    return X[train_idx],X_mask[train_idx],X[test_idx],X_mask[test_idx],y[train_idx],y[test_idx]
-
-def gait_loader(seed=0): 
-    y = np.load("../gait_exp/dataset/y.npy")
-    X = np.load("../gait_exp/dataset/X.npy")
-    X_mask = np.load("../gait_exp/dataset/X_mask.npy")
-    train_idx,test_idx = train_test_split(np.arange(X.shape[0]),test_size=0.33,shuffle=True,random_state=seed)
[…]

### 2024-05-24 gait_exp/gait_experiment.py

@@ -1,67 +1,0 @@
-import sys 
-sys.path.insert(0,"../")
-
-import numpy as np 
-import jax.numpy as jnp
-import json 
-import optax
-from optax.schedules import warmup_cosine_decay_schedule
-
-
-
-from src.utils import batch_dataset
-from src.kernel import VFTSGaussKernel, TSGaussGaussKernel,GaussKernel
-from src.barycenter import batch_varifold_barycenter_registration, batch_barycenter_registration
-from src.loss import SumVarifoldLoss
-
-
-BATCH_SIZE = 75
-NITER = 400
-
-
-scheduler = warmup_cosine_decay_schedule(0,0.3,30,400,0.)
-optimizer = optax.adam(scheduler)
-
-
-
-Kv = VFTSGaussKernel(1,0.1,40,1,1)
-Kl1 = TSGaussGaussKernel(2,2,2,2)
-Kl2 = TSGaussGaussKernel(1,1,1,1)
-Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
-Kls=[Kl1,Kl2,Kl3]
-dataloss = SumVarifoldLoss(Kls)
-
-
-
-
-
-SAVE_PATH = "./results/SS_2.json"
-
-
-if __name__ == "__main__":
-
-    # Import data 
-    X = np.load("./dataset/X.npy")
-    X_mask = np.load("./dataset/X_mask.npy")
-    batched_dataset, batched_masks = batch_dataset(X,BATCH_SIZE,X_mask)
-
-    print("Running experiment")
-    ps ,qb, qb_mask= batch_barycenter_registration(batched_dataset,batched_masks,Kv,dataloss,niter=NITER,optimizer=optimizer,gamma_loss=0.00)
-    
-    print("Exporting results")
-    dct =dict(
-        ps = json.dumps(ps.tolist()),
-        qb = json.dumps(qb.tolist()),
-        qb_mask = json.dumps(qb_mask.tolist())
-    )
-    
-    with open(SAVE_PATH, "w") as f: 
-        f.write(json.dumps(dct))
-
-    print("Done")
-
-
-
-
-
-

### 2024-05-24 mouse/experiment_1.py

@@ -1,50 +1,0 @@
-import pandas as pd
-import numpy as np 
-import matplotlib.pyplot as plt
-import sys
-import optax
-from optax.schedules import warmup_cosine_decay_schedule
-sys.path.insert(0,"../")
-
-
-from src.kernel import VFTSGaussKernel,TSGaussGaussKernel
-from src.loss import VarifoldLoss,SumVarifoldLoss
-from src.barycenter import batch_barycenter_registration
-from src.utils import batch_dataset
-
-np.random.seed(0)
-
-y =pd.read_csv("./dataset/y.csv",index_col=0)
-X = np.load("./dataset/X.npy")
-X_mask = np.load("./dataset/X_mask.npy")
-X = X[:,::2,:]
-X_mask = X_mask[:,::2,:]
-
-lst = []
-for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
-    idx = y[(y.filename == filename)*(y.before == "Y")].sample(50).index
-    lst.append(idx)
-idxs = np.concatenate(lst)
-
-bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
-y = y.iloc[idxs]
-
-Kv = VFTSGaussKernel(1,0.1,100,1,1)
-Kl1 = TSGaussGaussKernel(2,2,2,2)
-Kl2 = TSGaussGaussKernel(1,1,1,1)
-Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
-Kls=[Kl1,Kl2,Kl3]
-dataloss = SumVarifoldLoss(Kls)
-schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
-optimizer = optax.adabelief(schedule)
-
-
-if __name__ == "__main__": 
-
-  
-    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
-
-    np.save("./results/exp_1_0/idxs.npy",idxs)
-    np.save("./results/exp_1_0/p0s.npy",p0s)
-    np.save("./results/exp_1_0/q0.npy",q0)
-    np.save("./results/exp_1_0/q0_mask.npy",q0_mask)

### 2024-05-24 mouse/experiment_1_bis.py

@@ -1,56 +1,0 @@
-import pandas as pd
-import numpy as np 
-import matplotlib.pyplot as plt
-import sys
-import optax
-from optax.schedules import warmup_cosine_decay_schedule
-sys.path.insert(0,"../")
-
-
-from src.kernel import VFTSGaussKernel,TSGaussGaussKernel
-from src.loss import VarifoldLoss,SumVarifoldLoss
-from src.barycenter import batch_barycenter_registration
-from src.utils import batch_dataset
-
-np.random.seed(0)
-
-y =pd.read_csv("./dataset/y.csv",index_col=0)
-X = np.load("./dataset/X.npy")
-X_mask = np.load("./dataset/X_mask.npy")
-idxs = np.load("./results/exp_1_0/idxs.npy")
-
-X = X[:,::2,:]
-X_mask = X_mask[:,::2,:]
-bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
-y = y.iloc[idxs]
-
-print(y.before.unique())
-print(y.genotype.unique())
-
-
-if __name__ == "__main__": 
-
-    Kv = VFTSGaussKernel(1,0.1,100,1,1)
-    Kv = VFTSGaussKernel(1,0.1,100,1,1)
-    Kl1 = TSGaussGaussKernel(2,2,2,2)
-    Kl2 = TSGaussGaussKernel(1,1,1,1)
-    Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
-    Kls=[Kl1,Kl2,Kl3]
-    dataloss = SumVarifoldLoss(Kls)
-    schedule = warmup_cosine_decay_schedule(0,0.3,40,400,0)
-    optimizer = optax.adabelief(schedule)
-
-
-
-    for i,filename in enumerate(y.filename.unique()): 
-       
-        print(f"Mouse: {i+1}/{y.filename.unique().shape[0]} -- {filename}")
-        tidxs = y[y.filename == filename].index
-        bX,bX_mask = batch_dataset(X[tidxs],1,X_mask[tidxs])
-        p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
-
-        np.save("./results/exp_1_0/"+f"{filename[:-4]}_p0s.npy",p0s)
[…]

### 2024-05-24 mouse/experiment_1_lddmm.py

@@ -1,48 +1,0 @@
-import pandas as pd
-import numpy as np 
-import matplotlib.pyplot as plt
-import sys
-import optax
-from optax.schedules import warmup_cosine_decay_schedule
-sys.path.insert(0,"../")
-
-from src.masked.lddmm import registration,Shooting,Flowing
-from src.masked.kernel import VFTSGaussKernel,TSGaussGaussKernel, TSGaussKernel
-from src.masked.loss import VarifoldLoss
-from src.plotting import plot2Dfigure
-from src.masked.barycenter import batch_barycenter_registration
-from src.utils import batch_dataset
-
-np.random.seed(0)
-
-y =pd.read_csv("./dataset/inv_y.csv",index_col=0)
-X = np.load("./dataset/inv_X.npy")
-X_mask = np.load("./dataset/inv_X_mask.npy")
-X = X[:,::2,:]
-X_mask = X_mask[:,::2,:]
-
-lst = []
-for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
-    idx = y[(y.filename == filename)*(y.before == "Y")].sample(50).index
-    lst.append(idx)
-idxs = np.concatenate(lst)
-
-bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
-y = y.iloc[idxs]
-
-Kv = TSGaussKernel(100,1)
-Kl = TSGaussGaussKernel(2,1,2,0.6)
-dataloss = VarifoldLoss(Kl)
-schedule = warmup_cosine_decay_schedule(0,0.01,40,400,0)
-optimizer = optax.adabelief(schedule)
-
-
-if __name__ == "__main__": 
-
-  
-    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
-
-    np.save("./results/exp_1_lddmm_2/idxs.npy",idxs)
-    np.save("./results/exp_1_lddmm_2/p0s.npy",p0s)
-    np.save("./results/exp_1_lddmm_2/q0.npy",q0)
-    np.save("./results/exp_1_lddmm_2/q0_mask.npy",q0_mask)

### 2024-05-24 mouse/experiment_2.py

@@ -1,46 +1,0 @@
-import pandas as pd
-import numpy as np 
-import matplotlib.pyplot as plt
-import sys
-import optax
-from optax.schedules import warmup_cosine_decay_schedule
-sys.path.insert(0,"../")
-
-from src.masked.lddmm import registration,Shooting,Flowing
-from src.masked.kernel import VFTSGaussKernel,TSGaussGaussKernel
-from src.masked.loss import VarifoldLoss
-from src.plotting import plot2Dfigure
-from src.masked.barycenter import batch_barycenter_registration
-from src.utils import batch_dataset
-
-np.random.seed(0)
-y =pd.read_csv("./dataset/inv_y.csv",index_col=0)
-X = np.load("./dataset/inv_X.npy")
-X_mask = np.load("./dataset/inv_X_mask.npy")
-X = X[:,::2,:]
-X_mask = X_mask[:,::2,:]
-Kv = VFTSGaussKernel(1,0.1,150,1,1)
-Kl = TSGaussGaussKernel(5,1,5,0.6)
-dataloss = VarifoldLoss(Kl)
-schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
-optimizer = optax.adabelief(schedule)
-
-lst = []
-for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
-    idx = y[(y.before == "Y")*(y.filename == filename)].sample(25).index
-    lst.append(idx)
-    idx = y[(y.before == "N")*(y.filename == filename)].sample(50).index
-    lst.append(idx)
-idxs = np.concatenate(lst)
-
-bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
-y = y.iloc[idxs]
-
-if __name__ == "__main__": 
-
-    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=800,optimizer=optimizer,gamma_loss=1e-4)
-
-    np.save("./results/exp_2_3/idxs.npy",idxs)
-    np.save("./results/exp_2_3/p0s.npy",p0s)
-    np.save("./results/exp_2_3/q0.npy",q0)
-    np.save("./results/exp_2_3/q0_mask.npy",q0_mask)

### 2024-05-24 mouse/experiment_2_lddmm.py

@@ -1,51 +1,0 @@
-import pandas as pd
-import numpy as np 
-import matplotlib.pyplot as plt
-import sys
-import optax
-from optax.schedules import warmup_cosine_decay_schedule
-sys.path.insert(0,"../")
-
-from src.masked.lddmm import registration,Shooting,Flowing
-from src.masked.kernel import VFTSGaussKernel,TSGaussGaussKernel, TSGaussKernel
-from src.masked.loss import VarifoldLoss
-from src.plotting import plot2Dfigure
-from src.masked.barycenter import batch_barycenter_registration
-from src.utils import batch_dataset
-
-np.random.seed(0)
-
-y =pd.read_csv("./dataset/inv_y.csv",index_col=0)
-X = np.load("./dataset/inv_X.npy")
-X_mask = np.load("./dataset/inv_X_mask.npy")
-X = X[:,::2,:]
-X_mask = X_mask[:,::2,:]
-
-lst = []
-for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
-    idx = y[(y.before == "Y")*(y.filename == filename)].sample(25).index
-    lst.append(idx)
-    idx = y[(y.before == "N")*(y.filename == filename)].sample(50).index
-    lst.append(idx)
-idxs = np.concatenate(lst)
-
-bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
-y = y.iloc[idxs]
-
-
-Kv = TSGaussKernel(100,1)
-Kl = TSGaussGaussKernel(2,1,2,0.6)
-dataloss = VarifoldLoss(Kl)
-schedule = warmup_cosine_decay_schedule(0,0.01,40,400,0)
-optimizer = optax.adabelief(schedule)
-
-
-if __name__ == "__main__": 
-
-  
-    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=400,optimizer=optimizer,gamma_loss=1e-3)
-
-    np.save("./results/exp_2_lddmm_1/idxs.npy",idxs)
-    np.save("./results/exp_2_lddmm_1/p0s.npy",p0s)
-    np.save("./results/exp_2_lddmm_1/q0.npy",q0)
-    np.save("./results/exp_2_lddmm_1/q0_mask.npy",q0_mask)

### 2024-05-24 mouse/script_lddmm.py

@@ -1,0 +1,101 @@
+import pandas as pd
+import numpy as np 
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+from src.kernel import TSGaussGaussKernel, TSGaussKernel
+from src.loss import SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("./dataset/y_600.csv",index_col=0)
+X = np.load("./dataset/X_600.npy")
+X_mask = np.load("./dataset/X_600_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.filename == filename)*(y.before == "Y")].sample(50).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+Kv = TSGaussKernel(150,2)
+Kl1 = TSGaussGaussKernel(5,2,5,1)
+Kl2 = TSGaussGaussKernel(2,1,2,0.6)
+Kl3 = TSGaussGaussKernel(1,0.6,1,0.6)
+Kls=[Kl1,Kl2,Kl3]
+dataloss = SumVarifoldLoss(Kls)
+schedule = warmup_cosine_decay_schedule(0,0.01,80,800,0)
+optimizer = optax.adabelief(schedule)
+  
+p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=600,optimizer=optimizer,gamma_loss=1e-3)
+np.save("./results/exp_1_lddmm_0/idxs.npy",idxs)
+np.save("./results/exp_1_lddmm_0/p0s.npy",p0s)
+np.save("./results/exp_1_lddmm_0/q0.npy",q0)
+np.save("./results/exp_1_lddmm_0/q0_mask.npy",q0_mask)
+
+
+
+################################################################################################################
+################################################################################################################
+
+import pandas as pd
+import numpy as np 
[…]

### 2024-05-24 mouse/script_lddmm_exp_1.py

@@ -1,0 +1,49 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+from src.kernel import TSGaussGaussKernel, TSGaussKernel
+from src.loss import SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("./dataset/y_600.csv",index_col=0)
+X = np.load("./dataset/X_600.npy")
+X_mask = np.load("./dataset/X_600_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.filename == filename)*(y.before == "Y")].sample(15).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+Kv = TSGaussKernel(150,2)
+Kl1 = TSGaussGaussKernel(5,2,5,1)
+Kl2 = TSGaussGaussKernel(2,1,2,0.6)
+Kl3 = TSGaussGaussKernel(1,0.6,1,0.6)
+Kls=[Kl1,Kl2,Kl3]
+dataloss = SumVarifoldLoss(Kls)
+schedule = warmup_cosine_decay_schedule(0,0.01,80,800,0)
+optimizer = optax.adabelief(schedule)
+
+
+if __name__ == "__main__": 
+
+  
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=200,optimizer=optimizer,gamma_loss=1e-3)
+
+    np.save("./results/exp_1_lddmm_0/idxs.npy",idxs)
+    np.save("./results/exp_1_lddmm_0/p0s.npy",p0s)
+    np.save("./results/exp_1_lddmm_0/q0.npy",q0)
+    np.save("./results/exp_1_lddmm_0/q0_mask.npy",q0_mask)

### 2024-05-24 mouse/script_lddmm_exp_2.py

@@ -1,0 +1,55 @@
+import pandas as pd
+import numpy as np 
+import jax.numpy as jnp
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+
+from src.kernel import TSGaussGaussKernel, TSGaussKernel
+from src.loss import SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+y =pd.read_csv("./dataset/y_600.csv",index_col=0)
+X = np.load("./dataset/X_600.npy")
+X_mask = np.load("./dataset/X_600_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.before == "Y")*(y.filename == filename)].sample(5).index
+    lst.append(idx)
+    idx = y[(y.before == "N")*(y.filename == filename)].sample(10).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+
+Kv = TSGaussKernel(150,2)
+Kl0 = TSGaussGaussKernel(30,2,30,1)
+Kl1 = TSGaussGaussKernel(5,2,5,1)
+Kl2 = TSGaussGaussKernel(2,1,2,0.6)
+Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
+Kls=[Kl0,Kl1,Kl2,Kl3]
+dataloss = SumVarifoldLoss(Kls)
+schedule = warmup_cosine_decay_schedule(0,0.01,80,800,0)
+optimizer = optax.adabelief(schedule)
+
+qi = jnp.load("./results/exp_1_1/q0.npy")
+qi_mask = jnp.load("./results/exp_1_1/q0_mask.npy")
+
+if __name__ == "__main__": 
+
+  
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,qi,qi_mask,niter=200,optimizer=optimizer,gamma_loss=1e-3)
+
+    np.save("./results/exp_2_lddmm_1/idxs.npy",idxs)
[…]

### 2024-05-24 mouse/script_ts-lddmm_exp_1.py

@@ -1,0 +1,49 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel
+from src.loss import VarifoldLoss,SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("../datasets/Mouse/y_600.csv",index_col=0)
+X = np.load("../datasets/Mouse/X_600.npy")
+X_mask = np.load("../datasets/Mouse/X_600_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.filename == filename)*(y.before == "Y")].sample(50).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+Kv = VFTSGaussKernel(1,0.1,150,1,2)
+Kl1 = TSGaussGaussKernel(5,2,5,1)
+Kl2 = TSGaussGaussKernel(2,1,2,0.6)
+Kl3 = TSGaussGaussKernel(1,0.6,1,0.6)
+Kls=[Kl1,Kl2,Kl3]
+dataloss = SumVarifoldLoss(Kls)
+schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
+optimizer = optax.adabelief(schedule)
+
+
+if __name__ == "__main__": 
+  
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=800,optimizer=optimizer,gamma_loss=1e-3)
+
+    np.save("./results/ts-lddmm_exp_1/idxs.npy",idxs)
+    np.save("./results/ts-lddmm_exp_1/p0s.npy",p0s)
+    np.save("./results/ts-lddmm_exp_1/q0.npy",q0)
+    np.save("./results/ts-lddmm_exp_1/q0_mask.npy",q0_mask)

### 2024-05-24 mouse/script_ts-lddmm_exp_1_barycenter.py

@@ -1,0 +1,55 @@
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel
+from src.loss import VarifoldLoss,SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+
+y =pd.read_csv("./dataset/y_600.csv",index_col=0)
+X = np.load("./dataset/X_600.npy")
+X_mask = np.load("./dataset/X_600_mask.npy")
+idxs = np.load("./results/ts_lddmm_exp_1/idxs.npy")
+
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+print(y.before.unique())
+print(y.genotype.unique())
+
+
+if __name__ == "__main__": 
+
+    Kv = VFTSGaussKernel(1,0.1,150,1,2)
+    Kl1 = TSGaussGaussKernel(5,2,5,1)
+    Kl2 = TSGaussGaussKernel(2,1,2,0.6)
+    Kl3 = TSGaussGaussKernel(1,0.6,1,0.6)
+    Kls=[Kl1,Kl2,Kl3]
+    dataloss = SumVarifoldLoss(Kls)
+    schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
+    optimizer = optax.adabelief(schedule)
+
+
+
+    for i,filename in enumerate(y.filename.unique()): 
+       
+        print(f"Mouse: {i+1}/{y.filename.unique().shape[0]} -- {filename}")
+        tidxs = y[y.filename == filename].index
+        bX,bX_mask = batch_dataset(X[tidxs],1,X_mask[tidxs])
+        p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,niter=500,optimizer=optimizer,gamma_loss=1e-3)
+
+        np.save("./results/ts-lddmm_exp_1/"+f"{filename[:-4]}_p0s.npy",p0s)
+        np.save("./results/ts-lddmm_exp_1/"+f"{filename[:-4]}_q0.npy",q0)
[…]

### 2024-05-24 mouse/script_ts-lddmm_exp_2.py

@@ -1,0 +1,55 @@
+import pandas as pd
+import numpy as np 
+import jax.numpy as jnp
+import sys
+import optax
+from optax.schedules import warmup_cosine_decay_schedule
+sys.path.insert(0,"../")
+
+
+from src.kernel import VFTSGaussKernel,TSGaussGaussKernel
+from src.loss import SumVarifoldLoss
+from src.barycenter import batch_barycenter_registration
+from src.utils import batch_dataset
+
+np.random.seed(0)
+y =pd.read_csv("./dataset/y_600.csv",index_col=0)
+X = np.load("./dataset/X_600.npy")
+X_mask = np.load("./dataset/X_600_mask.npy")
+X = X[:,::2,:]
+X_mask = X_mask[:,::2,:]
+
+lst = []
+for filename in y[y.genotype.isin(["colq","wt"])].filename.unique(): 
+    idx = y[(y.before == "Y")*(y.filename == filename)].sample(5).index
+    lst.append(idx)
+    idx = y[(y.before == "N")*(y.filename == filename)].sample(10).index
+    lst.append(idx)
+idxs = np.concatenate(lst)
+
+bX,bX_mask = batch_dataset(X[idxs],1,X_mask[idxs])
+y = y.iloc[idxs]
+
+Kv = VFTSGaussKernel(1,0.1,150,1,2)
+Kl0 = TSGaussGaussKernel(30,2,30,1)
+Kl1 = TSGaussGaussKernel(5,2,5,1)
+Kl2 = TSGaussGaussKernel(2,1,2,0.6)
+Kl3 = TSGaussGaussKernel(1,0.1,1,0.1)
+Kls=[Kl0,Kl1,Kl2,Kl3]
+dataloss = SumVarifoldLoss(Kls)
+schedule = warmup_cosine_decay_schedule(0,0.3,80,800,0)
+optimizer = optax.adabelief(schedule)
+
+qi = jnp.load("./results/ts-lddmm_exp_1/q0.npy")
+qi_mask = jnp.load("./results/ts_lddmm_exp_1/q0_mask.npy")
+
+
+
+if __name__ == "__main__": 
+
+    p0s,q0,q0_mask = batch_barycenter_registration(bX,bX_mask,Kv,dataloss,qi,qi_mask,niter=100,optimizer=optimizer,gamma_loss=1e-3)
+
[…]

### 2024-05-24 mouse/segmentation.py

@@ -1,181 +1,0 @@
-import ruptures as rpt
-from scipy import signal
-import numpy as np
-from sklearn.base import BaseEstimator,TransformerMixin
-from itertools import groupby 
-
-class PhaseSeg(BaseEstimator,TransformerMixin):
-
-    def __init__(self, sampfreq:int, prominence:float, wlen:float, cycle_minimum_duration:float,cycle_maximum_duration:float,trainig_size_per_interval=None,interval=None) -> None:
-        """Initialization 
-
-        Args: 
-            sampfreq (int): Sampling frequency 
-            prominence (float): Peak prominence for volume, in ml.
-            wlen (float): Peak prominence search window, in seconds. 
-            cycle_minimum_duration (float): minimum cycle duration, in seconds. 
-            cycle_maximum_duration (float): maximum cycle duration, in seconds.  
-            trainig_size_per_interval (int): number of cycles slected randomly on each interval. 
-            interval (float): interval duration, in seconds. 
-        """
-        super().__init__()
-        self.sampfreq = sampfreq
-        self.prominence = prominence
-        self.wlen = wlen 
-        self.cycle_minimum_duration = cycle_minimum_duration
-        self.cycle_maximum_duration = cycle_maximum_duration
-        self.trainig_size_per_interval = trainig_size_per_interval
-        self.interval = interval
-
-    def fit(self,X): 
-        """Compute detrended volume
-
-        Args:
-            X (np.ndarray): original sequence, shape (N,).
-
-        """
-        self.flow_ = X
-        self._event_detection()
-        self._filtration()
-        if (self.trainig_size_per_interval is not None) & (self.interval is not None): 
-            self._training_selection()
-
-    def _volume(self,X:np.ndarray)->np.ndarray:
-        """Compute detrended volume
-
-        Args:
-            X (np.ndarray): original sequence, shape (N,).
-
-        Returns:
-            np.ndarray: Detrended volume, shape(N,)
-        """
-        arr = signal.detrend(np.cumsum(X*1/self.sampfreq))
-        arr = arr.astype(float)
-        return arr
-
-    def _event_detection(self)->None:
-        """
-        Detect inhalation and exhalation start_time
-        """
-        #removing unwanted time
-        self.flow_ = self.flow_.astype(float)
-
-        #getting volume
-        self.volume_ = self._volume(self.flow_)
-
-        #Getting start inspiration
[…]

### 2024-05-24 src/classifier.py

@@ -2,7 +2,7 @@ import numpy as np
 import optax
 
 from sklearn.svm import SVC
-from sklearn.metrics import accuracy_score
+from sklearn.metrics import accuracy_score,f1_score
 from src.utils import MaskMomentaGramMatrix,batch_dataset,unbatch_dataset
 from src.lddmm import batch_one_to_many_registration
 from src.barycenter import batch_barycenter_registration
@@ -42,8 +42,29 @@ class MomentaSVC(SVC):
         batched_ps,self.qbar_,self.qbar_mask_ = batch_barycenter_registration(batched_X,batched_X_mask,self.Kv,self.dataloss,init,init_mask,self.barycenter_initializer,self.niter,self.optimizer,self.gamma_loss,self.nt,self.deltat,self.verbose)
         self.ps_ = self._unbached(batched_ps)
         self._mgm = MaskMomentaGramMatrix(self.Kv,self.qbar_,self.qbar_mask_)
-        X_svc = self._mgm(self.ps_,self.ps_)
-        super().fit(X_svc,y)
+        self.X_svc_ = self._mgm(self.ps_,self.ps_)
+        self.y_ = y
+        super().fit(self.X_svc_,self.y_)
+        return self
+    
+    def set_C(self,X,X_mask,y,C_lst):
+        batched_X,batched_X_mask = self._batch(X,X_mask)
+        batched_ps,_,_ = batch_one_to_many_registration(self.qbar_,self.qbar_mask_,batched_X,batched_X_mask,self.Kv,self.dataloss,self.registration_initializer,self.gamma_loss,self.niter,self.optimizer,self.nt,self.deltat,self.verbose)
+        ps = self._unbached(batched_ps)
+        X_svc = self._mgm(ps,self.ps_)
+        best_C = 0
+        best_score = -np.inf
+        for C in C_lst: 
+            self.C = C
+            super().fit(self.X_svc_,self.y_)
+            y_pred = super().predict(X_svc)
+            score = f1_score(y,y_pred,average="macro")
+            print(f"Validation -- C: {C} -- f1score: {score}")
+            if score > best_score: 
+                best_C = C
+                best_score = score.copy()
+        self.C = best_C
+        print(f"Best C :{self.C}")
         return self
     
     def predict(self,X,X_mask):
@@ -56,6 +77,9 @@ class MomentaSVC(SVC):
     def score(self,X,X_mask,y):
[…]

### 2024-09-03 NEURIPS24/appendix.tex
Revert "Delete NEURIPS24 directory"
This reverts commit e09ecaea218c43964bb3eb1e2a9a2c5bfd236ded.

@@
 
-\section{Appendix for experiment: Analysis of respiratory behavior in mice}
-\subsection{Settings}
-\label{appendix: mice_exp_setting}
-
-This experiment involves TS-LDDMM and LDDMM~\cite{glaunes2008large} methods.
-Both methods are run twice, first on respiratory cycles before exposure to the irritant molecule to capture mice breathing behavior at rest and on all respiratory cycles to capture the influence of the irritant molecule.
-Exposure to the irritant molecule leads to significant shape deformation in the respiratory cycles, and the terms must be added to the varifold loss to capture deformations at a large time scale.
-
-\paragraph{TS-LDDMM parameters.}
-\begin{itemize}
-  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,150,1,2)$.
-The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
-  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,220,1,2)$.
-The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
-\end{itemize}
-
-\paragraph{LDDMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM.
-Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
-\begin{itemize}
-  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =150$ for the time dimension and $\sigma_x = 2$ for space dimensions. 
-  The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
-  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =220$ for the time dimension and $\sigma_x = 2$ for space dimensions.
-  The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
-The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
-\end{itemize}
-
 \section{Appendix for experiment: TS-LDDMM representation identifiability}
+\label{appendix: identifiability}
 In this experiment, we evaluate the ability of TS-LDDMM to retrieve the parameter $v_0^*$ that encodes the deformation $\varphi^{\{v_0^*\}}$ acting on a time series graph $\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation} between $\msg$ and $\varphi^{\{v_0^*\}}.\msg$.
@@
 \section{Appendix for experiment: Classification benchmark on regularly sampled datasets}
+\label{appendix: shape_classification}
 
@@
 \end{table}
+
+\section{Appendix for experiment: Analysis of respiratory behavior in mice}
+\label{appendix: mice_exp_setting}
+
+\subsection{Settings}
+
+This experiment involves TS-LDDMM and LDDMM~\cite{glaunes2008large} methods.
+Both methods are run twice, first on respiratory cycles before exposure to the irritant molecule to capture mice breathing behavior at rest and on all respiratory cycles to capture the influence of the irritant molecule.
+Exposure to the irritant molecule leads to significant shape deformation in the respiratory cycles, and the terms must be added to the varifold loss to capture deformations at a large time scale.
+
+\paragraph{TS-LDDMM parameters.}
+\begin{itemize}
+  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,150,1,2)$.
+The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
+  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,220,1,2)$.
+The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.3.
+\end{itemize}
+
+\paragraph{LDDMM parameters.} Note that varifold losses are unchanged between TS-LDDMM and LDDMM.
+Compared to TS-LDDMM, the convergence of LDDMM is more sensitive to the maximum stepsize $\eta_m$, which must remain small for LDDMM to guarantee the convergence.
+\begin{itemize}
+  \item \textbf{Before exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =150$ for the time dimension and $\sigma_x = 2$ for space dimensions. 
+  The varifold loss is the sum of three varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3): $\left((5,2,5,1),(2,1,2,0.6),(1,0.6,1,0.6)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
+  \item \textbf{Before/after exposure:} The velocity field kernel $K_G$ is an anysotropic Gaussian kernel with parameters $\sigma_{T} =220$ for the time dimension and $\sigma_x = 2$ for space dimensions.
+  The varifold loss is the sum of four varifolds to capture shapes variations at different scales with parameters: (Varifold 1,Varifold 2,Varifold 3, Varifold 4): $\left((30,2,30,1),(5,2,5,1),(2,1,2,0.6),(1,0.1,1,0.1)\right)$ and the mapper $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x})$.
+The optimizer has 800 steps with a maximum stepsize $\eta_M$ of 0.01.
+\end{itemize}

### 2024-09-03 NEURIPS24/experiment.tex
Revert "Delete NEURIPS24 directory"
This reverts commit e09ecaea218c43964bb3eb1e2a9a2c5bfd236ded.

@@
 \begin{enumerate}
-    \item \textbf{TS-LDDMM representation identifiability, \Cref{appendix: settings_identifiability}:} On synthetic data, we evaluate the ability of our method to retrieve the parameter $v_0^*$ that encodes the deformation $\varphi^{\{v_0^*\}}$ acting on a time series graph $\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation} between $\msg$ and $\varphi^{\{v_0^*\}}.\msg$.
+    \item \textbf{TS-LDDMM representation identifiability, \Cref{appendix: identifiability}:} On synthetic data, we evaluate the ability of our method to retrieve the parameter $v_0^*$ that encodes the deformation $\varphi^{\{v_0^*\}}$ acting on a time series graph $\msg$ by solving the geodesic shooting problem \eqref{eq:relaxation} between $\msg$ and $\varphi^{\{v_0^*\}}.\msg$.
      \textbf{Results} show that TS-LDDMM representations are identifiable or weakly identifiable depending on the velocity field kernel $K_G$ specification.
 
-    \item \textbf{Robustness to irregular sampling, \Cref{appendix: benchmark robustness}:} We compare the robustness of TS-LDDMM representation with 9 URL methods handling irregularly sampled multivariate time series on 15 shape-based datasets (7 univariates \& 8 multivariates).
+    \item \textbf{Robustness to irregular sampling, \Cref{appendix: robustness}:} We compare the robustness of TS-LDDMM representation with 9 URL methods handling irregularly sampled multivariate time series on 15 shape-based datasets (7 univariates \& 8 multivariates).
      We assess methods' classification performances under regular sampling (0\% missing rate) and three irregular sampling regimes (30\%, 50\%, and 70\% missing rates), according to the protocol depicted in \cite{kidger2020neural}.
@@
 
-    \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix: settings_classification}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates).
+    \item \textbf{Classification benchmark on regularly sampled datasets, \Cref{appendix: shape_classification}:} We compare performances of a kernel support vector machine (SVC) algorithm based on TS-LDDMM representation with 3 state-of-the-art classification methods from shape analysis on 15 shape-based datasets (7 univariates \& 8 multivariates).
 \textbf{Results} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.

### 2024-09-03 NEURIPS24/paper.tex
Revert "Delete NEURIPS24 directory"
This reverts commit e09ecaea218c43964bb3eb1e2a9a2c5bfd236ded.

@@
 
- All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix: settings_classification}.
+ All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix: robustness} and \Cref{appendix: shape_classification}.
  \vspace{-1ex}

### 2024-09-04 NEURIPS24/paper.tex
update

@@
   First, this optimization problem is ill-posed, and a penalty term is needed to obtain a unique solution.
-   In addition, we have to find a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, which allows us to solve this minimization problem efficiently.
-
+   In addition, a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, is sought to efficiently solve this minimization problem. 
+\paragraph*{From deformations to geodesics.}
 %\paragraph*{Regularizing the problem to derive uniqueness.}
@@
 By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy-Lipschitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$ \cite{miller2006geodesic}, as straight lines in Euclidean space.
-Denoting by $w(v_0)$ the geodesic starting from the identity with inital velocity $v_0$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \phi^v$ and the previous matching problem becomes a \textit{geodesic shooting problem}:
+Denoting by $\tau \mapsto \rho_{v_0}(\tau)\in\msh$ the geodesic starting from the identity with inital velocity $v_0\in \msv$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \rho_{v_0}(1)$.
+Using $\varphi^{\{v_0\}}$ instead of $\phi^v$, the previous matching problem becomes a \textit{geodesic shooting problem}:
  \begin{equation}
@@
    In our applications, the control points $(x_{k,0})_{k\in[\mathbf{n}_0]}$ can be understood as the discretized graph $(t_k,\mathbf{s}_0(t_k))_{k\in[\mathbf{n}_0]}$ of a starting time series $\mathbf{s}_0$.
-  With this parametrization of $v_0$, \citet{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same
+  With this parametrization of $v_0$, \cite{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same
   structure along time, such that for any $x\in\Rset^{d'}$ and $\tau\in[0,1]$, 
@@
 \vspace{-1ex}
+\sam{Lien avec functional data analysis ? Insister sur le fait que concaténer temps et espace c'est moisi}
 We consider in this paper observations which consist in a population of $N$ multivariate time series, for any $j\in[N]$, $s^j \in \rmC^1(\msi_j,\Rset^{d})$. 
@@
 \paragraph{Loss}
-
+\sam{Simplifier cette partie ?}
 This section specifies the distance function $\scrl$ introduced in the loss function defined in \eqref{eq:relaxation}.

### 2024-10-14 NEURIPS24/neurips_2024_to_submit.tex
update

@@
 % ready for submission
-\usepackage{neurips_2024}
+\usepackage[final]{neurips_2024}
 
@@
 \author{%
-  David S.~Hippocampus\thanks{Use footnote for providing further information
-    about author (webpage, alternative address)---\emph{not} for acknowledging
-    funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
+  Thibaut Germain \thanks{Corresponding author. Contact at \texttt{thibaut.germain@ens-paris-saclay.fr}} \\
+  Centre Borelli, ENS Paris-Saclay\\
+  4 av. des sciences, 91190 \\
+  %\texttt{thibaut.germain@ens-paris-saclay.fr} \\
   % examples of more authors
-  % \And
-  % Coauthor \\
-  % Affiliation \\
+  \And
+  Samuel Gruffaz \\
+  Centre Borelli, ENS Paris-Saclay\\
+  4 av. des sciences, 91190 \\
   % Address \\
-  % \texttt{email} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
+  %\texttt{samuel.gruffaz@ens-paris-saclay.fr} \\
+  \AND
+  Charles Truong \\
+  Centre Borelli, ENS Paris-Saclay\\
+  4 av. des sciences, 91190 \\
+  %\texttt{charles.truong@ens-paris-saclay.fr} \\
+  \And
+  Laurent Oudre \\
+  Centre Borelli, ENS Paris-Saclay\\
+  4 av. des sciences, 91190 \\
+  \And
+  Alain Durmus \\
+  CMAP, CNRS, Ecole polytechnique\\
+  Institut Polytechnique de Paris\\
+  91120 Palaiseau, France \\
 }
@@
 \maketitle
-%In this paper, we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
+%In this paper we propose an unsupervised representation method (TS-LDDMM) to analyze inter-individual variability of time series dataset.
     %Inspired by the shape analysis literature, we extend Large Deformation Diffeomorphic Metric Mapping (LDDMM) to time series, considering the temporal evolution as their distinctive feature.  
@@
 \input{paper.tex}
-
-\bibliographystyle{plain}
-\bibliography{ref.bib}
-
 \begin{ack}
@@
 
+\bibliographystyle{plain}
+\bibliography{ref.bib}
+
+\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2024-10-14 NEURIPS24/paper.tex
update
up

@@
 \label{section:LDDMM}
-
+\sam{Shape-FPCA employs the Square Root Velocity Field (SRVF) representation to separate space and time.
+However, this method is designed for continuous objects and only applies to time series of the same length.
+This issue can be addressed through interpolation, but this approach is not always reliable in sparse and irregular sampling scenarios.
+Most FDA papers [2,3,4] that we are aware of cope with the same issue using interpolation or basis function expansion.
+In a nutshell, FDA methods have to deal with continuous objects, while LDDMM algorithms can keep a discrete-to-discrete analysis.
+}
 In this part, we expose how to
@@
   \label{eq:geodesics_original}
-    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv\dd \tau
+    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
  \end{equation}
@@
 
+A pedagogical \href{https://tslddmmapp.streamlit.app}{\underline{online application}} is available to inspect the effect of hyperprameters on geodesic shooting \eqref{eq:integration} and registration \eqref{eq:relaxation}.
+
 %\sam{TO improve by Thibaut
@@
 \vspace{-1ex}
+
 Shape analysis focuses on statistical analysis of mathematical objects invariant under some deformations like rotations, dilations, or time parameterization.
@@
 They include temporal evolution, but not for the same purpose: the aim is to perform longitudinal modeling of brain images.
+
+ LDDMM framework has weaker assumptions than the Metamorphosis framework \cite{blanz2003face}.
+The 3DMM framework requires that each mesh be re-parametrized into a consistent form where the number of vertices, triangulation, and the anatomical meaning of each vertex are consistent across all meshes, as stated in the introduction of \cite{booth20163d}.
+ In our context, we do not need such pre-processing; the time series graph can have different sizes.
 

### 2024-10-14 NEURIPS24/paper.tex
update

@@
 \label{section:LDDMM}
-\sam{Shape-FPCA employs the Square Root Velocity Field (SRVF) representation to separate space and time.
-However, this method is designed for continuous objects and only applies to time series of the same length.
-This issue can be addressed through interpolation, but this approach is not always reliable in sparse and irregular sampling scenarios.
-Most FDA papers [2,3,4] that we are aware of cope with the same issue using interpolation or basis function expansion.
-In a nutshell, FDA methods have to deal with continuous objects, while LDDMM algorithms can keep a discrete-to-discrete analysis.
-}
+
 In this part, we expose how to
@@
 \vspace{-1ex}
-\sam{Lien avec functional data analysis ? Insister sur le fait que concaténer temps et espace c'est moisi}
+%\sam{Lien avec functional data analysis ? Insister sur le fait que concaténer temps et espace c'est moisi}
 We consider in this paper observations which consist in a population of $N$ multivariate time series, for any $j\in[N]$, $s^j \in \rmC^1(\msi_j,\Rset^{d})$. 
@@
     \begin{remark}
+      \label{remark:spaceandtime}
       With this choice of kernel, the features associated with the time transformation can be extracted from the momentums $(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}$ in \eqref{eq:def_v0} by taking the coordinates related to time.
       However, the features related to the space transformation are not only in the space coordinates since the related kernel $K_{\text{space}}$ depends on time as well.
+      The kernel's representation has been carefully designed to integrate both space and time, while ensuring that time remains independent of space.
+       Initially, we considered separating the spatial and temporal components.
+        However, post-hoc analysis of such a representation proved to be challenging.
+         The separated spatial and temporal representations are correlated, and understanding this correlation is essential for interpreting the data.
+          As a result, concatenating the two representations becomes necessary, though there is no straightforward method for doing so, as they are not commensurable.
+           Consequently, we opted for a representation that inherently integrates both space and time.
       \end{remark}
@@
 
- LDDMM framework has weaker assumptions than the Metamorphosis framework \cite{blanz2003face}.
-The 3DMM framework requires that each mesh be re-parametrized into a consistent form where the number of vertices, triangulation, and the anatomical meaning of each vertex are consistent across all meshes, as stated in the introduction of \cite{booth20163d}.
- In our context, we do not need such pre-processing; the time series graph can have different sizes.
-
  Leaving the LDDMM representation, the results of \cite{srivastava2010shape,heo2024logistic} address the representation of curves with the Square-Root Velocity (SRV) representation.
@@
  Consequently, the graph structure of the time series is not respected, and the original time evolution of the time series is not encoded in the final representation.
- Very recently, in a functional data analysis framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution.
- Nevertheless, this method is made for \textit{continuous objects} and only applies to time series of \textit{same length}, making the estimation more sensitive to noise.
+ Very recently, in a functional data analysis (FDA) framework, a paper \cite{wu2024shape} (Shape-FPCA) improved by representing the original time evolution.
+ However, the space and time representations remain correlated, complicating post-hoc analysis, as discussed in  \Cref{remark:spaceandtime}.
+ Additionally, this method is tailored for \textit{continuous objects} and applies only to time series of the \textit{same length}, making the estimation more sensitive to noise. 
+ This issue can be addressed through interpolation, but this approach is not always reliable in sparse and irregular sampling scenarios.
+ Most FDA approaches, as seen in \cite{shang2014survey,yu2017principal,warmenhoven2021pca}, address this challenge using interpolation or basis function expansion
+ In summary, FDA methods typically separate space and time representations for continuous objects, whereas TS-LDDMM algorithm maintain a discrete-to-discrete analysis, inherently integrating both space and time representations.
 
@@
  All these state-of-the-art methods previously mentionned \cite{glaunes2008large,oh2024stable,wu2024shape,heo2024logistic} are compared to TS-LDDMM in \Cref{appendix: robustness} and \Cref{appendix: shape_classification}.
+
+ Compared to the Metamorphosis framework \cite{blanz2003face}, LDDMM framework has weaker assumptions.
+ The 3DMM framework requires that each mesh be re-parametrized into a consistent form where the number of vertices, triangulation, and the anatomical meaning of each vertex are consistent across all meshes, as stated in the introduction of \cite{booth20163d}.
+In our context, we do not need such pre-processing; the time series graph can have different sizes.
  \vspace{-1ex}

### 2024-10-23 NEURIPS24/appendix.tex
update

@@
 
-    Let $\bar{f}\in \rmC^0(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^0(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
+    Let $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^1(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
     We have :
@@
     \end{align}
-    By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^0(\gamma(\msi), \Rset^d) $ by composition of continuous functions
+    By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^1(\gamma(\msi), \Rset^d) $ by composition of $C^1$ functions
     and $ \msg(\bar{s})= \Pi_{\gamma,f}.\msg(\mathbf{s}_0)$ by \eqref{eq:proof1_last_eq}, which concludes the proof.

### 2024-10-23 NEURIPS24/experiment.tex
update
improve exp

@@
 \label{section:experiments}
-
+\sam{Mettre expérience interprétation et sensibilité au bruit en appendice, améliorer l'interprétation des résultats}
 For conciseness, several experiments are relegated in appendix:
@@
 \end{enumerate}
-\begin{figure*}[t]
+
+\begin{figure}
   \centering
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[width = \textwidth]{pictures/ts_ddmm_shooting.pdf}
-    \caption{TS-LDDMM}
-    \label{fig:ts-lddmm shooting}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \includegraphics[width = \textwidth]{pictures/lddmm_shooting.pdf}
-    \caption{LDDMM}
-    \label{fig:lddmm shooting}
-  \end{subfigure}
-  \caption{Analysis of the two principal components (PC) related to mice' respiratory cycles before exposure for TS-LDDMM \Cref{fig:ts-lddmm shooting}, and LDDMM \Cref{fig:lddmm shooting}.
-  In both cases, I) displays PC densities according to mice genotype and II) displays  deformations of the reference graph $S_0$ along each PC.}
+  \includegraphics[width=\linewidth]{pictures/shooting before/shooting before.pdf}
+  \caption{Analysis of the two principal components (PC) related to mice ventilation before exposure with TS-LDDMM representations \textbf{(a)}, and LDDMM \textbf{(b)}.
+In both cases and for all PC, the left plot displays PC densities according to mice genotype and right plot displays deformations of the reference graph $\mathbf{c}_0$ along each PC.}
   \label{fig:exp1}
-\end{figure*}
+  \vspace{-1.5em}
+\end{figure}
 
-\vspace{-1ex}
 \begin{figure*}[t]
@@
     \includegraphics[width = \textwidth]{pictures/exp_1_cycle_example.jpeg}
-    \caption{colq cycle}
+    \caption{ColQ cycle}
     \label{fig:colq-cycle}
@@
     \includegraphics[width = \textwidth]{pictures/exp_1_wt.jpeg}
-    \caption{wt cycle}
+    \caption{WT cycle}
     \label{fig:wt-reference}
   \end{subfigure}
-  \caption{\Cref{fig:colq-cycle} is a \textbf{colq} respiratory cycle.
-\Cref*{fig: pcs-scatter} displays individual mouse reference respiratory cycle in the TS-LDDMM PC1-PC2 coordinates.
-\Cref{fig:wt-reference} is a referent respiratory cycle of a \textbf{wt} mouse learned with TS-LDDMM.
-}
-  \vspace{-1em}
+  \caption{\textbf{(a)} a ColQ respiratory cycle sample.
+\textbf{(b)} Referent respiratory cycle of individual mouse $\mathbf{c}_0^j$ in the TS-LDDMM PC1-PC2 coordinates system of $\mathbf{c}_0$.
+\textbf{(c)} a WT respiratory cycle sample.}
+  \vspace{-1.5em}
 \end{figure*}
-\vspace{-1ex}
-\subsection{Interpretability: analysis of respiratory behavior in mice}
+
+\subsection{Interpretability: mice ventilation analysis}
 \label{section:interpretability}
-\vspace{-1ex}
-%We consider a dataset composed of mouse respiratory time series before and after a drug injection.
-% A complete presentation of this dataset is given in \Cref{appendix:mouse_dataset}.
-%  The mouse are divided in two groups depending on their genotypes : \textit{colq} and \textit{wt}.
-%  We aim to study the difference of respiratory shapes according to the genotype.
-%   That is why, TS-LDDMM \eqref{eq:general_optimization_problem} is applied to derive the features representations $(\mathbf{\alpha}_j)_{j\in[N]}\in (\Rset^2)^N$ related to $N$=\sam{fill} respiratory time series coming from 14 mouse (7 \textit{colq} and \textit{wt}) before drug injection.
-%   Then, a Principal Components Analysis (PCA) is performed on $(\mathbf{\alpha}_j)_{j\in[N]}$
 
-%the densities of each genotype according to each PC are displayed.
-In Figure B), the deformations of the reference graph $S_0$ along each PC are given.
-In Figure D), the graph of reference $S^j$, also called barycenter, related to each mouse, is displayed according to their coordinates on PC1 and PC3.
-In Figure C) et E), illustrations of respiratory cycles related to mice coming from the \textbf{wt} and \textbf{colq} group are displayed.
+This experiment highlights the \textit{interpretability} of TS-LDDMM representation for studying the inter-individual variability in biomedical applications.
+We consider a time series dataset monitoring the evolution of mice's nasal and thoracic airflow when exposed to a drug altering respiration \cite{nervo2019respiratory}.
+The dataset includes recordings of 7 control mice (WT) and 7 mutant mice (ColQ) with an enzyme deficiency.
+The enzyme is involved in the respiration regulation, and the drug inhibits its activity.
+For each mouse, airflows were monitored for 15 to 20 minutes before the drug exposure and then for 35 to 40 minutes.
+A complete description of the dataset is given in the \Cref{appendix:mouse_dataset}.
 
-%\begin{figure*}[t]
-%  \centering
-%  \includegraphics[width=0.95\linewidth]{"./pictures/exp_1_bis.pdf"}
-%  \vspace{-1.5em}
-%  \caption{Analysis of the two principal components (PC) related to the mice respiratory cycles before exposure for TS-LDDMM \Cref{fig:ts-lddmm shooting}.
-%  In Figure A), the densities of each genotype according to each PC are displayed.
-In Figure B), the deformations of the reference graph $S_0$ along each PC are given.
-In Figure D), the graph of reference $S^j$, also called barycenter, related to each mouse, is displayed according to their coordinates on PC1 and PC3.
-In Figure C) et E), illustrations of respiratory cycles related to mice coming from the \textbf{wt} and \textbf{colq} group are displayed.
-}
-%  \label{fig:exp_1_PCA}
-%  %\vspace{-1em}
-%\end{figure*}
+\paragraph{Experimental protocol.} 
+We considered two experimental scenarios; the first focuses on mice ventilation before exposure to explore the inter-individual and genotype-specific variabilities.
+The second focuses on whole recordings to analyze the evolution of mice's ventilation after drug exposure.
+In both cases, the baseline protocol consists of first extracting $N$ respiratory cycles from the datasets with the procedure described in \cite{germain2023unsupervised}.
+Then, learning the referent respiratory cycle $\mathbf{c}_0$ and the representations  of respiratory cycles $(\bm{\alpha}_0^j)_{j \in [N]}$ by solving \eqref{eq:general_optimization_problem} using TS-LDDMM.
+$\bm{\alpha}_0^j$ being the momentum of the initial velocity field of the geodesic encodings the diffeomorphisms mapping $\mathbf{c}_0$ to the $j^{th}$ respiratory cycle.
+Finally, performing a Kernel-PCA on the initial velocity fields~\eqref{eq:def_v0} belonging to $\msv$ and encoded by the pairs $(\bm{\alpha}_0^j,\mathbf{c}_0)_{j \in [N]}$.
+The first experiment includes $N_1 = 700$ cycles collected before exposure.
+The second experiment includes $N_2 = 1400$ cycles with 25\% (resp.
+75\%) before (resp.
+after) exposure.
+We also performed the first experimental scenario with LDDMM representation, and \Cref{appendix: mice_exp_setting} describes the settings of both methods.
+Essentially, varifold losses are identical for both methods, and the velocity field kernels are set to encompass time and space scales.
 
-This experiment highlights the \textit{interpretability} of TS-LDDMM representation for studying the inter-individual variability in time series biomedical datasets.
-We consider a multivariate time series dataset that accounts for mice's nasal and thoracic airflow evolution when exposed to an irritant molecule altering respiratory functions \cite{nervo2019respiratory}.
-The dataset is divided into two groups, one composed of 7 control mice (\textbf{wt}) and the other of 7 mice (\textbf{colq}) deficient in an enzyme involved in the control of respiration.
-For each mouse, airflows were recorded for 15 to 20 minutes before exposure to the irritant molecule and then for 35 to 40 minutes.
-A complete
-description of the dataset is given in the \Cref{appendix:mouse_dataset}.
+\paragraph{Geodesic shooting along principal component directions.}
+Any principal component (PC), noted $v_0^{pc}$, from a kernel-PCA in $\msv$, is itself an initial velocity field encoded by a pair $(\mathbf{c}_0, \bm{\alpha}_0^{pc})$.
+PCs encode the principal axis of deformations, and it is possible to shoot along the geodesic they encode with the differential equations \eqref{eq:integration}, enabling interpretation of the main sources of deformations.
 
+\paragraph{Mice ventilation before exposure.}
+We focus on the analysis of the two first Principal Components (PC) for TS-LDDMM (\Cref{fig:exp1}a) and LDDMM (\Cref{fig:exp1}b).
+Looking at the geodesic shooting along PCs, \Cref{fig:exp1} shows that principal components learned with TS-LDMM lead to deformations that remain respiratory cycles.
+In contrast, deformations learned with LDDMM are challenging to interpret as respiratory cycles.
+The LDDMM velocity field kernel is a Gaussian anisotropic kernel that accounts for time and space scales; however, the entanglement of time and space dimensions in the kernel does not guarantee the graph structure, and it makes the convergence of the method complex (relative varifold loss error: TS-LDDMM: 0.06, LDDMM: 0.11).
+
+Regarding TS-LDDMM \Cref{fig:exp1}a, its PCs refer to deformations directions carrying different physiological meanings.
+Indeed, the geodesic shooting along these directions indicates that PC1 accounts for variations of the total duration of a respiratory cycle, while PC2 expresses the trade-off between inspiration and expiration duration.
+In addition, the distribution of ColQ respiratory cycles along PC1 is wider than in WT mice, indicating that the adaptation of mutant mice to their enzyme deficiency is variable.
+This observation can also be seen in \Cref{fig: pcs-scatter} where a referent respiratory cycle $\mathbf{c}_0^j$ is learned by solving \eqref{eq:general_optimization_problem} for each mouse and is encoded in the (PC1,PC2) coordinate system of $\mathbf{c}_0$ by registration \eqref{eq:geodesics_shooting}.
+Indeed, the average respiratory cycles of ColQ mice are more spread out than WT mice's.
+Going back to the densities of PC1, ColQ mice distribution has a heavier tail toward negative values compared to WT mice.
+When shooting in the opposite direction of PC1, we can observe that the inspiration is divided into two steps.
+Congruently with \cite{germain2023unsupervised}, such inspirations indicate motor control difficulties due to enzyme deficiency.
+\Cref{fig:colq-cycle} is an example of ColQ respiratory cycle with negative PC1 coordinate.
+
+\begin{figure*}[t]
+  \centering
+  \includegraphics[width=\linewidth]{pictures/exp after/exp after.pdf}
+  \caption{Analysis of the first Principal Component (PC1) related to mice ventilation before and after exposure with TS-LDDMM representations.
+\textbf{(a)} displays PC densities per mice genotype, \textbf{(b)} illustrates deformations of the reference respiratory cycle $\mathbf{c}_0$ along PC1, and \textbf{(c)} displays all respiratory cycles with respect to time in PC1 and PC3 coordinates}
+  \label{fig:exp_2_PCA}
+  \vspace{-1.5em}
+\end{figure*}
+
+\paragraph{Mice ventilation evolution after drug exposure.} 
+This experiment focuses on the first principal components learned from TS-LDDDM representations of respiratory cycles randomly sampled before and after drug exposure.
+\Cref{fig:exp_2_PCA}a illustrates the geodesic shootings along PC1.
+Again, PC1 accounts for variations in respiratory cycle duration, but more importantly, it can be observed on the deformation at -1.5 $\sigma_{\text{PC}}$ the apparition of a long pause after inspiration.
+Congruently, \Cref{fig:exp_2_PCA}c indicates that pauses appear after drug exposure as cycles with negative PC1 values mainly occur after 20 minutes and present more variability along PC3.
+In addition, \Cref{fig:exp_2_PCA}b shows a bimodal distribution for WT mice with one of the peaks in the negative values.
+This peak was not observed in the previous experiment \Cref{fig:exp1}a.
+It indicates that pauses after inspiration are prevalent in WT mice after drug exposure.
+On the other hand, the distributions of ColQ mice's respiratory cycles along PC1 in both experiments are similar and account for the same deformation, suggesting that ColQ mice weakly react to the drug exposure as they already adapt their enzyme deficiency.
+
+\paragraph{Experiment Conclusion.}
+Analyzing mice ventilation with TS-LDDMM representation highlights the method's ability to create meaningful interaction between experts and the data.
+Indeed, combining statistical and visual results shows that main deformations carry physiological meaning, enabling the characterization of some mice genotypes and the effects of drug exposure.
+
+\iffalse
 \paragraph{Protocol.} For both experiments and representations (TS-LDDMM and LDDMM), we derive the reference respiratory cycle's graph $S_0$ and the representations $(\alpha_j)_{j\in[N]}$ related to $N$ respiratory cycles extracted according to the procedure \cite{germain2023unsupervised} by solving  \eqref{eq:general_optimization_problem}.
@@
   \centering
-  \includegraphics[width=0.95\linewidth]{"./pictures/exp2.pdf"}
-  \caption{Analysis of the first Principal Component (PC1) related to mice' respiratory cycles before and after exposure for TS-LDDMM.
-Figure A) displays PC densities according to mice genotype, Figure B) displays  deformations of the reference graph $S_0$ along PC1, and Figure C) displays respiratory cycles with respect to time in PC1 and PC3 coordinates}
-  \label{fig:exp_2_PCA}
+  \includegraphics[width=0.95\linewidth]{pictures/exp after/exp after.pdf}
+  \caption{Analysis of the first Principal Component (PC1) related to mice ventilation before and after exposure with TS-LDDMM representations.
+\textbf{(a)} displays PC densities per mice genotype, \textbf{(b)} illustrates deformations of the reference respiratory cycle $\mmathbf{c}_0$ along PC1, and \textbf{(c)} displays all respiratory cycles with respect to time in PC1 and PC3 coordinates}
+  \label{fig:exp_2_PCAold}
   \vspace{-1.5em}
@@
 %  Additionally, we learn each mouse's reference respiratory cycle and represent them in the first and third PC coordinates system in Figure REF.
+
+\fi

### 2024-10-24 NEURIPS24/appendix.tex
add appendix

@@
 \begin{equation}
-   |\mu_X-\mu_Y |_{\msw^*}^2 = \sum_{i,j = 1}^{T_0-1}l_i k((x_i,\overrightarrow{v_i}),(x_i,\overrightarrow{v_i}^0))l_j - 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l_i k((x_i,\overrightarrow{v_i}),(y_i,\overrightarrow{w_i}))l_j' + \sum_{i,j = 1}^{T_1-1}l_i' k((y_i,\overrightarrow{w_i}),(y_i,\overrightarrow{w_i}))l_j' \eqsp.
+  \begin{array}{ll}
+    |\mu_X-\mu_Y |_{\msw^*}^2 = & \sum_{i,j = 1}^{T_0-1}l_i k((x_i,\overrightarrow{v_i}),(x_i,\overrightarrow{v_i}^0))l_j + \sum_{i,j = 1}^{T_1-1}l_i' k((y_i,\overrightarrow{w_i}),(y_i,\overrightarrow{w_i}))l_j' \\
+    & - 2 \sum_{i=1}^{T_0-1}\sum_{j=1}^{T_1-1}l_i k((x_i,\overrightarrow{v_i}),(y_i,\overrightarrow{w_i}))l_j' \eqsp.
+  \end{array}
 \end{equation}
@@
 
-\subsection{Benchmark metohds}
+\subsection{Benchmark methods}
 \begin{itemize}
@@
 
-\paragraph{Protocole}
+\subsection{Protocol}
 For each dataset and method, the evaluation protocol is a simple train,validation test with hyperparameter tuning:
@@
 
+\section{Appendix for the experiment: Noise sensitivity}
+This experiment evaluates the influence of noise on the learning of the reference sequence for TS-LDDMM and SRVF Kacher-mean, a subroutine of Shape-FPCA~\cite{wu2024shape}.
+
+\subsection{Protocol}
+The dataset includes 100 sine waves with randomly generated time parametrization by following the procedure described in \Cref{appendix: settings_identifiability} with $\text{Gen}(50,1,0)$ and uniformly resampled.
+The dataset has been altered under four scenarios with an additive Gaussian noise centered and with standard deviation $\sigma_\epsilon \in \{0,0.05,0.1,0.2\}$.
+The referent sequence is learned for each scenario, and the $L_2$-norm error between the exact and the learned barycenter is computed.
+
+\subsection{Method settings}
+For SRVF, the number of steps to learn the Kacher-mean is set to 20.
+Regarding TS-LDDMM, the velocity field kernel $K_G$ is set to $ (c_0,c_1,\sigma_{T,0},\sigma_{T,1},\sigma_x) = (1,0.1,65,1,1)$, and the varifold loss kernels $(k_{pos},k_{dir})$ are set to $(\sigma_{\pos,t}, \sigma_{\pos,t}, \sigma_{\dir,t}, \sigma_{\dir,x}) = (5,1,1,1)$.
+For TS-LDDMM, the optimizer is set with 400 epochs and a maximum learning rate $\eta_M = 0.1$.
+In all cases, the initial reference graph is selected in the dataset as a time series with the median length.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = 0.9\linewidth]{pictures/noise sensitivity/noise sensitivity.pdf}
+  \caption{Illustration of the learned barycenter (red) compared to the exact barycenter (green) for both TS-LDDMM \textbf{(a)} and Shape-FPCA \textbf{(b)}.
+The computation has been done for different level of noise $\epsilon \sim \mathcal{N}(0,\sigma_\epsilon)$ with $\sigma_\epsilon \in \{0,0.05,0.1,0.2\}$.}
+  \label{fig: noise sensitivity}
+\end{figure}
+
+\subsection{Results}
+\Cref{fig: noise sensitivity} illustrates the results.
+Noise level affects the learning reference graph in both cases, as depicted by the increasing error and the illustrations.
+However, the overall sine wave shape is better preserved by TS-LDDMM compared to SRVF Kacher-mean, for which the sine wave amplitude decreases as the noise increases.
+In addition, for TS-LDDMM, the regularity of the reference graph can be controlled by penalizing the norm of the velocity fields in the loss function.
+Further work on penalization will be conducted to handle noisy data better.
+
 \section{Appendix for experiment: Analysis of respiratory behavior in mice}
@@
 
-This experiment involves TS-LDDMM and LDDMM~\cite{glaunes2008large} methods.
-Both methods are run twice, first on respiratory cycles before exposure to the irritant molecule to capture mice breathing behavior at rest and on all respiratory cycles to capture the influence of the irritant molecule.
-Exposure to the irritant molecule leads to significant shape deformation in the respiratory cycles, and the terms must be added to the varifold loss to capture deformations at a large time scale.
+This experiment involves TS-LDDMM, LDDMM~\cite{glaunes2008large} and Shape-FPCA~\cite{wu2024shape} methods.
+Two scenarios are investigated: before drug exposure and before/after drug exposure.
+All methods are investigated on both scenarios.
 
@@
 \end{itemize}
+
+\paragraph{Shape-FPCA parameters.}
+For both scenarios, respiratory cycles are linearly interpolated and resampled to 200 points, and the length of the original time interval is kept.
+The computation of the Kacher-mean is done in a maximum of 50 iterations, and srv representations of the realigned time series and time parametrization are concatenated with cycle durations.
+When concatenating these vectors, the choice of amplitude factors is made to minimize the reconstruction error from the principal components analysis by following the procedure described in \cite{wu2024shape}.
+Shape-FPCA does not handle multivariate data, and we only kept the nasal airflow for this method.
+
+\subsection{Addiotinal results}
+\Cref{fig: after tslddmm} presents results for TS-LDDMM and \Cref{fig: after shapefpca}  presents results for Shape-FPCA.
+The main components look similar.
+However, a subtle difference, yet important, can be noticed.
+With Shape-FPCA, the deformation tends to be a uniform time scaling, whereas, with TS-LDDMM, the time dilatation mainly occurs during the pause between inspiration and expiration.
+Qualitatively, this last deformation fits the physiological phenomenon: Mice's muscles cannot relax after exposure to the irritant molecule, leading to pauses between inspiration and expiration~\cite{nervo2019respiratory}.
+Qualitatively, contrary to Shape-FPCA, which manages to represent the main phenomena in the data, the deformations of TS-LDDMM capture subtle physiological behaviors essential for understanding the phenomenon at hand.
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \linewidth]{pictures/exp after/exp after.pdf}
+  \caption{Analysis of the first Principal Component (PC1) related to mice ventilation before and after exposure with TS-LDDMM representations.
+\textbf{(a)} displays PC densities per mice genotype, \textbf{(b)} illustrates deformations of the reference respiratory cycle $\mathbf{c}_0$ along PC1, and \textbf{(c)} displays all respiratory cycles with respect to time in PC1 and PC3 coordinates}
+  \label{fig: after tslddmm}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width = \linewidth]{pictures/exp after/exp after fpca.pdf}
+  \caption{Analysis of the first Principal Component (PC1) related to mice ventilation before and after exposure with shape-FPCA representations.
+\textbf{(a)} displays PC densities per mice genotype, \textbf{(b)} illustrates deformations of the reference respiratory cycle $\mathbf{c}_0$ along PC1, and \textbf{(c)} displays all respiratory cycles with respect to time in PC1 and PC2 coordinates}
+  \label{fig: after shapefpca}
+\end{figure}

### 2024-10-24 NEURIPS24/experiment.tex
add appendix

@@
 \end{figure*}
-
+\vspace{-1ex}
 \subsection{Interpretability: mice ventilation analysis}
 \label{section:interpretability}
-
+\vspace{-1ex}
 This experiment highlights the \textit{interpretability} of TS-LDDMM representation for studying the inter-individual variability in biomedical applications.
@@
 A complete description of the dataset is given in the \Cref{appendix:mouse_dataset}.
+\vspace{-1ex}
 
@@
 Essentially, varifold losses are identical for both methods, and the velocity field kernels are set to encompass time and space scales.
+in addition, In addition, \Cref{appendix: mice_exp_setting}  presents a comparison between TS-LDDMM and Shape-FPCA on the second scenario.
+\vspace{-1ex}
 
@@
 PCs encode the principal axis of deformations, and it is possible to shoot along the geodesic they encode with the differential equations \eqref{eq:integration}, enabling interpretation of the main sources of deformations.
+\vspace{-1ex}
 
@@
 \end{figure*}
+\vspace{-1ex}
 
@@
 On the other hand, the distributions of ColQ mice's respiratory cycles along PC1 in both experiments are similar and account for the same deformation, suggesting that ColQ mice weakly react to the drug exposure as they already adapt their enzyme deficiency.
+\vspace{-1ex}
 

### 2024-10-25 NEURIPS24/appendix.tex
add last experiments

@@
 \section{Appendix for the experiment: Noise sensitivity}
+\label{appendix: noise sensitivity}
 This experiment evaluates the influence of noise on the learning of the reference sequence for TS-LDDMM and SRVF Kacher-mean, a subroutine of Shape-FPCA~\cite{wu2024shape}.

### 2024-10-25 NEURIPS24/experiment.tex
add last experiments

@@
 \textbf{Results} show that the TS-LDDMM-based method outperforms other methods (best performances over 13 datasets), making TS-LDDMM representation relevant for time series shape analysis.
+
+    \item \textbf{Noise sensitivity for learning the reference graph, \Cref{appendix: noise sensitivity}:} We evaluate the noise sensitivity of TS-LDDMM and Shape-FPCA \cite{wu2024shape} for learning the reference graph on a synthetic dataset and for several levels of additive Gaussian noise.
+\textbf{Results} show that both methods are sensitive to noise.
+However, TS-LDDMM preserves the overall shape while shape-FPCA alters the shape depending on the noise level.
 \end{enumerate}

### 2024-10-25 NEURIPS24/neurips_2024_to_submit.tex
update
add last experiments

@@
 
-\title{Shapes analysis for time series.}
+\title{Shapes analysis for time series}
 
@@
 \author{%
-  Thibaut Germain \thanks{Corresponding author. Contact at \texttt{thibaut.germain@ens-paris-saclay.fr}} \\
+  Thibaut Germain$^1$\thanks{Corresponding author. Contact at \texttt{thibaut.germain@ens-paris-saclay.fr}}\\
   Centre Borelli, ENS Paris-Saclay\\
@@
   \And
-  Samuel Gruffaz \\
+  Samuel Gruffaz$^1$\\
   Centre Borelli, ENS Paris-Saclay\\
@@
   \AND
-  Charles Truong \\
+  Charles Truong$^1$\\
   Centre Borelli, ENS Paris-Saclay\\
@@
   \And
-  Laurent Oudre \\
+  Laurent Oudre$^1$\\
   Centre Borelli, ENS Paris-Saclay\\
@@
 }
+
 \input{header.tex}
 \input{def.tex}
+
 \newcommand\sam[1]{\textcolor{blue}{#1}}
 \newcommand\thi[1]{\textcolor{orange}{#1}}
+
 \begin{document}
-
+\footnotetext[1]{Complete affiliation: Université Paris Saclay, Université Paris Cité, ENS Paris Saclay, CNRS, SSA, INSERM, Centre
+Borelli, F-91190, Gif-sur-Yvette, France.}
 \maketitle
@@
 \begin{ack}
-  Use unnumbered first level headings for the acknowledgments. All acknowledgments
-  go at the end of the paper before the list of references. Moreover, you are required to declare
-  funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-  More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.
-
-  Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
+This work was supported by grants from Région Ile-de-France (DIM MathInnov).
+Charles Truong is funded by the PhLAMES chair of ENS Paris-Saclay.
 \end{ack}

### 2024-10-25 NEURIPS24/paper.tex
improve exp
add appendix
add last experiments

@@
 by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
- by $||u||_\infty=\inf_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
+ by $||u||_\infty=\sup_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
 and by $\Nset_{>0}$ is the set of positive integers. 
@@
 \paragraph{Loss}
-\sam{Simplifier cette partie ?}
 This section specifies the distance function $\scrl$ introduced in the loss function defined in \eqref{eq:relaxation}.

### 2024-10-25 NEURIPS24/experiment.tex
add github

@@
 \label{section:experiments}
-\sam{Mettre expérience interprétation et sensibilité au bruit en appendice, améliorer l'interprétation des résultats}
-For conciseness, several experiments are relegated in appendix:
+The source code is available on Github\footnote{\url{https://github.com/thibaut-germain/TSLDDMM}}. For conciseness, several experiments are relegated in appendix:
 
@@
 \textbf{Results} show that our method, TS-LDDMM, outperforms all methods for sampling regimes with missing rates: 0\%, 30\%, and 50\%.
-      %{\color{red} J'enlèverai cette phrase ou je développerai en Appendix : The performance decrease of the Kernel-SVC based on TS-LDDMM representation is sensibly due to the misspecification of its regularization parameter.}
 
@@
 \end{figure*}
+
 \vspace{-1ex}
@@
 Indeed, combining statistical and visual results shows that main deformations carry physiological meaning, enabling the characterization of some mice genotypes and the effects of drug exposure.
-
-\iffalse
-\paragraph{Protocol.} For both experiments and representations (TS-LDDMM and LDDMM), we derive the reference respiratory cycle's graph $S_0$ and the representations $(\alpha_j)_{j\in[N]}$ related to $N$ respiratory cycles extracted according to the procedure \cite{germain2023unsupervised} by solving  \eqref{eq:general_optimization_problem}.
- Then, we perform a kernel PCA on the initial velocity field $\left(v_0(\alpha_j,S_0)\right)_{j\in[N]}\in \msv^{N_1}$ defined in \eqref{eq:def_v0} to breathing behavior.
-  The first experiment includes $N_1 = 700$ respiratory cycles collected before exposure.
-The second experiment includes $N_2 = 1400$ respiratory cycles with 25\% (resp.
-75\%) before (resp.
-after) exposure.
-   \Cref{appendix: mice_exp_setting} describes methods settings.
-    Essentially, varifold losses are identical for both methods, and the velocity field kernels are set to encompass time and space scales.
-
-    \vspace{-1ex}
-\paragraph{Breathing behaviors before exposure.}
- We focus on the analysis of the two first Principal Components (PC) for TS-LDDMM (\Cref{fig:ts-lddmm shooting}), and LDDMM (\Cref{fig:lddmm shooting}).
-  A respiratory cycle is an inspiration (positive flow) followed by an expiration (negative flow); \Cref{fig:wt-reference} displays an example of \textbf{wt} mouse respiratory cycle.
-   \Cref{fig:exp1} shows that principal components learned with TS-LDMM lead to deformations that remain respiratory cycles, while deformations learned with LDDMM are challenging to interpret as respiratory cycles.
-   The LDDMM velocity field kernel is a Gaussian anisotropic kernel that accounts for time and space scales; however, the entanglement of time and space dimensions in the kernel does not guarantee the graph structure, and it makes the convergence of the method difficult (relative varifold loss error: TS-LDDMM: 0.06, LDDMM: 0.11).
-    Regarding TS-LDDMM, \Cref{fig:ts-lddmm shooting} shows its principal components refer to different types of deformations.
-     By interpreting \Cref{fig:ts-lddmm shooting}: PC1 accounts for time warping, PC2 expresses the trade-off between inspiration and expiration duration.
-      Compared to \textbf{wt} mice, the distribution of \textbf{colq} mice feature along the PC1 axis has a heavy left tail, and the associated deformation (-2 $\sigma_{\text{PC}}$) shows an inspiration with two peaks.
-       \Cref{fig:colq-cycle} shows an example of such respiratory cycles, which may be caused by motor impairment due to enzyme deficiency \cite{germain2023unsupervised}.
-        Finally, \Cref{fig: pcs-scatter} shows that PC1 and PC2 capture the main differences between the two groups as their respective reference graphs $S^j$ are located in different parts of the space.
-
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=0.95\linewidth]{pictures/exp after/exp after.pdf}
-  \caption{Analysis of the first Principal Component (PC1) related to mice ventilation before and after exposure with TS-LDDMM representations.
-\textbf{(a)} displays PC densities per mice genotype, \textbf{(b)} illustrates deformations of the reference respiratory cycle $\mmathbf{c}_0$ along PC1, and \textbf{(c)} displays all respiratory cycles with respect to time in PC1 and PC3 coordinates}
-  \label{fig:exp_2_PCAold}
-  \vspace{-1.5em}
-\end{figure*}
-\vspace{-1ex}
-\paragraph{Breathing behaviors' evolution after exposure to the irritant molecule} 
-TS-LDDMM representation features are learned on a dataset that includes respiratory cycles before and after exposure.
- \Cref{fig:exp_2_PCA} displays the first principal component PC since it encodes the effect of the irritant molecule as depicted in \Cref{fig:exp_2_PCA}-C) (the exposure occurs at 20 minutes).
-\Cref{fig:exp_2_PCA}-B) shows that the deformation (-1.5 $\sigma_{\text{PC}}$) leads to longer respiratory cycles that include a pause between inspiration and expiration as observed in \cite{germain2023unsupervised}.
-  Conjointly, \Cref{fig:exp_2_PCA}-A) shows a bimodal distribution for \textbf{wt} mice whereas \textbf{wt}' distributions before exposure were unimodal (\Cref{fig:ts-lddmm shooting}). Indeed, the irritant molecule inhibits the action of the deficient enzyme, \textbf{wt} mice strongly react to the irritant molecule, whereas \textbf{colq} mice better adapt due to their deficiency \cite{germain2023unsupervised}.
-%   \vspace{-1ex}
-% \paragraph{Result summary.} By comparing the shape of individual respiratory cycles, we show that TS-LDDMM features can encode genotype distinctive breathing behaviors and their evolution after exposure to the irritant molecule contrary to LDDMM \cite{glaunes2008large}.
-% \vspace{-1ex}
-%Secondly, we compare breathing behaviors before and after exposure to observe the impact of the irritant molecule.
-% We follow the same procedure as for before exposure, but we take $N_2=1400$ respiratory cycles extracted according 
-% to the procedure CITE....
-% In \Cref{fig:exp_2_PCA}, we focus on the first Principal Component (PC) since it encodes the effect of the irritant molecule as demonstrated on Figure \ref{fig:exp_2_PCA}-C) (exposure at time 20).
-% We observe on \Cref{fig:exp_2_PCA}-B) that after exposure the mouse have longuer breath such that the expiration is longuer than inspiration.
-%  Moreover, we deduce from Figure \ref{fig:exp_2_PCA}-A) that \textbf{colq} are more constant in their breath compared to \textbf{wt} after exposure.
-%
-% In Figure REF, we 
-% also display the reference respiratory cycle S0 and its deformations in the principal component directions.
-%  Additionally, we learn each mouse's reference respiratory cycle and represent them in the first and third PC coordinates system in Figure REF.
-
-\fi

### 2024-10-25 NEURIPS24/paper.tex
add github

@@
-% citep
-% \section{TODO}
-% %De ce que je comprends , les gens font surtout des variational auto encoder, ou du masking
-% %Mentionner que la méthode proposé peut être utilisé avec des réseaux
-% %\sam{Approche NTK, fédéré personalisation, projet avec Axel, Lien entre continuous normalizing flow}
-% Plan intro: 
-% \begin{itemize}
-%     \item Representation of patterns in time series has several applications in biology, but halso for classification, deep learning finds features automaticly.
-%      Taking or not taking into account time can be done with DTW. However, either the representation is the result of a black box and need post analysis to be understand,
-%       which is difficult to apply statiscal tools on it, or the representation is desgned by hand to select relevant features, but the capacity of analysis is limited by the knowledge of the designer.
-%      \item In shape analysis, representation of shapes for statistical analysis in a long time issue, which has been solved mainly by the LDDMM framework.
-Representing shapes as diffeomorpshims of a common shape.
-%      \item However, this tool has not been yet applied on time series, even if there is already some tools developped for oriented curves. 
-%      \item In this work, we propose to apply LDDMM to time series and to show its interest to derive an interpretable and unsupervised representation of patterns in time series.
-%      We proposed some tools developped in the shape analysis community to the researcher interested in machine learning and time series ones.
-%      \item First, we introduce our assumptions regarding the dataset and the main problem of diffeomorpshim learning. Then, we show how to solve it with LDDMM .
-%       We expose how to apply it to time series by selecting wisely a RKHS kernel, motivating it with a representation theorem.
-%       experiments on synthetic data are presented to show that the parameter of representation are identifiable when the time seriess are not too sharp and the kernels parameter well tuned. 
-%       Then, we sh-ow how to apply the method on an unsupervised case of moouse respeitaory dataset and a supervised one on ... .
-
-% \end{itemize}
-
-% One option to tackle this issue is to derive ...
-% feature representation of time series which depends on the problem at hand.
-% which is parti... -> which is of prime interest in ...
 \vspace{-1ex}
@@
 \vspace{-1ex}
-%https://arxiv.org/abs/1811.00075
-%Appliqué sur les "UWaveGestureLibrary","ArticularyWordRecognition","Cricket" pour l'archive UEA
-%"ArrowHead", "BME", "ECG200","FacesUCR","GunPoint" dans le cas univarié
-
-%Idée de parler d'inter individualité, spécialiser davantage
-%k shape, shapelets , ye2009time
 
@@
 
- Although community structures with representatives can be learned in an unsupervised manner \cite{trirat2024universal, meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised, tonekaboni2021unsupervised, meng2023unsupervised} or similarity measures \cite{asgari2023clustering, germain2023unsupervised, paparrizos2015k, ye2009time},
-  the study of inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies, shirato2023identifying} remains an open problem in unsupervised representation learning (URL), particularly for \textit{irregularly sampled} time series with \textit{variable lengths}.
+Although community structures with representatives can be learned in an unsupervised manner \cite{trirat2024universal, meng2023unsupervised} using contrastive loss \cite{franceschi2019unsupervised, tonekaboni2021unsupervised, meng2023unsupervised} or similarity measures \cite{asgari2023clustering, germain2023unsupervised, paparrizos2015k, ye2009time}, the study of inter-individual variability of shapes within a cluster \cite{niennattrakul2007inaccuracies, shirato2023identifying} remains an open problem in unsupervised representation learning (URL), particularly for \textit{irregularly sampled} time series with \textit{variable lengths}.
 
@@
 $\msg(s^j) \sim \phi_j .
-\msg(\mathbf{s}_0)$ .
+\msg(\mathbf{s}_0)$.
 The diffeomorphism $\phi_j$ is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
@@
 This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
-   %In this regard, this work is not an application of LDDMM to 1D curves \cite{glaunes2008large}.
 
-     %In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
-
-    %  Our contributions can be summarized as follows:
-    %  \begin{itemize}
-    %    \item We propose an unsupervised method (TS-LDDMM) to analyze inter-individual variability of shapes in a time series dataset.
-In particular, the method can handle \textit{irregularly sampled} time series with \textit{variable sizes}.
-    %    \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-    %    \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method with respect to its hyperparameters, also providing guidelines for tuning.
-    %    We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-    %     We illustrate the quantitative interest of the representation on classification tasks on real shape-based datasets.
-
-    %  \end{itemize}
 Our contributions can be summarized as follows:
@@
 In particular, the method can handle multivariate time series \textit{irregularly sampled} and with \textit{variable sizes}.
+
   \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
+
   \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters (\Cref{appendix: settings_identifiability}), also providing guidelines for tuning (\Cref{appendix:kernel_TS_LDDMM}).
+
   \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset (\Cref{section:experiments}).
-        \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets with regular and irregular sampling (\Cref{appendix: robustness}-\ref{appendix: settings_classification}).
+
+  \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets with regular and irregular sampling (\Cref{appendix: robustness,appendix: shape_classification}).
 \end{itemize}
-
-    %To the best of our knowledge, these theoretical and practical works have never been done before.
-      % On synthetic data, we show the sensibility of the results according to the kernel parameters and give guidelines to tune it.
-      % In an unsupervised framework, we vectorize respitory patterns of a mouse dataset and analyze the variations depending on the drug taken by the mouse.
-      % Finally, we compare to the state-of-the-art methods in classification an ... classifier taking as inputs the representation given by LDDMM.
-
-  %    \paragraph*{Related papers}
-  %    %https://arxiv.org/pdf/2106.00750.pdf look to the related papers
-  %  In the shape analysis community, a famous paper \cite{srivastava2010shape} addresses the problem of representation of curves $C:\msi \to \Rset^d$ by using the Square-Root Velocity (SRV) representation.
-  %      This structure is particularly relevant to compute geodesics and distances between curves when the space is quotiented by time reparametrization.
-  %       However, in our case the time axis deserves a particular attention, since the curves are derived from time-series. 
-  %       %While \cite{srivastava2010shape} tackles the computations of geodesics and distances on quotient curves space, our aim is to get a vector representation of each pattern. 
-  %       Very recently, in \cite{heo2024logistic}, this representation was used to compute PCA and to perform classification on quotient curves space, we follow the same ideas, but we focus on time series and we use LDDMM instead of SRV.
- % optimization framework with a well chosen neural architecture in order to learn automatically the features, which has been very 
-% To understand these complex data, we are looking to simpler representations toperform classical statiscal analysis or to perform efficient downstream tasks of machine learning, the time series representation res time series to a real valued vector representation.
-in order to apply classical statiscal or machine learning tools depending on the needs such as Principal Component Analysis (PCA), logistic regression or deep learning methods for classification or anomaly detection.
-
-% \thi{
-%   Our goal is to analyze the inter-individual variability within a time series dataset, an approach of prime interest in physiological contexts \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
-More specifically, we aim to find an unsupervised features representation method that encodes individual time series specificities compared to a reference one.
-% }
-
-% \thi{
-%   Studying shape differences between time series related to biological mechanisms is a common practice in physiology to characterize healthy and pathological functioning CITE.
-For instance, the shapes of heartbeats in electrocardiograms are discriminant for some cardiovascular pathologies CITE.
-Several approaches have been proposed for such comparison.
-Some employ shape-based similarity measures between time series \cite{asgari2023clustering,germain2023unsupervised,paparrizos2015k,ye2009time},  others embed time series as vectors of predefined features CITE, and, with the rise of deep neural networks, unsupervised learning representation of time series~\cite{trirat2024universal,meng2023unsupervised} has shown to be a valuable approach CITE notably with contrastive learning~\cite{franceschi2019unsupervised,tonekaboni2021unsupervised,meng2023unsupervised}.
-However, shape-based representation of time series within cohorts~\cite{niennattrakul2007inaccuracies,shirato2023identifying} remains an open problem in URL.
-% }
-
-% \thi{
-%   Our work focuses explicitly on learning shape-based representation of time series.
-First, we propose not to see the shape of a time series through its curve $\{s_t:\eqsp t\in\msi\}$, but rather through its graph $\msg(s)=\{(t,s(t)):\eqsp t\in \msi\}$.
-%    Then, building on the shape analysis literature \cite{beg2005computing,vaillant2004statistics}, we follow the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework \cite{beg2005computing,vaillant2004statistics} to analyze these graphs.
-The idea is to represent each element $\msg(s^j)$ of a dataset $(s^j)_{j\in[N]}$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$ by a diffeomorphism $\phi_j$ (ie $\msg(s^j) \sim \phi_j .
-\msg(\mathbf{s}_0)$ ).
-%     The diffeomorphism $\phi_j$ is learned by integrating an ordinary differential equation parameterized by a Reproducing Kernel Hilbert Space (RKHS).
-%      The parameters $(\alpha_j)_{j\in[N]}$ encoding the diffemorphisms $(\phi_j)_{j\in[N]}$ yield the representation features of the graphs $(\msg(s^j))_{j\in[N]}$.
-Finally, these shape-encoding features can feed any statistical or machine-learning model.
-%   }
-
-% \thi{
-%   However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
-\Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
-Our contributions arise from this observation: we specify the class of diffeomorphisms to consider and show how to learn them.
-This change is fruitful in representing transformations of time series graphs as illustrated in \Cref{fig:transport}.
-% }
-
-% \thi{
-%   Our contributions can be summarized as follows:
-%     \begin{itemize}
-%        \item We propose an unsupervised method (TS-LDDMM) to analyze the inter-individual variability of shapes in a time series dataset.
-In particular, the method can handle multivariate time series \textit{irregularly sampled} and with \textit{variable sizes}.
-%        \item We motivate our extension of LDDMM to time series by introducing a theoretical framework with a representation theorem for time series graph (\Cref{theorem:representation}) and kernels related to their structure (\Cref{lemma:choice_of_kernel_V}).
-%        \item We demonstrate the identifiability of the model by estimating the true generating parameter of synthetic data, and we highlight the sensitivity of our method concerning its hyperparameters, also providing guidelines for tuning.
-%        \item We highlight the \textit{interpretability} of TS-LDDMM for studying the inter-individual variability in a clinical dataset.
-%         \item We illustrate the quantitative interest of such representation on classification tasks on real shape-based datasets.
-%  \end{itemize}
-% }
 
@@
   \includegraphics[width=0.7\linewidth]{"./pictures/diffeo.jpeg"}
-
   \caption{A time series' graph $\msg=\{(t,s(t)): \eqsp t\in\msi\} $ can lose its structure after applying a general diffeomorphism $\phi.\msg$: a time value can be related to two values on the space axis.}
   \label{fig:diffeo}
-
 \end{figure}
@@
 \section{Notations}
-We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$,
-by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$,
- by $||u||_\infty=\sup_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$,
-and by $\Nset_{>0}$ is the set of positive integers. 
-%We are in a sub class of curve quotiented by temporal reparametrisation since a time series' graph can be understood as the equivalent class of t ->(t,s(t))
+We denote by integer ranges by $[k:l]=\{k,\ldots,l\}\subset \mathcal{P}(\Zset)$ and $ [l]=[1:l]$ with $k,l\in \Nset$, by $\rmC^m(\msi,\mse)$ the set of $m$-times continously differentiable function defined on an open set $\msu$ to a normed vector space $\mse$, by $||u||_\infty=\sup_{x\in \msu} |u(x)| $ for any bounded function $u:\msu \to \mse$, and by $\Nset_{>0}$ is the set of positive integers.
 
-%Reproducing Kernel Hilbert Space (RKHS)
 \vspace{-1ex}
@@
 
-In this part, we expose how to
- learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
+In this part, we expose how to learn the diffeomorphisms $(\phi_j)_{j\in[N]}$ using LDDMM, initially introduced in \cite{beg2005computing}.
 In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
@@
 Given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$\footnote{Note that we denote by $d'\in\nset$ the ambient space}, a set of starting points $\mathbf{x}=(x_{i})_{i\in[T_1]}$ in $\Rset^{d'}$, we aim to find a diffeomorphism $\phi$ such that the finite set of points $\mathbf{y}$ is similar in a certain sense to the set of finite sets of transformed points $\phi \cdot \mathbf{x} =(\phi(x_i))_{i\in[T_1]} $.
- The function $\phi$ is occasionally referred to as a \textit{deformation}. In general, these sets $\mathbf{x},\mathbf{y}$ are meshes of continuous objects, e.g., surfaces, curves, images, and so on.
+The function $\phi$ is occasionally referred to as a \textit{deformation}.
+In general, these sets $\mathbf{x},\mathbf{y}$ are meshes of continuous objects, e.g., surfaces, curves, images, and so on.
 
 \vspace{-1ex}
-% LDDMM is designed to analyzed an existing dataset, while NF and CNF are made to generalize a dataset for data augmentation.
 \paragraph{Representing diffeomorpshims as deformations.}
@@
 \end{equation}
-where the velocity field is $v:\tau\in [0,1]\mapsto v_\tau\in \msv $
-and $\msv$ is a Hilbert space of continuously differentiable function
-on $\Rset^{d'}$.  If
-$||\dd u ||_{\infty}+|| u ||_{\infty}\leq ||u ||_\msv $ for any
-$u\in \msv$ and
-$v\in \rml^2([0,1],\msv)=\{v\in \rmC^0([0,1],\msv): \int_0^1 ||
-v_\tau||^2_\msv \dd \tau<\infty \} $, by \citep[Theorem 5]{glaunes2005transport}
-$\phi^v$ exists and belongs to $\mcd(\Rset^{d'})$, where we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
+where the velocity field is $v:\tau\in [0,1]\mapsto v_\tau\in \msv $ and $\msv$ is a Hilbert space of continuously differentiable function on $\Rset^{d'}$.
+If $||\dd u ||_{\infty}+|| u ||_{\infty}\leq ||u ||_\msv $ for any $u\in \msv$ and $v\in \rml^2([0,1],\msv)=\{v\in \rmC^0([0,1],\msv): \int_0^1 ||v_\tau||^2_\msv \dd \tau<\infty \} $, by \citep[Theorem 5]{glaunes2005transport} $\phi^v$ exists and belongs to $\mcd(\Rset^{d'})$, where we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
 Therefore, for any choice of $v$, $\phi^v$ defines a valid deformation.
@@
 
-With this in mind, the velocity field $v$ fitting the data can be
-estimated by minimizing 
-$v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
+With this in mind, the velocity field $v$ fitting the data can be estimated by minimizing $v \in \rml^2([0,1],\msv) \mapsto \mathscr{L}(\phi^{v}.\mathbf{x},\mathbf{y})$, where $\mathscr{L}$ is an appropriate loss function.
 However, two computational challenges arise.
@@
 In addition, a parametric family $\msv_{\Theta} \subset \rml^2([0,1],\msv)$, parameterized by $\Theta$, is sought to efficiently solve this minimization problem.
-\paragraph*{From deformations to geodesics.}
-%\paragraph*{Regularizing the problem to derive uniqueness.}
+
+\vspace{-1ex}
+\paragraph{From deformations to geodesics.}
 It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$.
@@
 
-% More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
-%  the following squared norm can be defined
-%  \begin{equation}
-%   \label{eq:geodesics_original}
-%     \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
-%  \end{equation}
-%      as the minimal "energy" needed to perform the deformation $g$.
-% By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
-% However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
-% To circumvent this issue, another characterization of geodesics can be considered.
-%  As in Riemannian geometry, 
-%  instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
-%  \sam{To CHANGE}
-%  %In other words, an Exponential map on deformations is wanted:
-% given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
-% %  \begin{equation}
-% %   \label{eq:geodesics_shooting}
-% %   v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
-% %  \end{equation}
-% % since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
-%  By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
-% Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.
-%  Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
-%  Moreover, we have $\mathscr{R}(\phi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in L^2([0,1],\msv)$.
-%  %Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
-%  This is presented with more details in the next paragraph.
-
 \vspace{-1ex}
@@
 In this part, $\msv$ is chosen as an RKHS \cite{berlinet2011reproducing} generated by a smooth kernel $K$ (e.g., Gaussian).
-We follow \cite{durrleman2013sparse} and define a 
- discrete parameterization of the velocity fields to perform geodesics shooting \eqref{eq:geodesics_shooting}.  
+We follow \cite{durrleman2013sparse} and define a discrete parameterization of the velocity fields to perform geodesics shooting \eqref{eq:geodesics_shooting}.
 The initial velocity field $v_0$ is chosen as a finite linear combination of the RKHS basis vector fields,
@@
 In our applications, the control points $(x_{k,0})_{k\in[\mathbf{n}_0]}$ can be understood as the discretized graph $(t_k,\mathbf{s}_0(t_k))_{k\in[\mathbf{n}_0]}$ of a starting time series $\mathbf{s}_0$.
-  With this parametrization of $v_0$, \cite{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same
-  structure along time, such that for any $x\in\Rset^{d'}$ and $\tau\in[0,1]$, 
+With this parametrization of $v_0$, \cite{miller2006geodesic} show that the velocity field $v$ of the solution of \eqref{eq:geodesics_shooting} keeps the same structure along time, such that for any $x\in\Rset^{d'}$ and $\tau\in[0,1]$,
 \begin{equation}
@@
 \end{equation}
-  These equations are derived from the hamiltonian $H:(\alpha_k,x_k)_{k\in [\mathbf{n}_0]}\mapsto \sum_{k,l=1}^{\mathbf{n}_0} \alpha_{k}^\top K(x_k,x_l)\alpha_{l}  $, such that
-  the velocity norm is preserved $||v_\tau||_{\msv}=||v_0||_\msv $ for any $\tau\in [0,1]$.
+These equations are derived from the hamiltonian $H:(\alpha_k,x_k)_{k\in [\mathbf{n}_0]}\mapsto \sum_{k,l=1}^{\mathbf{n}_0} \alpha_{k}^\top K(x_k,x_l)\alpha_{l}  $, such that the velocity norm is preserved $||v_\tau||_{\msv}=||v_0||_\msv $ for any $\tau\in [0,1]$.
 By \eqref{eq:integration}, the velocity field related to a geodesic $v^*$ is fully parametrized by its initial control points and momentum $(x_{k,0},\alpha_{k,0})_{k\in[\mathbf{n}_0]}$.
-   Thus, given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$, a set of starting points $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$ in $\Rset^{d'}$, a RKHS's kernel $K:\Rset^{d'}\times \Rset^{d'}\to \Rset^{d'\times d'}$, a distance on sets $\mathscr{L}$,
- a numerical integration scheme of ODE and a penalty factor $\lambda>0$, the basic geodesic shooting step minimizes the following function using a gradient descent method:
+Thus, given a set of targets $\mathbf{y}=(y_i)_{i\in[T_2]}$ in $\Rset^{d'}$, a set of starting points $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$ in $\Rset^{d'}$, a RKHS's kernel $K:\Rset^{d'}\times \Rset^{d'}\to \Rset^{d'\times d'}$, a distance on sets $\mathscr{L}$, a numerical integration scheme of ODE and a penalty factor $\lambda>0$, the basic geodesic shooting step minimizes the following function using a gradient descent method:
 \begin{equation}
@@
 
-   %Before to specify the choice of the RKHS kernel $K$ and the loss on sets $\mathscr{L}$ in \Cref{section:time_series_specificity}, we first discuss the connection of LDDMM with normalizing flows.
-% to highlight the interest of the LDDMM approach.
-
-%Conversely to NF or CNF, by \eqref{eq:integration}, different momentums parameters raise different final deformations $\phi^{\{v_0\}}$, this injectivity properties offers interpretability of the representation and gives sens to PCA or LDA post-processing as exposed in \Cref{section:experiments}.
-
-   %If a diffeomorpshim $\Phi$ is given and we want to represent it as a deformation $\phi^v$ with $v$ of minimal ernergy, then
-  %$(\alpha_{k,0})_k$ is the only parameters to learn since $(x_{k,0})_k$ is fixed.
-
-%   \begin{algorithm}[tb]
-%     \caption{Geodesic shooting with LDDMM}
-%     \label{alg:example}
-%  \begin{algorithmic}
-%     \STATE {\bfseries Input:} a RKHS's kernel $K$, a loss on sets $L$, $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$, $\mathbf{y}=(y_i)_{i\in[T_2]}$ two sets of $\Rset^{d'}$, initial momentums 
-%     $(\alpha_{k,0})_{k\in [T_1]} \in (\Rset^{d'})^{T_1}$ and a number of gradient iterations $n_{\text{iter}}$.
-
-%     \STATE $f: (\alpha_k)_k\mapsto L\left(\phi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2  $ with $v_0$ and $\phi^{\{v_0\}}$
-%      respectively defined by \eqref{eq:def_v0} and \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_k $. 
-%      \sam{Peut être à définir dans un autre algo ?}
-%     \STATE $\alpha_{k,0}^*\gets \alpha_{k,0}, \eqsp k\in[T_1]$
-%     \FOR{$i=1$ {\bfseries to} $n_{\text{iter}}$}
-%     \STATE $\alpha_{k,0}^*\gets \text{AdamStep}(f,\alpha_{k,0}^*$) , where $\text{AdamStep}$ is given in appendix (mettre ref)
-%     \ENDFOR
-%     \STATE {\bfseries Return:} $(\alpha_{k,0}^*)_{k\in [T_1]}$
-
-%  \end{algorithmic}
-%  \end{algorithm}
 \vspace{-1ex}
 \paragraph{Relation to Continuous Normalizing Flows.}
-
 One particular popular choice to address the problem of learning a diffeomorphism or a velocity field is Normalizing Flows \cite{rezende2015variational,kobyzev2020normalizing} (NF) or their continuous counterpart \cite{chen2018neural,grathwohl2019scalable,salman2018deep} (CNF).
@@
 In addition,  they are primarily designed to have tractable Jacobian functions, while we do not require such property in our applications.
-Finally, the use of a differential flow solution of an ODE
-\eqref{eq:LDDMM_dynamic} trick is also at the basis of CNF, which
-then consists of learning a velocity field to address in fitting the data through a loss aiming to address the problem
-at hand. Nevertheless, the main difference between CNF and LDDMM lies in the
-parametrization of the velocity field. LDDMM uses kernels to
-derive closed form formula and enhance interpretability while
-NF and CNF take advantage of deep neural networks to scale with
-large dataset in high dimensions.
+Finally, the use of a differential flow solution of an ODE \eqref{eq:LDDMM_dynamic} trick is also at the basis of CNF, which then consists of learning a velocity field to address in fitting the data through a loss aiming to address the problem at hand.
+Nevertheless, the main difference between CNF and LDDMM lies in the parametrization of the velocity field.
+LDDMM uses kernels to derive closed form formula and enhance interpretability while NF and CNF take advantage of deep neural networks to scale with large dataset in high dimensions.
+
 \vspace{-1ex}
@@
 \label{section:methodology}
-%\paragraph{Assumptions}
-%Denoting by $C^0(\msi,\msj) $ the space of continous functions between two sets $\msi,\msj$,
 \vspace{-1ex}
-%\sam{Lien avec functional data analysis ? Insister sur le fait que concaténer temps et espace c'est moisi}
+
 We consider in this paper observations which consist in a population of $N$ multivariate time series, for any $j\in[N]$, $s^j \in \rmC^1(\msi_j,\Rset^{d})$.
@@
 Intuitively speaking, the "shape" of a time series $s:\msi\to \Rset^d$ is encoded in its graphs $\msg(s)$ defined as the set $\{(t,s(t)):\eqsp t\in\msi \} $ and not only in its values $s(\msi)=\{s(t):\eqsp t\in\msi \} $ since the time axis is crucial.
-% on the shape as illustrated in Figure\alain{ (mettre ref)}.
 As a motivating use-case, $s^j$ can be the time series of a heartbeat extracted from an individual's electrocardiogram (ECG), see \Cref{fig:transport}.
 The homogeneity in a resulting dataset comes from the fact that humans have similar shapes of heartbeat \cite{ye2012heartbeat,madona2021pqrst}.
- % the shape is also referred as morphology of the time series. 
-  %  [More specifically, the heterogeneity in such dataset is often due to local or glocal dilatation of the time series regarding time, the heartbeat is quicker or slower depending on the individual.
-  %  In maths, denoting by $\mcd(\msi,\msj) $ the set of diffeomorpshim defined on the sets $\msi$ to $\msj$,
-  %   there exist a common pattern $\mathbf{s}_0: \msi \to \Rset^d$ and individual time reparametrisations $ \psi_j\in\mcd(\msi,\msi_j) $ such that,
-  %   for any $j\in[N]$, $s^j=\mathbf{s}_0\circ \psi_j^{-1} $.]
-%  \paragraph{Goals}%\sam{potentiellement changer l'ordre de time and space}
+
 \vspace{-1ex}
@@
 In this paper, we aim to study the inter-individual variability in the dataset by finding a relevant representation of each time series.
-Inspired from the framework of shape analysis \cite{vaillant2004statistics}, addressing similar problems in morphology,
- we suggest to represent each time series' graph $\msg(s^j)$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$, related to a time series $\mathbf{s}_0:\msi \to\Rset^d$, by a diffeomorphism $\phi_j$ on $\Rset^{d+1}$, for any $j\in[N]$,
+Inspired from the framework of shape analysis \cite{vaillant2004statistics}, addressing similar problems in morphology, we suggest to represent each time series' graph $\msg(s^j)$ as the transformation of a reference graph $\msg(\mathbf{s}_0)$, related to a time series $\mathbf{s}_0:\msi \to\Rset^d$, by a diffeomorphism $\phi_j$ on $\Rset^{d+1}$, for any $j\in[N]$,
 \begin{equation}
@@
 
-%First, we introduce the Large Deformation Diffeomorphic Metric Mapping (LDDMM) framework.
-% Then, we explain how to learn a discretization of the graph $\msg(\bfs_0)$ and the diffeomorphisms $(\phi_j)_{j\in[N]} $ by using LDDMM and a gradient descent minimization.
-  %Finally, we tackle the specificity of graph time series by deriving a representation theorem on diffeomorphisms, which enables us to select the kernel needed in LDDMM, and thus, we propose TS-LDDMM.
-
-     %\sam{Parler de l'unicté?}
 \vspace{-1ex}
@@
 Finally, the vectors $(\mathbf{\alpha}_j)_{j\in[N]}$ can be analyzed with any statistical or machine learning tools such as Principal Components Analysis (PCA), Latent Discriminant Analysis (LDA), longitudinal data analysis and so on.
-%As $\bfs_0$ is supposed to be fixed, then the representation of the time series $(s_j)_{j\in[N]}$ boils down to the one of the transformation $(\phi_j)_{j\in[N]}$.
- %In \Cref{section:optimization} the choice of the common time series $\mathbf{s}_0$ will be optimized to have $\phi_j$ of minimal norm. %T is given in the following part.
-
-%  To learn the diffeomorphisms $(\phi_j)_{j\in[N]}$, we propose to use in this paper Large Deformation Diffeomorphic Metric Mapping (LDDMM).
-%   In a nutshell, for any $j\in [N]$, $\phi_j$ corresponds to a differential flow related to a learnable velocity field belonging to a well-chosen Reproducing Kernel Hilbert Space (RKHS).
-% We now describe more precisely LDDMM and in a next step the procedure to learn $\mathbf{s}_0$.
-
-% \sam{to improve ?}
-% Compared to the litterature of Unsupervised Representation Learning (URL) \cite{meng2023unsupervised}, the homogeneity assumption can be quite restrictive, however this is quite common in Shape Analysis \cite{vaillant2004statistics} (SA).
-%     While in URL the goal is to find features accounting for the clustering structure in the heterogene data and to transfer this representation for different ML tasks, in SA the goal is to analyze the inter-individual variability with precision to derive clinical conclusions.
-%     The generalization of the method presented in this paper to heterogene dataset is possible, but relagated for future works.
-  %These velocity fields are then learned by minimizing the norm of this RKHS.
-
-%In all the following, we denote by $\mcd(\mso) $ the set of diffeomorpshim defined on an open set $\mso$ to $\mso$.
-%\sam{Transformer , en : dans les ensembles}
-
-   %\sam{Parler de l'unicté?}
-
-   %Before to specify the choice of the RKHS kernel $K$ and the loss on sets $\mathscr{L}$ in \Cref{section:time_series_specificity}, we first discuss the connection of LDDMM with normalizing flows.
-% to highlight the interest of the LDDMM approach.
-
-%Conversely to NF or CNF, by \eqref{eq:integration}, different momentums parameters raise different final deformations $\phi^{\{v_0\}}$, this injectivity properties offers interpretability of the representation and gives sens to PCA or LDA post-processing as exposed in \Cref{section:experiments}.
-
-   %If a diffeomorpshim $\Phi$ is given and we want to represent it as a deformation $\phi^v$ with $v$ of minimal ernergy, then
-  %$(\alpha_{k,0})_k$ is the only parameters to learn since $(x_{k,0})_k$ is fixed.
-
-%   \begin{algorithm}[tb]
-%     \caption{Geodesic shooting with LDDMM}
-%     \label{alg:example}
-%  \begin{algorithmic}
-%     \STATE {\bfseries Input:} a RKHS's kernel $K$, a loss on sets $L$, $\mathbf{x}=(x_{i,0})_{i\in[T_1]}$, $\mathbf{y}=(y_i)_{i\in[T_2]}$ two sets of $\Rset^{d'}$, initial momentums 
-%     $(\alpha_{k,0})_{k\in [T_1]} \in (\Rset^{d'})^{T_1}$ and a number of gradient iterations $n_{\text{iter}}$.
-
-%     \STATE $f: (\alpha_k)_k\mapsto L\left(\phi^{\{v_0\}}.\mathbf{x},\mathbf{y}\right)+\lambda||v_0||_\msv^2  $ with $v_0$ and $\phi^{\{v_0\}}$
-%      respectively defined by \eqref{eq:def_v0} and \eqref{eq:integration} using control points $\mathbf{x}$ and initial momentums $(\alpha_k)_k $. 
-%      \sam{Peut être à définir dans un autre algo ?}
-%     \STATE $\alpha_{k,0}^*\gets \alpha_{k,0}, \eqsp k\in[T_1]$
-%     \FOR{$i=1$ {\bfseries to} $n_{\text{iter}}$}
-%     \STATE $\alpha_{k,0}^*\gets \text{AdamStep}(f,\alpha_{k,0}^*$) , where $\text{AdamStep}$ is given in appendix (mettre ref)
-%     \ENDFOR
-%     \STATE {\bfseries Return:} $(\alpha_{k,0}^*)_{k\in [T_1]}$
-
-%  \end{algorithmic}
-%  \end{algorithm}
 
 Nevertheless, \eqref{eq:general_optimization_problem} asks to define a kernel and a loss in order to perform geodesic shooting \eqref{eq:relaxation}, which is the purpose of the following subsection.
+
 \vspace{-1ex}
@@
 \vspace{-1ex}
-%\tcr{Mettre l'applicaiton clair pour les times series}
-
 \label{section:time_series_specificity}
@@
 Consequently, we introduce the oriented Varifold distance.
+
 \vspace{-1ex}
@@
     \end{equation}
-     Moreover, for any $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\bar{\gamma}\in  \mcd(\Rset) $, there exists a continously differentiable time series $\bar{s}$ such that 
-     $\msg(\bar{s})= \Pi_{\bar{\gamma},\bar{f}}.\msg(\mathbf{s}_0)$
+    Moreover, for any $\bar{f}\in \rmC^1(\Rset^{d+1},\Rset^d)$ and $\bar{\gamma}\in  \mcd(\Rset) $, there exists a continously differentiable time series $\bar{s}$ such that $\msg(\bar{s})= \Pi_{\bar{\gamma},\bar{f}}.\msg(\mathbf{s}_0)$
 \end{theorem}
-    %\sam{preuve todo, classement des variétés à bord, donne un homeomorphisme sur la restriction}
-    % \begin{proof}
-    %   Let $s:  \msj \to \Rset^d  $ and $\mathbf{s}_0: \msi\to \Rset^d $ be two continuously differentiable time seriess with $\msi=(a,b),\msj=(\alpha,\beta)$ two intervals of $\Rset$.
-    %   % $\msg(\mathbf{s}_0)$ and $\msg(s)$ are two smooth connected 1-dimensional manifold of $\Rset^{d+1}$, thus, by \cite[Appendix
-    %   % Classfying 1-Manifolds]{milnor1965topology}, there exists $\Pi\in \mcd(\Rset^{d+1})$ such that $\msg(s)=\Pi.\msg(\mathbf{s}_0)$.
-    %   %  Therefore, denoting by $\Pi_1: \Rset^{d+1}\mapsto \Rset$, we define $\gamma(t)\triangleq \Pi_1(t,\mathbf{s}_0(t))$ for any $t\in \msi$, then for any $t\in [b,+\infty)$,
-    %   %   \begin{equation}
-    %   %     \gamma(t)=\\Pi_1(b,\mathbf{s}_0(b))+(t-b)\frac{\dd\gamma}{\dd^- t}(b) \eqsp,
-    %   %   \end{equation}
-    %   %   and for any $t\in (-\infty,a]$,
-    %   %   \begin{equation}
-    %   %      \gamma(t)=\Pi_1(a,\mathbf{s}_0(a))+(t-a)\frac{\dd\gamma}{\dd^+ t}(a)\eqsp ,
-    %   %   \end{equation}
-    %   %    where $ \frac{\dd\gamma}{\dd^- t},\frac{\dd\gamma}{\dd^+ t}$ are respectively the left and right derivative.
 
-    %   By setting $\gamma: t\in \Rset \mapsto (\beta-\alpha)(t-a)/(b-a)+\alpha\in \Rset $, we have $ \gamma(\msi)=\msj$ and $\gamma \in \mcd(\Rset) $.
-    %    By defining $f:(t,x)\in\Rset^{d+1}\mapsto x-\mathbf{s}_0(t)+s\circ \gamma(t) $, the map $\Phi_f\in \mcd(\Rset^{d+1})$,
-    %     indeed, its inverse is $\Phi_f^{-1}:(t,x)\in\Rset^{d+1}\mapsto (t,x+\mathbf{s}_0(t)-s(t)) $ and is continuously differentiable.
-    %      Moreover, we have $\Pi_{\gamma,f}.\msg(\mathbf{s}_0)=\{(\gamma(t),s\circ \gamma(t)):\eqsp t\in\msi \}=\msg(s) $.
-
-    %     Let $\bar{f}\in \rmC^0(\Rset^{d+1},\Rset^d)$, $\bar{\gamma}\in  \mcd(\Rset) $ and $\mathbf{s}_0\in \rmC^0(\msi,\Rset^d)$ with $\msi$ an interval of $\Rset$.
-    %     We have :
-    %     \begin{align}
-    %       \Pi_{\gamma,f}.\msg(\mathbf{s}_0)&=\{(\gamma(t),f(t,\mathbf{s}_0(t))),\eqsp t\in \msi \} \\
-    %       &\label{eq:proof1_last_eq}=\{(t,f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right),\eqsp t\in \gamma(\msi) \} \eqsp .
-    %     \end{align}
-    %     By defining $\bar{s}:t\in \gamma(\msi)\to f\left(\gamma^{-1}(t),\mathbf{s}_0(\gamma^{-1}(t))\right) $, we have $\bar{s}\in \rmC^0(\gamma(\msi), \Rset^d) $ by composition of continuous functions
-    %     and $ \msg(\bar{s})= \Pi_{\gamma,f}.\msg(\mathbf{s}_0)$ by \eqref{eq:proof1_last_eq}, which concludes the proof.
-    % \end{proof}
 \begin{remark}
@@
   \label{eq:kernel_TAS}
-      &K_{\msg}((t,x),(t',x'))=\begin{pmatrix}
+  &K_{\msg}((t,x),(t',x'))=
+  \begin{pmatrix}
     c_0K_{\text{time}} & 0 \\
@@
   \label{lemma:choice_of_kernel_V}
-      If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0},
-       there exist $\gamma \in \msd(\Rset) $ and $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ such that $\phi^v=\Psi_\gamma\circ\Phi_f $.
+  If we denote by $\msv$ the RKHS associated with the kernel $K_{\msg}$, then for any vector field $v$ generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0}, there exist $\gamma \in \msd(\Rset) $ and $f\in \rmC^1(\Rset^{d+1},\Rset^d)$ such that $\phi^v=\Psi_\gamma\circ\Phi_f $.
 \end{lemma}
 Instead of Gaussian kernels, other types of smooth kernels can be selected as long as the structure \eqref{eq:kernel_TAS} is respected.
-    % \begin{proof}
-    %   Let $v$ be a vector field generated by \eqref{eq:integration} with $v_0$ satisfying \eqref{eq:def_v0}.
-    %  We remark that the first coordinate of the velocity field $v_\tau$ denoted by $v_\tau^{\text{time}}$ only depends on the time variable $t$ for any $\tau\in[0,1]$.
-    %  Thus, when computing the first coordinate of the deformation $\phi^v$, denoted by $\gamma$, we integrate \eqref{eq:LDDMM_dynamic} with $v_\tau$ replaced by $v_\tau^{\text{time}}$,
-    %   thus $\gamma$ is independant of the variable $x$.
-Moreover, $\gamma\in \mcd(\Rset)$ since a Gaussian kernel induced an Hilbert space $\msv$ satisfying $|f|_V\leq |f|_\infty+ |\dd f|_\infty  $ for any $f\in \msv$ by \citep[Theorem 9]{glaunes2005transport}.
-    %   For the same reason, we have $\phi^v\in \mcd(\Rset^{d+1})$, and thus its last coordinates denoted by $f$ belongs to $\rmC^1(\Rset^{d+1},\Rset^d)$, and by construction $\phi^v=\Psi_\gamma\circ\Phi_f $.
-    % \end{proof
+
 \begin{remark}
@@
   With this choice of kernel, the features associated with the time transformation can be extracted from the momentums $(\alpha_{k,0})_{k\in[\mathbf{n}_0]}\in (\Rset^{d+1})^{\mathbf{n}_0}$ in \eqref{eq:def_v0} by taking the coordinates related to time.
-      However, the features related to the space transformation are not only in the space coordinates since the related kernel $K_{\text{space}}$ depends on time as well.
-      The kernel's representation has been carefully designed to integrate both space and time, while ensuring that time remains independent of space.
+However, the features related to the space transformation are not only in the space coordinates since the related kernel $K_{\text{space}}$ depends on time as well.The kernel's representation has been carefully designed to integrate both space and time, while ensuring that time remains independent of space.
 Initially, we considered separating the spatial and temporal components.
@@
 
-    %  and then we can perform a statistical analysis on the representations $(\psi_j,\phi^j)_{j\in[N]}$,
-    % such as applying PCA decomposition \cite{vaillant2004statistics}, carring out longitudinal analysis \cite{guigui2022parallel} and so on (\sam{mettre ref}).
-    % % Cette hypothèse intéressante car ça permet de gérer des signaux de taille variables, tout les signaux se repréentent comme ça 
-    % % Espace homogène, injectivité de t ,s(t), permet de garder la structure temporel en compte
-    % \sam{Can we show unicity by assuming that $\phi,\psi$ minimize a norm ?}
- % KEOPS tp pour lddmm
-
-%and their empiric samples version $\tilde{G}_j=(\{(t_i,s^j_i), \eqsp i\in [n_i]\})_j $
-
-%  We translate the notation from the continous to discrete by adding a tilde :
-%   $\tilde{G}^j=\triangleq \{ (t_i,\tilde{s}_i^j),\eqsp i\in[n_i] \}$.
-
-% \paragraph{A problem of minimization}
-% In the next section, we will present that diffeomorpshim on space an time can be parametrized as $\phi^{v^S},\phi^{v^T}$
-%  with parameters $(v^S,v^T) \in \Theta_S\times \Theta_T$. Then, pattern will be represented by the parameters $v^S,v^T$ solutions of the following minimization :
-% \begin{align}
-%   \label{eq:minimization}
-%   &\argmin_{v^S \in \Theta_S,\eqsp v^T\in \Theta_T } \mathbf{D}(v^S,v^T)+\lambda \mathbf{N}(v^S,v^T)\eqsp , \\
-%   & \mathbf{D}(v^S,v^T)=L(\phi^{v^S}.G^\msj(\mathbf{s_0}\circ (\psi^{v^T})^{-1}),G^\msj(s)) \eqsp , 
-% \end{align} 
-% where $L$ is a loss on graphs and $\mathbf{N}$ is a term of regularization.
-
-% $s_i: t\in[0,T_0]\mapsto s(t)\in \Rset^d $ is only known through its samples $(S_i^j(t_i)=s_i)_{i\in[N]} $ at timestamps $(t_i)_{i\in[T]}$.
-
-%\sam{améliorer transition}
 \vspace{-1ex}
@@
 
-In practice, we can only access discretized graphs of time series, $(t_i^j,\tsig^j_i)_{i\in[n_j]}$ for any $j\in[N]$, that are potentially of 
-different sizes $n_j$ and sampled at different timestamps $(t_i^j)_{i\in[n_j]}$ for any $j\in[N]$.
- Usual metrics, such as the Euclidean distance, are not appealing as they 
-make the underlying assumptions of equal size sets and the existence of a pairing between points.
- Distances between measures on 
-sets (taking the empirical distribution), such as Maximum Mean Discaprency (MMD) \cite{dziugaite2015training,borgwardt2006integrating}, alleviate those issues; however, MMD only accounts for positional information
-and lacks information about the time evolution between sampled points.
- A classical data fidelity metric from shape analysis 
-corresponding to the distance between \textit{oriented varifolds} associated with curves alleviates this last issue \cite{kaltenmark2017general}.  
-Intuitively, an oriented varifold is a measure that accounts for positional and tangential information about the underlying 
-curves at sample points. More details and information about \textit{oriented varifolds} can be found in \Cref{appendix:varifold}.
+In practice, we can only access discretized graphs of time series, $(t_i^j,\tsig^j_i)_{i\in[n_j]}$ for any $j\in[N]$, that are potentially of different sizes $n_j$ and sampled at different timestamps $(t_i^j)_{i\in[n_j]}$ for any $j\in[N]$.
+Usual metrics, such as the Euclidean distance, are not appealing as they make the underlying assumptions of equal size sets and the existence of a pairing between points.
+Distances between measures on sets (taking the empirical distribution), such as Maximum Mean Discaprency (MMD) \cite{dziugaite2015training,borgwardt2006integrating}, alleviate those issues; however, MMD only accounts for positional information and lacks information about the time evolution between sampled points.
+A classical data fidelity metric from shape analysis corresponding to the distance between \textit{oriented varifolds} associated with curves alleviates this last issue \cite{kaltenmark2017general}.
+Intuitively, an oriented varifold is a measure that accounts for positional and tangential information about the underlying curves at sample points.
+More details and information about \textit{oriented varifolds} can be found in \Cref{appendix:varifold}.
 
-More precisely, given two sets $\msg_0=(g_i^0)_{i\in[T_0]},\msg_1=(g_i^1)_{i\in[T_1]}\in (\Rset^{d+1})^{T_1}$ and a kernel\footnote{$\mathbb{S}^d=\{x\in\Rset^{d+1}:\eqsp |x|=1\}$} $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$
-verifying \citep[Proposition 2 \& 4]{kaltenmark2017general}, for any $\xi\in\{0,1\}$ and $i\in[T_\xi-1]$, denoting the center and length of the $i^{th}$ segment $[g_i^\xi,g_{i+1}^\xi]$ by
-$c_i^\xi = (g_i^\xi + g_{i+1}^\xi)/2$, $l_i^\xi = \| g_{i+1}^\xi-g_{i}^\xi\|$, and 
+More precisely, given two sets $\msg_0=(g_i^0)_{i\in[T_0]},\msg_1=(g_i^1)_{i\in[T_1]}\in (\Rset^{d+1})^{T_1}$ and a kernel\footnote{$\mathbb{S}^d=\{x\in\Rset^{d+1}:\eqsp |x|=1\}$} $k:(\Rset^{d+1} \times \mathbb{S}^d)^2\to \Rset$ verifying \citep[Proposition 2 \& 4]{kaltenmark2017general}, for any $\xi\in\{0,1\}$ and $i\in[T_\xi-1]$, denoting the center and length of the $i^{th}$ segment $[g_i^\xi,g_{i+1}^\xi]$ by $c_i^\xi = (g_i^\xi + g_{i+1}^\xi)/2$, $l_i^\xi = \| g_{i+1}^\xi-g_{i}^\xi\|$, and
 $\overrightarrow{v_i}^\xi = (g_{i+1}^\xi-g_{i}^\xi)/l_i^\xi$, the varifold distance between $\msg_0$ and $\msg_1$  is defined as,
@@
 A pedagogical \href{https://tslddmmapp.streamlit.app}{\underline{online application}} is available to inspect the effect of hyperprameters on geodesic shooting \eqref{eq:integration} and registration \eqref{eq:relaxation}.
-
-%\sam{TO improve by Thibaut
-%\thi{parler du fait qu'on a pas pairing, grace au difféo on conserve la structure}
-%To relax the geodesic problem \Cref{eq:geodesics_original}, the end point of the geodesics is compared to the target through a loss $\mathscr{L}$ as proposed in \eqref{eq:relaxation}.
-%Since we are dealing with discrete time series $\tsig^j$ of different samples size $n_j$, we need a loss on set measures and not only on vectors.
-% A natural choice would be the Maximum Mean Discrepency (MMD) \cite{dziugaite2015training,borgwardt2006integrating}, for any $X=(x_i)_{i\in[T_1]},Y=(y_i)_{i\in[T_2]}\subset Rset^{d'}$, given a kernel $k:\Rset^{d'}\times \Rset^{d'} \to \Rset$,
-% \begin{align}
-%  &\operatorname{MMD}^k(\mu_X,\mu_Y)=\frac{1}{T_1^2}\sum_{i,j=1}^{T_1} k(x_i,x_j)+\frac{1}{T_2^2}\sum_{i,j=1}^{T_2} k(y_i,y_j) \\
-%  &-\frac{2}{T_1T_2}\sum_{i=1}^{T_1}\sum_{j=1}^{T_2} k(x_i,y_j) \eqsp , 
-% \end{align}
-% where $\mu_X=\sum_{i=1}^{T_1} \delta_{x_i}/T_1$, $\mu_Y=\sum_{j=1}^{T_2} \delta_{y_j}/T_2$.
-%However, this distance is agnostic of the time evolutions in time series' graph $(t_i^j,\tsig^j_i)_{i\in[n_j]}$, there is no orientation of the set taken into account.
-% That is why, we prefer to choose the Varifold distance $d_V(\cdot,\cdot)$ \cite{kaltenmark2017general} which was specifically designed for oriented curves.
-%  The Varifold distance extends the MMD by adding tangent vectors to the input:
-%   instead of considering a set $X=(x_i)_{i\in[T_1]}$, we extend the input as $\overrightarrow{X}=(x_i,v_i)_{i\in[T_1]}$ where $v_i$ is the tangent vector at $x_i$ such that
-%    \begin{equation}
-%      d_V(X,Y)=\operatorname{MMD}^{K_p}(\mu_{\overrightarrow{X}},\mu_{\overrightarrow{X}}) \eqsp ,
-%    \end{equation}
-%    where in practice the tangent vectors are taken as discrete derivative and $K_p$ is a kernel on $\Rset^{2d'}$ such that for any $(x,v),(y,w)\in (\Rset^{2d})^2$, 
-%    we have, 
-%    \begin{equation}
-%      K_p((x,v),(y,w))=k(x,v)k_T(v,w) \eqsp ,
-%    \end{equation}
-%   where $k_T $ is a kernel on the tangent vector space.
-
-%\sam{Parler de méthode adaptatif ici}
-
-%Wavelet-Based Multiscale Flow For Realistic Image Deformation in the Large Diffeomorphic Deformation Model Framework
-% 3.1 +3.2 + specialité de la décomposition (t,f(t,x))
-% \section{Specificity to time series}
-% We are looking to time seriess 
-% Each time series $s: t\in[0,T]\mapsto s(t) $ is only known through its samples $(s(t_i)=s_i)_{i\in[N]} $ at timestamps $(t_i)_{i\in[N]}$.
-%is described by its graph $G(s)=\{(y,s(t)), \} $
-
-% \section{optimzation
-% % We are not doing Pre-ESPAce car l'on s'intéresse à la reparamétrisation en temps, au lieu de calculer la transfo en temps après avoir
-% % push en espace, on le fait avant pour que tout soit défini depuis s_0, méthode itérative car ça marche mieux, moins de calcul de gradient
-% %Natural smoothing with TTS
-% \label{section:optimization}
-%     \begin{center}
-%         \begin{tikzpicture}
-%             \node at (0,0) {$G_f^I$};
-%             \draw [-stealth](0.3,0) -- (1.3,0);
-%             \node at (1.6,0) {$G_f^J$};
-%             \draw [-stealth](1.6,-0.3) -- (1.6,-1.3);
-%             \node at (1.6,-1.6) {$G_g^J$};
-%             \draw [-stealth](1.3,-1.6) -- (0.3,-1.6);
-%             \node at (0,-1.6) {$G_g^I$};
-%             \draw [-stealth](0,-1.3) -- (0,-0.3);
-
-%             \node at (0.8,0.3) {$T$};
-%             \node at (0.8,-1.9) {$T^{-1}$};
-%             \node at (1.9,-0.8) {$S_J$};
-%             \node at (-0.4,-0.8) {$S_I^{-1}$};
-%         \end{tikzpicture}
-%         \end{center}
-
-%         where $S_I^{-1} = T^{-1} \circ S_J^{-1} \circ T$
 
@@
 This issue can be addressed through interpolation, but this approach is not always reliable in sparse and irregular sampling scenarios.
- Most FDA approaches, as seen in \cite{shang2014survey,yu2017principal,warmenhoven2021pca}, address this challenge using interpolation or basis function expansion
+Most FDA approaches, as seen in \cite{shang2014survey,yu2017principal,warmenhoven2021pca}, address this challenge using interpolation or basis function expansion.
 In summary, FDA methods typically separate space and time representations for continuous objects, whereas TS-LDDMM algorithm maintain a discrete-to-discrete analysis, inherently integrating both space and time representations.
@@
 Extending Neural ODEs, Neural Stochastic Differential Equations (Neural SDEs) introduce regularization effects \cite{liu2019neural}, although optimization remains challenging.
- Leveraging techniques from continuous-discrete filtering theory, Ansari et al. \cite{ansari2023neural} applied successfully Neural SDEs to irregular time series.
+Leveraging techniques from continuous-discrete filtering theory, Ansari et al.
+\cite{ansari2023neural} applied successfully Neural SDEs to irregular time series.
 Oh \textit{et al.} \cite{oh2024stable} improved these results by incorporating the concept of controlled paths into the drift term, similar to how Neural CDEs outperform Neural ODEs.
@@
 
-Although TS-LDDMM 
-is already competitive for classification, its performances can be leveraged on 
-more heterogeneous datasets using a hierarchical clustering extension, which is relegated for future work. 
-%   While TS-LDDMM performs well on shape-based datasets, it assumes a certain degree of homogeneity with the existence of a reference time series graph$\msg_0$.
-%  Extending TS-LDDMM to more heterogeneous datasets using clustering is ongoing work.
+Although TS-LDDMM is already competitive for classification, its performances can be leveraged on more heterogeneous datasets using a hierarchical clustering extension, which is relegated for future work.
 
@@
 In future work, adaptive methods are expected to be developed to provide a more user-friendly interface.
-  %\color{red} rappeler les limites de la méthodes : petite dimension à cause des kernels, tuner les hyper paramètres correctement, version adaptative}
-% \paragraph*{Conclusion}
-% \vspace{-1ex}
-
-%\section{Societal impact}
-%This paper presents work whose goal is to advance the field of Machine Learning. There are many potential societal consequences of our work, none which we feel must be specifically highlighted here.
-% \section{Acknowledgements}
-
-%The TS-LDDMM feature representation proposed in this paper enables us to study the inter-individual variability of shapes in a time series dataset with high interpretability.
-%Its unsupervised aspect is user-friendly and enables knowledge transfer for different supervised tasks such as classification.
-%Although the method is already competitive,
-% its capacity can be leveraged on more heterogeneous datasets using a hierarchical clustering extension.
-% This is relegated to future works.
-
-%Applying NF or CNF to our
- %  framework with high dimensional time series is relagated for future works.
-
-% \begin{figure}[ht]
-% \vskip 0.2in
-% \begin{center}
-% \centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
-% \caption{Historical locations and number of accepted papers for International
-% Machine Learning Conferences (ICML 1993 -- ICML 2008) and International
-% Workshops on Machine Learning (ML 1988 -- ML 1992). At the time this figure was
-% produced, the number of accepted papers for ICML 2008 was unknown and instead
-% estimated.}
-% \label{icml-historical}
-% \end{center}
-% \vskip -0.2in
-% \end{figure}
-
-% \begin{algorithm}[tb]
-%    \caption{Bubble Sort}
-%    \label{alg:example}
-% \begin{algorithmic}
-%    \STATE {\bfseries Input:} data $x_i$, size $m$
-%    \REPEAT
-%    \STATE Initialize $noChange = true$.
-%    \FOR{$i=1$ {\bfseries to} $m-1$}
-%    \IF{$x_i > x_{i+1}$}
-%    \STATE Swap $x_i$ and $x_{i+1}$
-%    \STATE $noChange = false$
-%    \ENDIF
-%    \ENDFOR
-%    \UNTIL{$noChange$ is $true$}
-% \end{algorithmic}
-% \end{algorithm}

### 2024-10-25 Robustness--Classification/url/model_run_time.py
clean the git

@@ -1,0 +1,254 @@
+#-*- coding:utf-8 -*-
+
+import os
+import sys
+import copy
+import pickle
+import random
+import numpy as np
+import pandas as pd
+import sklearn
+from sklearn.metrics import *
+import time
+
+
+from tqdm import tqdm
+
+import torch
+import torch.optim as optim
+from torch import nn
+from torch.utils.data import  DataLoader
+
+# setup seed
+def seed_everything(seed):
+    os.environ['PYTHONHASHSEED'] = str(seed)
+    np.random.seed(seed)
+    random.seed(seed)
+    torch.manual_seed(seed)
+    torch.random.manual_seed(seed)
+    torch.cuda.manual_seed(seed)
+    torch.cuda.manual_seed_all(seed)
+    torch.backends.cudnn.deterministic = True
+    torch.backends.cudnn.benchmark = True
+
+SEED = 0
+seed_everything(SEED)
+
+use_cuda = torch.cuda.is_available()
+device = torch.device("cuda" if use_cuda else "cpu")
+
+
+from utils import get_data, get_coeffs
+from model import ists_dataset, ists_classifier, train, evaluate 
+
+
+model_name_list = [
+    'lstm',
+]
+
+
+if not os.path.exists('../time_results'):
+    os.mkdir('../time_results')
+    
+    
+def run_model(data_name, missing_rate, model_name, model_config, EPOCHS=100, SEED=SEED, CHECK=False):
+    print(data_name, missing_rate, model_name)
+
+    # check exist
+    out_name = '_'.join([data_name, str(missing_rate), model_name, str(SEED)])
+    out_path = '../time_results/{}/{}/{}'.format(data_name, str(missing_rate), out_name)
+
+    if CHECK & os.path.exists(out_path):
+        return None
+    
+    # setup out path
+    if not os.path.exists('../time_results/{}'.format(data_name)):
+        os.mkdir('../time_results/{}'.format(data_name))
+    
+    # load data
+    X_missing, X_mask, X_delta,Y,train_idx,valid_idx,test_idx = get_data(data_name,missing_rate=missing_rate)
+    X_missing,X_mask,X_delta = torch.tensor(X_missing),torch.tensor(X_mask),torch.tensor(X_delta)
+    num_data = X_missing.shape[0]
+    seq_len = X_missing.shape[1]
+    num_dim = X_missing.shape[2]
+    num_class = len(np.unique(Y))
+    
+    # set batch_size by the number of data
+    batch_size = 2**4
+    for i in range(4,5): #8
[…]

### 2024-10-25 src/barycenter.py

@@ -4,9 +4,9 @@ from jax import vmap,grad,jit
 from jax.tree_util import Partial
 import optax
 
-from src.lddmm import LDDMMLoss,TimeLDDMMLoss
+from src.lddmm import LDDMMLoss,TimeLDDMMLoss,Shooting
 from src.loss import VarifoldLoss
-from src.optimizer import Optimizer, BatchBarycenterOptimizer,BatchBarycenterTimeOptimizer
+from src.optimizer import Optimizer, BatchBarycenterOptimizer,BatchBarycenterTimeOptimizer,BatchIteratedBarycenterOptimizer
 
 ####################################################################################################################################
 ####################################################################################################################################
@@ -64,6 +64,24 @@ def batch_varifold_barycenter_registration(batched_sigs,batched_sigs_masks,Kv,Kl
     dataloss = VarifoldLoss(Kl)
     return batch_barycenter_registration(batched_sigs,batched_sigs_masks,Kv,dataloss,init,init_mask,time_initializer,niter,optimizer,gamma_loss,nt,deltat,verbose)
 
+def batch_iterated_barycenter_registration(batched_sigs,batched_sigs_masks,Kv,dataloss:callable,init=None,init_mask=None,niter=400, update_interval = 100, optimizer = optax.adam(learning_rate=0.1),gamma_loss =0.,nt=10,deltat=1.0,verbose=True):
+    if init is None: 
+        sigs_size = np.sum(batched_sigs_masks,axis=2)
+        median_length_idx = np.unravel_index(np.argsort(sigs_size.flatten(),kind='stable')[sigs_size.shape[0]//2],sigs_size.shape)[:-1]
+        q0 = batched_sigs[median_length_idx]
+        q0_mask = batched_sigs_masks[median_length_idx]
+    else: 
+        q0 = init
+        q0_mask = init_mask
+
+    #batched_p0 = jnp.zeros_like(batched_sigs,dtype = jnp.float32)
+    batched_p0 = jnp.zeros((batched_sigs.shape[0],batched_sigs.shape[1],*q0.shape))
+    bloss = BarycenterLDDMMLoss(Kv,dataloss,gamma_loss,nt,deltat)
+    shoot = Shooting(Kv,nt,deltat)
+    if niter <= update_interval:
+        raise ValueError("niter must be greater than update interval")
[…]

### 2024-10-25 src/optimizer.py

@@ -192,4 +192,44 @@ def BatchBarycenterOptimizer(loss:callable,niter=100,optimizer = optax.adam(lear
     return f 
 
 
+def BatchIteratedBarycenterOptimizer(shoot:callable,loss:callable,niter=200, update_interval =100, optimizer = optax.adam(learning_rate=0.1),verbose=True):
+    def f(batched_p0,q0,q0_mask,batched_q1,batched_q1_mask):
+        
+        n_batches,batch_size = batched_p0.shape[0],batched_p0.shape[1]
+        batched_p,q = batched_p0, q0
+        opt_state = optimizer.init((batched_p))
+        jit_loss = jit(value_and_grad(loss))
+
+        def step(batched_p,q,q_mask,batched_q1,batched_q1_mask,opt_state):
+
+            p_grad = np.zeros_like(batched_p)
+            loss_value = 0
+
+            for i,(p,q1,q1_mask) in enumerate(zip(batched_p,batched_q1,batched_q1_mask)):
+                t_loss_value,grads = jit_loss(p,q,q_mask,q1,q1_mask)
+                p_grad[i] = grads[0]
+                loss_value += t_loss_value
+
+            loss_value /= (n_batches*batch_size)
+            updates, opt_state = optimizer.update(p_grad, opt_state,batched_p)
+            batched_p = optax.apply_updates(batched_p,updates)
+
+            return batched_p, opt_state, loss_value
+        
+
+        for i in range(niter): 
+            batched_p,opt_state,loss_value = step(batched_p,q,q0_mask,batched_q1,batched_q1_mask,opt_state)
+            if verbose: 
+                if (i==0)+((i+1)%10 == 0):
+                    print("iteration: ",i+1,"/",niter, " -- loss: ", "{:0.2f}".format(loss_value))
+                if ((i+1)!=niter)*(i!=0)*((i+1)%update_interval == 0):
+                    print("...Updating barycenter...")
+                    pbar = np.mean(batched_p,axis=(0,1))
+                    _,q = shoot(pbar,q,q0_mask)
+                    batched_p = batched_p - pbar[None,:,:]        
+        return batched_p,q
+    return f 
+
+
+
 

### 2024-10-25 NEURIPS24/paper.tex
change time seriesgraph

@@
 
-However, a graph time series transformation by a general diffeomorphism is not always a graph time series, see e.g.
-\Cref{fig:diffeo}, thus a graph time series is more than a simple curve \cite{glaunes2008large}.
+However, a time series graph transformation by a general diffeomorphism is not always a time series graph, see e.g.
+\Cref{fig:diffeo}, thus a time series graph is more than a simple curve \cite{glaunes2008large}.
 Our contributions arise from this observation: we specify the class of diffeomorphisms to consider and show how to learn them.
